### 2014-01-30 Images/Lineups/LineupGeneration.Rnw
RNW files for stimuli to generate a single file with all stimuli.

\documentclass{article}
\usepackage{fullpage}

\begin{document}

<<setup, out.width="5in", out.height="4in", fig.width=10, fig.height=8, fig.align="center", echo=FALSE, results='asis', cache=TRUE, message=FALSE, warning=TRUE>>=
setwd("../../")
library(nullabor)
library(ggplot2)
library(plyr)
library(Hmisc)

source("./Code/theme_lineup.R")
source("./Code/GenerateData.R")

set.seed(31415926)

# Linear trend alone - no outliers
N <- 10
seeds <- c(699181, 190785, 746705, 938857, 256246, 
           130991, 829750, 107019, 310685, 968407, 
           982098, 247471, 658815, 828038, 956567, 
           706934,  74904, 743373, 955470, 339582, 
           230370, 594429, 959280, 677468, 301141, 
           523820, 397938, 952596, 726989, 247199)
# seeds <- sample(0:1e6, N) + sample(0:1e3, N) + sample(0:100, N)
answers <- data.frame()

colors <- c("#882E72", "#D92120", "#4EB265", "#3F4EA1", 
            "#E7742F", "#1965B0", "#B1B343", "#4683C1", "#DFA53A")

pics.set <- as.character(1:(N*2))
pics.set[nchar(pics.set)==1] <- paste("0", pics.set[nchar(pics.set)==1], sep="")
names <- data.frame(filename = sample(paste("./Images/Lineups/", rep(1:3, each=N*2), "-", rep(pics.set, times=3), ".png", sep=""), 6*N), idx=1:(6*N), stringsAsFactors=FALSE)
names$title <- gsub(".png", "", gsub("./Images/Lineups/", "", names$filename, fixed=TRUE), fixed=TRUE)

plots <- as.list(rep(NA, length(names$filename)))
names(plots) <- names$filename

permute.var <- c("x", "y")
for(i in 1:N){
  j <- 3*i-2
  set.seed(seeds[i])
  dframe <- linear.trend()

  # Slope alone
  pos.x <- sample(1:20, 1)
  filename <- names$filename[j]
  pv <- sample(permute.var, 1)
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y), size=3) + 
                  facet_wrap(~.sample) +
                  theme_lineup() + 
                  ggtitle(names$title[j])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=NA, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="Slope"))

  # Color alone
  pos.x <- sample(1:20, 1)
  pv <- "group"
  filename <- names$filename[j+1]
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j+1]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y, color=factor(group)), size=3) + 
                  facet_wrap(~.sample) + 
                  scale_color_manual(values=colors[c(2,8)]) +
                  theme_lineup()+ 
                  ggtitle(names$title[j+1])
  answers <- rbind(answers, data.frame(name=filename, target1=NA, target2=pos.x, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="Color"))

  # Slope vs. Color
  pos.x <- sample(1:20, 1)
  pos.y <- sample(c(1:20)[which(1:20!=pos.x)], 1)
  pv <- sample(permute.var, 1)
  filename <- names$filename[j+2]
  lineupdata <- permute.groups2(lineup(null_permute(pv), dframe, pos=pos.x), 
                                ngroups=3, pos=pos.y)

plots[[j+2]] <- ggplot(data=lineupdata) + 
                geom_point(aes(x=x, y=y, color=factor(group.k)), size=3) + 
                facet_wrap(~.sample) + 
                scale_color_manual(values=colors[c(3,5,6)]) +
                theme_lineup() + 
                ggtitle(names$title[j+2])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=pos.y, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="ShapeColor"))
}

# write.csv(answers, "./Images/Lineups/NoOutliersKey.csv", row.names=FALSE)

# Linear trend with Outliers
N <- 10
set.seed(271828)
seeds <- c(488462,  89012,  68718, 875098, 790630,
           995942, 307213, 635549, 380455, 971611,
            90250, 687530, 557990, 873571, 611075,
           100338, 214593, 365595, 609200, 944618,
           578770, 710917, 117886, 458621, 307274,
           749423, 231016, 321914, 222946, 627623)
# seeds <- sample(0:1e6, N) + sample(0:1e3, N) + sample(0:100, N)
# answers <- data.frame()

colors <- c("#882E72", "#D92120", "#4EB265", "#3F4EA1", 
            "#E7742F", "#1965B0", "#B1B343", "#4683C1", "#DFA53A")
permute.var <- c("x", "y")
for(i in 1:N){
  set.seed(seeds[i])
  j <- 28+3*i
  dframe <- linear.trend(n=50, n.outliers=4, sd=4)

  # Slope alone
  pos.x <- sample(1:20, 1)
  filename <- names$filename[j]
  pv <- sample(permute.var, 1)
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y), size=3) + 
                  facet_wrap(~.sample) +
                  theme_lineup() + 
                  ggtitle(names$title[j])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=NA, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="Shape"))

  # Color alone
  pos.x <- sample(1:20, 1)
  pv <- "group"
  filename <- names$filename[j+1]
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j+1]] <- ggplot(data=lineupdata) + 
                    geom_point(aes(x=x, y=y, color=factor(group)), size=3) + 
                    facet_wrap(~.sample) + 
                    scale_color_manual(values=colors[c(3,5,6)]) +
                    theme_lineup()+ 
                    ggtitle(names$title[j+1])
  answers <- rbind(answers, data.frame(name=filename, target1=NA, target2=pos.x, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="Color"))

  # Slope vs. Color
  pos.x <- sample(1:20, 1)
  pos.y <- sample(c(1:20)[which(1:20!=pos.x)], 1)
  pv <- sample(permute.var, 1)
  filename <- names$filename[j+2]
  lineupdata <- permute.groups2(lineup(null_permute(pv), dframe, pos=pos.x), 
                                ngroups=4, pos=pos.y)

  plots[[j+2]] <- ggplot(data=lineupdata) + 
                    geom_point(aes(x=x, y=y, color=factor(group.k)), size=3) + 
                    facet_wrap(~.sample) + 
                    scale_color_manual(values=colors[c(3,5,6,1)]) +
                    theme_lineup() +
                    ggtitle(names$title[j+2])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=pos.y, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="ShapeColor"))
}
plots <- plots[order(names(plots), decreasing=F)]
answers$name <- as.character(answers$name)
answers <- answers[order(answers$name, decreasing=F),]

for(i in 1:length(plots)){
  print(plots[[i]])
#   cat("/clearpage")
}
@

\end{document}

### 2014-01-30 Images/VisualSearch/VisualSearchGeneration.Rnw
RNW files for stimuli to generate a single file with all stimuli.

\documentclass[letterpaper]{article}
\usepackage{fullpage}

\begin{document}

<<setup, out.width=".45\\textwidth", out.height=".45\\textwidth", fig.width=6, fig.height=6,  echo=FALSE, results='asis', cache=TRUE, message=FALSE, warning=TRUE>>=
setwd("../../")
# Given a 5x5 grid with squares 1-25 
# i.e. 
# | 1| 2| 3| 4| 5|
# | 6| 7| 8| 9|10|
# |11|12|13|14|15|
# |16|17|18|19|20|
# |21|22|23|24|25|

# How quickly can someone locate the target in a group of 24 distractors?

library(ggplot2)
library(reshape2)
library(plyr)

# define coordinate grid
rect.coords <- data.frame(num=1:25, xstart=(0:24)%%5, xend=(0:24)%%5+1, ystart=5-floor((0:24)/5)-1, yend=5-floor((0:24)/5))

# function to prevent two neighboring squares from being filled in
is.adjacent <- function(samp){
  (abs(diff(samp))==1 & diff(floor((samp-1)/5))==0) || 
   (diff((samp-1)%%5)==0 & abs(diff(floor((samp-1)/5)))==1) || 
   (diff(samp)==0)
}

# List of all possible choices in which the squares are non-adjacent
plot.choices <- expand.grid(1:25, 1:25)
plot.choices <- plot.choices[!apply(plot.choices, 1, is.adjacent),]

# data frame for all pictures together (to make the key)
overall <- data.frame()

# generate 25 data sets
set.seed(90909837)
plots <- as.list(rep(NA, 25))
for(i in 1:25){
  # sample 25 different stimuli sets
  stimuli <- plot.choices[sample(1:nrow(plot.choices), 25, replace=FALSE),]

  # replace one of them with stimuli # 13
  ans <- sample(c(1:12, 14:25), 1)
  stimuli[ans,] <- stimuli[13,]

  # Make stimuli into a data frame with each square filled in as a separate row
  stimuli <- melt(stimuli)
  stimuli$variable <- rep(1:25, times=2)
  names(stimuli) <- c("plot", "coords")

  # save answer
  stimuli$ans <- ans

  # make nice labels
  stimuli$plot <- as.character(stimuli$plot)
  stimuli$plot[which(stimuli$plot=="13")] <- "TARGET"
  stimuli$plot[c(14:25,39:50)] <- rep(as.character(13:24), times=2)
  stimuli$plot <- factor(stimuli$plot,levels=c(as.character(1:12), "TARGET", as.character(13:24)))

  # add in coordinate information for plotting
  stimuli <- cbind(stimuli, rect.coords[stimuli$coords,])

  # save trial number for answer key later
  stimuli$trial <- i
  overall <- rbind(overall, stimuli)

  # save plot
  plots[[i]] <- ggplot() + 
    geom_rect(data=rect.coords, aes(xmin=xstart, ymin=ystart, xmax=xend, ymax=yend), fill="#FFFFFF", colour="#000000") +
    geom_rect(data=stimuli, aes(xmin=xstart, ymin=ystart, xmax=xend, ymax=yend), fill="black", inherit.aes=FALSE) +
    facet_wrap(~plot) +
    coord_fixed() + 
    theme(axis.ticks=element_blank(), 
          panel.grid=element_blank(), 
          plot.background=element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank()) +
    ggtitle(paste("VST-", i, sep=""))
}

key <- ddply(overall, .(trial), summarize, ans=unique(ans))
write.csv(key, "./Images/VisualSearch/VST-Key.csv", row.names=FALSE)

for(i in 1:length(plots)){
  plots[[i]]
#   cat("/clearpage")
}
@

\end{document}

### 2014-01-30 Code/theme_lineup.R

@@ -4,5 +4,5 @@ theme_lineup <- function(base_size = 12, base_family = ""){
   theme_bw(base_size = base_size, base_family = base_family) %+replace% 
     theme(legend.position="none", 
           axis.text=element_blank(),
-#           axis.title=element_blank(),
+          axis.title=element_blank(),
           axis.ticks=element_blank())}

### 2014-09-18 Code/Analysis.R
Dissertation Chapter, Analysis code, and necessary images

@@ -1,0 +1,85 @@
+# setwd("~/Dropbox/GraphicsGroup/TestingVisualAptitude/")
+
+ans <- read.csv(paste(datadir, "VisualGraphicsData.csv", sep=""), stringsAsFactors=F)
+
+key <- ans[1,]
+key[,1:19] <- NA
+# Visual Search
+key[,20:44] <- c(17, 12, 15, 14, 8, 21, 6, 3, 21, 22, 4, 23, 9, 8, 22, 13, 2, 19, 16, 18, 7, 16, 12, 1, 7)
+# Lineups 1
+key[,45:64] <- c(6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17)
+# Card  Rotation
+key[,65:224] <- c("d", "s", "s", "d", "d", "s", "d", "s",  "s", "s", "s", "d", "s", "s", "s", "s", 
+                  "s", "d", "d", "d", "s", "s", "s", "d",  "s", "s", "d", "s", "d", "d", "d", "s",
+                  "d", "s", "d", "d", "s", "s", "d", "s",  "s", "d", "s", "s", "s", "s", "d", "d", 
+                  "s", "d", "s", "d", "d", "s", "s", "s",  "d", "d", "s", "s", "d", "s", "d", "d", 
+                  "d", "d", "s", "s", "d", "s", "s", "d",  "s", "d", "d", "s", "d", "d", "s", "s",
+                  "s", "s", "d", "d", "s", "s", "d", "d",  "s", "d", "d", "d", "s", "s", "s", "s",
+                  "d", "d", "s", "s", "s", "s", "s", "d",  "s", "d", "s", "s", "d", "d", "d", "s",
+                  "s", "s", "s", "d", "d", "d", "s", "s",  "d", "s", "s", "d", "s", "d", "d", "d",
+                  "s", "s", "d", "d", "d", "d", "d", "s",  "s", "s", "s", "s", "d", "d", "s", "s",
+                  "s", "s", "d", "d", "d", "d", "d", "s",  "s", "d", "d", "d", "s", "s", "s", "s")
+# Lineups 2
+key[,225:244] <- c(15, 15, 15,  4,  6,   15,  6, 15,  4,  6,   6,  6,  4,  4, 15,   4, 15, 15, 15, 15)
+
+# Figure Classification
[…]

### 2014-09-18 Code/GenerateData.R

@@ -1,37 +1,0 @@
-require(plyr)
-linear.trend <- function(n=30, sd.data=5, n.outliers=0){
-  a <- sample(c(-1, 1), 1)*rnorm(1, 3)
-  b <- rnorm(1)
-  df <- data.frame(x=seq(-5, 5, length.out=n))
-  df$y <- rnorm(n, a*df$x+b+rnorm(n, sd=sd.data))
-  df$outlier <- FALSE
-  df$group <- as.numeric((df$x+rnorm(n, sd=sd.data/5))>0)+1   
-  df$x <- df$x+rnorm(n, sd=0.01)
-  df$z <- rnorm(1, 1)*df$x+rnorm(1, 1)*df$y+rnorm(n, sd=0.01)
-  if(n.outliers>0){
-    # select outlier from the first 1/5 of the dataset or the last 1/5 of the dataset
-    idx <- c(1:n)
-    cutoff <- sample(c("low", "high"), 1)
-    outlier.vals <- idx[(cutoff=="low" & idx<=n/5) | (cutoff=="high" & idx>=4*n/5)]
-    outlier <- sample(outlier.vals, n.outliers)
-    replace.vals <- idx[(cutoff!="low" & idx<=n/5) | (cutoff!="high" & idx>=4*n/5)]
-    replace <- sample(replace.vals, n.outliers)
-    df$outlier[outlier] <- TRUE
-    df$y[outlier] <- jitter(df$y[replace])
-    df$group[outlier] <- 3
-  }
-
-  return(df)
-}
-
-permute.groups2 <- function(lineupdata, ngroups=3, pos=sample(1:20, 1)){
-  ddply(lineupdata, .(.sample), function(df){
-    dst <- dist(df[,c("x", "y", "z")])
-    if(sum(df$.sample==pos)==0){
-      df$group.k = cutree(hclust(dst, method="complete"), round(nrow(df)/(ngroups)))%%ngroups+1
-    } else {
-      df$group.k = cutree(hclust(dst, method="complete"), ngroups)
-    }
-    df
-  })
-}

### 2014-09-18 Paper/Draft1/TestingVisualAptitude.Rnw
Dissertation Chapter, Analysis code, and necessary images

\documentclass{article}
%---------------------------------------------------
\usepackage{color}
\usepackage[dvipsnames,svgnames]{xcolor}
\usepackage{wrapfig,float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{url}
\usepackage{ulem}
\usepackage[section]{placeins}
\usepackage{sidecap}
\usepackage{multirow}
\usepackage{bbm}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}

% Bibliography without numbers or labels
\usepackage{natbib}
\bibliographystyle{apa}

%Optional Package to add PDF bookmarks and hypertext links
\usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
%---------------------------------------------------

%---------------------------------------------------
%                 Editing Commands
\newcommand{\done}[2][inline]{\todo[color=SpringGreen, #1]{#2}}  % for todos that have been seen and dealt with
\newcommand{\meh}[2][inline]{\todo[color=White, #1]{#2}}   % for todos that may no longer be relevant 
\newcommand{\comment}[2][inline]{\todo[color=SkyBlue, #1]{#2}} % for comments that may not be "to-do"s
\newcommand{\mcomment}[1]{\todo[color=SkyBlue]{#1}} % for margin comments
\newcommand{\newtext}[1]{\todo[inline, color=White]{ \color{OliveGreen}{#1}}} % new text - not necessarily something to be done
\newcommand{\move}[1]{\todo[inline, color=Lime]{#1}} % new to do item
%
%---------------------------------------------------

%---------------------------------------------------
%             Set Picture Locations
\graphicspath{{Figure/}{Images/}}
%
%---------------------------------------------------

<<VisualAptitude-setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE>>=
rm(list=ls())
wd <- getwd()

opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=TRUE, concordance=TRUE, autodep=TRUE, root.dir="../")
datadir <- "../../Data/"
imgdir <- "Figure/"
codedir <- "../../Code/"

options(replace.assign=TRUE,width=70,scipen=3)
require(knitr)

library(reshape2)
suppressMessages(library(ggplot2))
library(plyr)
suppressMessages(library(gridExtra))
source(paste(codedir, "Analysis.R", sep=""))
@

\title{Spatial Reasoning and Statistical Graphics}
\author{Susan VanderPlas, Heike Hofmann} % , Stephanie De Graaf, Hillary Cheney ??

\begin{document}
\maketitle

\section{Introduction}
Relevant literature: 
\begin{itemize}
\item \citet{shah1995conceptual} showed that spatial ability was not correlated with accuracy on a simple two-dimensional line graph description task, but that mathematical ability was correlated with accuracy. 
\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same. 
\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
\end{itemize}

Lineups depend on the ability to search for a signal amid distractors (Visual Search Task) and the ability to infer patterns from stimuli (Pattern Recognition task). Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation task) and mentally manipulate graphs (paper folding task). By breaking the lineup task down into component parts, we can correlate lineup performance with similar cognitive factor tests to determine where additional variation in skill level factors into performance differences. In addition, we can correlate previous experiences (science-based major, research experience, Auto-CAD skills) with performance to explore the effect that participant experience has on lineup performance.

\section{Methods}
Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete). Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section. This provides more discrimination between high scorers and prevents score compression at the top of the range. 
\begin{itemize}
\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors. An example is shown in figure \ref{fig:VST}. 
\begin{figure}[htbp]\centering
\includegraphics[width=.7\textwidth]{VisualSearch}
\caption[Visual Search Task]{Visual Search Task. Participants are instructed to find the plot numbered 1-24 which matches the plot labeled "Target". Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
\end{figure}
\item Paper Folding Task: tests participants' ability to visualize and mentally manipulate figures in three dimensions. Associated with the ability to extrapolate symmetry and reflection over multiple steps. An example is shown in figure \ref{fig:paperfolding}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{paperfolding}
\caption[Paper Folding Task]{Paper Folding Task. Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure. Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
\end{figure}
\item Card Rotation Task: tests participant's ability to rotate objects in two dimensions to distinguish between left-hand and right-hand versions of the same figure. Tests spatial reasoning ability and mental rotation skills. An example is shown in figure \ref{fig:cardrotation}.
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{cardrotation}
\caption[Card Rotation Task]{Card Rotation Task. Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line. Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
\end{figure}
\item Figure Classification Task: tests participant's ability to extrapolate rules from provided figures. This task is associated with visual reasoning capabilities and we expect that it should correlate with the ability to pick out a signal plot from a lineup. An example is shown in figure \ref{fig:figureclassification}. 
\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{figureclassification}
\caption[Figure Classification Task]{Figure Classification Task. Participants classify each figure in the second row as belonging to group 1, 2, or 3 (if applicable). Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
\end{figure}
\end{itemize}

Between cognitive tasks, participants will also complete three blocks of 20 lineups each. These lineups have been previously tested \citep{hofmann2012graphical} and include some null lineups (i.e. lineups without a target plot). Participants will have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.

\begin{figure}[htbp]\centering
\includegraphics[width=.9\textwidth]{lineup}
\caption[Sample Lineup]{A sample lineup. Participants are instructed to choose the plot which appears most different from the others. In this lineup, plot 13 is the target. }\label{fig:lineup}
\end{figure}

In addition to these tests, participants will complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major. These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.

\section{Results}
We have tested 18 students, and plan to increase sample size once the summer semester begins. Preliminary results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}. Scoring for all of the tests was calculated such that with random guessing, the expected score on a problem would be 0; thus, for a test with possible answers A, B, C, D, and E, the score would be calculated by subtracting $1/4*\text{\# wrong answers}$ from the total number of correct answers. This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.

<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=4>>=
lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Video Game Hrs/Wk", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
lineup.summary$variable <- factor(lineup.summary$variable, levels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Video Game Hrs/Wk"))

qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=4)

lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience"))
lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
lineup.summary.categorical$value[lineup.summary.categorical$value=="m"] <- "Male"
lineup.summary.categorical$value[lineup.summary.categorical$value=="y"] <- "TRUE"
lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"

qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw()
ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
@

\begin{figure}[htbp]\centering
\includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCts}
\caption[Visual Aptitude Study Results]{Preliminary results of continuous variables compared to lineup score.}\label{fig:visualaptitudects}
\end{figure}

\begin{figure}[htbp]\centering
\includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCat}
\caption[Visual Aptitude Study Results]{Preliminary results of categorical variables compared to lineup score.}\label{fig:visualaptitudecat}
\end{figure}

\end{document}

### 2014-09-18 Code/GenerateLineups.R

@@ -1,138 +1,0 @@
-library(nullabor)
-library(ggplot2)
-library(plyr)
-
-source("./Code/theme_lineup.R")
-source("./Code/GenerateData.R")
-
-set.seed(31415926)
-
-# Linear trend alone - no outliers
-N <- 10
-seeds <- c(699181, 190785, 746705, 938857, 256246, 
-           130991, 829750, 107019, 310685, 968407, 
-           982098, 247471, 658815, 828038, 956567, 
-           706934,  74904, 743373, 955470, 339582, 
-           230370, 594429, 959280, 677468, 301141, 
-           523820, 397938, 952596, 726989, 247199)
-# seeds <- sample(0:1e6, N) + sample(0:1e3, N) + sample(0:100, N)
-answers <- data.frame()
-
-colors <- c("#882E72", "#D92120", "#4EB265", "#3F4EA1", 
-            "#E7742F", "#1965B0", "#B1B343", "#4683C1", "#DFA53A")
-
-names <- data.frame(filename = sample(paste("./Images/Lineups/", rep(1:3, each=N*2), "-", rep(1:(N*2), times=3), ".png", sep=""), 6*N), idx=1:(6*N), stringsAsFactors=FALSE)
-
-permute.var <- c("x", "y")
-for(i in 1:N){
-  j <- 3*i-2
-  set.seed(seeds[i])
-  dframe <- linear.trend()
-  
-  # Slope alone
-  pos.x <- sample(1:20, 1)
-  filename <- names$filename[j]
-  pv <- sample(permute.var, 1)
-  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
-  ggplot(data=lineupdata) + 
-    geom_point(aes(x=x, y=y), size=3) + 
-    facet_wrap(~.sample) +
-    theme_lineup()
-  ggsave(filename, width=10, height=8, dpi=300, units="in")
-  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=NA, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="Slope"))
-  
-  # Color alone
-  pos.x <- sample(1:20, 1)
-  pv <- "group"
-  filename <- names$filename[j+1]
-  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
-  ggplot(data=lineupdata) + 
-    geom_point(aes(x=x, y=y, color=factor(group)), size=3) + 
-    facet_wrap(~.sample) + 
[…]

### 2014-10-05 Code/Analysis.R
Log video game hours

@@ -66,7 +66,7 @@ ans.summary <- merge(ans[,1:19], ans.summary)
 ans.summary2 <- dcast(longform.sum, id~testtype+testnum, value.var = "value")
 ans.summary2 <- merge(ans[,1:19], ans.summary2)
 pct.ans <- dcast(longform.sum, id~testtype+testnum, value.var="pct.answered")
-
+ans.summary$log_vidgame_hrs <- log10(ans.summary$vidgame_hrs+1)
 # 
 # qplot(data=ans.summary, x=card_rot, y=lineup, geom="point")
 # 
@@ -77,7 +77,7 @@ pct.ans <- dcast(longform.sum, id~testtype+testnum, value.var="pct.answered")
 # qplot(data=ans.summary, x=vis_search, y=lineup, geom="point")
 # 
 
-lineup.summary <- melt(ans.summary, id.vars=c(1:17, 19, 23))
+lineup.summary <- melt(ans.summary, id.vars=c(1:19, 23))
 
 
 # qplot(data=lineup.summary, x=value, y=lineup, geom="point") + facet_wrap(~variable, scales="free")

### 2014-10-05 Paper/Draft1/TestingVisualAptitude.Rnw
Fixing formatting/refs
Pictures updated with all data

@@
 \usepackage{color}
+\usepackage{fullpage}
 \usepackage[dvipsnames,svgnames]{xcolor}
@@
 
-<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=4>>=
-lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Video Game Hrs/Wk", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
-lineup.summary$variable <- factor(lineup.summary$variable, levels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Video Game Hrs/Wk"))
+<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=10,fig.height=5, out.width='\\linewidth'>>=
+lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Log10 Video Game Hrs/Wk"))
+lineup.summary$variable <- factor(lineup.summary$variable, levels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Log10 Video Game Hrs/Wk"))
 
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw()
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
@@
 
+\bibliography{references}
+
 \end{document}

### 2014-10-06 Paper/Draft1/TestingVisualAptitude.Rnw
Set ID to be the answer id, then conduct t-tests comparing results from Hillary to those from Stephanie

@@
 
+\section*{Appendix}
+T-tests of results for Hillary and Stephanie: 
+<<ttest, echo=T, include=T>>=
+t.test(ans.summary$card_rot[1:18], ans.summary$card_rot[-c(1:18)])
+t.test(ans.summary$folding[1:18], ans.summary$folding[-c(1:18)])
+t.test(ans.summary$lineup[1:18], ans.summary$lineup[-c(1:18)])
+t.test(ans.summary$vis_search[1:18], ans.summary$vis_search[-c(1:18)])
+@
+
 \end{document}

### 2014-10-06 Code/Analysis.R

@@ -1,6 +1,7 @@
 # setwd("~/Dropbox/GraphicsGroup/TestingVisualAptitude/")
 
 ans <- read.csv(paste(datadir, "VisualGraphicsData.csv", sep=""), stringsAsFactors=F)
+ans$id <- 1:nrow(ans)
 
 key <- ans[1,]
 key[,1:19] <- NA

### 2014-10-08 Paper/Draft1/TestingVisualAptitude.Rnw
Additional information about test scoring

@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
+The scoring mechanism is also used on many standardized tests (the SAT, ACT, and the battery of psychological tests which certain sections of this test are drawn from); the ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier.
+In general, the formula for computing the score would be $$1/(n-1)*(\text{\# wrong answers}) + (\text{\# correct answers})$$ where $n$ is the total number of possible answers.
 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments

--- 2014-10-06 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 lineups without a target plot).
-Participants will have 5 minutes to complete each block of 20 lineups.
-Figure \ref{fig:lineup} shows a sample lineup of box plots.
+Participants %will
+have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.
 
@@
 \section{Results}
-We have tested 18 students, and plan to increase sample size once the summer semester begins.
+\newtext{Results are based on evaluation of \Sexpr{nrow(lineup.summary)} undergraduate students at Iowa State University. }
+%We have tested students, and plan to increase sample size once the summer semester begins. 
 Preliminary results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}. 
+\comment{No woulds, shoulds or coulds}
 Scoring for all of the tests was calculated such that with random guessing, the expected score on a problem would be 0; thus, for a test with possible answers A, B, C, D, and E, the score would be calculated by subtracting $1/4*\text{\# wrong answers}$ from the total number of correct answers.
+\comment{re-phrasing the above sentence:}
+\newtext{Scoring of all test results was done such that random guessing will lead to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ possible answers and one correct answer each, the score is calculated as}
+\begin{eqnarray}\label{eq.scoring}
+\# \text{total correct answers } - 1/k \cdot \text{\#wrong answers}.
+\end{eqnarray}
+
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
+This scoring mechanism is also used on many standardized tests, such as the  SAT, ACT, and the battery of psychological tests which \newtext{parts} of this test are drawn from.
+The ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier. 
+%In general, the formula for computing the score would be $$1/(n-1)*(\text{\# wrong answers}) + (\text{\# correct answers})$$ where $n$ is the total number of possible answers.
 
@@
 
+\newtext{All results and data shown here are done in accordance with IRB \# XXX?}
+
 \bibliography{references}

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
next couple of steps laid out

@@
 \section{Results}
-\newtext{Results are based on evaluation of \Sexpr{nrow(lineup.summary)} undergraduate students at Iowa State University. }
+\newtext{Results are based on an evaluation of \Sexpr{nrow(lineup.summary)} undergraduate students at Iowa State University. }
 %We have tested students, and plan to increase sample size once the summer semester begins. 
-Preliminary results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}. 
 \comment{No woulds, shoulds or coulds}
@@
 \comment{re-phrasing the above sentence:}
-\newtext{Scoring of all test results was done such that random guessing will lead to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ possible answers and one correct answer each, the score is calculated as}
+\newtext{Scoring of all test results was done such that random guessing leads to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as}
 \begin{eqnarray}\label{eq.scoring}
-\# \text{total correct answers } - 1/k \cdot \text{\#wrong answers}.
+\# \text{total correct answers } - 1/k \cdot \# \text{wrong answers}.
 \end{eqnarray}
@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-This scoring mechanism is also used on many standardized tests, such as the  SAT, ACT, and the battery of psychological tests which \newtext{parts} of this test are drawn from.
-The ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier. 
+\newtext{Combining accuracy and speed into a single number does not only make a comparison of test scores easier, } this scoring mechanism is also used on many standardized tests, such as the  SAT, ACT, and the battery of psychological tests which \newtext{parts} of this test are drawn from.
+\comment{We need a few citations here. }
+%The ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier. 
 %In general, the formula for computing the score would be $$1/(n-1)*(\text{\# wrong answers}) + (\text{\# correct answers})$$ where $n$ is the total number of possible answers. 
+\comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown.
+It would be good to scale everything, so that we have the same theoretical range.
+That will actually allow a comparison across different tests.}
+\comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).
+}
+
+\comment{Split the next sentence up into two paragraph: what is shown in each of the figures, what are the immediate conclusions?
+chuck out the gaming hours from figure \ref{fig:visualaptitudects}, that's a demograph, so move it to the other plot (I am aware that nine demographics are easier to display than ten, but content trumps).
+What happened to the arts skills?
+I seem to have a positive association now - why is my data different from yours?}
+Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.
 
@@
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
+@
+
+<<correlations>>=
+cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
+# using scaled version right now, should be changed to unscaled once the scores are internally scaled.
+pca <- prcomp(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], scale=T)
+summary(pca)
+screeplot(pca)
 @

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
summaries
scaling discussion

@@
 \begin{eqnarray}\label{eq.scoring}
-\# \text{total correct answers } - 1/k \cdot \# \text{wrong answers}.
+\# \text{total correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
 \end{eqnarray}
@@
 \comment{We need a few citations here. }
+\newtext{Additionally, we have to ensure that the ranges and units of test scores are comparable. Assume $n$ questions with $k$ choices (including one correct answer) each. 
+This leads to a theoretical range of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of }
+\begin{eqnarray*}
+\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{n, k}) = \\
+&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
+&=& n^2/(k-1)
+\end{eqnarray*}
+
 %The ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier. 
@@
 That will actually allow a comparison across different tests.}
+<<means>>=
+library(plyr)
+ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
+@
+
 \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).

### 2014-10-11 Code/Analysis.R
Add penalty for guessing wrong on figure classification

@@ -48,15 +48,20 @@ scored[,1:19] <- ans[,1:19]
 scored[,20:396] <- apply(scored[,20:396], 2, as.numeric)
 
 library(reshape2)
+library(stringr)
 longform <- melt(scored[,c(1, 20:396)], id.vars=1)
 longform$testtype <- gsub("_q[0123456789]+[abcdefgh]?$", "", longform$variable)
 longform$testnum <- as.numeric(gsub("[[:alpha:]_]+", "", longform$testtype))
 longform$testtype <- gsub("[[:digit:]]", "", longform$testtype)
+longform$qnum <- as.numeric(str_replace(str_extract(longform$variable, "q\\d{1,2}"), fixed("q"), ""))
 longform$penalty <- 0
 longform$penalty[longform$testtype=="vis_search"] <- 1/23
 longform$penalty[longform$testtype=="lineup"] <- 1/19
 longform$penalty[longform$testtype=="card_rot"] <- 1
 longform$penalty[longform$testtype=="folding"] <- 1/4
+longform$penalty[longform$testtype=="fig_class" & longform$qnum<=8] <- 1
+longform$penalty[longform$testtype=="fig_class" & longform$qnum>8] <- 1/2
+
 
 longform.sum <- ddply(longform, .(id, testtype, testnum), summarize, 
                       value=ifelse(is.numeric(value), sum(value, na.rm=T)-sum(!value,na.rm=T)*penalty, unique(value)), 

### 2014-10-11 Code/Analysis.R
Use "susan standard" on the third set of lineups

@@ -66,8 +66,8 @@ longform$penalty[longform$testtype=="fig_class" & longform$qnum>8] <- 1/2
 longform.sum <- ddply(longform, .(id, testtype, testnum), summarize, 
                       value=ifelse(is.numeric(value), sum(value, na.rm=T)-sum(!value,na.rm=T)*penalty, unique(value)), 
                       pct.answered=sum(!is.na(value))/length(value))
-longform.sum$value[longform.sum$testtype=="lineup" & longform.sum$testnum==3] <- NA
-ans.summary <- dcast(longform.sum, id~testtype, value.var="value", fun.aggregate = mean, na.rm=TRUE)
+# longform.sum$value[longform.sum$testtype=="lineup" & longform.sum$testnum==3] <- NA
+ans.summary <- dcast(longform.sum, id~testtype, value.var="value", fun.aggregate = sum, na.rm=TRUE)
 ans.summary <- merge(ans[,1:19], ans.summary)
 ans.summary2 <- dcast(longform.sum, id~testtype+testnum, value.var = "value")
 ans.summary2 <- merge(ans[,1:19], ans.summary2)

### 2014-10-11 Code/Analysis.R
Include video games with categorical data

@@ -83,9 +83,9 @@ ans.summary$log_vidgame_hrs <- log10(ans.summary$vidgame_hrs+1)
 # qplot(data=ans.summary, x=vis_search, y=lineup, geom="point")
 # 
 
-lineup.summary <- melt(ans.summary, id.vars=c(1:19, 23))
+lineup.summary <- melt(ans.summary, id.vars=c(1:19, 23, 25))
 
 
 # qplot(data=lineup.summary, x=value, y=lineup, geom="point") + facet_wrap(~variable, scales="free")
 
-lineup.summary.categorical <- melt(ans.summary[,c(1:3, 12:17, 19, 23)], id.vars=c("id", "lineup"))
+lineup.summary.categorical <- melt(ans.summary[,c(1:3, 12:17, 19, 23, 25)], id.vars=c("id", "lineup"))

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Accept changes with variance calculations

@@
 \section{Results}
-\newtext{Results are based on an evaluation of \Sexpr{nrow(lineup.summary)} undergraduate students at Iowa State University. }
-%We have tested students, and plan to increase sample size once the summer semester begins. 
-\comment{No woulds, shoulds or coulds}
-Scoring for all of the tests was calculated such that with random guessing, the expected score on a problem would be 0; thus, for a test with possible answers A, B, C, D, and E, the score would be calculated by subtracting $1/4*\text{\# wrong answers}$ from the total number of correct answers.
-\comment{re-phrasing the above sentence:}
-\newtext{Scoring of all test results was done such that random guessing leads to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as}
+\newtext{Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. }
+Scoring of all test results was done such that random guessing leads to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
 \begin{eqnarray}\label{eq.scoring}
@@
 \end{eqnarray}
+This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests which parts of this test are drawn from \citep{diamond1973correction, ekstrom1976manual}.
 
-This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-\newtext{Combining accuracy and speed into a single number does not only make a comparison of test scores easier, } this scoring mechanism is also used on many standardized tests, such as the  SAT, ACT, and the battery of psychological tests which \newtext{parts} of this test are drawn from.
-\comment{We need a few citations here. }
-\newtext{Additionally, we have to ensure that the ranges and units of test scores are comparable. Assume $n$ questions with $k$ choices (including one correct answer) each. 
-This leads to a theoretical range of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of }
+Additionally, we have to ensure that the ranges and units of test scores are comparable. Assume $n$ questions with $k$ choices (including one correct answer) each. 
+This leads to a theoretical range of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
+
 \begin{eqnarray*}
@@
 
-%The ability to combine both accuracy and number of questions attempted into a single number makes comparing test scores much easier. 
-%In general, the formula for computing the score would be $$1/(n-1)*(\text{\# wrong answers}) + (\text{\# correct answers})$$ where $n$ is the total number of possible answers. 
+\comment{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3.
+Or we could leave it as an exercise for the reader :)}
+
 \comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Comment on variable split - modification

@@
 
-\comment{Split the next sentence up into two paragraph: what is shown in each of the figures, what are the immediate conclusions?
+\comment{
+Split the next sentence up into two paragraph: what is shown in each of the figures, what are the immediate conclusions?
 chuck out the gaming hours from figure \ref{fig:visualaptitudects}, that's a demograph, so move it to the other plot (I am aware that nine demographics are easier to display than ten, but content trumps).
-What happened to the arts skills?
-I seem to have a positive association now - why is my data different from yours?}
+\\
+
+Actually, they were split up like that because of continuous vs. categorical. It's not easy to facet when you have different $x$ scales. I've removed video games entirely for the time being. 
+}
 Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Art association modification

@@
 }
+
+\comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
+I have the same updated results you do.
+I see that Stephanie changed the data file (but didn't add any more rows)...
+is it possible something changed there?
+Otherwise, maybe my dropbox didn't update correctly the first time?}
 Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Update pictures to account for log10 video games moving to categorical graph.

@@
 <<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=10,fig.height=5, out.width='\\linewidth'>>=
-lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Log10 Video Game Hrs/Wk"))
-lineup.summary$variable <- factor(lineup.summary$variable, levels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search", "Log10 Video Game Hrs/Wk"))
+lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
 
-qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
+qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=4)
@@
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience"))
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Log10 Video Game Hours"))
 lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=subset(lineup.summary.categorical, variable!="Log10 Video Game Hours"), x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Explanation for continuous variable associations.

@@
 
+In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
+This suggests that skills associated with visual reasoning ability are related to lineup performance.
+As participants must use the same skills in lineups (mental rotation, classification and determining categorization schemes, and multi-step spatial reasoning) as in the factor-referenced tests, this is not particularly surprising.
 \begin{figure}[htbp]\centering

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Update comments to show my responses

@@
 \\
-
-Actually, they were split up like that because of continuous vs. categorical. It's not easy to facet when you have different $x$ scales. I've removed video games entirely for the time being. 
-}
+\textit{Actually, they were split up like that because of continuous vs.
+categorical.
+It's not easy to facet when you have different $x$ scales.
+I've removed video games entirely for the time being.
+I agree it didn't really belong there, but I can't get it to work with the facetting, either.
+I will probably just bin it eventually.
+}}
 
 \comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
-I have the same updated results you do.
+\textit{I have the same updated results you do.
 I see that Stephanie changed the data file (but didn't add any more rows)...
 is it possible something changed there?
-Otherwise, maybe my dropbox didn't update correctly the first time?}
+Otherwise, maybe my dropbox didn't update correctly the first time?
+Or maybe it was figure caching?}}
 Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.

### 2014-10-11 Code/Analysis.R
Key formatting, etc.

@@ -33,7 +33,10 @@ key[,245:356] <- c(2, 1, 1, 2, 1, 2, 2, 2,  2, 1, 2, 1, 2, 1, 2, 2,
                    2, 2, 1, 3, 3, 1, 1, 1,  3, 1, 2, 3, 1, 3, 2, 2)
 
 # Lineups 3
-key[,357:376] <- c(12, 5, 15, 4, 15, 1, 4, 16, 1, 19, 13, 19, 12, 11, 12, 5, 7, 12, 19, 3)
+key[,357:376] <- c(12, 19, 14,  4, 13, 
+                    4, 14,  7,  8, 19, 
+                    8, 19, 12, 18, 12, 
+                    7,  7,  6,  6,  5)
 
 # Paper Folding
 key[,377:396] <- c("a", "d", "b", "d", "b", 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
additional comments

@@
 
-\comment{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3.
-Or we could leave it as an exercise for the reader :)}
+\comment{\textit{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3.
+Or we could leave it as an exercise for the reader :)}}
+
+\comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}}
 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
PCA plot caption

@@
 
-<<correlations>>=
+<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|">>=
 cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
 # using scaled version right now, should be changed to unscaled once the scores are internally scaled.
-pca <- prcomp(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], scale=T)
+pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T)
 summary(pca)

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Additional comment
IRB #

@@
 
+\comment{\textit{I did change the lineup scoring - for some reason, I was averaging before (and I hadn't included the third battery of lineups b/c of scoring issues).}}
+
 In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
@@
 
-\newtext{All results and data shown here are done in accordance with IRB \# XXX?}
+All results and data shown here are done in accordance with IRB \# 13-581.
 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Categorical video game hours/wk variable

@@
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Log10 Video Game Hours"))
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Video Game Hrs/Wk"))
 lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
@@
 
-qplot(data=subset(lineup.summary.categorical, variable!="Log10 Video Game Hours"), x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=3) + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)

### 2014-10-11 Code/Analysis.R

@@ -75,7 +75,9 @@ ans.summary <- merge(ans[,1:19], ans.summary)
 ans.summary2 <- dcast(longform.sum, id~testtype+testnum, value.var = "value")
 ans.summary2 <- merge(ans[,1:19], ans.summary2)
 pct.ans <- dcast(longform.sum, id~testtype+testnum, value.var="pct.answered")
-ans.summary$log_vidgame_hrs <- log10(ans.summary$vidgame_hrs+1)
+
+ans.summary$vidgame_hrs_factor <- ordered(factor(rowSums(sapply(c(-1, 0, 1.99, 4.99), function(i) ans$vidgame_hrs>i)), labels=c("0", "[1, 2)", "[2, 5)", "5+"), levels=1:4))
+# ans.summary$log_vidgame_hrs <- log10(ans.summary$vidgame_hrs+1)
 # 
 # qplot(data=ans.summary, x=card_rot, y=lineup, geom="point")
 # 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Additional clarification

@@
 
-\comment{\textit{I did change the lineup scoring - for some reason, I was averaging before (and I hadn't included the third battery of lineups b/c of scoring issues).}}
+\comment{\textit{I did change the lineup scoring - for some reason, I was averaging before (and I hadn't included the third battery of lineups b/c of scoring issues).
+That explains the scale, but not the change in art skills association.}}
 

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Visual Search discussion

@@
 As participants must use the same skills in lineups (mental rotation, classification and determining categorization schemes, and multi-step spatial reasoning) as in the factor-referenced tests, this is not particularly surprising.
+In addition,there seems to be some positive relationship between a participant's score on the visual search task and their score on lineups: the visual search task represents a baseline of a participant's ability to find a matching pattern, while lineups require that task as well as the ability to determine what the pattern is for a particular graph.
+Even excluding the one low visual search score that is a high-leverage point, there seems to be a positive relationship between a participant's score on lineups and their score for visual search.
+
 \begin{figure}[htbp]\centering

### 2014-10-11 Code/Analysis.R
Scale test scores

@@ -1,6 +1,6 @@
 # setwd("~/Dropbox/GraphicsGroup/TestingVisualAptitude/")
 
-ans <- read.csv(paste(datadir, "VisualGraphicsData.csv", sep=""), stringsAsFactors=F)
+ans <- read.csv(paste0(datadir, "VisualGraphicsData.csv"), na.strings=c(" ", ""), stringsAsFactors=F)
 ans$id <- 1:nrow(ans)
 
 key <- ans[1,]
@@ -45,7 +45,8 @@ key[,377:396] <- c("a", "d", "b", "d", "b",
                    "a", "e", "d", "d", "c")
 
 library(plyr)
-scored <- ddply(ans, .(id), function(j) j==key)
+scored <- ddply(ans, .(id), function(j) as.numeric(j==key) + c(0, NA)[1+is.na(j)])
+names(scored) <- names(ans)
 scored[,1:19] <- ans[,1:19]
 # scored[,357:376] <- NA # can't find answers for Lineup3?
 scored[,20:396] <- apply(scored[,20:396], 2, as.numeric)
@@ -65,16 +66,22 @@ longform$penalty[longform$testtype=="folding"] <- 1/4
 longform$penalty[longform$testtype=="fig_class" & longform$qnum<=8] <- 1
 longform$penalty[longform$testtype=="fig_class" & longform$qnum>8] <- 1/2
 
+qrange <- ddply(subset(longform, id==1), .(testtype), summarize, min.score=sum(penalty), max.score=length(testtype))
 
-longform.sum <- ddply(longform, .(id, testtype, testnum), summarize, 
-                      value=ifelse(is.numeric(value), sum(value, na.rm=T)-sum(!value,na.rm=T)*penalty, unique(value)), 
+longform.sum <- ddply(longform, .(id, testtype), summarize, 
+                      pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
+                      neg.pts = ifelse(is.numeric(value), sum((1-value)*penalty, na.rm=T), unique(value)),
                       pct.answered=sum(!is.na(value))/length(value))
+longform.sum <- merge(longform.sum, qrange)
+
+longform.sum$value <- with(longform.sum, (pos.pts - neg.pts + min.score)/(min.score+max.score)*100)
+
[…]

### 2014-10-11 Code/Analysis.R
Change pic to be on same scale

@@ -98,6 +98,6 @@ ans.summary$vidgame_hrs_factor <- ordered(factor(rowSums(sapply(c(-1, 0, 1.99, 4
 lineup.summary <- melt(ans.summary, id.vars=c(1:19, 23, 25))
 
 
-# qplot(data=lineup.summary, x=value, y=lineup, geom="point") + facet_wrap(~variable, scales="free")
+# qplot(data=lineup.summary, x=value, y=lineup, geom="point") + facet_wrap(~variable) + xlim(c(0,100))
 
 lineup.summary.categorical <- melt(ans.summary[,c(1:3, 12:17, 19, 23, 25)], id.vars=c("id", "lineup"))

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Comment clarification for scaling
Remove "value" X label

@@
 
-\comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}}
+\comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
+Right now, I'm assuming that we're scaling between 0 and 100 by test range $[-a, b]$. So we take the guessing-fixed score and add $a$, then divide by $(a+b)$ and multiply by 100.}
 
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
-ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=8, height=6)
 @

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Categorical variable analysis

@@
 
+Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
+There is little difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class.
+There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+In addition, completion of Calculus I is associated with increased performance on lineups (though completion of calculus is also associated with sex).
+AutoCAD experience is also not significantly associated with lineup performance; there is a difference in the medians, but it does not rise to the level of significance.
+There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and may be at least partially a result of the large difference in performance due to sex.
+
+<<t.test.categorical, echo=T, comment="|">>=
+t.test(ans.summary$lineup[ans.summary$sex=="m"], ans.summary$lineup[ans.summary$sex=="f"])
+
+t.test(ans.summary$lineup[ans.summary$calc_1=="y"], ans.summary$lineup[ans.summary$calc_1=="n"])
+
+ans.summary$vidgame_hrs_factor_new <- factor(ans.summary$vidgame_hrs_factor, ordered=FALSE)
+summary(lm(data=ans.summary, lineup~vidgame_hrs_factor_new))
+
+summary(lm(data=ans.summary, lineup~sex+calc_1+vidgame_hrs_factor_new))
+@
+
 All results and data shown here are done in accordance with IRB \# 13-581.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Plot label change

@@
 
-qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
+qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Scaled Score") + ylab("Scaled Lineup Score") + geom_smooth(method="lm") + theme_bw()
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=4)
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=8, height=6)

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
first responses

--- 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
+\comment{The text makes still a lot of references to `preliminary results'.
+Don't just use copy \& paste for the changes!!!!
+Go through those sections and adjust.
+We are working on the final analysis now.
+Even if numbers might still change, they are not preliminary anymore.
+}
 \section{Introduction}
@@
 \begin{eqnarray}\label{eq.scoring}
-\# \text{total correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
+\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
 \end{eqnarray}
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests which parts of this test are drawn from \citep{diamond1973correction, ekstrom1976manual}.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} which parts of this test are drawn from.
 
@@
 \comment{\textit{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3.
-Or we could leave it as an exercise for the reader :)}}
+Or we could leave it as an exercise for the reader :)}
+It's fairly trivial to apply this to the mixture - the only assumption that goes into the variance discussion is that the questions are independent.
+We don't need to include it here, but you should apply it to the figure classification set.
+
+}
 
 \comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
-Right now, I'm assuming that we're scaling between 0 and 100 by test range $[-a, b]$. So we take the guessing-fixed score and add $a$, then divide by $(a+b)$ and multiply by 100.}
+Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}
 
@@
 
-\comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
-\textit{I have the same updated results you do.
+% \comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
+% \textit{I have the same updated results you do.
 I see that Stephanie changed the data file (but didn't add any more rows)...
@@
 
-<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=10,fig.height=5, out.width='\\linewidth'>>=
+<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=8, out.width='0.9\\textwidth'>>=
 lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
-ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=8, height=6)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
 @
@@
 \begin{figure}[htbp]\centering
-\includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCts}
-\caption[Visual Aptitude Study Results]{Preliminary results of continuous variables compared to lineup score.}\label{fig:visualaptitudects}
+\includegraphics[width=0.9\textwidth]{fig-VisReasoningAnalysisCts}
+\caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
+\comment{What is the main message of these plots?
+Relatively strong positive correlation to all established test scores?}}\label{fig:visualaptitudects}
 \end{figure}

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
mini stuff

@@
 qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
-ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=4)
+# same aspect ratio
+ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=6)
 
@@
 \begin{figure}[htbp]\centering
-\includegraphics[width=0.9\textwidth]{fig-VisReasoningAnalysisCts}
+\includegraphics[width=0.75\textwidth]{fig-VisReasoningAnalysisCts}
 \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Fixing comments

@@
 
-\comment{The text makes still a lot of references to `preliminary results'.
+% \comment{The text makes still a lot of references to `preliminary results'.
 Don't just use copy \& paste for the changes!!!!
@@
 \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
-\comment{What is the main message of these plots?
-Relatively strong positive correlation to all established test scores?}}\label{fig:visualaptitudects}
+There is a relatively strong positive correlation between lineup score and scores on visuospatial reasoning tests.}\label{fig:visualaptitudects}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCat}
-\caption[Visual Aptitude Study Results]{Preliminary results of categorical variables compared to lineup score.}\label{fig:visualaptitudecat}
+\caption[Visual Aptitude Study Results]{Demographic characteristics compared with lineup score.}\label{fig:visualaptitudecat}
 \end{figure}

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Removing warning/error message

@@
 
-<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=8, out.width='0.9\\textwidth'>>=
+<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=8, out.width='0.9\\textwidth', message=F, warning=F>>=
 lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
@@
 
-<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|">>=
+<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|", warning=F, message=F>>=
 cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Deleted text and caption changes

--- 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + geom_smooth(method="lm") + theme_bw()
+qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Scaled Score") + ylab("Scaled Lineup Score") + geom_smooth(method="lm") + theme_bw()
 # same aspect ratio
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("Value") + ylab("Lineup Score (# Correct - 1/19 # Incorrect)") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
@@
 
+Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
+There is little difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class.
+There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+In addition, completion of Calculus I is associated with increased performance on lineups (though completion of calculus is also associated with sex).
+AutoCAD experience is also not significantly associated with lineup performance; there is a difference in the medians, but it does not rise to the level of significance.
+There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and may be at least partially a result of the large difference in performance due to sex.
+
+<<t.test.sex, echo=T, comment="|">>=
+t.test(ans.summary$lineup[ans.summary$sex=="m"], ans.summary$lineup[ans.summary$sex=="f"])
+@
+<<t.test.calculus, echo=T, comment="|">>=
+t.test(ans.summary$lineup[ans.summary$calc_1=="y"], ans.summary$lineup[ans.summary$calc_1=="n"])
+@
+<<lm.vid.games, echo=T, comment="|">>=
+ans.summary$vidgame_hrs_factor_new <- factor(ans.summary$vidgame_hrs_factor, ordered=FALSE)
+summary(lm(data=ans.summary, lineup~vidgame_hrs_factor_new))
+# summary(lm(data=ans.summary, lineup~vidgame_hrs))
+@
+<<lm.all, echo=T, comment="|">>=
+summary(lm(data=ans.summary, lineup~sex+calc_1+vidgame_hrs_factor_new))
+@
 All results and data shown here are done in accordance with IRB \# 13-581.

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Remove preliminary

@@
 \includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCat}
-\caption[Visual Aptitude Study Results]{Preliminary results of categorical variables compared to lineup score.}\label{fig:visualaptitudecat}
+\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.}\label{fig:visualaptitudecat}
 \end{figure}

### 2014-10-11 Code/Analysis.R
Scale by variance

@@ -65,8 +65,10 @@ longform$penalty[longform$testtype=="card_rot"] <- 1
 longform$penalty[longform$testtype=="folding"] <- 1/4
 longform$penalty[longform$testtype=="fig_class" & longform$qnum<=8] <- 1
 longform$penalty[longform$testtype=="fig_class" & longform$qnum>8] <- 1/2
+ 
+qrange <- ddply(subset(longform, id==1), .(testtype), summarize, min.score=sum(penalty), max.score=length(testtype), n=length(testtype), k=mean(1/penalty+1))
+qrange$var <- with(qrange, n^2/(k-1))
 
-qrange <- ddply(subset(longform, id==1), .(testtype), summarize, min.score=sum(penalty), max.score=length(testtype))
 
 longform.sum <- ddply(longform, .(id, testtype), summarize, 
                       pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
@@ -74,7 +76,13 @@ longform.sum <- ddply(longform, .(id, testtype), summarize,
                       pct.answered=sum(!is.na(value))/length(value))
 longform.sum <- merge(longform.sum, qrange)
 
-longform.sum$value <- with(longform.sum, (pos.pts - neg.pts + min.score)/(min.score+max.score)*100)
+longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
+tmp <- ddply(longform.sum, .(testtype), summarize, test.mean = mean(value, na.rm=T))
+longform.sum <- merge(longform.sum, tmp)
+longform.sum$value <- with(longform.sum, value - test.mean)
+longform.sum$value <- with(longform.sum, value/sqrt(var))
+
+# longform.sum$value <- with(longform.sum, (pos.pts - neg.pts + min.score)/(min.score+max.score)*100)
 
 # longform.sum$value[longform.sum$testtype=="lineup" & longform.sum$testnum==3] <- NA
 ans.summary <- dcast(longform.sum, id~testtype, value.var="value", na.rm=TRUE)

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Scale by variance changes

@@
 
-<<VisualAptitude-setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE>>=
+<<VisualAptitude-setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
 rm(list=ls())
@@
 \includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCat}
-\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.}\label{fig:visualaptitudecat}
+\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.
+Sex (male), Calculus completion, and hours spent playing video games per week are all associated with a difference in lineup score.
+}\label{fig:visualaptitudecat}
 \end{figure}

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Remove comment about preliminary results as the comment no longer applies.
Rearrange paper text and comment out comments that are no longer new/relevant.

@@
 
-% \comment{The text makes still a lot of references to `preliminary results'.
-Don't just use copy \& paste for the changes!!!!
-Go through those sections and adjust.
-We are working on the final analysis now.
-Even if numbers might still change, they are not preliminary anymore.
-}
 \section{Introduction}
@@
 
-\comment{\textit{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3.
-Or we could leave it as an exercise for the reader :)}
-It's fairly trivial to apply this to the mixture - the only assumption that goes into the variance discussion is that the questions are independent.
-We don't need to include it here, but you should apply it to the figure classification set.
-
-}
-
-\comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
-Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}
-
-\comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown.
+% \comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
+% Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}
+% 
+% \comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown.
 It would be good to scale everything, so that we have the same theoretical range.
@@
 
+<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|", warning=F, message=F>>=
+cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
+# using scaled version right now, should be changed to unscaled once the scores are internally scaled.
+pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T)
+summary(pca)
+screeplot(pca)
+@
+
 \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).
-}
+\textit{I'm not sure where to find the ``normal scores" in the factor-referenced manual.
+I'll check through it again, but I don't recall seeing it.}}
 
-\comment{
-Split the next sentence up into two paragraph: what is shown in each of the figures, what are the immediate conclusions?
-chuck out the gaming hours from figure \ref{fig:visualaptitudects}, that's a demograph, so move it to the other plot (I am aware that nine demographics are easier to display than ten, but content trumps).
-\\
-\textit{Actually, they were split up like that because of continuous vs.
-categorical.
-It's not easy to facet when you have different $x$ scales.
-I've removed video games entirely for the time being.
-I agree it didn't really belong there, but I can't get it to work with the facetting, either.
-I will probably just bin it eventually.
-}}
-
-% \comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
-% \textit{I have the same updated results you do.
-I see that Stephanie changed the data file (but didn't add any more rows)...
-is it possible something changed there?
-Otherwise, maybe my dropbox didn't update correctly the first time?
-Or maybe it was figure caching?}}
-Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.
+In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
+This suggests that skills associated with visual reasoning ability are related to lineup performance.
+As participants must use the same skills in lineups (mental rotation, classification and determining categorization schemes, and multi-step spatial reasoning) as in the factor-referenced tests, this is not particularly surprising.
+In addition,there seems to be some positive relationship between a participant's score on the visual search task and their score on lineups: the visual search task represents a baseline of a participant's ability to find a matching pattern, while lineups require that task as well as the ability to determine what the pattern is for a particular graph.
+Even excluding the one low visual search score that is a high-leverage point, there seems to be a positive relationship between a participant's score on lineups and their score for visual search.
 
@@
 
-<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|", warning=F, message=F>>=
-cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
-# using scaled version right now, should be changed to unscaled once the scores are internally scaled.
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T)
-summary(pca)
-screeplot(pca)
-@
-
 \begin{figure}[htbp]\centering
@@
 
-\comment{\textit{I did change the lineup scoring - for some reason, I was averaging before (and I hadn't included the third battery of lineups b/c of scoring issues).
-That explains the scale, but not the change in art skills association.}}
-
-In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
-This suggests that skills associated with visual reasoning ability are related to lineup performance.
-As participants must use the same skills in lineups (mental rotation, classification and determining categorization schemes, and multi-step spatial reasoning) as in the factor-referenced tests, this is not particularly surprising.
-In addition,there seems to be some positive relationship between a participant's score on the visual search task and their score on lineups: the visual search task represents a baseline of a participant's ability to find a matching pattern, while lineups require that task as well as the ability to determine what the pattern is for a particular graph.
-Even excluding the one low visual search score that is a high-leverage point, there seems to be a positive relationship between a participant's score on lineups and their score for visual search.
+Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
 
@@
 
-Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
-There is little difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class.
+There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class.
 There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.

### 2014-10-12 Code/Analysis.R
Don't center scores

@@ -1,8 +1,12 @@
 # setwd("~/Dropbox/GraphicsGroup/TestingVisualAptitude/")
 
+# Read in data
 ans <- read.csv(paste0(datadir, "VisualGraphicsData.csv"), na.strings=c(" ", ""), stringsAsFactors=F)
+# Create anon. IDs that are unique
 ans$id <- 1:nrow(ans)
 
+
+# Create a key that has the same structure as the answers
 key <- ans[1,]
 key[,1:19] <- NA
 # Visual Search
@@ -44,13 +48,17 @@ key[,377:396] <- c("a", "d", "b", "d", "b",
                    "c", "b", "a", "e", "b", 
                    "a", "e", "d", "d", "c")
 
+
 library(plyr)
+# Score each answer by whether it is the same as the key. Make sure NA values stay as NA. 
 scored <- ddply(ans, .(id), function(j) as.numeric(j==key) + c(0, NA)[1+is.na(j)])
 names(scored) <- names(ans)
+# Transfer demographic data
 scored[,1:19] <- ans[,1:19]
 # scored[,357:376] <- NA # can't find answers for Lineup3?
 scored[,20:396] <- apply(scored[,20:396], 2, as.numeric)
 
+# Melt scored data into long format, calculate penalties, and compute scores using ddply
 library(reshape2)
 library(stringr)
 longform <- melt(scored[,c(1, 20:396)], id.vars=1)
@@ -66,46 +74,40 @@ longform$penalty[longform$testtype=="folding"] <- 1/4
 longform$penalty[longform$testtype=="fig_class" & longform$qnum<=8] <- 1
 longform$penalty[longform$testtype=="fig_class" & longform$qnum>8] <- 1/2
  
+# Compute penalty and positive points for each test type, as well as quantities needed for variance calculation
 qrange <- ddply(subset(longform, id==1), .(testtype), summarize, min.score=sum(penalty), max.score=length(testtype), n=length(testtype), k=mean(1/penalty+1))
 qrange$var <- with(qrange, n^2/(k-1))
 
-
+# Compute penalty and positive points for each participant
 longform.sum <- ddply(longform, .(id, testtype), summarize, 
                       pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
[…]

### 2014-10-12 Paper/Draft1/TestingVisualAptitude.Rnw

@@
 \section{Results}
-\newtext{Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. }
-Scoring of all test results was done such that random guessing leads to an expected value of 0; thus for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
+Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. 
+Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$.
+Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
 \begin{eqnarray}\label{eq.scoring}
@@
 
-Additionally, we have to ensure that the ranges and units of test scores are comparable. Assume $n$ questions with $k$ choices (including one correct answer) each. 
-This leads to a theoretical range of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
+Additionally, we have to ensure that the ranges and units of test scores are comparable.
+Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
 
 \begin{eqnarray*}
-\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{n, k}) = \\
+\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
 &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-&=& n^2/(k-1)
+&=& n^2/(k-1).
 \end{eqnarray*}
+
+\newtext{The above consideration only assumes independence between questions, which is granted.
+While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.}
 

### 2014-10-13 Code/Analysis.R
Calc unscaled scores as well for comparison with previous values

@@ -87,6 +87,9 @@ longform.sum <- merge(longform.sum, qrange)
 # Calculate raw score
 longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
 
+# Save unscaled version
+longform.sum.unscaled <- longform.sum
+
 # Calculate test-wise means (for centering) and center scores
 # tmp <- ddply(longform.sum, .(testtype), summarize, test.mean = mean(value, na.rm=T))
 # longform.sum <- merge(longform.sum, tmp)
@@ -102,6 +105,10 @@ longform.sum$value <- with(longform.sum, value/sqrt(var))
 ans.summary <- dcast(longform.sum, id~testtype, value.var="value", na.rm=TRUE)
 ans.summary <- merge(ans[,1:19], ans.summary)
 pct.ans <- dcast(longform.sum, id~testtype, value.var="pct.answered")
+
+# Cast unscaled version
+ans.summary.unscaled <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)
+ans.summary.unscaled <- merge(ans[,1:19], ans.summary.unscaled)
 
 # Create video game factor variable
 ans.summary$vidgame_hrs_factor <- ordered(factor(rowSums(sapply(c(-1, 0, 1.99, 4.99), function(i) ans$vidgame_hrs>i)), labels=c("0", "[1, 2)", "[2, 5)", "5+"), levels=1:4))

### 2014-10-13 Paper/Draft1/TestingVisualAptitude.Rnw
Define k a bit earlier and also add the standard deviation.

@@
 Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. 
-Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$.
+Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
 Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
@@
 &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-&=& n^2/(k-1).
+&=& n^2/(k-1).\\
+\text{Sd}(X_{n, k}) &=& n/sqrt{k-1}
 \end{eqnarray*}
@@
 While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.}
+\newtext{In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
+This approach also allows us to compare the test scores using similar orders of magnitude.}
 

### 2014-10-13 Paper/Draft1/TestingVisualAptitude.Rnw
Looking at the relationship between cognitive tests sans lineups: we don't have 4 separate factors here!

@@
 # using scaled version right now, should be changed to unscaled once the scores are internally scaled.
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T)
+pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=F)
 summary(pca)
 screeplot(pca)
+pca$rotation
+
+# Just using the 4 cognitive tests and ignoring the lineups...
+pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
+summary(pca.cog.tests)
+screeplot(pca.cog.tests)
+pca.cog.tests$rotation
 @

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Table of comparative results

@@
 \usepackage{rotating}
+\usepackage{footnote} % table footnotes
+\makesavenoteenv{tabular}
 
@@
 
-opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=TRUE, concordance=TRUE, autodep=TRUE, root.dir="../")
+opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, concordance=TRUE, autodep=TRUE, root.dir="../")
 datadir <- "../../Data/"
@@
 
-Additionally, we have to ensure that the ranges and units of test scores are comparable.
-Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+\begin{savenotes}
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlll}
+  & Card Rotation & Paper Folding & Figure Classification \\\hline
+ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 2)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
+             & \Sexpr{round(mean(ans.summary.unscaled$folding), 2)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
+             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 2)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
+Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
+              & 13.8 (4.5)
+              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 2)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617/2), 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
+Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
+                & 13.8 (4.5)
+                & M:~120.0~(30.0), F:~114.9~(27.8)\\
+{\footnotesize Population}    
+              & \parbox[t]{.2\textwidth}{\footnotesize $\approx$ 550 male Naval recruits} 
+              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.3\textwidth}{\footnotesize Suburban 11-12th grade students\\(288-300 males, 317-329 females)}\\\hline
+\end{tabular}
+
+\caption{Comparison of scores from Iowa State students and scores reported in \citet{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison}}
+\end{table}
+\end{savenotes}
+
+\newtext{Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations compared using the Kit of Factor-Referenced Cognitive Tests \citet{ekstrom1976manual}.
+Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
+In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.}\footnote{That is, if comparison data was available for 9th grade students and 12th grade students, we have compared Iowa State students' scores with the 12th grade students.}
+
+\newtext{For the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.} Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
 This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
@@
 
-\newtext{The above consideration only assumes independence between questions, which is granted.
-While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.}
-\newtext{In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
-This approach also allows us to compare the test scores using similar orders of magnitude.}
+The above consideration only assumes independence between questions, which is reasonable.
+While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
+In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
+This approach also allows us to compare the test scores using similar orders of magnitude.
 

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Grammar fix.

@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} which parts of this test are drawn from.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
 

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Move comment, add more explanation

@@
 
+\newtext{The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual} and comparison data are available from the validation of these factors.
+The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).}
+
+% \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).}
 \begin{savenotes}
@@
 
-\newtext{Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations compared using the Kit of Factor-Referenced Cognitive Tests \citet{ekstrom1976manual}.
+\newtext{Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
 Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
-In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.}\footnote{That is, if comparison data was available for 9th grade students and 12th grade students, we have compared Iowa State students' scores with the 12th grade students.}
+In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.}\footnote{That is, if comparison data was available for 9th grade students and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
+When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.}
 
-\newtext{For the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.} Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+\newtext{As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.
+To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.} Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
 This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
@@
 &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-&=& n^2/(k-1).\\
-\text{Sd}(X_{n, k}) &=& n/sqrt{k-1}
+&=& n^2/(k-1).
 \end{eqnarray*}
@@
 pca$rotation
-
-# Just using the 4 cognitive tests and ignoring the lineups...
-pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
-summary(pca.cog.tests)
-screeplot(pca.cog.tests)
-pca.cog.tests$rotation
 @
 
-\comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).
-\textit{I'm not sure where to find the ``normal scores" in the factor-referenced manual.
-I'll check through it again, but I don't recall seeing it.}}
+<<pca, fig.cap="Plots of the principle components with observations.
+Visual Search and Paper Folding strongly contribute to both PC1 and PC2, while Figure Classification and Paper Folding strongly contributes to PC3 and Card Rotation strongly contributes to PC4", comment="|", fig.height=4.5, fig.width=4.5, warning=F, message=F, out.width='.45\\textwidth'>>=
+# Just using the 4 cognitive tests and ignoring the lineups...
+pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], retx=T)
+summary(pca.cog.tests)
+pca.cog.tests$rotation
+biplot(pca.cog.tests, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
+biplot(pca.cog.tests, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
+@
+
+<<pca-reg, comment="|", warning=F, message=F>>=
+# Linear regression using PC's. 
+ans.pca <- cbind(ans.summary[,c(1:19)], ans.summary["lineup"], pca.cog.tests$x)
+summary(lm(lineup~PC1+PC2+PC3+PC4, data=ans.pca))
+@
 

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Nicer t-test tables - getting rid of raw R output

@@
 
-<<t.test.sex, echo=T, comment="|">>=
-t.test(ans.summary$lineup[ans.summary$sex=="m"], ans.summary$lineup[ans.summary$sex=="f"])
+<<ttests.categ, echo=F, results='asis'>>=
+tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
+    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp)[1,]
+    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
+    rownames(coefs) <- NULL
+    coefs <- coefs[,c(6, 1:5)]
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
+    coefs
+})
+
+tmp <- rbind.fill(tmp)
+tmp <- tmp[order(tmp$p.value, decreasing=F),]
+library(xtable)
+print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3)), include.rownames=F)
 @
-<<t.test.calculus, echo=T, comment="|">>=
-t.test(ans.summary$lineup[ans.summary$calc_1=="y"], ans.summary$lineup[ans.summary$calc_1=="n"])
-@
-<<lm.vid.games, echo=T, comment="|">>=
-ans.summary$vidgame_hrs_factor_new <- factor(ans.summary$vidgame_hrs_factor, ordered=FALSE)
-summary(lm(data=ans.summary, lineup~vidgame_hrs_factor_new))
-# summary(lm(data=ans.summary, lineup~vidgame_hrs))
-@
-<<lm.all, echo=T, comment="|">>=
-summary(lm(data=ans.summary, lineup~sex+calc_1+vidgame_hrs_factor_new))
-@
+
 All results and data shown here are done in accordance with IRB \# 13-581.
@@
 T-tests of results for Hillary and Stephanie: 
-<<ttest, echo=T, include=T>>=
-t.test(ans.summary$card_rot[1:18], ans.summary$card_rot[-c(1:18)])
-t.test(ans.summary$folding[1:18], ans.summary$folding[-c(1:18)])
-t.test(ans.summary$lineup[1:18], ans.summary$lineup[-c(1:18)])
-t.test(ans.summary$vis_search[1:18], ans.summary$vis_search[-c(1:18)])
+<<ttest, echo=F, include=T, results='asis'>>=
+library(xtable)
+ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
+card.rotation <- t.test(card_rot~tester, data=ans.summary)
+paper.folding <- t.test(folding~tester, data=ans.summary)
+fig.class <- t.test(fig_class~tester, data=ans.summary)
+lineup <- t.test(lineup~tester, data=ans.summary)
+
+res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
+  tmp <- t.test(i~ans.summary$tester)
+  data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
+             t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
+})
+
+res$.id <- c("Card Rotation", "Paper Folding", "Figure Classification", "Lineups")
+names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
+
+print(xtable(res), include.rownames = F)
 @

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Change categorical variable interpretation as new data has been added.

@@
 
-There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class.
-There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
-In addition, completion of Calculus I is associated with increased performance on lineups (though completion of calculus is also associated with sex).
-AutoCAD experience is also not significantly associated with lineup performance; there is a difference in the medians, but it does not rise to the level of significance.
-There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and may be at least partially a result of the large difference in performance due to sex.
+There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
+% There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+Completion of Calculus I is associated with increased performance on lineups.
+This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
+There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
 

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Allow tidying of code, load xtable at the beginning.

@@
 
-opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, concordance=TRUE, autodep=TRUE, root.dir="../")
+opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|")
 datadir <- "../../Data/"
@@
 suppressMessages(library(gridExtra))
+library(xtable)
 source(paste(codedir, "Analysis.R", sep=""))

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Consistency with other tables

@@
 \begin{tabular}{rlll}
+\hline
   & Card Rotation & Paper Folding & Figure Classification \\\hline

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
xtable captions

@@
 
-<<pca-reg, comment="|", warning=F, message=F>>=
+<<pca-reg, comment="|", warning=F, message=F, results='asis', echo=F>>=
 # Linear regression using PC's. 
 ans.pca <- cbind(ans.summary[,c(1:19)], ans.summary["lineup"], pca.cog.tests$x)
-summary(lm(lineup~PC1+PC2+PC3+PC4, data=ans.pca))
+xtable(summary(lm(lineup~PC1+PC2+PC3+PC4, data=ans.pca)), caption="Lineup score, as predicted by principal components. Only the first principal component is a significant predictor of lineup score.")
 @
@@
 tmp <- tmp[order(tmp$p.value, decreasing=F),]
-library(xtable)
-print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3)), include.rownames=F)
+print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each demographic variable compared with lineup score.
+Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F)
 @
@@
 <<ttest, echo=F, include=T, results='asis'>>=
-library(xtable)
 ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
Remove comment that's no longer relevant.

@@
 cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
-# using scaled version right now, should be changed to unscaled once the scores are internally scaled.
+
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=F)
@@
 
-<<pca, fig.cap="Plots of the principle components with observations.
+<<pca, fig.cap="Plots of the principal components with observations.
 Visual Search and Paper Folding strongly contribute to both PC1 and PC2, while Figure Classification and Paper Folding strongly contributes to PC3 and Card Rotation strongly contributes to PC4", comment="|", fig.height=4.5, fig.width=4.5, warning=F, message=F, out.width='.45\\textwidth'>>=

### 2014-10-16 Paper/Draft1/TestingVisualAptitude.Rnw
additional explanation of significance of calculus completion.

@@
 This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
+\newtext{This result is consistent with \citep{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.}  There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
 

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
We don't know if the lineups are null or not...
scaling

@@
 Between cognitive tasks, participants will also complete three blocks of 20 lineups each.
-These lineups have been previously tested \citep{hofmann2012graphical} and include some null lineups (i.e.
+These lineups have been previously tested \citep{hofmann2012graphical} %and include some null lineups (i.e.
 lineups without a target plot).
-Participants %will
+. Participants %will 
 have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.
@@
 
-\newtext{The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual} and comparison data are available from the validation of these factors.
-The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).}
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual}
+\comment{the bib file is not updated on the repo}
+and comparison data are available from the validation of these factors.
+The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
 
@@
   & Card Rotation & Paper Folding & Figure Classification \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 2)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary.unscaled$folding), 2)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
-             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 2)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
+ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
+             & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
+             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
 Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
               & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 2)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617/2), 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
+              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
 Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
@@
 {\footnotesize Population}    
-              & \parbox[t]{.2\textwidth}{\footnotesize $\approx$ 550 male Naval recruits} 
+              & \parbox[t]{.2\textwidth}{\footnotesize approx, 550 male \\
+Naval recruits} 
               & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
@@
 
-\newtext{Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
+Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
 Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
-In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.}\footnote{That is, if comparison data was available for 9th grade students and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
+In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.\footnote{If comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, to closer in age to college students.
 When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.}
-
-\newtext{As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.
-To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.} Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+\comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions?
+I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation?
+$\text{Var}(aX) = a^2 \text{Var}(X)$      I just don't quite follow the code for the combined scores ...
+it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper.
+If you want to, you could instead extend the footnotes a bit for now.
+...
+so just let me try this: Let's assume that $X_M$ and $X_F$ are the scores based on 294 men, and 323 women (are those midscore s of the intervals reported?).
+with corresponding $s_M$ and $s_F$ for the reported standard deviation of the full test.
+Then half of the test would have half of the scores with half of the standard deviation.
+we have a breakdown of 17 girls and 19 boys.
+$
+X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
+$}
+As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.
+To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.
+Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
 This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
still tryying to understand the change in sd for the second set

@@
 I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation?
-$\text{Var}(aX) = a^2 \text{Var}(X)$      I just don't quite follow the code for the combined scores ...
+$\text{Var}(aX) = a^2 \text{Var}(X)$  ok I got the first one - because it is $X_1 + X_2$ not $2 X_1$.
+still working on understanding the second one, though ...
 it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper.
@@
 X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
-$}
+$. For the Variance of $X$ we then have:
+$\text{Var}(X) = \text{Var}(X_M)/2^2 \cdot (294/617)^2 + \text{Var}(X_F)/2^2 \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2^2 * (294/617)^2 + 27.8^2/2^2 * (323/617)^2),1)}^2$
+}
 As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
ok, I think I got it

@@
 {\footnotesize Population}    
-              & \parbox[t]{.2\textwidth}{\footnotesize approx, 550 male \\
-Naval recruits} 
+              & \parbox[t]{.2\textwidth}{\footnotesize approx. 550 male \\
+naval recruits} 
               & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\textwidth}{\footnotesize Suburban 11-12th grade students\\(288-300 males, 317-329 females)}\\\hline
+              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
 \end{tabular}
@@
 $. For the Variance of $X$ we then have:
-$\text{Var}(X) = \text{Var}(X_M)/2^2 \cdot (294/617)^2 + \text{Var}(X_F)/2^2 \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2^2 * (294/617)^2 + 27.8^2/2^2 * (323/617)^2),1)}^2$
+$\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
 }

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
outline of next couple of steps

@@
 {\footnotesize Population}    
-              & \parbox[t]{.2\textwidth}{\footnotesize approx. 550 male \\
+              & \parbox[t]{.2\textwidth}{\footnotesize approx.\ 550 male \\
 naval recruits} 
@@
 As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.
-To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.
-Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
-This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
+To facilitate this goal, for the remainder of this analysis, we will scale the test results \newtext{by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
+This scaling ensures that test scores across different types are comparable.
+All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1.
+In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.
+}
+%so that the ranges and units of test scores are comparable across the different types of tests.
+%Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+%This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
 
+\paragraph{Variance consideration of scaled test score:}
+Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+\newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
 
 The above consideration only assumes independence between questions, which is reasonable.
-While we only consider a test consisting of questions with the same number of choices, $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
-In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
+While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
+
+%In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
 This approach also allows us to compare the test scores using similar orders of magnitude.
 
-% \comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
-% Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}
-% 
-% \comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown.
-It would be good to scale everything, so that we have the same theoretical range.
-That will actually allow a comparison across different tests.}
-<<means>>=
-library(plyr)
-ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
-@
+\comment{OK, so on to the next steps ...
+we would like to figure out, how the lineups play into the established test scores.
+I think we need to talk about how to best talk through the visual search score.
+We have a lot of information on the other three scores - is there an established test in the naval test that is `like' the visual search?
+- i.e.
+could we use those scores to compare to?
+
+so outline for the next few steps:
+\begin{enumerate}
+\item some general words on behavior of lineup scores by themselves - I think the whole discussion that is at the moment  at the end of the paper should move up here.
+The levels of the video gaming are a bit messed up - they should be ordered according to number of games played rather than alphabetically :) - and I think that then the relationship actually is a monotonic one ...
+ In any case, sort the figures in the table according to their $p$-values in the table. Gender shows up as significantly different in my table of $p$ values, but in the text you say that it is not. 
+ The table with the t-tests should probably go into an appendix .... 
+\item figure 8 with either three or four scatterplots depending on how the visual search scores fit in
+\item a paragraph on discussing the multicollinearity between the test scores based on correlations (either, again, including visual search or not) - this should lead to the conclusion that we cannot directly use a linear model of lineup in test scores.
+\item PCA of test scores, with a bit of discussion
+\item linear model of lineup scores in PCA with discussion and tying it back to raw scores (figure classification really is the best predictor for lineups, but it only explains some of the variance).
+\end{enumerate}
+ }
+
+% <<means>>=
+% library(plyr)
+% ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
+% @
 

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
fix incorrect reference key.

@@
 
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual}
-\comment{the bib file is not updated on the repo}
-and comparison data are available from the validation of these factors.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude} and comparison data are available from the validation of these factors.
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
little bit more on methods section

@@
 \section{Methods}
+\comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected.
+part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)}
 Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete).

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
Appendix updates with more explicit equations.

--- 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-% \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).}
 \begin{savenotes}
@@
 \section*{Appendix}
+\subsection{Examining potential testing biases}
 T-tests of results for Hillary and Stephanie: 
@@
 @
+\subsection{Scaling Scores}
+To calculate ``scaled" comparison scores between tests which included different numbers of sections, we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
+The variance calculation is a bit more complicated: In the case described above, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
+
+\begin{savenotes}
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlll}
+\hline
+  & Card Rotation & Paper Folding & Figure Classification \\\hline
+ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
+             & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
+             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
+Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
+              & 13.8 (4.5)
+              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
+Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
+                & 13.8 (4.5)
+                & M:~120.0~(30.0), F:~114.9~(27.8)\\
+{\footnotesize Population}    
+              & \parbox[t]{.2\textwidth}{\footnotesize approx.\ 550 male \\
+naval recruits} 
+              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
+\end{tabular}
+
+\caption{Comparison of scores from Iowa State students and scores reported in \citet{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison}}
+\end{table}
+\end{savenotes}
+
+This scaling gets slightly more complicated for scores which have two sub-groups, as with the Figure Classification test.
+To get a single unified score with standard deviation, we did the following calculations:
+\begin{eqnarray}
+\mu_{all} &= (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
+\sigma_{all} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)}
+\end{eqnarray}
+Then, in order to account for the fact that ISU students took only part I of two parts to the Figure Classification test (and thus completed half of the questions), we adjusted the transformation as follows
+\begin{eqnarray}
+\mu_{all} &= 1/2 (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
+\sigma_{all} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(2(N_F + N_M))}
+\end{eqnarray}
 

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments

--- 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. 
+\comment{Teensy overview of the demographics - this might end up in an appendix, but then we need to refer to from here.
+I'd like to know breakdown of men/women, classification, maybe an overview of different majors (not necessarily a breakdown, that might be leading to problems with identifiability)}
 Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
@@
 
+\comment{could we also refer to the fact that these tests have been long established (maybe with references to earlier tests or updates?)}
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude} and comparison data are available from the validation of these factors.
+\comment{the well studied reference list needs to be a bit more extensive ... do you know this book by Michel Hersen?}
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
Fix  label repetition

--- 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison}}
+Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison2}}
 \end{table}

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
More comment responses.

@@
 
-\comment{could we also refer to the fact that these tests have been long established (maybe with references to earlier tests or updates?)}
+\comment{could we also refer to the fact that these tests have been long established (maybe with references to earlier tests or updates?) \textit{Yes, definitely.
+I'll pull up some references for that too...}}
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude} and comparison data are available from the validation of these factors.
-\comment{the well studied reference list needs to be a bit more extensive ... do you know this book by Michel Hersen?}
+\comment{the well studied reference list needs to be a bit more extensive ...
+do you know this book by Michel Hersen?\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
+It's definitely a work in progress.
+I haven't seen that book before, but I'm not at all surprised it exists.}}
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
More references...

@@
 \comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected.
-part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)}
+part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)\\
+\textit{I don't know if the tests are supposed to be orthogonal.
+It would be hard to construct a set of tests that was orthogonal, because the abilities themselves aren't easily separable - math reasoning and spatial reasoning are related, for instance, and verbal abilities are correlated with creativity too.}}
 Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete).
@@
 Between cognitive tasks, participants will also complete three blocks of 20 lineups each.
-These lineups have been previously tested \citep{hofmann2012graphical} %and include some null lineups (i.e.
-lineups without a target plot).
-. Participants %will 
-have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.
+These lineups have been previously tested \citep{hofmann2012graphical}.
+Participants have 5 minutes to complete each block of 20 lineups.
+Figure \ref{fig:lineup} shows a sample lineup of box plots.
 
@@
 I'll pull up some references for that too...}}
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude} and comparison data are available from the validation of these factors.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude,schaie1998longitudinal,hampson1990variations} and comparison data are available from the validation of these factors.
 \comment{the well studied reference list needs to be a bit more extensive ...
@@
 There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
-% There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
 Completion of Calculus I is associated with increased performance on lineups.

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
References for VST

@@
 Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison}}
+Data shown are from the population most similar to ISU students, out of the data available.
+The Visual Search task \citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+\label{tab:scorecomparison}}
 \end{table}

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
Move footnote into a full paragraph describing comparison populations.

@@
 Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
-In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.\footnote{If comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, to closer in age to college students.
-When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.}
+Values have been adjusted to accommodate differences in test procedures: some data is reported for a single part of a two-part test (Adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
+Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
+In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.
+
+Comparison populuation data was chosen in the following manner: if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
+When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.
+
 \comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions?
@@
 }
+
 As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
Table - add VST to table column

@@
 \centering
-\begin{tabular}{rlll}
+\begin{tabular}{rllll}
 \hline
-  & Card Rotation & Paper Folding & Figure Classification \\\hline
+  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
 ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
              & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
-             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
+             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
+             & \Sexpr{round(mean(ans.summary.unscaled$vis_search), 1)} (\Sexpr{round(sd(ans.summary.unscaled$vis_search), 1)})\\
 Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
               & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
+              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}
+              & N/A \\
 Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
                 & 13.8 (4.5)
-                & M:~120.0~(30.0), F:~114.9~(27.8)\\
+                & M:~120.0~(30.0), F:~114.9~(27.8)
+                & N/A\\
 {\footnotesize Population}    
-              & \parbox[t]{.2\textwidth}{\footnotesize approx.\ 550 male \\
-naval recruits} 
-              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
+              & \parbox[t]{.15\textwidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
+              & \parbox[t]{.15\textwidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
+              & N/A\\\hline
 \end{tabular}

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
Change appendix numbering

@@
 \section*{Appendix}
+\renewcommand{\thesubsection}{\Alph{subsection}}
 \subsection{Examining potential testing biases}
@@
 @
-\subsection{Scaling Scores}
+\subsection{Scaling Scores}\label{app:ScoreAdj}
 To calculate ``scaled" comparison scores between tests which included different numbers of sections, we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
Justify use of the VST, addition of references.

@@
 \begin{itemize}
-\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors. An example is shown in figure \ref{fig:VST}. 
+\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors.
+An example is shown in figure \ref{fig:VST}.
+The visual search task is similar in concept to lineups: it tests one's ability to find the target plot.
+Historically, it has been used as a measure of brain damage\citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in \citet{anderson1983interactive}).
+The similarity to lineup protocol as well as the simiplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
 \begin{figure}[htbp]\centering

### 2014-10-19 Paper/Draft1/TestingVisualAptitude.Rnw
clarify appendix form

@@
 It would be hard to construct a set of tests that was orthogonal, because the abilities themselves aren't easily separable - math reasoning and spatial reasoning are related, for instance, and verbal abilities are correlated with creativity too.}}
-Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once testing is complete).
+Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once we are sure there is no need for follow-up experiments).
 Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section.

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Lit review update

@@
 \item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
+\item \citet{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
+They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
+Test administration differences may account for the change.
+They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
+\comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
+Alternate scoring procedures still need to be investigated...
+they cite a poster, but it's not available anywhere online as far as I can tell.}}
+\item \citet{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
+They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
+\comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
+They also cite a bunch of studies (including \citet{voyer1995magnitude}) that I haven't had time to completely explore.}}
+\item \citet{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
+The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
+that enable viewers to compare and observe relations among variables".
+\begin{itemize}
+\item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
+Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
+In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
+The review also specifically discusses visuospatial ability, citing a speculation from \citet{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
+\comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
+\item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
+That is, graphics are a form of ``external cognition" \citep{scaife1996external} that guide, constrain, and facilitate cognitive behavior\citep{zhang1997nature}.
+This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\citep{zhang1994representations}.
+Graphical displays have higher search and computational efficiency than text \citep{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \citep{shah1999graphs}.
+\comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
+\item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
+Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
+This is pretty unlikely, and the literature is not so relevant to our cause.
+Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
+\end{itemize}
+\item \citet{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
+They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
+Their problem solving test was more practical (engineering-based), so isn't applicable to our study.
+They were testing instruction using visual + audio in sync, visual + audio out of sync, and then no instruction as a control.
+High-ability students could make more use of the in-sync presentation of visual and auditory information than low-ability students, but did not differ significantly in the out-of-sync information condition.
+Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
 \item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Subsection for scoring/scaling above the results section

@@
 I'd like to know breakdown of men/women, classification, maybe an overview of different majors (not necessarily a breakdown, that might be leading to problems with identifiability)}
+\subsection{Scoring and Scaling of Answers}
+
 Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
@@
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
+
+\paragraph{Variance consideration of scaled test score:}
+Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+\begin{eqnarray*}
+\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
+&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
+&=& n^2/(k-1).
+\end{eqnarray*}
+\newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
+
+The above consideration only assumes independence between questions, which is reasonable.
+While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
+
+To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
+This scaling ensures that test scores across different types are comparable.
+All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1.
+In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.
 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Rearrangement of the results section to better accommodate the newly outlined flow.

@@
 
-\section{Results}
-Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University. 
-\comment{Teensy overview of the demographics - this might end up in an appendix, but then we need to refer to from here.
-I'd like to know breakdown of men/women, classification, maybe an overview of different majors (not necessarily a breakdown, that might be leading to problems with identifiability)}
 \subsection{Scoring and Scaling of Answers}
@@
 I haven't seen that book before, but I'm not at all surprised it exists.}}
-The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
 
@@
 In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.
+
+\section{Results}
+Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
+About one-half of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, education, and liberal arts.
+Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
+This is reasonably representative of the university as a whole; ISU graduates are about 20\% engineering, 16\% business, and 11\% agriculture students, however, these numbers are for completed degrees, not current majors.
+\comment{Heike, can you get better enrollment numbers for current students? My guess is there are more people enrolled in engineering than actually graduate :). }
+% \comment{Teensy overview of the demographics - this might end up in an appendix, but then we need to refer to from here.
+I'd like to know breakdown of men/women, classification, maybe an overview of different majors (not necessarily a breakdown, that might be leading to problems with identifiability)}
+
+The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
+We compare Iowa State students' unscaled scores in table \ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
 
@@
 Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Values have been adjusted to accommodate differences in test procedures: some data is reported for a single part of a two-part test (Adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
+Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (Adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
 Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
@@
 }
-
-As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study.
-To facilitate this goal, for the remainder of this analysis, we will scale the test results \newtext{by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
-This scaling ensures that test scores across different types are comparable.
-All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1.
-In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.
-}
-%so that the ranges and units of test scores are comparable across the different types of tests.
-%Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
-%This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
-
-\paragraph{Variance consideration of scaled test score:}
-Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
-\begin{eqnarray*}
-\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
-&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-&=& n^2/(k-1).
-\end{eqnarray*}
-\newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
-
-The above consideration only assumes independence between questions, which is reasonable.
-While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Figure tweaks.
Transition text/comments and new paragraph heading.

@@
 
-\begin{figure}[htbp]\centering
-\includegraphics[width=.9\textwidth]{lineup}
+\begin{figure}\centering
+\includegraphics[width=.8\textwidth]{lineup}
 \caption[Sample Lineup]{A sample lineup. Participants are instructed to choose the plot which appears most different from the others. In this lineup, plot 13 is the target. }\label{fig:lineup}
@@
 \comment{Heike, can you get better enrollment numbers for current students? My guess is there are more people enrolled in engineering than actually graduate :). }
+
+\newtext{The testing was conducted by two research assistants; the first tested 18 participants, the second tested 20 participants.
+A comparison of results from each researcher is available in appendix \ref{app:ratests}.
+}
 % \comment{Teensy overview of the demographics - this might end up in an appendix, but then we need to refer to from here.
@@
 
+\paragraph{Comparison of Spatial Tests with Previously Validated Results}
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Tweaks to paragraph structure.

@@
 Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (Adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
+Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
 Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
@@
 
-Comparison populuation data was chosen in the following manner: if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
+Comparison populuation data was chosen to most closely match ISU undergraduate population demographics.
+Thus, if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
 When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.
 
-\comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions?
+\comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
+\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.}
+
+% 
+% \comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions?
 I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation?
@@
 $
-X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
-$. For the Variance of $X$ we then have:
-$\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
-}
+% X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
+% $. For the Variance of $X$ we then have:
+% $\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
+% }
 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
New paragraph about demographic characteristics.

@@
 
+\paragraph{Lineup Performance and Demographic Characteristics}
+
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
+lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
+lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Video Game Hrs/Wk"))
+lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
+lineup.summary.categorical$value[lineup.summary.categorical$value=="m"] <- "Male"
+lineup.summary.categorical$value[lineup.summary.categorical$value=="y"] <- "TRUE"
+lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"
+
+tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
+    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp)[1,]
+    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
+    rownames(coefs) <- NULL
+    coefs <- coefs[,c(6, 1:5)]
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
+    coefs
+})
+
+tmp <- rbind.fill(tmp)
+tmp <- tmp[order(tmp$p.value, decreasing=F),]
+
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
+
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+@
+\begin{figure}[htbp]\centering
+\includegraphics[width=\textwidth]{fig-VisReasoningCategorical-1}
+\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.
+Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a difference in lineup score.
+}\label{fig:visualaptitudecat}
+\end{figure}
+
+Completion of Calculus I is associated with increased performance on lineups.
+This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
+This association is consistent with \citep{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
+There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
+These demographic characteristics were chosen to account for life experience and personal skills which may have influenced the results.
+Statistical test results are available in appendix \ref{app:categoricalresults}.
+
 % <<means>>=

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Moving demographic information and code around.

@@
 
-<<VisReasoningAnalysis,echo=FALSE, include=FALSE, fig.width=8,fig.height=8, out.width='0.9\\textwidth', message=F, warning=F>>=
+<<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
 lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
@@
 qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Scaled Score") + ylab("Scaled Lineup Score") + geom_smooth(method="lm") + theme_bw()
-# same aspect ratio
-ggsave(paste(imgdir, "fig-VisReasoningAnalysisCts.png", sep=""), width=6, height=6)
-
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Video Game Hrs/Wk"))
-lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="m"] <- "Male"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="y"] <- "TRUE"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"
-
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
-ggsave(paste(imgdir, "fig-VisReasoningAnalysisCat.png", sep=""), width=6, height=6)
 @
@@
 \begin{figure}[htbp]\centering
-\includegraphics[width=0.75\textwidth]{fig-VisReasoningAnalysisCts}
+\includegraphics[width=0.75\textwidth]{fig-VisReasoningCts-1}
 \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
@@
 Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
-
-\begin{figure}[htbp]\centering
-\includegraphics[width=\textwidth]{fig-VisReasoningAnalysisCat}
-\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.
-Sex (male), Calculus completion, and hours spent playing video games per week are all associated with a difference in lineup score.
-}\label{fig:visualaptitudecat}
-\end{figure}
-
-There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
-There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
-Completion of Calculus I is associated with increased performance on lineups.
-This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-\newtext{This result is consistent with \citep{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.}  There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-
-<<ttests.categ, echo=F, results='asis'>>=
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each demographic variable compared with lineup score.
-Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F)
-@
 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Label RA tests in appendix

@@
 \renewcommand{\thesubsection}{\Alph{subsection}}
-\subsection{Examining potential testing biases}
+\subsection{Examining potential testing biases}\label{app:ratests}
 T-tests of results for Hillary and Stephanie: 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
New lineup demographic characteristics appendix w/ model data.

@@
 
+\subsection{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
+
+<<ttests.categ, echo=F, results='asis'>>=
+tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
+    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp)[1,]
+    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
+    rownames(coefs) <- NULL
+    coefs <- coefs[,c(6, 1:5)]
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
+    coefs
+})
+
+tmp <- rbind.fill(tmp)
+tmp <- tmp[order(tmp$p.value, decreasing=F),]
+print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each demographic variable compared with lineup score.
+Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F)
+@
+
 \end{document}

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Additional reference for factor referenced kit

@@
 I'll pull up some references for that too...}}
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude,schaie1998longitudinal,hampson1990variations} and comparison data are available from the validation of these factors.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude,schaie1998longitudinal,hampson1990variations,mayer1994whom} and comparison data are available from the validation of these factors.
 \comment{the well studied reference list needs to be a bit more extensive ...

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Label for scaling section
Graph to show scaled score results

@@
 
-\subsection{Scoring and Scaling of Answers}
+\subsection{Scoring and Scaling of Answers}\label{sec:scaling}
 
@@
 \comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
-\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.}
+\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Figure \ref{fig:ScaledScores} shows the range of possible scores and the observed score distribution.
+}
+<<ResultsScaledScores, echo=F, include=T, fig.cap="Scaled test scores for lineups and visuospatial tests.
+As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
+\\label{fig:ScaledScores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=
 
+score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
+score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
+qrange$sqrt.k <- sqrt(qrange$k-1)
+qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
+qplot(data=score.summary, x=Test, y=Score, geom="boxplot") + 
+  geom_rect(data=qrange, aes(xmin=as.numeric(Test)-.4, xmax=as.numeric(Test)+.4, ymin=-sqrt.k, ymax=sqrt.k, colour="red"), fill=NA, inherit.aes=F) + 
+  scale_colour_discrete(name="", labels="Theoretical Range") + 
+  theme(legend.position="bottom") + 
+  coord_flip()
+               # + theme_bw() #+ geom_jitter(aes(x=Test, y=Score))
+@
 % 

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Validity explanation

@@
 \end{figure}
+Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally (for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel).
+
 \item Card Rotation Task: tests participant's ability to rotate objects in two dimensions to distinguish between left-hand and right-hand versions of the same figure.
@@
 \end{figure}
+As with the paper folding test, two-dimensional comparisons are an important component of lineup performance.
+In some lineup situations, these comparisons involve translation, in other lineups, rotation is required.
+
 \item Figure Classification Task: tests participant's ability to extrapolate rules from provided figures.
@@
 \end{figure}
+The figure classification test requires the same types of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
+In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
+As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
 \end{itemize}

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Scaled PC's

@@
 # Just using the 4 cognitive tests and ignoring the lineups...
-pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], retx=T)
+pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], retx=T, scale=T)
 summary(pca.cog.tests)

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Unscaled answers

@@
 
-\subsection{Scoring and Scaling of Answers}\label{sec:scaling}
+\subsection{Test Scoring}\label{sec:scaling}
 
@@
 
-\paragraph{Variance consideration of scaled test score:}
-Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
-\begin{eqnarray*}
-\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
-&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-&=& n^2/(k-1).
-\end{eqnarray*}
-\newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
-
-The above consideration only assumes independence between questions, which is reasonable.
+% 
+% \paragraph{Variance consideration of scaled test score:}
+% Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
+% \begin{eqnarray*}
+% \text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
+% &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
+% &=& n^2/(k-1).
+% \end{eqnarray*}
+% \newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
+% 
+% The above consideration only assumes independence between questions, which is reasonable.
 While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
-
-To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
+% 
+% To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
 This scaling ensures that test scores across different types are comparable.
@@
 \comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
-\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
-Figure \ref{fig:ScaledScores} shows the range of possible scores and the observed score distribution.
+\newtext{Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Figure \ref{fig:Scores} shows the range of possible scores and the observed score distribution.
 }
-<<ResultsScaledScores, echo=F, include=T, fig.cap="Scaled test scores for lineups and visuospatial tests.
+<<ResultsScaledScores, echo=F, include=T, fig.cap="Test scores for lineups and visuospatial tests.
 As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
-\\label{fig:ScaledScores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=
+\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=
 

### 2014-10-20 Code/Analysis.R

@@ -102,9 +102,9 @@ longform.sum$value <- with(longform.sum, value/sqrt(var))
 # longform.sum$value <- with(longform.sum, (pos.pts - neg.pts + min.score)/(min.score+max.score)*100)
 
 # Cast back to wide-ish form with just test totals
-ans.summary <- dcast(longform.sum, id~testtype, value.var="value", na.rm=TRUE)
+ans.summary <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)
 ans.summary <- merge(ans[,1:19], ans.summary)
-pct.ans <- dcast(longform.sum, id~testtype, value.var="pct.answered")
+pct.ans <- dcast(longform.sum.unscaled, id~testtype, value.var="pct.answered")
 
 # Cast unscaled version
 ans.summary.unscaled <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Remove scaling from paper

@@
   & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
-             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
-             & \Sexpr{round(mean(ans.summary.unscaled$vis_search), 1)} (\Sexpr{round(sd(ans.summary.unscaled$vis_search), 1)})\\
+ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
+             & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
+             & \Sexpr{round(mean(ans.summary$fig_class), 1)} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
+             & \Sexpr{round(mean(ans.summary$vis_search), 1)} (\Sexpr{round(sd(ans.summary$vis_search), 1)})\\
 Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
@@
 
-<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".5\\linewidth", fig.height=4.5, fig.width=7.5, comment="|", warning=F, message=F>>=
+<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".45\\linewidth", fig.height=4.5, fig.width=4.5, comment="|", warning=F, message=F>>=
 cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
 
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=F)
+pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
 summary(pca)
@@
 
-<<pca, fig.cap="Plots of the principal components with observations.
-Visual Search and Paper Folding strongly contribute to both PC1 and PC2, while Figure Classification and Paper Folding strongly contributes to PC3 and Card Rotation strongly contributes to PC4", comment="|", fig.height=4.5, fig.width=4.5, warning=F, message=F, out.width='.45\\textwidth'>>=
-# Just using the 4 cognitive tests and ignoring the lineups...
-pca.cog.tests <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], retx=T, scale=T)
-summary(pca.cog.tests)
-pca.cog.tests$rotation
-biplot(pca.cog.tests, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
-biplot(pca.cog.tests, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
-@
-
-<<pca-reg, comment="|", warning=F, message=F, results='asis', echo=F>>=
-# Linear regression using PC's. 
-ans.pca <- cbind(ans.summary[,c(1:19)], ans.summary["lineup"], pca.cog.tests$x)
-xtable(summary(lm(lineup~PC1+PC2+PC3+PC4, data=ans.pca)), caption="Lineup score, as predicted by principal components. Only the first principal component is a significant predictor of lineup score.")
+<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\textwidth', comment="|", fig.cap="Plots of the principal components with observations.
+PC1 seems to represent a general ``visual intelligence'' factor.
+Figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
+Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.">>=
+biplot(pca, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
+biplot(pca, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
 @
@@
 
-qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Scaled Score") + ylab("Scaled Lineup Score") + geom_smooth(method="lm") + theme_bw()
+qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Score") + ylab("Lineup Score") + geom_smooth(method="lm") + theme_bw()
 @
@@
   & Card Rotation & Paper Folding & Figure Classification \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
-             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
+ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
+             & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
+             & \Sexpr{round(mean(ans.summary$fig_class), 1)} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
 Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})

### 2014-10-22 Code/Analysis.R
Remove scaling

@@ -88,7 +88,7 @@ longform.sum <- merge(longform.sum, qrange)
 longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
 
 # Save unscaled version
-longform.sum.unscaled <- longform.sum
+# longform.sum.unscaled <- longform.sum
 
 # Calculate test-wise means (for centering) and center scores
 # tmp <- ddply(longform.sum, .(testtype), summarize, test.mean = mean(value, na.rm=T))
@@ -96,19 +96,19 @@ longform.sum.unscaled <- longform.sum
 # longform.sum$value <- with(longform.sum, value - test.mean)
 
 # Scale scores
-longform.sum$value <- with(longform.sum, value/sqrt(var))
+# longform.sum$value <- with(longform.sum, value/sqrt(var))
 
 # Alt. Scaling method by range. 
 # longform.sum$value <- with(longform.sum, (pos.pts - neg.pts + min.score)/(min.score+max.score)*100)
 
 # Cast back to wide-ish form with just test totals
-ans.summary <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)
+ans.summary <- dcast(longform.sum, id~testtype, value.var="value", na.rm=TRUE)
 ans.summary <- merge(ans[,1:19], ans.summary)
-pct.ans <- dcast(longform.sum.unscaled, id~testtype, value.var="pct.answered")
+pct.ans <- dcast(longform.sum, id~testtype, value.var="pct.answered")
 
 # Cast unscaled version
-ans.summary.unscaled <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)
-ans.summary.unscaled <- merge(ans[,1:19], ans.summary.unscaled)
+# ans.summary.unscaled <- dcast(longform.sum.unscaled, id~testtype, value.var="value", na.rm=TRUE)
+# ans.summary.unscaled <- merge(ans[,1:19], ans.summary.unscaled)
 
 # Create video game factor variable
 ans.summary$vidgame_hrs_factor <- ordered(factor(rowSums(sapply(c(-1, 0, 1.99, 4.99), function(i) ans$vidgame_hrs>i)), labels=c("0", "[1, 2)", "[2, 5)", "5+"), levels=1:4))
[…]

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Code to sanitize latex table names

@@
 source(paste(codedir, "Analysis.R", sep=""))
+
+sanitize <- function(x){
+  str_replace(gsub("_", ".", x, fixed=T), "Pr\\((.{1,}))", "Pr$(\\1)$")
+}
+
 @

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
ETS Version

@@
 I'll pull up some references for that too...}}
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995magnitude,schaie1998longitudinal,hampson1990variations,mayer1994whom} and comparison data are available from the validation of these factors.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by \citet{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \citep{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \citep{educational1963kit}.
 \comment{the well studied reference list needs to be a bit more extensive ...

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Fix scaled score picture - unscaled scores

@@
 As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
-\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=
+We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
+\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.95\\textwidth', warning=F, message=F>>=
 
 score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
-score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
+score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
 qrange$sqrt.k <- sqrt(qrange$k-1)
-qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
-qplot(data=score.summary, x=Test, y=Score, geom="boxplot") + 
-  geom_rect(data=qrange, aes(xmin=as.numeric(Test)-.4, xmax=as.numeric(Test)+.4, ymin=-sqrt.k, ymax=sqrt.k, colour="red"), fill=NA, inherit.aes=F) + 
+qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
+ggplot() + 
+  geom_bar(data=qrange, aes(width=.9, x=Test, y=-n/(k-1), fill="red"), alpha=.5, colour="red", stat="identity", inherit.aes=F) + 
+  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, colour="red"), stat="identity", fill=NA, inherit.aes=F) + 
+  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F) +
+  scale_fill_discrete(name="", labels="Scores indicating guessing") +
   scale_colour_discrete(name="", labels="Theoretical Range") + 
-  theme(legend.position="bottom") + 
+  theme(legend.position="bottom", legend.box="horizontal") + 
+  ylab("Score") + 
   coord_flip()

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Nicer tables with xtable

@@
 
-<<correlations, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".45\\linewidth", fig.height=4.5, fig.width=4.5, comment="|", warning=F, message=F>>=
-cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")])
-
+<<correlations, echo=F, comment="|", warning=F, message=F, results='asis'>>=
+print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(3, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
+@
+<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".45\\linewidth", fig.height=4.5, fig.width=4.5, warning=F, message=F>>=
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 screeplot(pca)
-pca$rotation
+@
+<<pcarotation,echo=F, results='asis'>>=
+print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
 @
@@
 Figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
-Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.">>=
+Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}">>=
 biplot(pca, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Scatterplot matrix

@@
 
+<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
+Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
+Paper folding and card rotation scores are also highly correlated.">>=
+GGally::ggpairs(ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")], lower=list(continuous='smooth'), upper=list(continuous="cor"))
+@
+
 \begin{figure}[htbp]\centering

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Sanitize table names
Exploring partial regression using visual search scores.

@@
 
-print(xtable(res), include.rownames = F)
+print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
 @
@@
 @
+\subsection{Lineup Performance: Accounting for ``Visual Intelligence''}
+As figure \ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
+As the visual search task (VST) is the most generic of visual tasks (find the matching pattern) and requires no spatial manipulation, rotation, or inference, we might use it as an indicator of general visual intelligence (or cognitive tempo and attention to detail).
+
+<<partialregression, results='asis', echo=1>>=
+model1 <- lm(lineup ~ vis_search, data=ans.summary)
+print(xtable(summary(model1), caption="Results of the linear regression of visual search score on lineup score. \\label{vst-reg}."), sanitize.text.function=sanitize)
+@
+<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\textwidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
+Adjusted lineup scores are most associated with figure classification in the second principal component.
+In the third and fourth components, adjusted scores are almost entirely orthogonal.
+\\label{fig:biplot-adjusted}", tidy=T, echo=1:7>>=
+# partial regression data set
+ans.part <- ans.summary
+ans.part$lineup <- resid(lm(lineup ~ vis_search, data=ans.summary))
+ans.part$card_rot <- resid(lm(card_rot ~ vis_search, data=ans.summary))
+ans.part$fig_class <- resid(lm(fig_class ~ vis_search, data=ans.summary))
+ans.part$folding <- resid(lm(folding ~ vis_search, data=ans.summary))
+
+pca2 <- prcomp(ans.part[,c("lineup", "card_rot", "fig_class", "folding")], scale=T, retx=T)
+biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
+biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
+@
+<<partialregression2, echo=F, results='asis'>>=
+model2 <- lm(lineup~fig_class+card_rot+folding, data=ans.part)
+print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize)
+@
+Removing paper folding from the analysis due to its' obvious nonsignificance and collinearity with card rotation, we get a slightly better regression result. 
+<<partialregression3, echo=F, results='asis'>>=
+model3 <- lm(lineup~fig_class+card_rot, data=ans.part)
+print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize)
+print(xtable(anova(model3, model2), caption="Model comparison between full model (including paper folding) and reduced model (excluding paper folding), when accounting for scores on the VST in both independent and dependent variables.
+"), sanitize.text.function=sanitize)
+@
+
+Comparing the full and reduced models, we see that the reduction does not significantly reduce the explained sum of squares.
+
+Even using VST as a ``general intelligence" factor, the first PC still does not discriminate between the tests. This approach is probably not worth pursuing further.
 

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Nicer table for pca summary

@@
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-summary(pca)
 screeplot(pca)
@@
 <<pcarotation,echo=F, results='asis'>>=
+print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
 print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Move lit review list to the back
Add discuussion of spatial reasoning and statistical graphics.

@@
 \section{Introduction}
-Relevant literature: 
-\begin{itemize}
-\item \citet{shah1995conceptual} showed that spatial ability was not correlated with accuracy on a simple two-dimensional line graph description task, but that mathematical ability was correlated with accuracy.
-\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-\item \citet{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
-They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
-Test administration differences may account for the change.
-They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
-\comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
-Alternate scoring procedures still need to be investigated...
-they cite a poster, but it's not available anywhere online as far as I can tell.}}
-\item \citet{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
-They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
-\comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
-They also cite a bunch of studies (including \citet{voyer1995magnitude}) that I haven't had time to completely explore.}}
-\item \citet{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
-The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
-that enable viewers to compare and observe relations among variables".
-\begin{itemize}
-\item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
-Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
-In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
-The review also specifically discusses visuospatial ability, citing a speculation from \citet{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
-\comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
-\item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-That is, graphics are a form of ``external cognition" \citep{scaife1996external} that guide, constrain, and facilitate cognitive behavior\citep{zhang1997nature}.
-This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\citep{zhang1994representations}.
-Graphical displays have higher search and computational efficiency than text \citep{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \citep{shah1999graphs}.
-\comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
-\item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
-Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
-This is pretty unlikely, and the literature is not so relevant to our cause.
-Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
-\end{itemize}
-\item \citet{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
-They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
-Their problem solving test was more practical (engineering-based), so isn't applicable to our study.
-They were testing instruction using visual + audio in sync, visual + audio out of sync, and then no instruction as a control.
-High-ability students could make more use of the in-sync presentation of visual and auditory information than low-ability students, but did not differ significantly in the out-of-sync information condition.
-Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
-\end{itemize}
+\comment{Intro about statistical graphics, lineups, ...}
+\subsection{Spatial Reasoning and Statistical Graphics}
+Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
+Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
+Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli).
+Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}.
+Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
+Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \citep{lowrie2007solving}.
 
-Lineups depend on the ability to search for a signal amid distractors (Visual Search Task) and the ability to infer patterns from stimuli (Pattern Recognition task).
-Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation task) and mentally manipulate graphs (paper folding task).
-By breaking the lineup task down into component parts, we can correlate lineup performance with similar cognitive factor tests to determine where additional variation in skill level factors into performance differences.
-In addition, we can correlate previous experiences (science-based major, research experience, Auto-CAD skills) with performance to explore the effect that participant experience has on lineup performance.
+\subsection{The Lineup Protocol}
+\comment{More lineup introduction here}
+
+\newtext{Statistical lineups \citep{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
+By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
+These assessments will then be used to account for some of the variation in lineup performance differences.
+In addition, we will also examine the effect that previous experiences (science-based major, research experience, Auto-CAD skills) have on lineup performance and the implications for the use of lineups to assess the graphical display of statistical information.
+}
 
 \section{Methods}
+\comment{Use the setup in the lineup protocol section to justify each of the tests chosen. }
 \comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected.
@@
 
+\subsection{Lit Review Rejects}
+\begin{itemize}
+\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
+\item \citet{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
+They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
+Test administration differences may account for the change.
+They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
+\comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
+Alternate scoring procedures still need to be investigated...
+they cite a poster, but it's not available anywhere online as far as I can tell.}}
+\item \citet{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
+They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
+\comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
+They also cite a bunch of studies (including \citet{voyer1995magnitude}) that I haven't had time to completely explore.}}
+\item \citet{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
+The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
+that enable viewers to compare and observe relations among variables".
+\begin{itemize}
+\item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
+Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
+In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
+The review also specifically discusses visuospatial ability, citing a speculation from \citet{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
+\comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
+\item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
+That is, graphics are a form of ``external cognition" \citep{scaife1996external} that guide, constrain, and facilitate cognitive behavior\citep{zhang1997nature}.
+This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\citep{zhang1994representations}.
+Graphical displays have higher search and computational efficiency than text \citep{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \citep{shah1999graphs}.
+\comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
+\item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
+Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
+This is pretty unlikely, and the literature is not so relevant to our cause.
+Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
+\end{itemize}
+\item \citet{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
+They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
+Their problem solving test was more practical (engineering-based), so isn't applicable to our study.
+They were testing instruction using visual + audio in sync, visual + audio out of sync, and then no instruction as a control.
+High-ability students could make more use of the in-sync presentation of visual and auditory information than low-ability students, but did not differ significantly in the out-of-sync information condition.
+Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
+\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
+\end{itemize}
 \end{document}

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Fix reference

@@
 Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli).
+Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994whom} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli).
 Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}.

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Revision of list of tests into paragraphs

@@
 \section{Methods}
+\subsection{Measures of visuospatial ability}
 \comment{Use the setup in the lineup protocol section to justify each of the tests chosen. }
@@
 part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)\\
-\textit{I don't know if the tests are supposed to be orthogonal.
-It would be hard to construct a set of tests that was orthogonal, because the abilities themselves aren't easily separable - math reasoning and spatial reasoning are related, for instance, and verbal abilities are correlated with creativity too.}}
+}
+
 Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once we are sure there is no need for follow-up experiments).
@@
 This provides more discrimination between high scorers and prevents score compression at the top of the range.
-\begin{itemize}
-\item Visual Search Task: designed to test participants' ability to find a target stimulus in a field of distractors.
-An example is shown in figure \ref{fig:VST}.
+
+The \textbf{visual searching task} (VST), shown in figure \ref{fig:VST}, is designed to test participants' ability to find a target stimulus in a field of distractors.
 The visual search task is similar in concept to lineups: it tests one's ability to find the target plot.
@@
 \end{figure}
-\item Paper Folding Task: tests participants' ability to visualize and mentally manipulate figures in three dimensions.
-Associated with the ability to extrapolate symmetry and reflection over multiple steps.
-An example is shown in figure \ref{fig:paperfolding}.
-\begin{figure}[htbp]\centering
-\includegraphics[width=.9\textwidth]{paperfolding}
-\caption[Paper Folding Task]{Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure.
-Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
-\end{figure}
-Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally (for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel).
 
-\item Card Rotation Task: tests participant's ability to rotate objects in two dimensions to distinguish between left-hand and right-hand versions of the same figure.
-Tests spatial reasoning ability and mental rotation skills.
-An example is shown in figure \ref{fig:cardrotation}.
-\begin{figure}[htbp]\centering
-\includegraphics[width=.9\textwidth]{cardrotation}
-\caption[Card Rotation Task]{Card Rotation Task.
-Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line.
-Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
-\end{figure}
-As with the paper folding test, two-dimensional comparisons are an important component of lineup performance.
-In some lineup situations, these comparisons involve translation, in other lineups, rotation is required.
-
-\item Figure Classification Task: tests participant's ability to extrapolate rules from provided figures.
-This task is associated with visual reasoning capabilities and we expect that it should correlate with the ability to pick out a signal plot from a lineup.
+The \textbf{figure classification task} tests participant's ability to extrapolate rules from provided figures.
+This task is associated with inductive reasoning ablities (factor I in \citet{ekstrom1976manual}).
 An example is shown in figure \ref{fig:figureclassification}.
@@
 As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
-\end{itemize}
 
-Between cognitive tasks, participants will also complete three blocks of 20 lineups each.
-These lineups have been previously tested \citep{hofmann2012graphical}.
+The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
+It tests mental rotation skills, and is classified as a test of spatial orientation in \citet{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+An example is shown in figure \ref{fig:cardrotation}.
+\begin{figure}[htbp]\centering
+\includegraphics[width=.9\textwidth]{cardrotation}
+\caption[Card Rotation Task]{Card Rotation Task.
+Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line.
+Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
+\end{figure}
+Two-dimensional comparisons are an important component of lineup performance.
+In some lineup situations, these comparisons involve translation, but in other lineups, rotation is required.
+Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
+
+The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
+It is classified as part of the visualization factor in \citet{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure mentally; a more complex and demanding task than simple rotation.
+It is associated with the ability to extrapolate symmetry and reflection over multiple steps.
+An example is shown in figure \ref{fig:paperfolding}.
+\begin{figure}[htbp]\centering
+\includegraphics[width=.9\textwidth]{paperfolding}
+\caption[Paper Folding Task]{Paper Folding Task.
+Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure.
+Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
+\end{figure}
+Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally (for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel).
+
+Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies \citep{hofmann2012graphical,majumder2013validation}.
 Participants have 5 minutes to complete each block of 20 lineups.

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Showing history of ETS Kit tests

@@
 Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-
-\comment{could we also refer to the fact that these tests have been long established (maybe with references to earlier tests or updates?) \textit{Yes, definitely.
-I'll pull up some references for that too...}}
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by \citet{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \citep{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \citep{educational1963kit}.

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
More accurate numbers for ISU students

@@
 Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
-About one-half of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, education, and liberal arts.
+About one-half of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
-This is reasonably representative of the university as a whole; ISU graduates are about 20\% engineering, 16\% business, and 11\% agriculture students, however, these numbers are for completed degrees, not current majors.
-\comment{Heike, can you get better enrollment numbers for current students? My guess is there are more people enrolled in engineering than actually graduate :). }
-
+This is reasonably representative of the university as a whole; in the fall 2012 semester, 23\% of students were associated with the college of engineering, 23\% were associated with the college of liberal arts and sciences, 23\% were associated with the college of human sciences, 12\% with the business school, and 14\% with the school of agriculture.
+\footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}}
 \newtext{The testing was conducted by two research assistants; the first tested 18 participants, the second tested 20 participants.

### 2014-10-27 Paper/Draft1/TestingVisualAptitude.Rnw
Table fixes

@@
 \centering
-\begin{tabular}{rllll}
+\begin{tabular}{rlllc}
 \hline
@@
              & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
-             & \Sexpr{round(mean(ans.summary$fig_class), 1)} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
+             & \Sexpr{sprintf("%.1f", mean(ans.summary$fig_class))} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
              & \Sexpr{round(mean(ans.summary$vis_search), 1)} (\Sexpr{round(sd(ans.summary$vis_search), 1)})\\
@@
               & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}
-              & N/A \\
+              & -- \\
 Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
@@
                 & M:~120.0~(30.0), F:~114.9~(27.8)
-                & N/A\\
+                & --\\
 {\footnotesize Population}    
@@
               & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
-              & N/A\\\hline
+              & \\\hline
 \end{tabular}

### 2014-10-27 Paper/Draft1/TestingVisualAptitude.Rnw
remove unneccesary comment
Fix theming on score plot

@@
 
-\comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
 \newtext{Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
@@
   geom_bar(data=qrange, aes(width=.9, x=Test, y=n, colour="red"), stat="identity", fill=NA, inherit.aes=F) + 
-  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F) +
+  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
   scale_fill_discrete(name="", labels="Scores indicating guessing") +
   scale_colour_discrete(name="", labels="Theoretical Range") + 
+  theme_bw()  + 
   theme(legend.position="bottom", legend.box="horizontal") + 
@@
   coord_flip()
-               # + theme_bw() #+ geom_jitter(aes(x=Test, y=Score))
 @

### 2014-10-27 Paper/Draft1/TestingVisualAptitude.Rnw
Fix caption

@@
 <<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\textwidth', comment="|", fig.cap="Plots of the principal components with observations.
-PC1 seems to represent a general ``visual intelligence'' factor.
+PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here).
 Figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
 Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}">>=
-biplot(pca, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
-biplot(pca, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), adj=.75)
 @

### 2014-10-27 Paper/Draft1/TestingVisualAptitude.Rnw
Remove Appendix A discussing testing of research assistant differences

@@
 \footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}}
-\newtext{The testing was conducted by two research assistants; the first tested 18 participants, the second tested 20 participants.
-A comparison of results from each researcher is available in appendix \ref{app:ratests}.
-}
-% \comment{Teensy overview of the demographics - this might end up in an appendix, but then we need to refer to from here.
-I'd like to know breakdown of men/women, classification, maybe an overview of different majors (not necessarily a breakdown, that might be leading to problems with identifiability)}
 
@@
 \renewcommand{\thesubsection}{\Alph{subsection}}
-\subsection{Examining potential testing biases}\label{app:ratests}
-T-tests of results for Hillary and Stephanie: 
-<<ttest, echo=F, include=T, results='asis'>>=
-ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
-card.rotation <- t.test(card_rot~tester, data=ans.summary)
-paper.folding <- t.test(folding~tester, data=ans.summary)
-fig.class <- t.test(fig_class~tester, data=ans.summary)
-lineup <- t.test(lineup~tester, data=ans.summary)
-
-res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
-  tmp <- t.test(i~ans.summary$tester)
-  data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
-             t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
-})
-
-res$.id <- c("Card Rotation", "Paper Folding", "Figure Classification", "Lineups")
-names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
-
-print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
-@
+% \subsection{Examining potential testing biases}\label{app:ratests}
+% T-tests of results for Hillary and Stephanie: 
+% <<ttest, echo=F, include=T, results='asis'>>=
+% ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
+% card.rotation <- t.test(card_rot~tester, data=ans.summary)
+% paper.folding <- t.test(folding~tester, data=ans.summary)
+% fig.class <- t.test(fig_class~tester, data=ans.summary)
+% lineup <- t.test(lineup~tester, data=ans.summary)
+% 
+% res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
+%   tmp <- t.test(i~ans.summary$tester)
+%   data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
+%              t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
+% })
+% 
+% res$.id <- c("Card Rotation", "Paper Folding", "Figure Classification", "Lineups")
+% names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
+% 
+% print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
+% @
 \subsection{Scaling Scores}\label{app:ScoreAdj}

### 2014-10-27 Paper/Draft1/TestingVisualAptitude.Rnw
Fix caption for appendix table

@@
 
-<<ttests.categ, echo=F, results='asis'>>=
+<<ttests.categ, echo=F, eval=T, results='asis'>>=
 tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
@@
 tmp <- tmp[order(tmp$p.value, decreasing=F),]
-print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each demographic variable compared with lineup score.
+print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each single demographic variable compared with lineup score.
 Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F)

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
teensy bit of editing throughout the first section

@@
 
+library(knitr)
 opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|")
@@
 Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994whom} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli).
+Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994whom} hypothesize that individuals with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
 Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}.
@@
 
-\newtext{Statistical lineups \citep{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+Statistical lineups \citep{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
 Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
 By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-These assessments will then be used to account for some of the variation in lineup performance differences.
-In addition, we will also examine the effect that previous experiences (science-based major, research experience, Auto-CAD skills) have on lineup performance and the implications for the use of lineups to assess the graphical display of statistical information.
-}
+These assessments are then used to account for some of the variation in lineup performance differences.
+In addition, we also examine the effect that previous experiences (science-based major, research experience, Auto-CAD skills) have on lineup performance and the implications for the use of lineups to assess the graphical display of statistical information.
 
@@
 The \textbf{visual searching task} (VST), shown in figure \ref{fig:VST}, is designed to test participants' ability to find a target stimulus in a field of distractors.
-The visual search task is similar in concept to lineups: it tests one's ability to find the target plot.
+The visual search task is similar in concept to lineups: it tests an individual's ability to find the target plot.
 Historically, it has been used as a measure of brain damage\citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in \citet{anderson1983interactive}).
-The similarity to lineup protocol as well as the simiplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
+The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
 \begin{figure}[htbp]\centering
@@
 \caption[Visual Search Task]{Visual Search Task.
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled "Target".
+Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
 Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
@@
 
-The \textbf{figure classification task} tests participant's ability to extrapolate rules from provided figures.
+The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
 This task is associated with inductive reasoning ablities (factor I in \citet{ekstrom1976manual}).
@@
 \end{figure}
-The figure classification test requires the same types of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
+The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
 In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-It is classified as part of the visualization factor in \citet{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure mentally; a more complex and demanding task than simple rotation.
-It is associated with the ability to extrapolate symmetry and reflection over multiple steps.
+It is classified as part of the visualization factor in \citet{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure mentally, which makes it a more complex and demanding task than simple rotation.
+The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
 An example is shown in figure \ref{fig:paperfolding}.
@@
 \end{figure}
-Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally (for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel).
+Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
 
@@
 
-In addition to these tests, participants will complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
-These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
+\comment{XXX should we classify majors into STEM/non-STEM?
+- there's a ton of literature and interest that we could potentially tap into} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
 
@@
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by \citet{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \citep{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \citep{educational1963kit}.
-\comment{the well studied reference list needs to be a bit more extensive ...
-do you know this book by Michel Hersen?\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
+\comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
 It's definitely a work in progress.

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
fixed order of levels in last chart explicitly

@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
-\begin{figure}[htbp]\centering
-\includegraphics[width=.7\textwidth]{VisualSearch}
+\begin{figure}[hbpt]\centering
+\includegraphics[width=.65\textwidth]{VisualSearch}
 \caption[Visual Search Task]{Visual Search Task.
@@
 An example is shown in figure \ref{fig:figureclassification}.
-\begin{figure}[htbp]\centering
+\begin{figure}[hbpt]\centering
 \includegraphics[width=.9\textwidth]{figureclassification}
@@
 An example is shown in figure \ref{fig:cardrotation}.
-\begin{figure}[htbp]\centering
+\begin{figure}[hbpt]\centering
 \includegraphics[width=.9\textwidth]{cardrotation}
@@
 An example is shown in figure \ref{fig:paperfolding}.
-\begin{figure}[htbp]\centering
+\begin{figure}[hbpt]\centering
 \includegraphics[width=.9\textwidth]{paperfolding}
@@
 }
-<<ResultsScaledScores, echo=F, include=T, fig.cap="Test scores for lineups and visuospatial tests.
+<<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests.
 As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
@@
   theme(legend.position="bottom", legend.box="horizontal") + 
-  ylab("Score") + 
+  ylab("Test Scores") + 
+  xlab("") +
   coord_flip()
@@
 lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
+lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "18-20", "21-24", "25-28", "5+", "FALSE", "TRUE", "Female", "Male"))
 
@@
 @
-\begin{figure}[htbp]\centering
+\begin{figure}[hbpt]\centering
 \includegraphics[width=\textwidth]{fig-VisReasoningCategorical-1}
@@
 
-\begin{figure}[htbp]\centering
+\begin{figure}[hbpt]\centering
 \includegraphics[width=0.75\textwidth]{fig-VisReasoningCts-1}

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
couple more small edits

@@
 Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
+Values have been adjusted to accommodate for differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
 Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
@@
 
-Comparison populuation data was chosen to most closely match ISU undergraduate population demographics.
-Thus, if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students.
-When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.
+Comparison population data was chosen to most closely match ISU undergraduate population demographics.
+Thus, if comparison data was available for 9th and 12th grade students, scores of Iowa State students were compared to scores of 12th grade students, who are closer in age to college students.
+When data was available from college students and Army enlistees, comparisons of scores were based on other college students, as college students are more likely to have a similar gender distribution to ISU students.
 
-\newtext{Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
 Figure \ref{fig:Scores} shows the range of possible scores and the observed score distribution.
-}
+
 <<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests.
@@
 %
-% \comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions?
-I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation?
-$\text{Var}(aX) = a^2 \text{Var}(X)$  ok I got the first one - because it is $X_1 + X_2$ not $2 X_1$.
-still working on understanding the second one, though ...
-it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper.
-If you want to, you could instead extend the footnotes a bit for now.
-...
-so just let me try this: Let's assume that $X_M$ and $X_F$ are the scores based on 294 men, and 323 women (are those midscore s of the intervals reported?).
-with corresponding $s_M$ and $s_F$ for the reported standard deviation of the full test.
-Then half of the test would have half of the scores with half of the standard deviation.
-we have a breakdown of 17 girls and 19 boys.
-$
-% X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
-% $. For the Variance of $X$ we then have:
-% $\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
-% }
 
-%In order to ensure all tests have similar variance, we scaled test scores by the standard deviation under random guessing.
-This approach also allows us to compare the test scores using similar orders of magnitude.
+\comment{
 
-\comment{OK, so on to the next steps ...
-we would like to figure out, how the lineups play into the established test scores.
-I think we need to talk about how to best talk through the visual search score.
-We have a lot of information on the other three scores - is there an established test in the naval test that is `like' the visual search?
-- i.e.
-could we use those scores to compare to?
-
-so outline for the next few steps:
 \begin{enumerate}

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
first step to TVCG style file

@@
-\documentclass{article}
+\documentclass[10pt,journal,compsoc]{IEEEtran}
 %---------------------------------------------------
@@
 \begin{figure}[hbpt]\centering
-\includegraphics[width=.65\textwidth]{VisualSearch}
+\includegraphics[width=\linewidth]{VisualSearch}
 \caption[Visual Search Task]{Visual Search Task.
@@
 An example is shown in figure \ref{fig:figureclassification}.
-\begin{figure}[hbpt]\centering
-\includegraphics[width=.9\textwidth]{figureclassification}
+\begin{figure*}[hbpt]\centering
+\includegraphics[width=.9\linewidth]{figureclassification}
 \caption[Figure Classification Task]{Figure Classification Task.
@@
 Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
-\end{figure}
+\end{figure*}
 The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
@@
 An example is shown in figure \ref{fig:cardrotation}.
-\begin{figure}[hbpt]\centering
-\includegraphics[width=.9\textwidth]{cardrotation}
+\begin{figure*}[hbpt]\centering
+\includegraphics[width=.9\linewidth]{cardrotation}
 \caption[Card Rotation Task]{Card Rotation Task.
@@
 Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
-\end{figure}
+\end{figure*}
 Two-dimensional comparisons are an important component of lineup performance.
@@
 An example is shown in figure \ref{fig:paperfolding}.
-\begin{figure}[hbpt]\centering
-\includegraphics[width=.9\textwidth]{paperfolding}
+\begin{figure*}[hbpt]\centering
+\includegraphics[width=.9\linewidth]{paperfolding}
 \caption[Paper Folding Task]{Paper Folding Task.
@@
 Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
-\end{figure}
+\end{figure*}
 Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
@@
 \begin{figure}\centering
-\includegraphics[width=.8\textwidth]{lineup}
+\includegraphics[width=.8\linewidth]{lineup}
 \caption[Sample Lineup]{A sample lineup. Participants are instructed to choose the plot which appears most different from the others. In this lineup, plot 13 is the target. }\label{fig:lineup}
@@
 {\footnotesize Population}    
-              & \parbox[t]{.15\textwidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
-              & \parbox[t]{.15\textwidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
+              & \parbox[t]{.15\linewidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
+              & \parbox[t]{.15\linewidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.3\linewidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
               & \\\hline
@@
 We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.95\\textwidth', warning=F, message=F>>=
+\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.95\\linewidth', warning=F, message=F>>=
 
@@
 \begin{figure}[hbpt]\centering
-\includegraphics[width=\textwidth]{fig-VisReasoningCategorical-1}
+\includegraphics[width=\linewidth]{fig-VisReasoningCategorical-1}
 \caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.
@@
 
-<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\textwidth', comment="|", fig.cap="Plots of the principal components with observations.
+<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Plots of the principal components with observations.
 PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here).
@@
 \begin{figure}[hbpt]\centering
-\includegraphics[width=0.75\textwidth]{fig-VisReasoningCts-1}
+\includegraphics[width=0.75\linewidth]{fig-VisReasoningCts-1}
 \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
@@
 {\footnotesize Population}    
-              & \parbox[t]{.2\textwidth}{\footnotesize approx.\ 550 male \\
+              & \parbox[t]{.2\linewidth}{\footnotesize approx.\ 550 male \\
 naval recruits} 
-              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\textwidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
+              & \parbox[t]{.2\linewidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.3\linewidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
 \end{tabular}
@@
 @
-<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\textwidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
+<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
 Adjusted lineup scores are most associated with figure classification in the second principal component.

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
changes to style

@@
 \documentclass[10pt,journal,compsoc]{IEEEtran}
+%---------------------------------------------------
+% TVCG packages and header
+% *** CITATION PACKAGES ***
+%
+\ifCLASSOPTIONcompsoc
+  % IEEE Computer Society needs nocompress option
+  % requires cite.sty v4.0 or later (November 2003)
+  \usepackage[nocompress]{cite}
+\else
+  % normal IEEE
+  \usepackage{cite}
+\fi
+% *** GRAPHICS RELATED PACKAGES ***
+%
+\ifCLASSINFOpdf
+  \usepackage{graphicx}
+  % declare the path(s) where your graphic files are
+  \graphicspath{{Figure/}{Images/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  \DeclareGraphicsExtensions{.pdf,.jpg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+% *** MATH PACKAGES ***
+%
+\usepackage[cmex10]{amsmath}
+\usepackage{amssymb}
+
+% *** SUBFigure~PACKAGES ***
+%\ifCLASSOPTIONcompsoc
+  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
+%\else
+%  \usepackage[caption=false,font=footnotesize]{subfig}
+%\fi
+
+% *** FLOAT PACKAGES ***
+%
+\usepackage{dblfloatfix}
+% The latest version can be found at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+\usepackage{url}
+
 %---------------------------------------------------
@@
 \usepackage[dvipsnames,svgnames]{xcolor}
-\usepackage{wrapfig,float}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{url}
+%\usepackage{wrapfig,float}
+%\usepackage{caption}
+%\usepackage{subcaption}
+%\usepackage{url}
 \usepackage{ulem}
@@
 % Bibliography without numbers or labels
-\usepackage{natbib}
-\bibliographystyle{apa}
+%\usepackage{natbib}
+%\bibliographystyle{apa}
 
@@
 \newcommand{\comment}[2][inline]{\todo[color=SkyBlue, #1]{#2}} % for comments that may not be "to-do"s
-\newcommand{\mcomment}[1]{\todo[color=SkyBlue]{#1}} % for margin comments
+%\newcommand{\mcomment}[1]{\todo[color=SkyBlue]{#1}} % for margin comments
 \newcommand{\newtext}[1]{\todo[inline, color=White]{ \color{OliveGreen}{#1}}} % new text - not necessarily something to be done
 \newcommand{\move}[1]{\todo[inline, color=Lime]{#1}} % new to do item
-%
-%---------------------------------------------------
-
-%---------------------------------------------------
-%             Set Picture Locations
-\graphicspath{{Figure/}{Images/}}
 %
@@
 
+\begin{document}
 \title{Spatial Reasoning and Statistical Graphics}
+
 \author{Susan VanderPlas, Heike Hofmann} % , Stephanie De Graaf, Hillary Cheney ??
 
-\begin{document}
+\author{Susan~VanderPlas,
+        and~Heike~Hofmann,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
+\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. VanderPlas and H. Hofmann are with the Department
+of Statistics and Statistical Laboratory, Iowa State University, Ames,
+IA, 50011.\protect\\
+% note need leading \protect in front of \\ to get a newline within \thanks as
+% \\ is fragile and will error, could use \hfil\break instead.
+E-mail: {srvanderplas, hofmann}@iastate.edu}% <-this % stops an unwanted space
+\thanks{Manuscript received Month XX, 2015; revised Month XX, XXXX.}}
+
+% The paper headers
+\markboth{IEEE TRANSACTIONS ON VISUALIZATION AND COMPUTER GRAPHICS, No.~X, Month~YYYY}%
+{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Statistical Graphics}
+
+\IEEEtitleabstractindextext{%
+\begin{abstract}
+The abstract goes here.
+\end{abstract}
+
+% Note that keywords are not normally used for peerreview papers.
+\begin{IEEEkeywords}
+Computer Society, IEEEtran, journal, \LaTeX, paper, template.
+\end{IEEEkeywords}}
+
+% make the title area
 \maketitle
 
-\section{Introduction}
+\IEEEdisplaynontitleabstractindextext
+% \IEEEdisplaynontitleabstractindextext has no effect when using
+% compsoc or transmag under a non-conference mode.
+
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
+%
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
+
+\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
+
+% The very first letter is a 2 line initial drop letter followed
+% by the rest of the first word in caps (small caps for compsoc).
+% 
+% form to use if the first word consists of a single letter:
+% \IEEEPARstart{A}{demo} file is ....
+% 
+% Here we have the typical use of a "T" for an initial drop letter
+% and "HIS" in caps to complete the first word.
+\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
+for IEEE Computer Society journal papers produced under \LaTeX\ using
+IEEEtran.cls version 1.8a and later.
+
 \comment{Intro about statistical graphics, lineups, ...}
@@
 Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
-Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994whom} hypothesize that individuals with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}.
+Mathematical ability is important, even for the simplest graphs: XXX in \cite{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
+Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: XXX in \cite{mayer1994whom} hypothesize that individuals with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \cite{vekiri2002value}.
 Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \citep{lowrie2007solving}.
+Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \cite{lowrie2007solving}.
 
@@
 
-Statistical lineups \citep{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+Statistical lineups \cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
 Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
@@
 
-The \textbf{visual searching task} (VST), shown in figure \ref{fig:VST}, is designed to test participants' ability to find a target stimulus in a field of distractors.
+The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test participants' ability to find a target stimulus in a field of distractors.
 The visual search task is similar in concept to lineups: it tests an individual's ability to find the target plot.
-Historically, it has been used as a measure of brain damage\citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in \citet{anderson1983interactive}).
+Historically, it has been used as a measure of brain damage\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in XXX in \cite{anderson1983interactive}).
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
@@
 
-The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning ablities (factor I in \citet{ekstrom1976manual}).
-An example is shown in figure \ref{fig:figureclassification}.
+The \textbf{figure~classification task} tests a participant's ability to extrapolate rules from provided figures.
+This task is associated with inductive reasoning ablities (factor I in XXX in \cite{ekstrom1976manual}).
+An example is shown in figure~\ref{fig:figureclassification}.
 \begin{figure*}[hbpt]\centering
-\includegraphics[width=.9\linewidth]{figureclassification}
-\caption[Figure Classification Task]{Figure Classification Task.
-Participants classify each figure in the second row as belonging to group 1, 2, or 3 (if applicable).
+\includegraphics[width=.75\linewidth]{figureclassification}
+\caption[Figure~Classification Task]{Figure~Classification Task.
+Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
 Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
 \end{figure*}
-The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
+The figure~classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
 In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
-As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
+As such, the figure~classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
 
 The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in \citet{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in figure \ref{fig:cardrotation}.
+It tests mental rotation skills, and is classified as a test of spatial orientation in XXX in \cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+An example is shown in figure~\ref{fig:cardrotation}.
 \begin{figure*}[hbpt]\centering
-\includegraphics[width=.9\linewidth]{cardrotation}
+\includegraphics[width=.75\linewidth]{cardrotation}
 \caption[Card Rotation Task]{Card Rotation Task.
-Participants mark each figure on the right hand side as either the same or different than the figure on the left hand side of the dividing line.
+Participants mark each figure~on the right hand side as either the same or different than the figure~on the left hand side of the dividing line.
 Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-It is classified as part of the visualization factor in \citet{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure mentally, which makes it a more complex and demanding task than simple rotation.
+It is classified as part of the visualization factor in XXX in \cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
-An example is shown in figure \ref{fig:paperfolding}.
+An example is shown in figure~\ref{fig:paperfolding}.
 \begin{figure*}[hbpt]\centering
-\includegraphics[width=.9\linewidth]{paperfolding}
+\includegraphics[width=.75\linewidth]{paperfolding}
 \caption[Paper Folding Task]{Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown in the left-hand figure.
+Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
 Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
@@
 
-Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies \citep{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies \cite{hofmann2012graphical,majumder2013validation}.
 Participants have 5 minutes to complete each block of 20 lineups.
-Figure \ref{fig:lineup} shows a sample lineup of box plots.
+Figure~\ref{fig:lineup} shows a sample lineup of box plots.
 
 \begin{figure}\centering
-\includegraphics[width=.8\linewidth]{lineup}
+\includegraphics[width=\linewidth]{lineup}
 \caption[Sample Lineup]{A sample lineup. Participants are instructed to choose the plot which appears most different from the others. In this lineup, plot 13 is the target. }\label{fig:lineup}
@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \citep{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by \citet{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \citep{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \citep{educational1963kit}.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests \cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by XXX in \cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \cite{educational1963kit}.
 \comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
@@
 % 
-% \paragraph{Variance consideration of scaled test score:}
+% \subsection{Variance consideration of scaled test score:}
 % Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
@@
 % The above consideration only assumes independence between questions, which is reasonable.
-While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
+While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure~classification score.
 % 
@@
 
-\paragraph{Comparison of Spatial Tests with Previously Validated Results}
-The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
+\subsection{Comparison of Spatial Tests with Previously Validated Results}
+The card rotation, paper folding, and figure~classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
 We compare Iowa State students' unscaled scores in table \ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
@@
 \begin{savenotes}
-\begin{table}[ht]
+\begin{table*}[ht]
 \centering
@@
 \hline
-  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
+  & Card Rotation & Paper Folding & Figure~Classification & Visual Search  \\\hline
 ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
@@
 
-\caption{Comparison of scores from Iowa State students and scores reported in \citet{ekstrom1976manual}.
+\caption{Comparison of scores from Iowa State students and scores reported in XXX in \cite{ekstrom1976manual}.
 Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
 Data shown are from the population most similar to ISU students, out of the data available.
-The Visual Search task \citep{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+The Visual Search task \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
 \label{tab:scorecomparison}}
-\end{table}
+\end{table*}
 \end{savenotes}
@@
 Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
-Figure \ref{fig:Scores} shows the range of possible scores and the observed score distribution.
+Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
 
@@
 score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
-score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
+score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
 qrange$sqrt.k <- sqrt(qrange$k-1)
-qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
+qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
 ggplot() + 
@@
  The table with the t-tests should probably go into an appendix .... 
-\item figure 8 with either three or four scatterplots depending on how the visual search scores fit in
+\item figure~8 with either three or four scatterplots depending on how the visual search scores fit in
 \item a paragraph on discussing the multicollinearity between the test scores based on correlations (either, again, including visual search or not) - this should lead to the conclusion that we cannot directly use a linear model of lineup in test scores.
 \item PCA of test scores, with a bit of discussion
-\item linear model of lineup scores in PCA with discussion and tying it back to raw scores (figure classification really is the best predictor for lineups, but it only explains some of the variance).
+\item linear model of lineup scores in PCA with discussion and tying it back to raw scores (figure~classification really is the best predictor for lineups, but it only explains some of the variance).
 \end{enumerate}
@@
 
-\paragraph{Lineup Performance and Demographic Characteristics}
+\subsection{Lineup Performance and Demographic Characteristics}
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8,fig.height=5, message=F, warning=F>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x") + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
 @
-\begin{figure}[hbpt]\centering
+\begin{figure*}[hbpt]\centering
 \includegraphics[width=\linewidth]{fig-VisReasoningCategorical-1}
@@
 }\label{fig:visualaptitudecat}
-\end{figure}
+\end{figure*}
 
@@
 This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with \citep{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
+This association is consistent with \cite{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
 There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \cite{hausmann2000sex}.
 There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
@@
 PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here).
-Figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
+Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
 Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}">>=
@@
 
-In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
+In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure~classification, and paper folding tasks.
 This suggests that skills associated with visual reasoning ability are related to lineup performance.
@@
 <<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure Classification", "Paper Folding", "Visual Search"))
+lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure~Classification", "Paper Folding", "Visual Search"))
 
@@
 <<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
+Lineup scores are most highly correlated with figure~classification scores, and are also highly correlated with card rotation scores.
 Paper folding and card rotation scores are also highly correlated.">>=
@@
 
-Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
+Figure~\ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
 
@@
 
+\bibliographystyle{IEEEtran}
 \bibliography{references}
@@
 % 
-% res$.id <- c("Card Rotation", "Paper Folding", "Figure Classification", "Lineups")
+% res$.id <- c("Card Rotation", "Paper Folding", "Figure~Classification", "Lineups")
 % names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
@@
 \hline
-  & Card Rotation & Paper Folding & Figure Classification \\\hline
+  & Card Rotation & Paper Folding & Figure~Classification \\\hline
 ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
@@
 
-\caption{Comparison of scores from Iowa State students and scores reported in \citet{ekstrom1976manual}.
+\caption{Comparison of scores from Iowa State students and scores reported in XXX in \cite{ekstrom1976manual}.
 Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
@@
 
-This scaling gets slightly more complicated for scores which have two sub-groups, as with the Figure Classification test.
+This scaling gets slightly more complicated for scores which have two sub-groups, as with the Figure~Classification test.
 To get a single unified score with standard deviation, we did the following calculations:
@@
 \end{eqnarray}
-Then, in order to account for the fact that ISU students took only part I of two parts to the Figure Classification test (and thus completed half of the questions), we adjusted the transformation as follows
+Then, in order to account for the fact that ISU students took only part I of two parts to the Figure~Classification test (and thus completed half of the questions), we adjusted the transformation as follows
 \begin{eqnarray}
@@
 \subsection{Lineup Performance: Accounting for ``Visual Intelligence''}
-As figure \ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
+As figure~\ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
 As the visual search task (VST) is the most generic of visual tasks (find the matching pattern) and requires no spatial manipulation, rotation, or inference, we might use it as an indicator of general visual intelligence (or cognitive tempo and attention to detail).
@@
 @
-<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
-Adjusted lineup scores are most associated with figure classification in the second principal component.
+<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure~classification, and paper folding scores principal components analysis.
+Adjusted lineup scores are most associated with figure~classification in the second principal component.
 In the third and fourth components, adjusted scores are almost entirely orthogonal.
@@
 \begin{itemize}
-\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-\item \citet{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
+\item XXX in \cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
+\item XXX in \cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
 They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
@@
 they cite a poster, but it's not available anywhere online as far as I can tell.}}
-\item \citet{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
+\item XXX in \cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
 They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
 \comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
-They also cite a bunch of studies (including \citet{voyer1995magnitude}) that I haven't had time to completely explore.}}
-\item \citet{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
+They also cite a bunch of studies (including XXX in \cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
+\item XXX in \cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
 The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
@@
 In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
-The review also specifically discusses visuospatial ability, citing a speculation from \citet{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
+The review also specifically discusses visuospatial ability, citing a speculation from XXX in \cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
 \comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
 \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-That is, graphics are a form of ``external cognition" \citep{scaife1996external} that guide, constrain, and facilitate cognitive behavior\citep{zhang1997nature}.
-This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\citep{zhang1994representations}.
-Graphical displays have higher search and computational efficiency than text \citep{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \citep{shah1999graphs}.
+That is, graphics are a form of ``external cognition" \cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
+Graphical displays have higher search and computational efficiency than text \cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \cite{shah1999graphs}.
 \comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
@@
 \end{itemize}
-\item \citet{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
+\item XXX in \cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
 They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
@@
 Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
+\item XXX in \cite{hofmann2012graphical} for lineup stimuli and general lineup performance
 \end{itemize}

### 2014-10-29 Paper/Draft1/TestingVisualAptitude.Rnw
starting to be in a readable format again

@@
 options(replace.assign=TRUE,width=70,scipen=3)
-require(knitr)
 
@@
 library(plyr)
+library(grid)
 suppressMessages(library(gridExtra))
@@
 Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
-Mathematical ability is important, even for the simplest graphs: XXX in \cite{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph.
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: XXX in \cite{mayer1994whom} hypothesize that individuals with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \cite{vekiri2002value}.
+Mathematical ability is important, even for the simplest graphs:  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph.
+Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of  individuals with low spatial ability  on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
 Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \cite{lowrie2007solving}.
+Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
 
@@
 
-Statistical lineups \cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+Statistical lineups~\cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
 Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
@@
 The visual search task is similar in concept to lineups: it tests an individual's ability to find the target plot.
-Historically, it has been used as a measure of brain damage\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in XXX in \cite{anderson1983interactive}).
+Historically, it has been used as a measure of brain damage\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in~\cite{anderson1983interactive}).
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
@@
 The \textbf{figure~classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning ablities (factor I in XXX in \cite{ekstrom1976manual}).
+This task is associated with inductive reasoning ablities (factor I  in~\cite{ekstrom1976manual}).
 An example is shown in figure~\ref{fig:figureclassification}.
@@
 The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in XXX in \cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
 An example is shown in figure~\ref{fig:cardrotation}.
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-It is classified as part of the visualization factor in XXX in \cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
+It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
@@
 
-Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies \cite{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
 Participants have 5 minutes to complete each block of 20 lineups.
@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests \cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests \cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis by XXX in \cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors \cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit \cite{educational1963kit}.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
 \comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
@@
 The card rotation, paper folding, and figure~classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
-We compare Iowa State students' unscaled scores in table \ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
+We compare Iowa State students' unscaled scores in table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
 
@@
 \begin{table*}[ht]
+\caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.
+The Visual Search task~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+\label{tab:scorecomparison}}
 \centering
@@
 \end{tabular}                
-
-\caption{Comparison of scores from Iowa State students and scores reported in XXX in \cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.
-The Visual Search task \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
-\label{tab:scorecomparison}}
 \end{table*}
@@
 
-Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Values have been adjusted to accommodate for differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}).
+Table~\ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
+Values have been adjusted to accommodate for differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix~\ref{app:ScoreAdj}).
 Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
@@
 
-Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
 Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
@@
 We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.95\\linewidth', warning=F, message=F>>=
+\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
 
@@
   geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
-  scale_fill_discrete(name="", labels="Scores indicating guessing") +
-  scale_colour_discrete(name="", labels="Theoretical Range") + 
+  scale_colour_discrete(name="", labels="Theoretical\nRange") +
+  scale_fill_discrete(name="", labels="Scores\nindicating\nguessing") +
   theme_bw()  + 
-  theme(legend.position="bottom", legend.box="horizontal") + 
+#  theme(legend.position="bottom", legend.box="horizontal") + 
   ylab("Test Scores") + 
   xlab("") +
-  coord_flip()
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(plot.margin=unit(c(0,0,0,0),"mm"))
+  #coord_flip()
 @
@@
 
-\comment{
-
-\begin{enumerate}
-\item some general words on behavior of lineup scores by themselves - I think the whole discussion that is at the moment  at the end of the paper should move up here.
-The levels of the video gaming are a bit messed up - they should be ordered according to number of games played rather than alphabetically :) - and I think that then the relationship actually is a monotonic one ...
- In any case, sort the figures in the table according to their $p$-values in the table. Gender shows up as significantly different in my table of $p$ values, but in the text you say that it is not. 
- The table with the t-tests should probably go into an appendix .... 
-\item figure~8 with either three or four scatterplots depending on how the visual search scores fit in
-\item a paragraph on discussing the multicollinearity between the test scores based on correlations (either, again, including visual search or not) - this should lead to the conclusion that we cannot directly use a linear model of lineup in test scores.
-\item PCA of test scores, with a bit of discussion
-\item linear model of lineup scores in PCA with discussion and tying it back to raw scores (figure~classification really is the best predictor for lineups, but it only explains some of the variance).
-\end{enumerate}
- }
-
 \subsection{Lineup Performance and Demographic Characteristics}
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8,fig.height=5, message=F, warning=F>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5, message=F, warning=F>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + ylab("Scaled Lineup Score") + theme_bw() + geom_point(aes(x=value, y=lineup), shape=1, size=3)
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + 
+  facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + 
+  ylab("Scaled Lineup Score") + 
+  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
+  theme_bw() + 
+  theme(plot.margins=unit(c(0,0,0,0), unit="mm"))
 @
 \begin{figure*}[hbpt]\centering
-\includegraphics[width=\linewidth]{fig-VisReasoningCategorical-1}
-\caption[Visual Aptitude Study Results]{Sample demographic characteristics compared with lineup score.
+\includegraphics[width=.9\linewidth]{fig-VisReasoningCategorical-1}
+\caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
 Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a difference in lineup score.
@@
 This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with \cite{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
+This association is consistent with~\cite{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
 There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \cite{hausmann2000sex}.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
 There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
 These demographic characteristics were chosen to account for life experience and personal skills which may have influenced the results.
-Statistical test results are available in appendix \ref{app:categoricalresults}.
+Statistical test results are available in appendix~\ref{app:categoricalresults}.
 
@@
 \begin{savenotes}
-\begin{table}[ht]
+\begin{table*}[hbt]
+\caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison2}}
+
 \centering
@@
 \end{tabular}                
-
-\caption{Comparison of scores from Iowa State students and scores reported in XXX in \cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison2}}
-\end{table}
+\end{table*}
 \end{savenotes}
@@
 \begin{itemize}
-\item XXX in \cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-\item XXX in \cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
+\item authors in~\cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
+\item authors in~\cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
 They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
@@
 they cite a poster, but it's not available anywhere online as far as I can tell.}}
-\item XXX in \cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
+\item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
 They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
 \comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
-They also cite a bunch of studies (including XXX in \cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
-\item XXX in \cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
+They also cite a bunch of studies (including authors in~\cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
+\item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
 The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
@@
 In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
-The review also specifically discusses visuospatial ability, citing a speculation from XXX in \cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
+The review also specifically discusses visuospatial ability, citing a speculation from authors in~\cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
 \comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
 \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-That is, graphics are a form of ``external cognition" \cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+That is, graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
 This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
-Graphical displays have higher search and computational efficiency than text \cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information \cite{shah1999graphs}.
+Graphical displays have higher search and computational efficiency than text~\cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information~\cite{shah1999graphs}.
 \comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
@@
 \end{itemize}
-\item XXX in \cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
+\item authors in~\cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
 They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
@@
 Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-\item XXX in \cite{hofmann2012graphical} for lineup stimuli and general lineup performance
+\item authors in~\cite{hofmann2012graphical} for lineup stimuli and general lineup performance
 \end{itemize}

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Formatting changes to accommodate two-column layout

@@
 
-options(replace.assign=TRUE,width=70,scipen=3)
+options(replace.assign=TRUE,width=40,scipen=3)
 
@@
   theme_bw()  + 
-#  theme(legend.position="bottom", legend.box="horizontal") + 
   ylab("Test Scores") + 
   xlab("") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position=c(.75,.8)) +
   theme(plot.margin=unit(c(0,0,0,0),"mm"))
@@
   theme_bw() + 
-  theme(plot.margins=unit(c(0,0,0,0), unit="mm"))
+  theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
 @
@@
 
-<<correlations, echo=F, comment="|", warning=F, message=F, results='asis'>>=
-print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(3, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
+<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize">>=
+print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9)
 @
-<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".45\\linewidth", fig.height=4.5, fig.width=4.5, warning=F, message=F>>=
+<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".8\\linewidth", fig.height=3, fig.width=4.5, warning=F, message=F, size="footnotesize">>=
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 <<pcarotation,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
+print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize")
+print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
 @
@@
 Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
-Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}">>=
+Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}", echo=F>>=
 biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
More slight formatting changes.
Additional cognitive load literature + paragraph break

@@
 
-options(replace.assign=TRUE,width=40,scipen=3)
+options(replace.assign=TRUE,width=30,scipen=3)
 
@@
 % \\ is fragile and will error, could use \hfil\break instead.
-E-mail: {srvanderplas, hofmann}@iastate.edu}% <-this % stops an unwanted space
+E-mail: {skoons, hofmann}@iastate.edu}% <-this % stops an unwanted space
 \thanks{Manuscript received Month XX, 2015; revised Month XX, XXXX.}}
@@
 Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of  individuals with low spatial ability  on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+\newtext{It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+This constraint reduces memory load and makes more cognitive resources available for other tasks, but also implicates visual ability as an important factor in graph comprehension.
+}
+
 Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
 Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
+In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Citations for Mahbub and Niladri's arxiv papers.

@@
 By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-These assessments are then used to account for some of the variation in lineup performance differences.
-In addition, we also examine the effect that previous experiences (science-based major, research experience, Auto-CAD skills) have on lineup performance and the implications for the use of lineups to assess the graphical display of statistical information.
+\newtext{Demographic factors are known to contribute to lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender affected the amount of time spent on lineups \cite{humanfactorslineups}.}\newtext{In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.}\newtext{In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability, with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
+}
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Remove unnecessary comments

@@
 \subsection{Measures of visuospatial ability}
-\comment{Use the setup in the lineup protocol section to justify each of the tests chosen. }
-\comment{Besides pointing out what we asked participants to do, we also need some why, and also why in particular these tests were selected.
-part of the validity consideration can go in here - and the classification from the naval test of the tests we picked into different classes (maybe already with a hint that for our data these weren't actually orthogonal classes, because we found a really high correlation between card rotation and paper folding.)\\
-}
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Wording change...
Comment addition

@@
 
-Participants will complete the following tasks (sample pictures included, full stimuli set will be added to the appendix once we are sure there is no need for follow-up experiments).
+Participants will complete several cognitive tests designed to measure spatial and reasoning ability.
 Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section.
@@
 \comment{XXX should we classify majors into STEM/non-STEM?
-- there's a ton of literature and interest that we could potentially tap into} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+- there's a ton of literature and interest that we could potentially tap into \\\textit{I'm not opposed, but we have to decide where ag-majors fall...
+not everything is easily classifiable.
+Psychology is another problem: I was very STEM-psychology, but you don't have to be.}} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Slight caption fix

@@
 \caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
-Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a difference in lineup score.
+Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a significant difference in lineup score.
 }\label{fig:visualaptitudecat}

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
tidying

@@
 
-<<partialregression, results='asis', echo=1>>=
+<<partialregression, results='asis', echo=1, tidy=T>>=
 model1 <- lm(lineup ~ vis_search, data=ans.summary)
@@
 \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-That is, graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-This constraint reduces memory load and makes more cognitive resources available for other tasks; displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
+Displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
 Graphical displays have higher search and computational efficiency than text~\cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information~\cite{shah1999graphs}.
-\comment{\textit{I am working on making my way through all of this literature - right now, I'm re-citing stuff; I will print the papers out and add to this ASAP once I get toner into the printer.}}
 \item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
small edits

@@
 IEEEtran.cls version 1.8a and later.
+
+\comment{Next few steps for the analysis: (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; (2)}
 
@@
 We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
+\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='.\\linewidth', warning=F, message=F>>=
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
next couple of steps

@@
 
-\comment{Next few steps for the analysis: (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; (2)}
+\comment{Next few steps for the analysis: (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; (2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
+not sure that that will pay off, but it's interesting.
+(3) in terms of why we would use lineups in a test comparison ...
+lineups give us a way to say 'this is correct' which regular graphics don't.
+So we should put a spin sentence on at the intro and the back that we are suggesting to use lineups as a tool to not only evaluate individuals (which we can see is possible) but also as a tool to evaluate charts ...
+which we've done before ...
+and tie it into the regular testing literature ...}
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Move figure around
Transition paragraph

--- 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
-Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
-
 <<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests.
@@
 %
+
+Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
 
@@
 \end{figure*}
+
+\newtext{Next, we examine the effect of demographic factors on lineup performance.
+Previous work found a relationship between lineup performance and factors such as education level, country of origin, and age\cite{humanfactorslineups}; our participant population is more homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
+}
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Caption placement and figure environment/location changes.

@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
-\begin{figure}[hbpt]\centering
+\begin{figure}[hbt]\centering
 \includegraphics[width=\linewidth]{VisualSearch}
@@
 An example is shown in figure~\ref{fig:figureclassification}.
-\begin{figure*}[hbpt]\centering
+\begin{figure*}[hbt]\centering
 \includegraphics[width=.75\linewidth]{figureclassification}
@@
 An example is shown in figure~\ref{fig:cardrotation}.
-\begin{figure*}[hbpt]\centering
+\begin{figure*}[hbt]\centering
 \includegraphics[width=.75\linewidth]{cardrotation}
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
+A sample question from the test is shown in figure~\ref{fig:paperfolding}.
 It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
-An example is shown in figure~\ref{fig:paperfolding}.
-\begin{figure*}[hbpt]\centering
+\begin{figure*}[hbt]\centering
 \includegraphics[width=.75\linewidth]{paperfolding}
@@
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
-\comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
+% \comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
 It's definitely a work in progress.
@@
 \begin{savenotes}
-\begin{table*}[ht]
+\begin{table*}[htb]
 \caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5, message=F, warning=F>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5, message=F, warning=F, fig.pos='htb'>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 @
-\begin{figure*}[hbpt]\centering
+\begin{figure*}[h!tb]\centering
 \includegraphics[width=.9\linewidth]{fig-VisReasoningCategorical-1}
@@
 <<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize">>=
-print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9)
+print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9, table.placement="htb", caption.placement="top")
 @
@@
 <<pcarotation,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize")
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
+print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
 @
@@
 Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
-Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}", echo=F>>=
+Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}", echo=F, fig.env="figure*">>=
 biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)
@@
 
-\begin{figure}[hbpt]\centering
+\begin{figure}[hbt]\centering
 \includegraphics[width=0.75\linewidth]{fig-VisReasoningCts-1}
@@
 print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each single demographic variable compared with lineup score.
-Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F)
+Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F, floating.environment="table*", caption.placement="top")
 @
@@
 model2 <- lm(lineup~fig_class+card_rot+folding, data=ans.part)
-print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize)
+print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
 @
@@
 model3 <- lm(lineup~fig_class+card_rot, data=ans.part)
-print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize)
+print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
 print(xtable(anova(model3, model2), caption="Model comparison between full model (including paper folding) and reduced model (excluding paper folding), when accounting for scores on the VST in both independent and dependent variables.
-"), sanitize.text.function=sanitize)
+"), sanitize.text.function=sanitize, caption.placement="top")
 @

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Fix age categories

@@
 lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
-lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "18-20", "21-24", "25-28", "5+", "FALSE", "TRUE", "Female", "Male"))
+lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
 

### 2014-10-30 Code/Analysis.R

@@ -4,7 +4,8 @@
 ans <- read.csv(paste0(datadir, "VisualGraphicsData.csv"), na.strings=c(" ", ""), stringsAsFactors=F)
 # Create anon. IDs that are unique
 ans$id <- 1:nrow(ans)
-
+# Fix age categories
+ans$age <- gsub("21-24", "21+", gsub("25-28", "21+", ans$age))
 
 # Create a key that has the same structure as the answers
 key <- ans[1,]

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Rearrange categorical plot text

@@
 
-Completion of Calculus I is associated with increased performance on lineups.
-This may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
+Figure~\ref{fig:visualaptitudecat} shows participants' lineup scores as a function of their responses to the questionnaire given at the beginning of the study; these demographic questions allow us to explore relationships between demographic characteristics (major, research experience, etc.) and test performance.
+
+Completion of Calculus I is associated with increased performance on lineups; this may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
 This association is consistent with~\cite{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
@@
 Statistical test results are available in appendix~\ref{app:categoricalresults}.
-
 % <<means>>=

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
New section, rearranging plots, commenting out unnecessary plots

@@
 
-<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize">>=
-print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9, table.placement="htb", caption.placement="top")
+\subsection{Understanding Visual Abilities and Lineup Performance}
+<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize", include=F>>=
+print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests.\\label{tab:corrmatrix5}"), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9, table.placement="htb", caption.placement="top")
 @
+
+<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
+Lineup scores are most highly correlated with figure~classification scores, and are also highly correlated with card rotation scores.
+Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
+GGally::ggpairs(ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")], lower=list(continuous='smooth'), upper=list(continuous="cor"))
+@
+
 <<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".8\\linewidth", fig.height=3, fig.width=4.5, warning=F, message=F, size="footnotesize">>=
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-screeplot(pca)
+# screeplot(pca)
 @
+<<pcasummary,echo=F, results='asis'>>=
+print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+@
+
 <<pcarotation,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.\\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
 @
@@
 
-<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure~classification scores, and are also highly correlated with card rotation scores.
-Paper folding and card rotation scores are also highly correlated.">>=
-GGally::ggpairs(ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")], lower=list(continuous='smooth'), upper=list(continuous="cor"))
-@
-
-\begin{figure}[hbt]\centering
-\includegraphics[width=0.75\linewidth]{fig-VisReasoningCts-1}
-\caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
+% 
+% \begin{figure}[hbt]\centering
+% \includegraphics[width=0.75\linewidth]{fig-VisReasoningCts-1}
+% \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
 There is a relatively strong positive correlation between lineup score and scores on visuospatial reasoning tests.}\label{fig:visualaptitudects}
-\end{figure}
-
-Figure~\ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.
+% \end{figure}
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
More references to stat graphics/visual learning papers.

@@
 An example is shown in figure~\ref{fig:cardrotation}.
+\newtext{The card rotation test is often used in studies of the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.}
 \begin{figure*}[hbt]\centering
@@
 Two-dimensional comparisons are an important component of lineup performance.
-In some lineup situations, these comparisons involve translation, but in other lineups, rotation is required.
+In some lineup situations, these comparisons sometimes involve translation, but in other lineups, rotation is required.
 Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
PCA section

@@
 @
+
+\newtext{Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are influenced by factors other than visual ability - general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few.
+In order to assess factors contributing to lineup performance, we first will examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores.}
+\newtext{Table \ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance.
+As expected, PC1 accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, that is, all of the general factors discussed above which influence test results.
+We will focus on the remaining components; these show differences in performance on the tests, and thus provide information about the skills needed to successfully examine lineups.
+}
+
+\newtext{Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}, and the corresponding rotation matrix is provided in \ref{tab:PCArotation5}.
+PC1 seems to represent a general ``visual intelligence'' factor.}
+
 <<pcasummary,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
 @
+
+\newtext{Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+}
 
@@
 
-<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Plots of the principal components with observations.
-PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here).
-Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
-Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}", echo=F, fig.env="figure*">>=
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), adj=.75)
+\newtext{The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.}
+\newtext{PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.}
+<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+@
+<<pcasummary2,echo=F, results='asis'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+@
+\newtext{If we examine the relationship between the four cognitive tests (excluding lineups), we can determine how lineup performance affects the variance partitions.
+Table \ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
+}
+
+\newtext{As with the previous principal components analysis, PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure \ref{fig:biplots4} and table \ref{tab:PCArotation4} confirm that PC1 is again a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test; this is incredibly similar to the role of PC2 in the 5-component analysis.
+}
+
+<<pcarotation2,echo=F, results='asis'>>=
+print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
 @
 
-In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure~classification, and paper folding tasks.
+\newtext{PC3 differentiates the figure classification and visual search tests from the paper folding test; this is true in the 5-component analysis as well, but in that analysis the role of the figure classification test was not as significant due to the inclusion of the lineup task.
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task, much like PC5 in the full PCA.}
+
+\newtext{Comparing the two principal component analyses in this manner suggests that lineups are a separate dimension (as the addition of lineups result in an additional principal component which accounts for at least 10\% of the variance), but that they are most closely related to the figure classification task, and less closely related to the visual searching task.}
+
+\newtext{This emphasizes the underpinnings of lineups: they are a visual medium, but they ultimately are a classification task, presented in a graphical manner.
+Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.}
+\comment{that last sentance is probably reaching a bit far...}
+
+<<pca2, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
+PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
+\\label{fig:biplots4}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
+@
+
+% In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure~classification, and paper folding tasks.
 This suggests that skills associated with visual reasoning ability are related to lineup performance.
@@
 Even excluding the one low visual search score that is a high-leverage point, there seems to be a positive relationship between a participant's score on lineups and their score for visual search.
-
-<<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure~Classification", "Paper Folding", "Visual Search"))
-
-qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Score") + ylab("Lineup Score") + geom_smooth(method="lm") + theme_bw()
-@
+% 
+% <<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
+% lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure~Classification", "Paper Folding", "Visual Search"))
+%  
+% qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Score") + ylab("Lineup Score") + geom_smooth(method="lm") + theme_bw()
+% @
 

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Phrasing of IRB statement
Fixing appendices

@@
 
-All results and data shown here are done in accordance with IRB \# 13-581.
+All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
 
@@
 
-\section*{Appendix}
-\renewcommand{\thesubsection}{\Alph{subsection}}
-% \subsection{Examining potential testing biases}\label{app:ratests}
+\appendices
+
+% \section{Examining potential testing biases}\label{app:ratests}
 % T-tests of results for Hillary and Stephanie: 
@@
 % @
-\subsection{Scaling Scores}\label{app:ScoreAdj}
-To calculate ``scaled" comparison scores between tests which included different numbers of sections, we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
-The variance calculation is a bit more complicated: In the case described above, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
 
-\begin{savenotes}
-\begin{table*}[hbt]
-\caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.\label{tab:scorecomparison2}}
-
-\centering
-\begin{tabular}{rlll}
-\hline
-  & Card Rotation & Paper Folding & Figure~Classification \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
-             & \Sexpr{round(mean(ans.summary$fig_class), 1)} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
-Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
-              & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
-Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
-                & 13.8 (4.5)
-                & M:~120.0~(30.0), F:~114.9~(27.8)\\
-{\footnotesize Population}    
-              & \parbox[t]{.2\linewidth}{\footnotesize approx.\ 550 male \\
-naval recruits} 
-              & \parbox[t]{.2\linewidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\linewidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}\\\hline
-\end{tabular}                
-\end{table*}
-\end{savenotes}
+\section{Scaling Scores}\label{app:ScoreAdj}
+To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table \ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
+The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
 
 This scaling gets slightly more complicated for scores which have two sub-groups, as with the Figure~Classification test.
-To get a single unified score with standard deviation, we did the following calculations:
+To get a single unified score with standard deviation, we completed the following calculations:
 \begin{eqnarray}
@@
 
-\subsection{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-
+\section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
+Table \ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
+Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).
+Due to sample size considerations, multiple testing corrections were not performed; in addition, the independent variables are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games.
+As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
+To better understand the effects of these variables, a much larger study would be required.
 <<ttests.categ, echo=F, eval=T, results='asis'>>=
@@
     coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "MeanSq", "F", "p.val")
     coefs
@@
 tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-print(xtable(tmp, digits=c(1, 1, 0, 3, 3, 2, 3), caption="Model results of each single demographic variable compared with lineup score.
-Multiple testing issues aside, it appears that very few demographic variables (if any) are significantly associated with score on lineups among Iowa State undergraduate students."), include.rownames=F, floating.environment="table*", caption.placement="top")
+tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
+tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
+print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Results of lineup score modeled by single demographic variables.
+\\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")
 @
-\subsection{Lineup Performance: Accounting for ``Visual Intelligence''}
-As figure~\ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
+% \section{Lineup Performance: Accounting for ``Visual Intelligence''}
+% As figure~\ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
 As the visual search task (VST) is the most generic of visual tasks (find the matching pattern) and requires no spatial manipulation, rotation, or inference, we might use it as an indicator of general visual intelligence (or cognitive tempo and attention to detail).
-
-<<partialregression, results='asis', echo=1, tidy=T>>=
-model1 <- lm(lineup ~ vis_search, data=ans.summary)
-print(xtable(summary(model1), caption="Results of the linear regression of visual search score on lineup score. \\label{vst-reg}."), sanitize.text.function=sanitize)
-@
-<<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure~classification, and paper folding scores principal components analysis.
+% 
+% <<partialregression, results='asis', echo=1, tidy=T>>=
+% model1 <- lm(lineup ~ vis_search, data=ans.summary)
+% print(xtable(summary(model1), caption="Results of the linear regression of visual search score on lineup score. \\label{vst-reg}."), sanitize.text.function=sanitize)
+% @
+% <<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure~classification, and paper folding scores principal components analysis.
 Adjusted lineup scores are most associated with figure~classification in the second principal component.
@@
 \\label{fig:biplot-adjusted}", tidy=T, echo=1:7>>=
-# partial regression data set
-ans.part <- ans.summary
-ans.part$lineup <- resid(lm(lineup ~ vis_search, data=ans.summary))
-ans.part$card_rot <- resid(lm(card_rot ~ vis_search, data=ans.summary))
-ans.part$fig_class <- resid(lm(fig_class ~ vis_search, data=ans.summary))
-ans.part$folding <- resid(lm(folding ~ vis_search, data=ans.summary))
+% # partial regression data set
+% ans.part <- ans.summary
+% ans.part$lineup <- resid(lm(lineup ~ vis_search, data=ans.summary))
+% ans.part$card_rot <- resid(lm(card_rot ~ vis_search, data=ans.summary))
+% ans.part$fig_class <- resid(lm(fig_class ~ vis_search, data=ans.summary))
+% ans.part$folding <- resid(lm(folding ~ vis_search, data=ans.summary))
+% 
+% pca2 <- prcomp(ans.part[,c("lineup", "card_rot", "fig_class", "folding")], scale=T, retx=T)
+% biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
+% biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
+% @
+% <<partialregression2, echo=F, results='asis'>>=
+% model2 <- lm(lineup~fig_class+card_rot+folding, data=ans.part)
+% print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
+% @
+% Removing paper folding from the analysis due to its' obvious nonsignificance and collinearity with card rotation, we get a slightly better regression result. 
+% <<partialregression3, echo=F, results='asis'>>=
+% model3 <- lm(lineup~fig_class+card_rot, data=ans.part)
+% print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
+% print(xtable(anova(model3, model2), caption="Model comparison between full model (including paper folding) and reduced model (excluding paper folding), when accounting for scores on the VST in both independent and dependent variables.
+"), sanitize.text.function=sanitize, caption.placement="top")
+% @
+% 
+% Comparing the full and reduced models, we see that the reduction does not significantly reduce the explained sum of squares. 
+% 
+% 
+% Even using VST as a ``general intelligence" factor, the first PC still does not discriminate between the tests. This approach is probably not worth pursuing further.
 
-pca2 <- prcomp(ans.part[,c("lineup", "card_rot", "fig_class", "folding")], scale=T, retx=T)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
-@
-<<partialregression2, echo=F, results='asis'>>=
-model2 <- lm(lineup~fig_class+card_rot+folding, data=ans.part)
-print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
-@
-Removing paper folding from the analysis due to its' obvious nonsignificance and collinearity with card rotation, we get a slightly better regression result. 
-<<partialregression3, echo=F, results='asis'>>=
-model3 <- lm(lineup~fig_class+card_rot, data=ans.part)
-print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
-print(xtable(anova(model3, model2), caption="Model comparison between full model (including paper folding) and reduced model (excluding paper folding), when accounting for scores on the VST in both independent and dependent variables.
-"), sanitize.text.function=sanitize, caption.placement="top")
-@
-
-Comparing the full and reduced models, we see that the reduction does not significantly reduce the explained sum of squares.
-
-Even using VST as a ``general intelligence" factor, the first PC still does not discriminate between the tests. This approach is probably not worth pursuing further.
-
-\subsection{Lit Review Rejects}
-\begin{itemize}
-\item authors in~\cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-\item authors in~\cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
+% \section{Lit Review Rejects}
+% \begin{itemize}
+% \item authors in~\cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
+% \item authors in~\cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
 They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
@@
 They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
-\comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
+% \comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
 Alternate scoring procedures still need to be investigated...
 they cite a poster, but it's not available anywhere online as far as I can tell.}}
-\item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
+% \item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
 They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
-\comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
-They also cite a bunch of studies (including authors in~\cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
-\item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
+% \comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
+% They also cite a bunch of studies (including authors in~\cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
+% \item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
 The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
 that enable viewers to compare and observe relations among variables".
-\begin{itemize}
-\item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
-Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
+% \begin{itemize}
+% \item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
+% Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
 In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
 The review also specifically discusses visuospatial ability, citing a speculation from authors in~\cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
-\comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
-\item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
+% \comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
+% \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
 Displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
 Graphical displays have higher search and computational efficiency than text~\cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information~\cite{shah1999graphs}.
-\item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
+% \item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
 Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
@@
 Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
-\end{itemize}
-\item authors in~\cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
+% \end{itemize}
+% \item authors in~\cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
 They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
@@
 Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-\item authors in~\cite{hofmann2012graphical} for lineup stimuli and general lineup performance
-\end{itemize}
+% \item authors in~\cite{hofmann2012graphical} for lineup stimuli and general lineup performance
+% \end{itemize}
 \end{document}

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw

@@
 
+\comment{Next few steps for the analysis: (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; (2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
+not sure that that will pay off, but it's interesting.
+(3) in terms of why we would use lineups in a test comparison ...
+lineups give us a way to say 'this is correct' which regular graphics don't.
+So we should put a spin sentence on at the intro and the back that we are suggesting to use lineups as a tool to not only evaluate individuals (which we can see is possible) but also as a tool to evaluate charts ...
+which we've done before ...
+and tie it into the regular testing literature ...}
+
 \comment{Intro about statistical graphics, lineups, ...}
@@
 
+Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
+
 <<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests.

### 2014-11-01 Paper/Draft1/TestingVisualAptitude.Rnw
Fixing "free" graphic
half way through the new text

@@
 Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-\newtext{It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
 This constraint reduces memory load and makes more cognitive resources available for other tasks, but also implicates visual ability as an important factor in graph comprehension.
-}
 
@@
 
-Statistical lineups~\cite{hofmann2012graphical,buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+Statistical lineups~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
 Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
 By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-\newtext{Demographic factors are known to contribute to lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender affected the amount of time spent on lineups \cite{humanfactorslineups}.}\newtext{In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.}\newtext{In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability, with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
-}
+Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups \cite{humanfactorslineups}.
+In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
+In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
 
@@
 
-Participants will complete several cognitive tests designed to measure spatial and reasoning ability.
-Tasks are designed so that participants are under time pressure; they are not expected to complete all of the problems in each section.
-This provides more discrimination between high scorers and prevents score compression at the top of the range.
+Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
+Tasks are times such  that participants are under  pressure to complete; participants are not expected to finish all of the problems in each section.
+This allows for a better discrimination between scores and prevents score compression at the top of the range.
 
-The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test participants' ability to find a target stimulus in a field of distractors.
-The visual search task is similar in concept to lineups: it tests an individual's ability to find the target plot.
-Historically, it has been used as a measure of brain damage\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks, have been used to measure cognitive performance in a variety of situations (under the influence of drugs, for example, in~\cite{anderson1983interactive}).
+The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
+Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
@@
 
-The \textbf{figure~classification task} tests a participant's ability to extrapolate rules from provided figures.
+The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
 This task is associated with inductive reasoning ablities (factor I  in~\cite{ekstrom1976manual}).
@@
 \includegraphics[width=.75\linewidth]{figureclassification}
-\caption[Figure~Classification Task]{Figure~Classification Task.
+\caption[figure classification Task]{figure classification Task.
 Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
@@
 \end{figure*}
-The figure~classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
+The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
 In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
-As such, the figure~classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
+As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
 
@@
 An example is shown in figure~\ref{fig:cardrotation}.
-\newtext{The card rotation test is often used in studies of the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.}
+The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
 \begin{figure*}[hbt]\centering
@@
 \caption[Card Rotation Task]{Card Rotation Task.
-Participants mark each figure~on the right hand side as either the same or different than the figure~on the left hand side of the dividing line.
+Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
 Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
@@
 % The above consideration only assumes independence between questions, which is reasonable.
-While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure~classification score.
+While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
 % 
@@
 \subsection{Comparison of Spatial Tests with Previously Validated Results}
-The card rotation, paper folding, and figure~classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
+The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
 We compare Iowa State students' unscaled scores in table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
@@
 \hline
-  & Card Rotation & Paper Folding & Figure~Classification & Visual Search  \\\hline
+  & Card Rotation & Paper Folding & figure classification & Visual Search  \\\hline
 ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
@@
 We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='.\\linewidth', warning=F, message=F>>=
+\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
 
@@
 <<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure~classification scores, and are also highly correlated with card rotation scores.
+Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
 Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
@@
 
-\newtext{Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
+\newtext{figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
 These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
@@
 
-% In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure~classification, and paper folding tasks.
+% In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
 This suggests that skills associated with visual reasoning ability are related to lineup performance.
@@
 % <<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-% lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "Figure~Classification", "Paper Folding", "Visual Search"))
+% lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "figure classification", "Paper Folding", "Visual Search"))
 %  
@@
 % 
-% res$.id <- c("Card Rotation", "Paper Folding", "Figure~Classification", "Lineups")
+% res$.id <- c("Card Rotation", "Paper Folding", "figure classification", "Lineups")
 % names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
@@
 
-This scaling gets slightly more complicated for scores which have two sub-groups, as with the Figure~Classification test.
+This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test.
 To get a single unified score with standard deviation, we completed the following calculations:
@@
 \end{eqnarray}
-Then, in order to account for the fact that ISU students took only part I of two parts to the Figure~Classification test (and thus completed half of the questions), we adjusted the transformation as follows
+Then, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjusted the transformation as follows
 \begin{eqnarray}
@@
 % @
-% <<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure~classification, and paper folding scores principal components analysis.
-Adjusted lineup scores are most associated with figure~classification in the second principal component.
+% <<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
+Adjusted lineup scores are most associated with figure classification in the second principal component.
 In the third and fourth components, adjusted scores are almost entirely orthogonal.

### 2014-11-01 Paper/Draft1/TestingVisualAptitude.Rnw
discussion of PCA of five

@@
 \usepackage{footnote} % table footnotes
+\usepackage{tablefootnote}
 \makesavenoteenv{tabular}
@@
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
-This is reasonably representative of the university as a whole; in the fall 2012 semester, 23\% of students were associated with the college of engineering, 23\% were associated with the college of liberal arts and sciences, 23\% were associated with the college of human sciences, 12\% with the business school, and 14\% with the school of agriculture.
-\footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}}
+This is reasonably representative\footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}} of the university as a whole; in the fall 2012 semester, 23\% of students were associated with the college of engineering, 23\% were associated with the college of liberal arts and sciences, 23\% were associated with the college of human sciences, 12\% with the business school, and 14\% with the school of agriculture.
 
@@
 
-\begin{savenotes}
+%\begin{savenotes}
 \begin{table*}[htb]
@@
 \end{table*}
-\end{savenotes}
+%\end{savenotes}
 
@@
   xlab("") +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position=c(.75,.8)) +
+  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position=c(.8,.8)) +
   theme(plot.margin=unit(c(0,0,0,0),"mm"))
@@
 
-\newtext{Next, we examine the effect of demographic factors on lineup performance.
-Previous work found a relationship between lineup performance and factors such as education level, country of origin, and age\cite{humanfactorslineups}; our participant population is more homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
-}
+%Next, we examine the effect of demographic factors on lineup performance. 
+Previous work found a relationship between lineup performance and demographic factors such as education level, country of origin, and age \cite{humanfactorslineups}; our participant population is very homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
 
-Figure~\ref{fig:visualaptitudecat} shows participants' lineup scores as a function of their responses to the questionnaire given at the beginning of the study; these demographic questions allow us to explore relationships between demographic characteristics (major, research experience, etc.) and test performance.
+Figure~\ref{fig:visualaptitudecat} shows participants' lineup scores in relationship to their responses in the questionnaire given at the beginning of the study; this allows us to explore effects of demographic characteristics (major, research experience, etc.) on test performance.
 
 Completion of Calculus I is associated with increased performance on lineups; this may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with~\cite{shah1995conceptual}, who found an association between  mathematical ability and performance on simple graph description tasks.
-There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
+This association is consistent with findings in~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
+There is also a significant relationship between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
 There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
@@
 
-\newtext{Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are influenced by factors other than visual ability - general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few.
-In order to assess factors contributing to lineup performance, we first will examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores.}
-\newtext{Table \ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance.
-As expected, PC1 accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, that is, all of the general factors discussed above which influence test results.
-We will focus on the remaining components; these show differences in performance on the tests, and thus provide information about the skills needed to successfully examine lineups.
-}
-
-\newtext{Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}, and the corresponding rotation matrix is provided in \ref{tab:PCArotation5}.
-PC1 seems to represent a general ``visual intelligence'' factor.}
+Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are measuring visual ability of individuals in some way.
+%are influenced by factors other than visual ability -- general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few. 
+What is of more interest to us, is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
+In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores.
+%
+Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance. 
+PC1
+accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
+As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor. %, that is, all of the general factors discussed above which influence test results. 
+We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
+Biplots of these principal components are shown in figure~\ref{fig:biplots5}.
 
@@
 
-\newtext{figure classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3).
+Figure classification is strongly related to lineups (PC2, PC3).
+Performance on the visual search task  is also related to lineup performance (PC3).
 These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-}
 
 <<pcarotation,echo=F, results='asis'>>=
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.\\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
+             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
 @
 
-\newtext{The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.}
-\newtext{PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.}
+The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+
+PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
 <<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
@@
 @
-\newtext{If we examine the relationship between the four cognitive tests (excluding lineups), we can determine how lineup performance affects the variance partitions.
-Table \ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
-}
+If we examine the relationship between the four cognitive tests (excluding lineups), we can determine how lineup performance affects the variance partitions.
+Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
 
-\newtext{As with the previous principal components analysis, PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure \ref{fig:biplots4} and table \ref{tab:PCArotation4} confirm that PC1 is again a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+\newtext{As with the previous principal components analysis, PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is again a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
 PC2 differentiates the figure classification test from the visual searching test; this is incredibly similar to the role of PC2 in the 5-component analysis.
@@
 \section{Scaling Scores}\label{app:ScoreAdj}
-To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table \ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
+To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
 The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
@@
 \section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-Table \ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
+Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
 Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).

### 2014-11-01 Paper/Draft1/TestingVisualAptitude.Rnw
figures

@@
 An example is shown in figure~\ref{fig:figureclassification}.
-\begin{figure*}[hbt]\centering
-\includegraphics[width=.75\linewidth]{figureclassification}
-\caption[figure classification Task]{figure classification Task.
+% \begin{figure*}[hbt]\centering
+% \includegraphics[width=.75\linewidth]{figureclassification}
+% \caption[figure classification Task]{figure classification Task.
 Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
 Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
-\end{figure*}
+% \end{figure*}
 The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
@@
 As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
+
+ \begin{figure*}[!ht]
+  \centering
+   \subfloat[Figure Classification Task.
+Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
+Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
+  \includegraphics[width=.75\linewidth]{figureclassification}
+}
+\hfill
+    \subfloat[Card Rotation Task.
+Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
+Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
+\includegraphics[width=.75\linewidth]{cardrotation}
+    }
+    \hfill
+    \subfloat[Paper Folding Task.
+Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
+Participants will complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
+\includegraphics[width=.75\linewidth]{paperfolding}
+   }
+    \caption{Visuospatial tests}
+    \label{fig:tests}
+  \end{figure*}
 
@@
 The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
-\begin{figure*}[hbt]\centering
-\includegraphics[width=.75\linewidth]{cardrotation}
-\caption[Card Rotation Task]{Card Rotation Task.
+
+% \begin{figure*}[hbt]\centering
+% \includegraphics[width=.75\linewidth]{cardrotation}
+% \caption[Card Rotation Task]{Card Rotation Task.
 Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
 Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
-\end{figure*}
+% \end{figure*}
 Two-dimensional comparisons are an important component of lineup performance.
@@
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
-\begin{figure*}[hbt]\centering
-\includegraphics[width=.75\linewidth]{paperfolding}
-\caption[Paper Folding Task]{Paper Folding Task.
+% \begin{figure*}[hbt]\centering
+% \includegraphics[width=.75\linewidth]{paperfolding}
+% \caption[Paper Folding Task]{Paper Folding Task.
 Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
 Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
-\end{figure*}
+% \end{figure*}
 Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
@@
 qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
+
+qr1 <- qrange
+qr1$Type <- "Theoretical\nRange"
+qr2 <- qrange
+qr2$Type <- "Scores\nindicating\nguessing"
+qr2$n <- with(qr2, -n/(k-1))
+
+qrange <- rbind(qr1, qr2)
+qrange$Type <- factor(qrange$Type)
+qrange$Type <- factor(qrange$Type, levels=levels(qrange$Type)[2:1])
+
 ggplot() + 
-  geom_bar(data=qrange, aes(width=.9, x=Test, y=-n/(k-1), fill="red"), alpha=.5, colour="red", stat="identity", inherit.aes=F) + 
-  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, colour="red"), stat="identity", fill=NA, inherit.aes=F) + 
+  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, fill=Type), alpha=.35, stat="identity", position="identity", colour="red", inherit.aes=F) +
+  scale_fill_manual("", values=c(NA, "red")) + 
   geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
-  scale_colour_discrete(name="", labels="Theoretical\nRange") +
-  scale_fill_discrete(name="", labels="Scores\nindicating\nguessing") +
   theme_bw()  + 
@@
   xlab("") +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position=c(.8,.8)) +
-  theme(plot.margin=unit(c(0,0,0,0),"mm"))
+  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+  theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
+  theme(legend.position=c(.85,.85),  legend.box.just = "left", 
+        legend.key.height=unit(10, "mm")) 
   #coord_flip()
@@
 
-Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
+% Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
 Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.

### 2014-11-01 Paper/Draft1/TestingVisualAptitude.Rnw
table footnotes

@@
 \usepackage{rotating}
+\usepackage[referable]{threeparttablex}
 \usepackage{footnote} % table footnotes
-\usepackage{tablefootnote}
-\makesavenoteenv{tabular}
+%\makesavenoteenv{tabular}
 
@@
 An example is shown in figure~\ref{fig:figureclassification}.
-% \begin{figure*}[hbt]\centering
-% \includegraphics[width=.75\linewidth]{figureclassification}
-% \caption[figure classification Task]{figure classification Task.
-Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
-Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.}\label{fig:figureclassification}
-% \end{figure*}
+
 The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
@@
    \subfloat[Figure Classification Task.
-Participants classify each figure~in the second row as belonging to group 1, 2, or 3 (if applicable).
-Participants will complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
+Participants classify each figure in the second row as belonging to one of the groups above.
+Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
   \includegraphics[width=.75\linewidth]{figureclassification}
@@
 Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
+Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
 \includegraphics[width=.75\linewidth]{cardrotation}
@@
     \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
-Participants will complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
+Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
+Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
 \includegraphics[width=.75\linewidth]{paperfolding}
@@
 \includegraphics[width=\linewidth]{lineup}
-\caption[Sample Lineup]{A sample lineup. Participants are instructed to choose the plot which appears most different from the others. In this lineup, plot 13 is the target. }\label{fig:lineup}
+\caption[Sample Lineup]{Sample lineup of boxplots.
+Participants are instructed to choose the plot which appears most different from the others.
+In this lineup, plot 13 is the target.
+}\label{fig:lineup}
 \end{figure}
@@
 \centering
+\begin{threeparttable}  
 \begin{tabular}{rlllc}
 \hline
-  & Card Rotation & Paper Folding & figure classification & Visual Search  \\\hline
+  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
 ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
@@
 \end{tabular}
+        \begin{tablenotes}
+            \item[2] ISU students took only Part I due to time constraints.
+            \item[3] Averages calculated assuming 294 males and 323 females.
+            \item[4] Data from Part I only.
+        \end{tablenotes}      
+    \end{threeparttable}
 \end{table*}

### 2014-11-02 Paper/Draft1/TestingVisualAptitude.Rnw
mini change to figure

@@
   theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.85,.85),  legend.box.just = "left", 
-        legend.key.height=unit(10, "mm")) 
+  theme(legend.position=c(.85,.825),  legend.box.just = "left", 
+        legend.key.height=unit(12, "mm")) 
   #coord_flip()

### 2014-11-02 Paper/Draft1/TestingVisualAptitude.Rnw
added linear model

@@
 % \end{itemize}
+
+<<linearmodel, results='hide'>>=
+ans.model <- data.frame(ans.summary, data.frame(pca2$x))
+# exclude all variables that are not supposed to go into the linear model
+# those are the unrotated tests; instead we have the PCs
+excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
+
+ans.model$math_science_research <- ans.model$math_science_research=="y"
+ans.model$stats_class <- ans.model$stats_class=="y"
+ans.model$calc_1 <- ans.model$calc_1=="y"
+pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills")])
+biplot(pcmath)
+biplot(pcmath, 3:4)
+# variables split into stats_class/math_science_research and calc_1/math_skills
+# leave out math_skills in favor of calc_1 - for interpretability
+
+xtabs(~stats_class+math_science_research, data=ans.model)
+# only one kid is doing research without a stats class
+# recode to sum of them
+ans.model$statsP <- with(ans.model, stats_class + math_science_research)
+
+excl <- c(excl, c("math_skills", "math_science_research", "stats_class"))
+incl <- setdiff(names(ans.model), excl)
+
+m0 <- lm(lineup~., data=ans.model[, incl])
+
+# use AIC in backward selection
+library(MASS)
+m1 <- stepAIC(m0, direction="both")
+@
+<<modelresults>>=
+anova(m1)
+summary(m1)
+@
 \end{document}

### 2014-11-02 Paper/Draft1/TestingVisualAptitude.Rnw
some re-structuring: PCA of four to appendix, discussion of differences to PCA of five still at the front

@@
 In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores.
+\comment{give a mini-paragraph summary of the PCA of four}
 %
@@
 PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
-<<pcasummary2,echo=F, results='asis'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-@
-If we examine the relationship between the four cognitive tests (excluding lineups), we can determine how lineup performance affects the variance partitions.
-Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
 
-\newtext{As with the previous principal components analysis, PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is again a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test; this is incredibly similar to the role of PC2 in the 5-component analysis.
-}
-
-<<pcarotation2,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
-
-\newtext{PC3 differentiates the figure classification and visual search tests from the paper folding test; this is true in the 5-component analysis as well, but in that analysis the role of the figure classification test was not as significant due to the inclusion of the lineup task.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task, much like PC5 in the full PCA.}
-
+\comment{The following sentences are great, but would need to move up a bit, I think}
 \newtext{Comparing the two principal component analyses in this manner suggests that lineups are a separate dimension (as the addition of lineups result in an additional principal component which accounts for at least 10\% of the variance), but that they are most closely related to the figure classification task, and less closely related to the visual searching task.}
@@
 \comment{that last sentance is probably reaching a bit far...}
-
-<<pca2, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
-PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
-\\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
-@
-
 % In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
@@
 \begin{eqnarray}
-\mu_{all} &= (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
-\sigma_{all} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)}
+\mu_{\text{all}} &= (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
+\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
 \end{eqnarray}
-Then, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjusted the transformation as follows
+\newtext{where $\mu_F$ and $\mu_M$ are the averages observed individually, ...}
+\comment{You need to define all elements in the calculations}
+
+\comment{In a next step, include the derivations using the actual numbers.}
+
+\comment{We can approach the next piece a bit more generally.}
+\newtext{Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
+We can achieve comparability by scaling the scores accordingly.}
+
+In order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
+
 \begin{eqnarray}
-\mu_{all} &= 1/2 (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
-\sigma_{all} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(2(N_F + N_M))}
+\mu_{\text{part I}} &= 1/2 \mu_{\text{all}}\\
+\sigma_{\text{part I}} &= 1/\sqrt{2} \sigma_{\text{all}}
 \end{eqnarray}
@@
 @
+
+\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca4}
+<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+@
+<<pcasummary2,echo=F, results='asis'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+@
+Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
+
+\newtext{PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test, while}
+
+<<pcarotation2,echo=F, results='asis'>>=
+print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+@
+
+\newtext{
+PC3 differentiates these two from the paper folding test.
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.}
+
+<<pca2, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
+PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
+\\label{fig:biplots4}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
+@
+
 % \section{Lineup Performance: Accounting for ``Visual Intelligence''}
@@
 <<linearmodel, results='hide'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
 ans.model <- data.frame(ans.summary, data.frame(pca2$x))

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments for PCA section

@@
 
-<<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests.
+<<ResultsScaledScores, echo=F, include=T, fig.pos="htb", fig.cap="Test scores for lineups and visuospatial tests.
 As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
@@
 %
+\newtext{A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about XXX\% of the total variability.
+PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
+The principal components beyond that span another two dimensions, while the last dimension is weak (at XXX\%).
+The second principal component differentiates the figure classification task from card rotation, PC3  ...
+PC4 ...}
+
+\comment{then move on to the PCA of five, but summarize strongly - what does this one have in common with the PCA of four. The comments from below would be great here.}
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance. 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Fix scatterplot matrix scales by making a better scatterplot matrix using old code from deprecated plotmatrix() function in ggplot2

--- 2014-11-02 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
-GGally::ggpairs(ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")], lower=list(continuous='smooth'), upper=list(continuous="cor"))
+data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
+grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
+grid <- subset(grid, x != y & x<y)
+
+all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
+  xcol <- grid[i, "x"]
+  ycol <- grid[i, "y"]
+
+  rbind.fill(
+    data.frame(
+      xvar = names(data)[ycol], 
+      yvar = names(data)[xcol],
+      x = data[, xcol], 
+      y = data[, ycol], 
+      points = T,
+      data
+    ),
+    data.frame(
+      yvar = names(data)[ycol], 
+      xvar = names(data)[xcol],
+      x = mean(range(data[, ycol])), 
+      y = mean(range(data[, xcol])),
+      label = round(cor(data[,xcol], data[,ycol]), 3),
+      points = F
+    )
+  )
+}))
+all$xvar <- factor(all$xvar, levels=names(data))
+all$yvar <- factor(all$yvar, levels=names(data))
+densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
+    tmp <- density(data[,i])
+    data.frame(
+      xvar = names(data)[i], 
+      yvar = names(data)[i],
+      x = tmp$x, 
+      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
+    )
+  }))
+
+ggplot()+
+  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
+  geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
+  geom_text(data=subset(all, !points), aes(x=x, y=y, label=paste0("Correlation = \n", label))) + 
+  geom_line(data=densities, aes(x=x, y=y)) + 
+  facet_grid(xvar~yvar, scales="free") + 
+  theme_bw() + 
+  theme(axis.title=element_blank())
 @

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Density parameter modifications

@@
 densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-    tmp <- density(data[,i])
+    tmp <- density(data[,i], cut = 1, adjust=.75)
     data.frame(

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
suggestions appendix B

@@
 
-\begin{figure}\centering
+In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
+\comment{XXX should we classify majors into STEM/non-STEM?
+- there's a ton of literature and interest that we could potentially tap into \\\textit{I'm not opposed, but we have to decide where ag-majors fall...
+not everything is easily classifiable.
+Psychology is another problem: I was very STEM-psychology, but you don't have to be.}} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+
+\begin{figure}[hbtp]
+\centering
 \includegraphics[width=\linewidth]{lineup}
@@
 \end{figure}
-
-In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
-\comment{XXX should we classify majors into STEM/non-STEM?
-- there's a ton of literature and interest that we could potentially tap into \\\textit{I'm not opposed, but we have to decide where ag-majors fall...
-not everything is easily classifiable.
-Psychology is another problem: I was very STEM-psychology, but you don't have to be.}} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
 
@@
 Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.}
-\comment{that last sentance is probably reaching a bit far...}
+\comment{that last sentence is probably reaching a bit far... \textit I think that sentence is just fine :)}
 % In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
@@
 \begin{eqnarray}
-\mu_{\text{part I}} &= 1/2 \mu_{\text{all}}\\
-\sigma_{\text{part I}} &= 1/\sqrt{2} \sigma_{\text{all}}
+\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
+\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
 \end{eqnarray}
@@
 Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).
-Due to sample size considerations, multiple testing corrections were not performed; in addition, the independent variables are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games.
+\comment{Comment on a few more things here: order of variables in the table, and what exactly you have done to compute the values (model), also, why did you use separate factors for the skills and the number of video game hours?
+All of these variables are ordinal, so a linear factor using a single degree of freedom might be the most suitable for a start.}
+%Due to sample size considerations, multiple testing corrections were not performed;
+Note that these covariates are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games. 
+\comment{the next sentence will need some re-phrasing according to the results of the linear model.}
 As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.

### 2014-11-03 Code/Analysis.R
Add STEM indicator

@@ -7,15 +7,20 @@ ans$id <- 1:nrow(ans)
 # Fix age categories
 ans$age <- gsub("21-24", "21+", gsub("25-28", "21+", ans$age))
 
+
+ans$stem <- str_detect(ans$major1, "(Engineering)|(Biology)|([cC]hemistry)|(Statistics)|(Genetics)|(Animal Science)|(Horticulture)")
+
+ans <- ans[,c(1:19, 397, 20:396)]
+
 # Create a key that has the same structure as the answers
 key <- ans[1,]
-key[,1:19] <- NA
+key[,1:20] <- NA
 # Visual Search
-key[,20:44] <- c(17, 12, 15, 14, 8, 21, 6, 3, 21, 22, 4, 23, 9, 8, 22, 13, 2, 19, 16, 18, 7, 16, 12, 1, 7)
+key[,21:45] <- c(17, 12, 15, 14, 8, 21, 6, 3, 21, 22, 4, 23, 9, 8, 22, 13, 2, 19, 16, 18, 7, 16, 12, 1, 7)
 # Lineups 1
-key[,45:64] <- c(6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17)
+key[,46:65] <- c(6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17, 6, 16, 13, 10, 17)
 # Card  Rotation
-key[,65:224] <- c("d", "s", "s", "d", "d", "s", "d", "s",  "s", "s", "s", "d", "s", "s", "s", "s", 
+key[,66:225] <- c("d", "s", "s", "d", "d", "s", "d", "s",  "s", "s", "s", "d", "s", "s", "s", "s", 
                   "s", "d", "d", "d", "s", "s", "s", "d",  "s", "s", "d", "s", "d", "d", "d", "s",
                   "d", "s", "d", "d", "s", "s", "d", "s",  "s", "d", "s", "s", "s", "s", "d", "d", 
                   "s", "d", "s", "d", "d", "s", "s", "s",  "d", "d", "s", "s", "d", "s", "d", "d", 
@@ -26,10 +31,10 @@ key[,65:224] <- c("d", "s", "s", "d", "d", "s", "d", "s",  "s", "s", "s", "d", "
                   "s", "s", "d", "d", "d", "d", "d", "s",  "s", "s", "s", "s", "d", "d", "s", "s",
[…]

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Add libraries for string manipulation, chaining

--- 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 library(xtable)
+library(stringr)
+library(magrittr)
 source(paste(codedir, "Analysis.R", sep=""))

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Fix variable labels

@@
 score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
-score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
+score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
 qrange$sqrt.k <- sqrt(qrange$k-1)
-qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
+qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
 
@@
 qr2 <- qrange
-qr2$Type <- "Scores\nindicating\nguessing"
+qr2$Type <- "Scores indicating\nguessing"
 qr2$n <- with(qr2, -n/(k-1))
@@
   theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.85,.825),  legend.box.just = "left", 
-        legend.key.height=unit(12, "mm")) 
+  theme(legend.position=c(.80,.825),  legend.box.just = "left", 
+        legend.key.height=unit(8, "mm")) 
   #coord_flip()

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
STEM plot changes

@@
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "Video Game Hrs/Wk"))
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "STEM Major", "Video Game Hrs/Wk"))
 lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
@@
 qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + 
-  facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + 
+  facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 
   ylab("Scaled Lineup Score") + 
@@
 \caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
-Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a significant difference in lineup score.
+Categories are ordered by effect size; majoring in a STEM field, calculus completion, hours spent playing video games per week, and sex are all associated with a significant difference in lineup score.
 }\label{fig:visualaptitudecat}

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
remove commented out mean table
Fix variable labels in scatterplot matrix

@@
 Statistical test results are available in appendix~\ref{app:categoricalresults}.
-% <<means>>=
-% library(plyr)
-% ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
-% @
 
@@
   geom_line(data=densities, aes(x=x, y=y)) + 
-  facet_grid(xvar~yvar, scales="free") + 
+  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
   theme_bw() + 
   theme(axis.title=element_blank())
+rm(list=c("data", "all", "densities"))
 @

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Clean up knitr chunk options

@@
 
-<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".8\\linewidth", fig.height=3, fig.width=4.5, warning=F, message=F, size="footnotesize">>=
+<<pca,echo=F, warning=F, message=F>>=
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
 # screeplot(pca)

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Clean up pca chunk labels

@@
 
-<<pcasummary,echo=F, results='asis'>>=
+<<pcasummary5,echo=F, results='asis'>>=
 print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
@@
 
-<<pcarotation,echo=F, results='asis'>>=
+<<pcarotation5,echo=F, results='asis'>>=
 print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
@@
 
-\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca4}
-<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
-<<pcasummary2,echo=F, results='asis'>>=
+\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
+\subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
+
+<<pcasummary4,echo=F, results='asis'>>=
 pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 
-<<pcarotation2,echo=F, results='asis'>>=
+<<pcarotation4,echo=F, results='asis'>>=
 print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
@@
 
-<<pca2, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
 PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
@@
 @
+
+\subsection{PCA of all Five Tests (Cognitive Tests and Lineups)}\label{app:pca5}
+PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
+Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}.
 
@@
 % Even using VST as a ``general intelligence" factor, the first PC still does not discriminate between the tests. This approach is probably not worth pursuing further. 
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+@
 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Remove "visual intelligence" regression
Move lit review summary to below end of document
Remove unnecessary lineup vs. other test plots, comments
Code to fix scatterplot matrix plot names

@@
   }))
+fix.names <- function(values){
+  values %>% 
+    str_replace("lineup", "Lineups") %>% 
+    str_replace("vis_search", "Visual Search") %>%
+    str_replace("fig_class", "Figure Classification") %>% 
+    str_replace("card_rot", "Card Rotation") %>%
+    str_replace("folding", "Paper Folding")
+}
 
@@
 \comment{that last sentance is probably reaching a bit far...}
-% In figure~\ref{fig:visualaptitudects}, we see that participant performance on lineups is positively correlated with performance on card rotation, figure classification, and paper folding tasks.
-This suggests that skills associated with visual reasoning ability are related to lineup performance.
-As participants must use the same skills in lineups (mental rotation, classification and determining categorization schemes, and multi-step spatial reasoning) as in the factor-referenced tests, this is not particularly surprising.
-In addition,there seems to be some positive relationship between a participant's score on the visual search task and their score on lineups: the visual search task represents a baseline of a participant's ability to find a matching pattern, while lineups require that task as well as the ability to determine what the pattern is for a particular graph.
-Even excluding the one low visual search score that is a high-leverage point, there seems to be a positive relationship between a participant's score on lineups and their score for visual search.
-% 
-% <<VisReasoningCts,echo=FALSE, include=FALSE, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-% lineup.summary$variable <- factor(lineup.summary$variable, labels=c("Card Rotation", "figure classification", "Paper Folding", "Visual Search"))
-%  
-% qplot(data=lineup.summary, x=value, y=lineup, geom="point", shape=I(1), size=I(3)) + facet_wrap(~variable, scales="free_x") + xlab("Score") + ylab("Lineup Score") + geom_smooth(method="lm") + theme_bw()
-% @
-
-% 
-% \begin{figure}[hbt]\centering
-% \includegraphics[width=0.75\linewidth]{fig-VisReasoningCts-1}
-% \caption[Visual Aptitude Study Results]{Scatterplots of all test scores compared to participants' scores in the lineup tests.
-There is a relatively strong positive correlation between lineup score and scores on visuospatial reasoning tests.}\label{fig:visualaptitudects}
-% \end{figure}
 
@@
 
-% \section{Lineup Performance: Accounting for ``Visual Intelligence''}
-% As figure~\ref{fig:biplots}(a) shows, the first principal component seems to primarily describe ``general intelligence", or ``visual intelligence" in a generic sense (as we have only tested visual skills).
-As the visual search task (VST) is the most generic of visual tasks (find the matching pattern) and requires no spatial manipulation, rotation, or inference, we might use it as an indicator of general visual intelligence (or cognitive tempo and attention to detail).
-% 
-% <<partialregression, results='asis', echo=1, tidy=T>>=
-% model1 <- lm(lineup ~ vis_search, data=ans.summary)
-% print(xtable(summary(model1), caption="Results of the linear regression of visual search score on lineup score. \\label{vst-reg}."), sanitize.text.function=sanitize)
-% @
-% <<partialregression1,fig.keep='all', fig.width=4, fig.height=4, out.width='.45\\linewidth', fig.cap="Biplots of the adjusted lineup, card rotation, figure classification, and paper folding scores principal components analysis.
-Adjusted lineup scores are most associated with figure classification in the second principal component.
-In the third and fourth components, adjusted scores are almost entirely orthogonal.
-\\label{fig:biplot-adjusted}", tidy=T, echo=1:7>>=
-% # partial regression data set
-% ans.part <- ans.summary
-% ans.part$lineup <- resid(lm(lineup ~ vis_search, data=ans.summary))
-% ans.part$card_rot <- resid(lm(card_rot ~ vis_search, data=ans.summary))
-% ans.part$fig_class <- resid(lm(fig_class ~ vis_search, data=ans.summary))
-% ans.part$folding <- resid(lm(folding ~ vis_search, data=ans.summary))
-% 
-% pca2 <- prcomp(ans.part[,c("lineup", "card_rot", "fig_class", "folding")], scale=T, retx=T)
-% biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), adj=.75)
-% biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), adj=.75)
-% @
-% <<partialregression2, echo=F, results='asis'>>=
-% model2 <- lm(lineup~fig_class+card_rot+folding, data=ans.part)
-% print(xtable(model2, caption="Results of the linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
-% @
-% Removing paper folding from the analysis due to its' obvious nonsignificance and collinearity with card rotation, we get a slightly better regression result. 
-% <<partialregression3, echo=F, results='asis'>>=
-% model3 <- lm(lineup~fig_class+card_rot, data=ans.part)
-% print(xtable(model3, caption="Results of the reduced linear regression of (the ETS tests $|$ VST) on (lineup score $|$ VST)"), sanitize.text.function=sanitize, caption.placement="top")
-% print(xtable(anova(model3, model2), caption="Model comparison between full model (including paper folding) and reduced model (excluding paper folding), when accounting for scores on the VST in both independent and dependent variables.
-"), sanitize.text.function=sanitize, caption.placement="top")
-% @
-% 
-% Comparing the full and reduced models, we see that the reduction does not significantly reduce the explained sum of squares. 
-% 
-% 
-% Even using VST as a ``general intelligence" factor, the first PC still does not discriminate between the tests. This approach is probably not worth pursuing further. 
 <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
@@
 @
+
+<<linearmodel, results='hide'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+ans.model <- data.frame(ans.summary, data.frame(pca2$x))
+# exclude all variables that are not supposed to go into the linear model
+# those are the unrotated tests; instead we have the PCs
+excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
+
+ans.model$math_science_research <- ans.model$math_science_research=="y"
+ans.model$stats_class <- ans.model$stats_class=="y"
+ans.model$calc_1 <- ans.model$calc_1=="y"
+pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills")])
+biplot(pcmath)
+biplot(pcmath, 3:4)
+# variables split into stats_class/math_science_research and calc_1/math_skills
+# leave out math_skills in favor of calc_1 - for interpretability
+
+xtabs(~stats_class+math_science_research, data=ans.model)
+# only one kid is doing research without a stats class
+# recode to sum of them
+ans.model$statsP <- with(ans.model, stats_class + math_science_research)
+
+excl <- c(excl, c("math_skills", "math_science_research", "stats_class"))
+incl <- setdiff(names(ans.model), excl)
+
+m0 <- lm(lineup~., data=ans.model[, incl])
+
+# use AIC in backward selection
+library(MASS)
+m1 <- stepAIC(m0, direction="both")
+@
+<<modelresults>>=
+anova(m1)
+summary(m1)
+@
+
+\end{document}
 
@@
 % \end{itemize}
-
-<<linearmodel, results='hide'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-ans.model <- data.frame(ans.summary, data.frame(pca2$x))
-# exclude all variables that are not supposed to go into the linear model
-# those are the unrotated tests; instead we have the PCs
-excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
-
-ans.model$math_science_research <- ans.model$math_science_research=="y"
-ans.model$stats_class <- ans.model$stats_class=="y"
-ans.model$calc_1 <- ans.model$calc_1=="y"
-pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills")])
-biplot(pcmath)
-biplot(pcmath, 3:4)
-# variables split into stats_class/math_science_research and calc_1/math_skills
-# leave out math_skills in favor of calc_1 - for interpretability
-
-xtabs(~stats_class+math_science_research, data=ans.model)
-# only one kid is doing research without a stats class
-# recode to sum of them
-ans.model$statsP <- with(ans.model, stats_class + math_science_research)
-
-excl <- c(excl, c("math_skills", "math_science_research", "stats_class"))
-incl <- setdiff(names(ans.model), excl)
-
-m0 <- lm(lineup~., data=ans.model[, incl])
-
-# use AIC in backward selection
-library(MASS)
-m1 <- stepAIC(m0, direction="both")
-@
-<<modelresults>>=
-anova(m1)
-summary(m1)
-@
-\end{document}

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
PCA explanatory text rearrange

@@
 
-Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are measuring visual ability of individuals in some way.
-%are influenced by factors other than visual ability -- general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few. 
-What is of more interest to us, is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
-In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores.
-\comment{give a mini-paragraph summary of the PCA of four}
-%
-\newtext{A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about XXX\% of the total variability.
+Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
+What is of more interest to us is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
+In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests using the same procedure.
+
+\newtext{
+A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-The principal components beyond that span another two dimensions, while the last dimension is weak (at XXX\%).
-The second principal component differentiates the figure classification task from card rotation, PC3  ...
-PC4 ...}
+The principal components beyond that span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
+PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test. 
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task. 
+More detailed results from the 4-test analysis are provided in Appendix \ref{app:pca}.
+}
 
-\comment{then move on to the PCA of five, but summarize strongly - what does this one have in common with the PCA of four. The comments from below would be great here.}
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance. 
-PC1
-accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
-As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor. %, that is, all of the general factors discussed above which influence test results. 
-We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
-Biplots of these principal components are shown in figure~\ref{fig:biplots5}.
+Incorporating the lineup task into the principal component analysis, we find the principal components to be fairly similar to the four-component analysis. 
+Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance; the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
 
@@
 @
-
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
 
@@
 
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor, and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data.
+We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
+Biplots of these principal components are provided in Appendix \ref{app:pca5}.
 
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
+\newtext{
+Figure classification is strongly related to lineups (PC2, PC3), and as in the four-component PCA, figure classification is strongly represented in both principal components. 
+Performance on the visual search task is also related to lineup performance (PC3), and is more strongly represented in PC3 in the five-component PCA than in the four-componet PCA. 
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.}
 
-\comment{The following sentences are great, but would need to move up a bit, I think}
-\newtext{Comparing the two principal component analyses in this manner suggests that lineups are a separate dimension (as the addition of lineups result in an additional principal component which accounts for at least 10\% of the variance), but that they are most closely related to the figure classification task, and less closely related to the visual searching task.}
+\newtext{
+Comparing the two principal component analyses in this manner suggests that lineups are a separate dimension (as the addition of lineups result in an additional principal component which accounts for at least 10\% of the variance), but that they are most closely related to the figure classification task, and less closely related to the visual searching task.
+}
+
+\newtext{
+The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
+PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
+}
 
@@
 
+Figure classification is strongly related to lineups (PC2, PC3).
+Performance on the visual search task  is also related to lineup performance (PC3).
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+
+The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+
+PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
+
 <<linearmodel, results='hide'>>=

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Remove STEM comment.

@@
 In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
-\comment{XXX should we classify majors into STEM/non-STEM?
-- there's a ton of literature and interest that we could potentially tap into \\\textit{I'm not opposed, but we have to decide where ag-majors fall...
-not everything is easily classifiable.
-Psychology is another problem: I was very STEM-psychology, but you don't have to be.}} These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
 
@@
 Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.}
-\comment{that last sentance is probably reaching a bit far...}
 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Fix wording

@@
 
-Scoring of all test results was done such that random guessing leads to an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
+All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
 Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Add STEM percentage

@@
 Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
-About one-half of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
+\Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Wording of PC analysis

@@
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-The principal components beyond that span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
-PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test. 
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task. 
+The other principal components span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
+PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two tests from the paper folding test. 
 More detailed results from the 4-test analysis are provided in Appendix \ref{app:pca}.

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Expanded Scaling Scores appendix

@@
 
-This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test.
+This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
 To get a single unified score with standard deviation, we completed the following calculations: 
-\begin{eqnarray}
-\mu_{\text{all}} &= (\mu_{F}N_F + \mu_{M}N_M)/(N_F + N_M)\\
+\begin{align}
+\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
 \sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
-\end{eqnarray}
-\newtext{where $\mu_F$ and $\mu_M$ are the averages observed individually, ...}
-\comment{You need to define all elements in the calculations}
+\end{align}
+\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
+Substituting in the provided numbers, }
+\begin{align}
+\mu_{\text{all}} &= \left(323\!*\!114.9\! +\! 294\!*\!120.0\right)/(323\!+\!294) \\
+& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
+\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
+& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}
+\end{align}
 
-\comment{In a next step, include the derivations using the actual numbers.}
-
-\comment{We can approach the next piece a bit more generally.}
-\newtext{Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
-We can achieve comparability by scaling the scores accordingly.}
-
-In order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
+Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
+We can achieve comparability by scaling the scores accordingly.
+For example, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
todos not comments!
Remove old commented-out stuff
un-squish boxplots for categorical data, since a new row was added...

@@
 
-\comment{Intro about statistical graphics, lineups, ...}
+\todo[inline]{Intro about statistical graphics, lineups, ...}
 \subsection{Spatial Reasoning and Statistical Graphics}
@@
 \subsection{The Lineup Protocol}
-\comment{More lineup introduction here}
+\todo[inline]{More lineup introduction here}
 
@@
 The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
-% \comment{maybe reference book by Michel Hersen\\\textit{I was working on it yesterday along with trying to validate the general idea that reading graphics requires certain spatial abilities.
-It's definitely a work in progress.
-I haven't seen that book before, but I'm not at all surprised it exists.}}
-
-% 
-% \subsection{Variance consideration of scaled test score:}
-% Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
-% \begin{eqnarray*}
-% \text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
-% &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
-% &=& n^2/(k-1).
-% \end{eqnarray*}
-% \newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
-% 
-% The above consideration only assumes independence between questions, which is reasonable.
-While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.
-% 
-% To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct).
-This scaling ensures that test scores across different types are comparable.
-All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1.
-In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.
 
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5, message=F, warning=F, fig.pos='htb'>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=6, message=F, warning=F, fig.pos='htb'>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
-% \comment{\textit{In the current analysis, we have scored the paper folding test out of 20.
+\textit{In the current analysis, we have scored the paper folding test out of 20.
 Alternate scoring procedures still need to be investigated...
-they cite a poster, but it's not available anywhere online as far as I can tell.}}
+they cite a poster, but it's not available anywhere online as far as I can tell.}
 % \item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
 They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
-% \comment{\textit{I can't understand their model results very well, so I'm not entirely sure what's going on here. \\
-% They also cite a bunch of studies (including authors in~\cite{voyer1995magnitude}) that I haven't had time to completely explore.}}
 % \item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
@@
 The review also specifically discusses visuospatial ability, citing a speculation from authors in~\cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
-% \comment{\textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.\\That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}}
+% \textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.
+That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}
 % \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Additional table diagnostic

@@
 
+xtabs(~stem+calc_1, data=ans.model)
+# stem major is more associated with performance on lineups than calc 1...
+
 xtabs(~stats_class+math_science_research, data=ans.model)
 # only one kid is doing research without a stats class
+
 # recode to sum of them

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Lineup types

@@
 
+\section{Lineup Types}
+The two lineup sections which contain boxplots are shown as separate density curves, as they appear to be testing different things (outliers vs. difference in medians). 
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.env="figure*", fig.width=6, fig.height=6, out.width=".9\\linewidth">>=
+lineup.type.sum$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rotated"))
+qplot(data=lineup.type.sum, x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
+  facet_wrap(~plot.type, ncol=5) + 
+  xlab("Percent of Lineups Correct") + 
+  ylab("Scaled Density") + 
+  geom_vline(aes(xintercept=0), linetype=3) + 
+  scale_colour_discrete("Lineup Test Number") + 
+  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
+  theme_bw() + theme(legend.position="bottom", legend.direction="horizontal", axis.text.x=element_text(angle=30))
+@
+Conclusion: stacked bar plots are awful. 
 \end{document}

### 2014-11-03 Code/Analysis.R

@@ -55,6 +55,22 @@ key[,378:397] <- c("a", "d", "b", "d", "b",
                    "a", "e", "d", "d", "c")
 
 
+keytype <- key
+keytype[,46:65] <- rep(c("boxplot", "density", "scatterplot", "stacked.bar"), each=5)
+keytype[,226:245] <- c("boxplot", "boxplot", "violinplot", "violinplot", 
+                       "boxplot", "violinplot", "jitter.boxplot", "jitter.boxplot", 
+                       "boxplot", "violinplot", "violinplot", "tree.boxplot", 
+                       "tree.boxplot", "violinplot", "violinplot", "jitter.boxplot",
+                       "tree.boxplot", "jitter.boxplot", "violinplot", "tree.boxplot")
+keytype[,358:377] <- c("qq.line.error", "qq.line", "qq.line.error", "qq.line", 
+                   "qq.line", "qq.line.error.rotated", "qq.line.error", "qq.line.error.rotated", 
+                   "qq.line.error", "qq.line.error.rotated", "qq.line", "qq.line.error", 
+                   "qq.line.error.rotated", "qq.line.error.rotated", "qq.line.error", "qq.line",
+                   "qq.line.error", "qq.line", "qq.line.error", "qq.line.error.rotated")
+
+keytype <- keytype[,str_detect(names(keytype), "lineup")]
+keytype <- data.frame(question=names(keytype), plot.type=as.character(keytype[1,]))
+
 library(plyr)
 # Score each answer by whether it is the same as the key. Make sure NA values stay as NA. 
 scored <- ddply(ans, .(id), function(j) as.numeric(j==key) + c(0, NA)[1+is.na(j)])
@@ -92,6 +108,16 @@ longform.sum <- ddply(longform, .(id, testtype), summarize,
 longform.sum <- merge(longform.sum, qrange)
 # Calculate raw score
 longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
+
+# calculate scores by question type
+longform <- merge(longform, keytype, by.x="variable", by.y="question", all.x=T, all.y=T)
+lineup.type.sum <- ddply(subset(longform, !is.na(plot.type)), .(id, plot.type, testnum), summarize, 
+                         pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
+                         neg.pts = ifelse(is.numeric(value), sum((1-value)*penalty, na.rm=T), unique(value)), 
[…]

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Binary and extra appendix section

@@
 
+\section{Linear Model}
 <<linearmodel, results='hide'>>=

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Plot tweaks

@@
 The two lineup sections which contain boxplots are shown as separate density curves, as they appear to be testing different things (outliers vs. difference in medians). 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.env="figure*", fig.width=6, fig.height=6, out.width=".9\\linewidth">>=
-lineup.type.sum$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rotated"))
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.width=3, fig.height=8, out.width=".9\\linewidth">>=
+lineup.type.sum$plot.type <- str_replace(lineup.type.sum$plot.type, "rotated", "rot")
+lineup.type.sum$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rot"))
 qplot(data=lineup.type.sum, x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=5) + 
-  xlab("Percent of Lineups Correct") + 
+  facet_wrap(~plot.type, ncol=2) + 
+  xlab("Proportion of Lineups Correct") + 
   ylab("Scaled Density") + 
@@
   scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  theme_bw() + theme(legend.position="bottom", legend.direction="horizontal", axis.text.x=element_text(angle=30))
+  scale_y_continuous(breaks=c(0, .5, 1)) +
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
 @

### 2014-11-03 Code/Analysis.R
Don't overwrite longform, tally other scores too.

@@ -110,13 +110,16 @@ longform.sum <- merge(longform.sum, qrange)
 longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
 
 # calculate scores by question type
-longform <- merge(longform, keytype, by.x="variable", by.y="question", all.x=T, all.y=T)
-lineup.type.sum <- ddply(subset(longform, !is.na(plot.type)), .(id, plot.type, testnum), summarize, 
+longform2 <- merge(longform, keytype, by.x="variable", by.y="question", all.x=T, all.y=T)
+lineup.type.sum <- ddply(longform2, .(id, testtype, plot.type, testnum), summarize, 
                          pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
                          neg.pts = ifelse(is.numeric(value), sum((1-value)*penalty, na.rm=T), unique(value)), 
                          total.pos = length(value),
                          total.neg = sum(rep(1, length(value))*penalty))
+lineup.type.sum$score <- with(lineup.type.sum, (pos.pts-neg.pts))
+lineup.type.sum2 <- ddply(subset(lineup.type.sum, testtype!="lineup"), .(id, testtype), summarize, score=sum(score), pos.pts=sum(pos.pts), neg.pts=sum(neg.pts), total.pos=sum(total.pos), total.neg=sum(total.neg))
 lineup.type.sum$score <- with(lineup.type.sum, (pos.pts-neg.pts)/total.pos)
+lineup.type.sum2 <- rbind.fill(lineup.type.sum2, subset(lineup.type.sum, testtype=="lineup"))
 
 
 # Save unscaled version

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Scatterplot of test score by plot type

--- 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 The two lineup sections which contain boxplots are shown as separate density curves, as they appear to be testing different things (outliers vs. difference in medians). 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.env="figure*", fig.width=6, fig.height=6, out.width=".9\\linewidth">>=
-lineup.type.sum$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rotated"))
-qplot(data=lineup.type.sum, x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=5) + 
-  xlab("Percent of Lineups Correct") + 
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.width=3, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
+lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rot"))
+qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
+      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
+  facet_wrap(~plot.type, ncol=2) + 
+  xlab("Proportion of Lineups Correct") + 
   ylab("Scaled Density") + 
@@
   scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  theme_bw() + theme(legend.position="bottom", legend.direction="horizontal", axis.text.x=element_text(angle=30))
+  scale_y_continuous(breaks=c(0, .5, 1)) +
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
 @
 Conclusion: stacked bar plots are awful.
+
+<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores", fig.width=6, fig.height=12, out.width=".65\\linewidth", fig.env="figure*">>=
+lineup.type.sum2$plot.type <- str_replace(lineup.type.sum$plot.type, "error", "err")
+lineup.type.sum2$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rot"))
+
+lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = sum)
+other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
+names(other.scores)[3] <- "testscore"
+lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
+names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
+lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
+
+qplot(data=lineup.type.sum3, 
+      x=testscore, y=lineup.pct, geom="jitter") + 
+  geom_smooth(method="lm") + 
+  facet_grid(lineup.type~testtype, scales="free_x") + 
+  ylab("Proportion of Lineups Correct") + 
+  xlab("Test Score")+
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
+@
 \end{document}

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
labels, list correlation on the plot

--- 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 The two lineup sections which contain boxplots are shown as separate density curves, as they appear to be testing different things (outliers vs. difference in medians). 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.env="figure*", fig.width=6, fig.height=6, out.width=".9\\linewidth">>=
-lineup.type.sum$plot.type <- factor(lineup.type.sum$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.error", "qq.line.error.rotated"))
-qplot(data=lineup.type.sum, x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=5) + 
-  xlab("Percent of Lineups Correct") + 
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.width=3, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
+lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
+lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.err", "qq.line.err.rot"))
+qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
+      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
+  facet_wrap(~plot.type, ncol=2) + 
+  xlab("Proportion of Lineups Correct") + 
   ylab("Scaled Density") + 
@@
   scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  theme_bw() + theme(legend.position="bottom", legend.direction="horizontal", axis.text.x=element_text(angle=30))
+  scale_y_continuous(breaks=c(0, .5, 1)) +
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
 @
-Conclusion: stacked bar plots are awful. 
+Conclusion: stacked bar plots are awful - the same as guessing.
+
+<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores", fig.width=6, fig.height=12, out.width=".65\\linewidth", fig.env="figure*">>=
+lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = sum)
+other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
+names(other.scores)[3] <- "testscore"
+lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
+names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
+lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
+
+cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(range(testscore)), y=max(range(lineup.pct)), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
+
+qplot(data=lineup.type.sum3, 
+      x=testscore, y=lineup.pct, geom="jitter", colour=I("grey30")) + 
+  geom_smooth(method="lm") + 
+  facet_grid(lineup.type~testtype, scales="free_x") + 
+  ylab("Proportion of Lineups Correct") + 
+  xlab("Test Score")+
+  geom_text(data=cor.labels, aes(x=x, y=1.1, label=label), hjust=0, vjust=1, size=3) + 
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
+@
 \end{document}

### 2014-11-03 Code/Analysis.R
Include 2nd major with STEM

@@ -8,7 +8,8 @@ ans$id <- 1:nrow(ans)
 ans$age <- gsub("21-24", "21+", gsub("25-28", "21+", ans$age))
 
 
-ans$stem <- str_detect(ans$major1, "(Engineering)|(Biology)|([cC]hemistry)|(Statistics)|(Genetics)|(Animal Science)|(Horticulture)")
+ans$stem <- str_detect(ans$major1, "(Engineering)|(Biology)|([cC]hemistry)|(Statistics)|(Genetics)|(Environmental Science)|(Animal Science)|(Horticulture)") 
+ans$stem[!is.na(ans$major2)] <- ans$stem[!is.na(ans$major2)] | str_detect(ans$major2[!is.na(ans$major2)], "(Engineering)|(Biology)|([cC]hemistry)|(Statistics)|(Genetics)|(Environmental Science)|(Animal Science)|(Horticulture)")
 
 ans <- ans[,c(1:19, 397, 20:396)]
 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
formula massages
Fix scaled score boxplot

@@
 
-score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
+score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
 score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
@@
 qr1 <- qrange
-qr1$Type <- "Theoretical\nRange"
+qr1$Type <- "Theoretical Range"
 qr2 <- qrange
-qr2$Type <- "Scores indicating\nguessing"
+qr2$Type <- "Scores indicating guessing"
 qr2$n <- with(qr2, -n/(k-1))
@@
   theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.80,.825),  legend.box.just = "left", 
+  theme(legend.position=c(.80,.125),  legend.box.just = "left", 
         legend.key.height=unit(8, "mm")) 
@@
 \newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
-Substituting in the provided numbers, }
-\begin{align}
+Substituting in the provided numbers, we get}
+\begin{align*}
 \mu_{\text{all}} &= \left(323\!*\!114.9\! +\! 294\!*\!120.0\right)/(323\!+\!294) \\
@@
 \sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
-& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}
-\end{align}
+& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
+\end{align*}
 
@@
 
-\begin{eqnarray}
-\mu_{\text{part I}} &= 1/2 \mu_{\text{all}}\\
-\sigma_{\text{part I}} &= 1/\sqrt{2} \sigma_{\text{all}}
-\end{eqnarray}
+\begin{eqnarray*}
+\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
+\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
+\end{eqnarray*}
 
@@
 # leave out math_skills in favor of calc_1 - for interpretability
+
+xtabs(~stem+sex, data=ans.model)
+# sad day!
 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Fix boxplot legend alignment

@@
   theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.80,.125),  legend.box.just = "left", 
-        legend.key.height=unit(8, "mm")) 
+  theme(legend.position=c(.68,.125),  legend.box.just = "left", 
+        legend.key.height=unit(6, "mm")) 
   #coord_flip()

### 2014-11-03 Code/Analysis.R
Plottype name changes

@@ -57,12 +57,12 @@ key[,378:397] <- c("a", "d", "b", "d", "b",
 
 
 keytype <- key
-keytype[,46:65] <- rep(c("boxplot", "density", "scatterplot", "stacked.bar"), each=5)
+keytype[,46:65] <- rep(c("boxplot", "density", "scatterplot", "histogram"), each=5)
 keytype[,226:245] <- c("boxplot", "boxplot", "violinplot", "violinplot", 
                        "boxplot", "violinplot", "jitter.boxplot", "jitter.boxplot", 
-                       "boxplot", "violinplot", "violinplot", "tree.boxplot", 
-                       "tree.boxplot", "violinplot", "violinplot", "jitter.boxplot",
-                       "tree.boxplot", "jitter.boxplot", "violinplot", "tree.boxplot")
+                       "boxplot", "violinplot", "violinplot", "bee.swarm.boxplot", 
+                       "bee.swarm.boxplot", "violinplot", "violinplot", "jitter.boxplot",
+                       "bee.swarm.boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot")
 keytype[,358:377] <- c("qq.line.error", "qq.line", "qq.line.error", "qq.line", 
                    "qq.line", "qq.line.error.rotated", "qq.line.error", "qq.line.error.rotated", 
                    "qq.line.error", "qq.line.error.rotated", "qq.line", "qq.line.error", 

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
Fix labels

@@
 lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
-lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "tree.boxplot", "scatterplot", "density", "stacked.bar", "qq.line", "qq.line.err", "qq.line.err.rot"))
+lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot", "scatterplot", "density", "histogram", "qq.line", "qq.line.err", "qq.line.err.rot"))
 qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
@@
 
-<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores", fig.width=6, fig.height=12, out.width=".65\\linewidth", fig.env="figure*">>=
-lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = sum)
+<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores", fig.width=6, fig.height=12, out.width=".65\\linewidth", fig.env="figure*", warning=F, message=F>>=
+lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="pos.pts", fun.aggregate = sum)
 other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
@@
   geom_smooth(method="lm") + 
-  facet_grid(lineup.type~testtype, scales="free_x") + 
+  facet_grid(lineup.type~testtype, scales="free") + 
   ylab("Proportion of Lineups Correct") + 

### 2014-11-04 Paper/Draft1/TestingVisualAptitude.Rnw
next section of analysis

@@
 
-\newtext{
 A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
@@
 More detailed results from the 4-test analysis are provided in Appendix \ref{app:pca}.
-}
 
 Incorporating the lineup task into the principal component analysis, we find the principal components to be fairly similar to the four-component analysis. 
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance; the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
+Table~\ref{tab:PCAvariance5} shows the importance of each principal component.
+\newtext{From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.} From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that
 
@@
 
-As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor, and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data.
-We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
-Biplots of these principal components are provided in Appendix \ref{app:pca5}.
+the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
+Biplots of the remaining components are provided in Appendix \ref{app:pca5}.
 
-\newtext{
 Figure classification is strongly related to lineups (PC2, PC3), and as in the four-component PCA, figure classification is strongly represented in both principal components. 
 Performance on the visual search task is also related to lineup performance (PC3), and is more strongly represented in PC3 in the five-component PCA than in the four-componet PCA. 
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.}
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
 
-\newtext{
-Comparing the two principal component analyses in this manner suggests that lineups are a separate dimension (as the addition of lineups result in an additional principal component which accounts for at least 10\% of the variance), but that they are most closely related to the figure classification task, and less closely related to the visual searching task.
+While lineups do span a separate dimension, but the PCA also suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
+
+The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
+PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
+
+This emphasizes the underpinnings of lineups: they are a visual medium, but they ultimately are a classification task, presented in a graphical manner.
+Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.
+
+\subsection{Linear model of demographic factors}
+\newtext{The demographic variables are highy correlated, e.g.~there is a high correlation between STEM majors and taking calculus 1 -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
+The situation between a statistics class and doing research is even more correlated.
+Only one student is doing research who has (not yet ?) taen a statistics course.
+A principal component of the five math/stats questions splits the variables into two areas: the first principal component is an average of math skills, calculaus 1 and STEM, while the second principal component is an average of XXX }
+<<linearmodel, results='hide'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+ans.model <- data.frame(ans.summary, data.frame(pca2$x))
+# exclude all variables that are not supposed to go into the linear model
+# those are the unrotated tests; instead we have the PCs
+excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
+
+ans.model$math_science_research <- ans.model$math_science_research=="y"
+ans.model$stats_class <- ans.model$stats_class=="y"
+ans.model$calc_1 <- ans.model$calc_1=="y"
+pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills", "stem")], scale=T)
+pcmath$rotation
+# first component of math PCA is average of math_skills, calc_1 and stem
+# second component is stat_class and research
+
+#pcm <- data.frame(pcmath$x)
+#names(pcm) <- paste("SM", 1:5, sep="")
+#ans.model <- data.frame(ans.model, pcm)
+biplot(pcmath)
+biplot(pcmath, 3:4)
+
+xtabs(~stem+sex, data=ans.model)
+# sad day!
+
+xtabs(~stem+calc_1, data=ans.model)
+# stem major is more associated with performance on lineups than calc 1...
+
+xtabs(~stats_class+math_science_research, data=ans.model)
+# only one kid is doing research without a stats class
+
+# recode to sum of them
+ans.model$STATS <- with(ans.model, stats_class + math_science_research)
+ans.model$MATH <- with(ans.model, calc_1+math_skills/5+stem)
+
+excl <- c(excl, c("math_science_research", "stats_class"))
+excl <- c(excl, c("calc_1", "math_skills", "stem"))
+incl <- setdiff(names(ans.model), excl)
+
+m0 <- lm(lineup~., data=ans.model[, incl])
+
+# use AIC in backward selection
+library(MASS)
+m1 <- stepAIC(m0, direction="both")
+@
+<<modelresults>>=
+anova(m1)
+summary(m1)
+@
+\subsection{Lineup Types}
+\comments{
+results need to go here
 }
 
-\newtext{
-The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
-PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
-}
-
-\newtext{This emphasizes the underpinnings of lineups: they are a visual medium, but they ultimately are a classification task, presented in a graphical manner.
-Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.}
+\section{Discussion and Conclusions}
 
@@
 
-\section{Linear Model}
-<<linearmodel, results='hide'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-ans.model <- data.frame(ans.summary, data.frame(pca2$x))
-# exclude all variables that are not supposed to go into the linear model
-# those are the unrotated tests; instead we have the PCs
-excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
-
-ans.model$math_science_research <- ans.model$math_science_research=="y"
-ans.model$stats_class <- ans.model$stats_class=="y"
-ans.model$calc_1 <- ans.model$calc_1=="y"
-pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills")])
-biplot(pcmath)
-biplot(pcmath, 3:4)
-# variables split into stats_class/math_science_research and calc_1/math_skills
-# leave out math_skills in favor of calc_1 - for interpretability
-
-xtabs(~stem+sex, data=ans.model)
-# sad day!
-
-xtabs(~stem+calc_1, data=ans.model)
-# stem major is more associated with performance on lineups than calc 1...
-
-xtabs(~stats_class+math_science_research, data=ans.model)
-# only one kid is doing research without a stats class
-
-# recode to sum of them
-ans.model$statsP <- with(ans.model, stats_class + math_science_research)
-
-excl <- c(excl, c("math_skills", "math_science_research", "stats_class"))
-incl <- setdiff(names(ans.model), excl)
-
-m0 <- lm(lineup~., data=ans.model[, incl])
-
-# use AIC in backward selection
-library(MASS)
-m1 <- stepAIC(m0, direction="both")
-@
-<<modelresults>>=
-anova(m1)
-summary(m1)
-@
-
 \section{Lineup Types}

### 2014-11-05 Paper/Draft1/TestingVisualAptitude.Rnw
linear model

@@
 The situation between a statistics class and doing research is even more correlated.
-Only one student is doing research who has (not yet ?) taen a statistics course.
-A principal component of the five math/stats questions splits the variables into two areas: the first principal component is an average of math skills, calculaus 1 and STEM, while the second principal component is an average of XXX }
-<<linearmodel, results='hide'>>=
+Only one student is doing research who has (not yet ?) taken a statistics course.
+A principal component of the five math/stats questions splits the variables into two areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
+We therefore decided to use sums of these variables to come up with a separate math and a stats score.
+Note, that the correlation between the math and the stats score is almost zero.
+}
+<<linearmodel, results='hide', echo=FALSE>>=
 pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 
-#pcm <- data.frame(pcmath$x)
-#names(pcm) <- paste("SM", 1:5, sep="")
-#ans.model <- data.frame(ans.model, pcm)
-biplot(pcmath)
-biplot(pcmath, 3:4)
+#biplot(pcmath)
+#biplot(pcmath, 3:4)
 
@@
 @
-<<modelresults>>=
-anova(m1)
-summary(m1)
+\newtext{We fit a linear model of lineup scores in all of the demographic variables and the test scores from the visuo-spatial tests.
+Then we use a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) to select the best model.
+The result is shown in table~\ref{tab:m1}.
+Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores.
+We can think of PC1 as an innate visual ability, while the MATH score is a matter of both ability and training.
+}
+<<modelresults, echo=FALSE, results='asis'>>=
+library(xtable)
+xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1")
 @
 \subsection{Lineup Types}
-\comments{
+\comment{
 results need to go here

### 2014-11-05 Paper/Draft1/TestingVisualAptitude.Rnw
latex issues fixed

@@
 
-<<ResultsScaledScores, echo=F, include=T, fig.pos="htb", fig.cap="Test scores for lineups and visuospatial tests.
-As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
-We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
+\begin{figure}[ht]
+<<ResultsScaledScores, echo=F, include=T, fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
 
@@
 @
+\caption{Test scores for lineups and visuospatial tests.
+As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
+We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
+\label{fig:Scores}}
+\end{figure}
 %
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component.
-\newtext{From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.} From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that
-
+From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.
 <<pcasummary5,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="ht", caption.placement="top")
 @
@@
 @
-
+From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that  
 the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
@@
 library(xtable)
-xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1")
+print(xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1"), caption.placement="top")
 @
@@
 \begin{align*}
-\mu_{\text{all}} &= \left(323\!*\!114.9\! +\! 294\!*\!120.0\right)/(323\!+\!294) \\
+\mu_{\text{all}} &= \left(323\!\cdot\!114.9\! +\! 294\!\cdot\!120.0\right)/(323\!+\!294) \\
 & = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
some re-phrasing

@@
 \subsection{Linear model of demographic factors}
-\newtext{The demographic variables are highy correlated, e.g.~there is a high correlation between STEM majors and taking calculus 1 -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
-The situation between a statistics class and doing research is even more correlated.
+\newtext{Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
+Similaryl, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
 Only one student is doing research who has (not yet ?) taken a statistics course.
-A principal component of the five math/stats questions splits the variables into two areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
+A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
 We therefore decided to use sums of these variables to come up with a separate math and a stats score.
@@
 @
-\newtext{We fit a linear model of lineup scores in all of the demographic variables and the test scores from the visuo-spatial tests.
-Then we use a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) to select the best model.
+\newtext{We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests.
+Using a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) we select the best model.
 The result is shown in table~\ref{tab:m1}.

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
Fix PCA discussion to emphasize lineups = figure classification

@@
 
-options(replace.assign=TRUE,width=30,scipen=3)
+options(replace.assign=TRUE,scipen=3)
 
@@
 
-Figure classification is strongly related to lineups (PC2, PC3), and as in the four-component PCA, figure classification is strongly represented in both principal components. 
-Performance on the visual search task is also related to lineup performance (PC3), and is more strongly represented in PC3 in the five-component PCA than in the four-componet PCA. 
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+Figure classification is strongly related to lineups, and as in the four-component PCA, figure classification is strongly represented in the first two principal components.
+While lineups do span a separate dimension, the PCA suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
 
-While lineups do span a separate dimension, but the PCA also suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
+% 
+% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
+% PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
+%
 
-The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
-PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
-
-This emphasizes the underpinnings of lineups: they are a visual medium, but they ultimately are a classification task, presented in a graphical manner.
+This emphasizes the underpinnings of lineups: the test utilizes a visual medium, but is ultimately are a classification task presented in a graphical manner.
 Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.
@@
 \newtext{Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
-Similaryl, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
+Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
 Only one student is doing research who has (not yet ?) taken a statistics course.
@@
 @
-\newtext{We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests.
-Using a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) we select the best model.
+\newtext{We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
 The result is shown in table~\ref{tab:m1}.
 Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores.
-We can think of PC1 as an innate visual ability, while the MATH score is a matter of both ability and training.
+We can think of PC1 as a measure of innate visual or intellectual ability, while the MATH score is a matter of both ability and training.
 }

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
Add example lineups

@@
   % declare the path(s) where your graphic files are
-  \graphicspath{{Figure/}{Images/}}
+  \graphicspath{{Figure/}{Images/}{../../Images/}}
   % and their extensions so you won't have to specify these with
@@
 @
+\section{Lineup Task Examples}
+\subsection{Lineup Set 1}
+\comment{Description of original purpose of the experiment}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
+\caption{Boxplots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
+\caption{Density plots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
+\caption{Dotplots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
+\caption{Histograms used to compare the two distributions.}
+\end{figure}
+\FloatBarrier
+\subsection{Lineup Set 2}
+\comment{Description of original purpose of the experiment. Note that this experiment appears to have more outliers.}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f94f6965d7}
+\caption{Boxplots.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
+\caption{Boxplots with jittered points.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9fc58d73}
+\caption{Bee swarm boxplots.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
+\caption{Violin plots.}
+\end{figure}
+\FloatBarrier
+\subsection{Lineup Set 3}
+\comment{Description of original purpose of the experiment.}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
+\caption{QQ Plot with guide line.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
+\caption{QQ Plot with acceptance bands.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
+\caption{QQ Plot rotated 45 degrees.}
+\end{figure}
 \end{document}

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
Add comment on ISU honors research explanation?

@@
 Only one student is doing research who has (not yet ?) taken a statistics course.
+}
+\comment{The honors program puts freshman in research labs in their first year; Ryan worked in a corn biology lab as a physics/engineering major.
+If they are LBAR, they may still be placed in science labs, or they could be doing LBAR research - may not require stats.}
+\newtext{
 A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
Updated binary

@@
 We can think of PC1 as a measure of innate visual or intellectual ability, while the MATH score is a matter of both ability and training.
+The remaining principal components were not sufficiently associated with lineup score to include in the model.
 }
@@
 \subsection{Lineup Types}
-\comment{
-results need to go here
+\newtext{
+We can also compare participants' performance on specific types of lineup plots compared with their scores on the visual aptitude tests, for instance, accuracy on lineups which require mental rotation may be related to performance on the card rotation task.
+Examples lineups for each type of plot are provided in Appendix \ref{app:lineuptypes}.
 }
+
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=3, fig.height=6, out.width=".9\\linewidth", warning=F, message=F>>=
+lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
+lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
+lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
+                                     levels=c("boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot", 
+                                              "scatterplot", "density", "histogram", 
+                                              "qq.line", "qq.line.err", "qq.line.err.rot"), 
+                                     labels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
+                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
+                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
+                                              "Rotated QQ plot +\nAcceptance Band")
+                                     )
+qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
+      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
+  facet_wrap(~plot.type, ncol=2) + 
+  xlab("Proportion of Lineups Correct") + 
+  ylab("Density") + 
+  geom_vline(aes(xintercept=0), linetype=3) + 
+  scale_colour_discrete("Lineup Test Number") + 
+  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
+  scale_y_continuous(breaks=c(0, .5, 1)) +
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"), 
+        axis.ticks.y=element_blank(), axis.text.y=element_blank())
+@
+\newtext{
+Figure \ref{fig:plottypesdensity} compares performance on each different type of plot.
+As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs.
+difference in medians), different tasks are shown as different colors, so that accuracy on tasks which are shown in blue can be compared to other blue density curves.
+}
+
+\newtext{
+Figure \ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
+Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks.
+In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots?
+}
+<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores.\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
+lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
+other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
+names(other.scores)[3] <- "testscore"
+lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
+names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
+lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
+lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
+                                     levels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
+                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
+                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
+                                              "Rotated QQ plot +\nAcceptance Band"), 
+                                     labels=c("Boxplot", "Boxplot\n + Points", "Violin Plot", "Bee Swarm \nBoxplot",
+                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
+                                              "QQ plot\n+ Line", "QQ plot\n+ Bands", 
+                                              "Rotated QQ\nplot + Bands")
+                                     )
+
+cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
+
+qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
+      x=testscore, y=lineup.pct, geom=c("smooth", "jitter"), colour=I("grey30"), method="lm") + 
+  facet_grid(lineup.type~testtype, scales="free_x") + 
+  ylab("Scaled Lineup Score") + 
+  xlab("Test Score")+
+  geom_text(data=subset(cor.labels, testtype!="vis_search"), aes(x=x, y=1, label=label), hjust=0, vjust=1, size=3) + 
+  theme_bw() + 
+  theme(legend.position="bottom", legend.direction="horizontal", 
+        axis.text.x=element_text(angle=30, hjust=1),
+        plot.margin=unit(c(0,0,0,0),"mm"))
+@
 
@@
 
-\section{Lineup Types}
-The two lineup sections which contain boxplots are shown as separate density curves, as they appear to be testing different things (outliers vs. difference in medians). 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups", fig.width=3, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
-lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
-lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
-lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, levels=c("boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot", "scatterplot", "density", "histogram", "qq.line", "qq.line.err", "qq.line.err.rot"))
-qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
-      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=2) + 
-  xlab("Proportion of Lineups Correct") + 
-  ylab("Scaled Density") + 
-  geom_vline(aes(xintercept=0), linetype=3) + 
-  scale_colour_discrete("Lineup Test Number") + 
-  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  scale_y_continuous(breaks=c(0, .5, 1)) +
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"))
-@
-Conclusion: stacked bar plots are awful - the same as guessing.
-
-<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores", fig.width=6, fig.height=12, out.width=".65\\linewidth", fig.env="figure*", warning=F, message=F>>=
-lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="pos.pts", fun.aggregate = sum)
-other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
-names(other.scores)[3] <- "testscore"
-lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
-names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
-lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
-
-cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(range(testscore)), y=max(range(lineup.pct)), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-
-qplot(data=lineup.type.sum3, 
-      x=testscore, y=lineup.pct, geom="jitter", colour=I("grey30")) + 
-  geom_smooth(method="lm") + 
-  facet_grid(lineup.type~testtype, scales="free") + 
-  ylab("Proportion of Lineups Correct") + 
-  xlab("Test Score")+
-  geom_text(data=cor.labels, aes(x=x, y=1.1, label=label), hjust=0, vjust=1, size=3) + 
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"))
-@
-\section{Lineup Task Examples}
+\section{Lineup Task Examples}\label{app:lineuptypes}
 \subsection{Lineup Set 1}

### 2014-11-07 Paper/Draft1/TestingVisualAptitude.Rnw
couple of edits

@@
 \subsection{Linear model of demographic factors}
-\newtext{Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
+Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
 Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
-Only one student is doing research who has (not yet ?) taken a statistics course.
-}
-\comment{The honors program puts freshman in research labs in their first year; Ryan worked in a corn biology lab as a physics/engineering major.
+Only one student is doing research who has (not yet?) taken a statistics course.
+%\comment{The honors program puts freshman in research labs in their first year; Ryan worked in a corn biology lab as a physics/engineering major.
 If they are LBAR, they may still be placed in science labs, or they could be doing LBAR research - may not require stats.}
-\newtext{
 A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
@@
 Note, that the correlation between the math and the stats score is almost zero.
-}
 <<linearmodel, results='hide', echo=FALSE>>=
@@
 @
-\newtext{We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
+We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
 The result is shown in table~\ref{tab:m1}.
@@
 We can think of PC1 as a measure of innate visual or intellectual ability, while the MATH score is a matter of both ability and training.
-The remaining principal components were not sufficiently associated with lineup score to include in the model.
-}
+The remaining principal components were not sufficiently associated with lineup score to be included in the model.
+
 <<modelresults, echo=FALSE, results='asis'>>=

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
lineup types
Working memory and cognitive load explanation

@@
 It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-This constraint reduces memory load and makes more cognitive resources available for other tasks, but also implicates visual ability as an important factor in graph comprehension.
+\newtext{Graphics utilize higher-bandwidth visual pathways to encode information\cite{baddeley1974working}, allowing viewers to quickly and intuitively encode multiple dimensions of numerical relationships; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability.
+Well-designed graphics present the most important features of the data while minimizing less-important features, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
+Under this framework, well-designed graphics reduce memory load and make more cognitive resources available for other tasks (such as drawing conclusions from the data), at the cost of depending on certain visuospatial reasoning abilities.
+}
 
@@
 \subsection{Lineup Types}
+\newtext{The survey consisted of three sets of different types of lines. Examples lineups for each type of plot are provided in Appendix \ref{app:lineuptypes}. 
+}
 \newtext{
 We can also compare participants' performance on specific types of lineup plots compared with their scores on the visual aptitude tests, for instance, accuracy on lineups which require mental rotation may be related to performance on the card rotation task.
-Examples lineups for each type of plot are provided in Appendix \ref{app:lineuptypes}.
 }

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Reorder boxplot levels so that "boxplot" is located between the two relevant experiments.

@@
 lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
-                                     levels=c("boxplot", "jitter.boxplot", "violinplot", "bee.swarm.boxplot", 
+                                     levels=c("violinplot", "jitter.boxplot", "bee.swarm.boxplot", "boxplot", 
                                               "scatterplot", "density", "histogram", 
                                               "qq.line", "qq.line.err", "qq.line.err.rot"), 
-                                     labels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
+                                     labels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
@@
 lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
-                                     levels=c("Boxplot", "Boxplot + \nOverlaid Data", "Violin Plot", "Bee Swarm Boxplot",
+                                     levels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
@@
                                               "Rotated QQ plot +\nAcceptance Band"), 
-                                     labels=c("Boxplot", "Boxplot\n + Points", "Violin Plot", "Bee Swarm \nBoxplot",
+                                     labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
                                               "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Reflow comment to emphasize specific items

@@
 
-\comment{Next few steps for the analysis: (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; (2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
-not sure that that will pay off, but it's interesting.
+\comment{Next few steps for the analysis: \\
+(1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; \\
+(2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
+not sure that that will pay off, but it's interesting.\\
 (3) in terms of why we would use lineups in a test comparison ...

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Move lineup picture, and use a more obvious sample lineup. Add some text explaining lineup protocol.

@@
 \subsection{The Lineup Protocol}
-\todo[inline]{More lineup introduction here}
+\todo[inline]{Standard lineup protocol introduction here.} 
+\newtext{Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants would be expected to determine that sub-plot 4 is the most different because the two groups have notably different medians.}
 
@@
 In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
+
+\begin{figure}\centering
+\includegraphics[width=\linewidth]{lineup}
+\caption[Sample Lineup]{Sample lineup of boxplots.
+Participants are instructed to choose the plot which appears most different from the others.
+In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.
+}\label{fig:lineup}
+\end{figure}
+
+\newtext{One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features most conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
+In addition, lineups provide researchers with a statistically valid methodology for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
+As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.}
+
 In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
@@
 Figure~\ref{fig:lineup} shows a sample lineup of box plots.
-
-\begin{figure}\centering
-\includegraphics[width=\linewidth]{lineup}
-\caption[Sample Lineup]{Sample lineup of boxplots.
-Participants are instructed to choose the plot which appears most different from the others.
-In this lineup, plot 13 is the target.
-}\label{fig:lineup}
-\end{figure}
 

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Explain score compression in VST and fixes in future experiments.

@@
 Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
+\newtext{Participants' scores on the VST indicate score compression; that is, both participants with medium and high visual search abilities scored at the extremely high end of the spectrum. 
+In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.}
 

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Additional sentances explaining lineup task relationships. Some tweaks to figure sizes.

@@
 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=3, fig.height=6, out.width=".9\\linewidth", warning=F, message=F>>=
+<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
                                               "qq.line", "qq.line.err", "qq.line.err.rot"), 
-                                     labels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
+                                     labels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
@@
 lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
-                                     levels=c("Violin Plot", "Boxplot + \nOverlaid Data", "Bee Swarm Boxplot", "Boxplot", 
+                                     levels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
@@
 
+\newtext{For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.}
+\newtext{While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
+}
+\newtext{Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in figure \ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
+}
 \section{Discussion and Conclusions}

### 2014-11-10 Paper/Draft1/TestingVisualAptitude.Rnw
Lineup Block PCA

@@
 \subsection{Linear model of demographic factors}
-Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
+Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.
+% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
 Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
-Only one student is doing research who has (not yet?) taken a statistics course.
+Only one student is doing research who has not taken a statistics course.
 %\comment{The honors program puts freshman in research labs in their first year; Ryan worked in a corn biology lab as a physics/engineering major.
@@
 \subsection{Lineup Types}
-\newtext{The survey consisted of three sets of different types of lines. Examples lineups for each type of plot are provided in Appendix \ref{app:lineuptypes}. 
+\newtext{Each of the three sets of 20 lineups was taken from previous experiments which tested different types of plots to determine which plot type most effectively conveyed important characteristics of the underlying data set.
 }
+\comment{Add experiment summary/refs here}
+\newtext{Examples lineups for each experiment are provided in Appendix \ref{app:lineuptypes}. 
+}
+
+\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
+\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
+Participants were much more accurate in this experiment ...
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
+\newtext{The final lineup section explored qq plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
+\comment{Add citations, more explanations, etc.}
+
+\newtext{Calculating participant scores on each block of lineups separately, we can use principal components analysis to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of this experiment.}
+
+<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Lineup block PCA rotation matrix in graphical form.">>=
+lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+rotation <- melt(lineup.block.pca$rotation)
+importance <- as.data.frame(summary(lineup.block.pca)$importance)[2,]
+rotation <- merge(rotation, melt(importance, variable.name="Var2", value.name="importance"))
+rotation$Var1 <- factor(rotation$Var1, 
+                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_3", "lineup_2", "lineup_1"),
+                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
+library(grid) # needed for arrow function
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("")
+@
+
+\newtext{We see that again, PC1 accounts for general visual aptitude.
+PC2, however, emphasizes the relationship between figure classification and Lineup task 1 (and to a lesser extent, on lineup task 2).
+PC3 differentiates the lineup task 3 from the visual search task; other tasks do not impact PC3 to a meaningful degree.
+PC4 separates lineup test 2 from the paper folding and card rotation tests and emphasizes the overlap with the visual search test.
+PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test.
+PC 6 and 7 account for \Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
+}
+<<lineupblockpca-biplot,echo=F, warning=F, message=F, fig.cap="This is awful, but at least shows everything", fig.env="figure*", fig.width=11, fig.height=10, out.width=".9\\linewidth">>=
+data <- as.data.frame(lineup.block.pca$rotation)
+grid <- expand.grid(x=1:7, y=1:7)
+grid <- subset(grid, x != y)
+
+all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
+  xcol <- grid[i, "x"]
+  ycol <- grid[i, "y"]
+
+  rbind.fill(
+    data.frame(
+      xvar = names(data)[xcol], 
+      yvar = names(data)[ycol],
+      x = 0,
+      y = 0,
+      xend = data[, xcol], 
+      yend = data[, ycol], 
+      label = rownames(data),
+      points = T
+    )
+  )
+}))
+all$xvar <- factor(all$xvar, levels=names(data))
+all$yvar <- factor(all$yvar, levels=names(data))
+
+tmp <- rotation[,1:3]
+names(tmp) <- c("xvar", "label", "rotation")
+tmp$xvar <- tmp$yvar <- factor(tmp$xvar, levels=names(data))
+
+all$label <- factor(all$label, 
+                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_1", "lineup_2", "lineup_3"),
+                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 1", "Lineup Test 2", "Lineup Test 3"))
+
+ggplot()+
+  geom_segment(data=subset(all, points), aes(x=x, y=y, xend=xend, yend=yend, colour=label), arrow=arrow(length=unit(.1, "cm"))) + 
+#   geom_point(data=tmp, aes(x=0*rotation, y=rotation, colour=label)) + 
+  geom_text(data=tmp, aes(x=ceiling(rotation)-.5, y=rotation, label=label, colour=label)) + 
+  facet_grid(xvar~yvar) + 
+  theme_bw() + 
+  theme(axis.title=element_blank()) + 
+  scale_colour_manual("Test", values = c("#f781bf", "#e41a1c", "#ff7f00", "#66a61e", "#1d91c0", "#225ea8", "#0c2c84"))
+rm(list=c("data", "all"))
+@
+\subsection{Lineup Plot Types}
+
 \newtext{

### 2014-11-10 Code/Analysis.R

@@ -110,6 +110,22 @@ longform.sum <- merge(longform.sum, qrange)
 # Calculate raw score
 longform.sum$value <- with(longform.sum, pos.pts-neg.pts)
 
+# calculate scores by lineup section
+lineup.sec.sum <- ddply(longform, .(id, testtype, testnum), summarize, 
+                         pos.pts = ifelse(is.numeric(value), sum(value, na.rm=T), unique(value)),
+                         neg.pts = ifelse(is.numeric(value), sum((1-value)*penalty, na.rm=T), unique(value)), 
+                         total.pos = length(value),
+                         total.neg = sum(rep(1, length(value))*penalty))
+lineup.sec.sum$score <- with(lineup.sec.sum, (pos.pts-neg.pts))
+lineup.sec.sum2 <- ddply(subset(lineup.sec.sum, testtype!="lineup"), .(id, testtype), summarize, score=sum(score), pos.pts=sum(pos.pts), neg.pts=sum(neg.pts), total.pos=sum(total.pos), total.neg=sum(total.neg))
+lineup.section.sum <- rbind.fill(lineup.sec.sum2, subset(lineup.sec.sum, testtype=="lineup"))
+lineup.section.sum$testnum[is.na(lineup.section.sum$testnum)] <- ""
+lineup.section.sum$testnum[nchar(lineup.section.sum$testnum)>0] <- paste0("_", lineup.section.sum$testnum[nchar(lineup.section.sum$testnum)>0])
+lineup.section.sum$testtype <- paste0(lineup.section.sum$testtype, lineup.section.sum$testnum)
+
+lineup.section.summary <- merge(ans[,1:20], dcast(lineup.section.sum, id~testtype, value.var="score", na.rm=TRUE))
+rm(lineup.sec.sum2, lineup.sec.sum, lineup.section.sum)
+
 # calculate scores by question type
 longform2 <- merge(longform, keytype, by.x="variable", by.y="question", all.x=T, all.y=T)
 lineup.type.sum <- ddply(longform2, .(id, testtype, plot.type, testnum), summarize, 
@@ -121,7 +137,7 @@ lineup.type.sum$score <- with(lineup.type.sum, (pos.pts-neg.pts))
[…]

### 2014-11-11 Paper/Draft1/TestingVisualAptitude.Rnw
mini edits during meeting

@@
 % Note that keywords are not normally used for peerreview papers.
-\begin{IEEEkeywords}
-Computer Society, IEEEtran, journal, \LaTeX, paper, template.
-\end{IEEEkeywords}}
+% \begin{IEEEkeywords}
+% Computer Society, IEEEtran, journal, \LaTeX, paper, template.
+% \end{IEEEkeywords}
+}
 
@@
 (1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; \\
+\textbf{Done; added to Results section}\\
 (2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
 not sure that that will pay off, but it's interesting.\\
+\textbf{Done, along with an examination of each experiment separately}\\
 (3) in terms of why we would use lineups in a test comparison ...
@@
 which we've done before ...
-and tie it into the regular testing literature ...}
+and tie it into the regular testing literature ...\\
+\textbf{A few sentences added, still need to tie into the existing literature}}
 
@@
 
-<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
-Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
+\begin{figure}[htp]
+<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
 data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
@@
 @
+\caption{Pairwise scatterplots of test scores.
+Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
+Paper folding and card rotation scores are also highly correlated.\label{fig:scatterplotmatrix}}
+\end{figure}
 
@@
 }
-<<lineupblockpca-biplot,echo=F, warning=F, message=F, fig.cap="This is awful, but at least shows everything", fig.env="figure*", fig.width=11, fig.height=10, out.width=".9\\linewidth">>=
-data <- as.data.frame(lineup.block.pca$rotation)
-grid <- expand.grid(x=1:7, y=1:7)
-grid <- subset(grid, x != y)
-
-all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-  xcol <- grid[i, "x"]
-  ycol <- grid[i, "y"]
-
-  rbind.fill(
-    data.frame(
-      xvar = names(data)[xcol], 
-      yvar = names(data)[ycol],
-      x = 0,
-      y = 0,
-      xend = data[, xcol], 
-      yend = data[, ycol], 
-      label = rownames(data),
-      points = T
-    )
-  )
-}))
-all$xvar <- factor(all$xvar, levels=names(data))
-all$yvar <- factor(all$yvar, levels=names(data))
-
-tmp <- rotation[,1:3]
-names(tmp) <- c("xvar", "label", "rotation")
-tmp$xvar <- tmp$yvar <- factor(tmp$xvar, levels=names(data))
-
-all$label <- factor(all$label, 
-                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_1", "lineup_2", "lineup_3"),
-                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 1", "Lineup Test 2", "Lineup Test 3"))
-
-ggplot()+
-  geom_segment(data=subset(all, points), aes(x=x, y=y, xend=xend, yend=yend, colour=label), arrow=arrow(length=unit(.1, "cm"))) + 
-#   geom_point(data=tmp, aes(x=0*rotation, y=rotation, colour=label)) + 
-  geom_text(data=tmp, aes(x=ceiling(rotation)-.5, y=rotation, label=label, colour=label)) + 
-  facet_grid(xvar~yvar) + 
-  theme_bw() + 
-  theme(axis.title=element_blank()) + 
-  scale_colour_manual("Test", values = c("#f781bf", "#e41a1c", "#ff7f00", "#66a61e", "#1d91c0", "#225ea8", "#0c2c84"))
-rm(list=c("data", "all"))
-@
 \subsection{Lineup Plot Types}

### 2014-11-11 Paper/Draft1/TestingVisualAptitude.Rnw
Slight modifications to comments; remove keywords as this is a peer-reviewed paper

--- 2014-11-11 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-\begin{figure}[htp]
-<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
+<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
+Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
+Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
 data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
@@
 @
-\caption{Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
-Paper folding and card rotation scores are also highly correlated.\label{fig:scatterplotmatrix}}
-\end{figure}
 

### 2014-11-14 Paper/Draft1/TestingVisualAptitude.Rnw
attempt at mean structure with residuals for pca - not sure that it is successful, it seems to be very similar to the other pca

@@
 % \end{itemize}
+
+<<>>=
+vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
+lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
+vmain <- rowSums(vistests)/4
+alltests <- data.frame(vmain, vistests-vmain, lineups)
+pctests <- prcomp(alltests)
+biplot(pctests)
+biplot(pctests, 3:4)
+
+@

### 2014-11-14 Paper/Draft1/TestingVisualAptitude.Rnw
few edits, nothing major

@@
 \comment{Next few steps for the analysis: \\
-(1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; \\
-\textbf{Done; added to Results section}\\
-(2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ...
-not sure that that will pay off, but it's interesting.\\
-\textbf{Done, along with an examination of each experiment separately}\\
+\\
 (3) in terms of why we would use lineups in a test comparison ...
@@
 and tie it into the regular testing literature ...\\
-\textbf{A few sentences added, still need to tie into the existing literature}}
+\textbf{A few sentences added, still need to tie into the existing literature}
+(4) I'll take care of the intro to lineups}
 
@@
 It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-\newtext{Graphics utilize higher-bandwidth visual pathways to encode information\cite{baddeley1974working}, allowing viewers to quickly and intuitively encode multiple dimensions of numerical relationships; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability.
-Well-designed graphics present the most important features of the data while minimizing less-important features, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
+Graphics utilize higher-bandwidth visual pathways to encode information\cite{baddeley1974working}, allowing viewers to quickly and intuitively encode multiple dimensions of numerical relationships; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability.
+Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
 Under this framework, well-designed graphics reduce memory load and make more cognitive resources available for other tasks (such as drawing conclusions from the data), at the cost of depending on certain visuospatial reasoning abilities.
-}
 
@@
 \todo[inline]{Standard lineup protocol introduction here.}
+
 \newtext{Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants would be expected to determine that sub-plot 4 is the most different because the two groups have notably different medians.}
@@
 
-\newtext{One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features most conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
-In addition, lineups provide researchers with a statistically valid methodology for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
+\newtext{One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
+In addition, lineups provide researchers with a rigorous framework for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
 As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.}

### 2014-11-15 Paper/Draft1/TestingVisualAptitude.Rnw
some more edits

@@
 \comment{Next few steps for the analysis: \\
-\\
 (3) in terms of why we would use lineups in a test comparison ...
@@
 and tie it into the regular testing literature ...\\
-\textbf{A few sentences added, still need to tie into the existing literature}
+\textbf{A few sentences added, still need to tie into the existing literature}\\
 (4) I'll take care of the intro to lineups}
@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
-\begin{figure}[hbt]\centering
+\begin{figure}[htp]\centering
 \includegraphics[width=\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task.
+\caption[Visual Search Task]{Visual Search Task (VST).
 Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
@@
 
- \begin{figure*}[!ht]
+ \begin{figure*}[ht]
   \centering
@@
 Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
-\newtext{Participants' scores on the VST indicate score compression; that is, both participants with medium and high visual search abilities scored at the extremely high end of the spectrum. 
-In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.}
+Participants' scores on the VST indicate score compression; that is, both participants with medium and high visual search abilities scored at the extremely high end of the spectrum. 
+In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.
 
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=6, message=F, warning=F, fig.pos='htb'>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=6, message=F, warning=F, fig.pos='ht'>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 
-<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
+<<VisReasoningSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
 Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
@@
 The other principal components span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
-PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two tests from the paper folding test. 
+PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
 More detailed results from the 4-test analysis are provided in Appendix \ref{app:pca}.
@@
 \subsection{Lineup Types}
-\newtext{Each of the three sets of 20 lineups was taken from previous experiments which tested different types of plots to determine which plot type most effectively conveyed important characteristics of the underlying data set.
-}
-\comment{Add experiment summary/refs here}
-\newtext{Examples lineups for each experiment are provided in Appendix \ref{app:lineuptypes}. 
-}
-
-\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
-}
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
-\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment ...
-}
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
-\newtext{The final lineup section explored qq plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
-\comment{Add citations, more explanations, etc.}
-
-\newtext{Calculating participant scores on each block of lineups separately, we can use principal components analysis to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of this experiment.}
+Each of the three sets of 20 lineups was taken from previous experiments which tested different types of plots to determine which plot type most effectively conveyed important characteristics of the underlying data set.
+Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix \ref{app:lineuptypes}.
+
+In order to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of an experiment, we employ principal component analysis on participant scores averaged across each block of lineups.
 
@@
 
-\newtext{We see that again, PC1 accounts for general visual aptitude.
+From figure~\ref{fig:lineupblockpca} we see that again, PC1 accounts for general visual aptitude.
 PC2, however, emphasizes the relationship between figure classification and Lineup task 1 (and to a lesser extent, on lineup task 2).
@@
 PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test.
-PC 6 and 7 account for \Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
-}
+PC6 and PC7 account for \newtext{joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.} %\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
+
+\comment{What is the overall conclusion?}
+
 \subsection{Lineup Plot Types}
-
+\comment{Susan, would you agree in the assessment that the results from this section are inconclusive at best?
+Unless I am overlooking anything, I don't see substantial findings.
+We could move this section to the appendix, and just refer to it as being inconclusive.}
 \newtext{
@@
 \subsection{Lineup Set 1}
+
+\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
 \comment{Description of original purpose of the experiment}
@@
 \comment{Description of original purpose of the experiment. Note that this experiment appears to have more outliers.}
+
+\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
+Participants were much more accurate in this experiment ...
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
 \begin{figure}[htbp]
@@
 \comment{Description of original purpose of the experiment.}
+\newtext{The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
+\comment{Add citations, more explanations, etc.}
+
 \begin{figure}[htbp]

### 2014-11-17 Paper/Draft1/TestingVisualAptitude.Rnw
Remove newtext since it's been there a while

@@
 
-\newtext{One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
+One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
 In addition, lineups provide researchers with a rigorous framework for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
-As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.}
+As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.
 

### 2014-11-17 Paper/Draft1/TestingVisualAptitude.Rnw
theme_bw() for PCA plot

@@
 library(grid) # needed for arrow function
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("")
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
 @

### 2014-11-17 Paper/Draft1/TestingVisualAptitude.Rnw
Re-do explanation of lineup-type PCA, with additional explanation of future work

@@
 From figure~\ref{fig:lineupblockpca} we see that again, PC1 accounts for general visual aptitude. 
-PC2, however, emphasizes the relationship between figure classification and Lineup task 1 (and to a lesser extent, on lineup task 2).
+% PC2 emphasizes the relationship between figure classification and lineup task 1 (and to a lesser extent, on lineup task 2).
 PC3 differentiates the lineup task 3 from the visual search task; other tasks do not impact PC3 to a meaningful degree.
@@
 PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test.
-PC6 and PC7 account for \newtext{joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.} %\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
-
-\comment{What is the overall conclusion?}
+PC6 and PC7 account for a joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.
+\newtext{PC2 emphasizes the effect of performance on lineup test 1, which is associated with all of the measures of visual aptitude.
+PC3 emphasizes the association between performance on lineup test 3 and the visual search test, while PC4 emphasizes the association between performance on lineup test 2 and performance on the paper folding test.
+All three lineups, plus the figure classification, paper folding, and visual search tests contribute to PC5.
+PC6 and PC7 jointly account for less than 10\% of the variance in the data and do not display any distinct patterns in the loadings.
+}
+
+\comment{If PC6-7 aren't important, then we still only have 5 PCs, or one additional PC from the grouped analysis. }
+%\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
 
@@
 We could move this section to the appendix, and just refer to it as being inconclusive.}
+% \comment{What is the overall conclusion?}
 \newtext{
@@
 }
+While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
+Overall, there is an additional principal component gained from separating the lineup blocks by experiment type.
+As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
+}
 
@@
                                      )
+\newtext{The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
+}
 
 cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
+\newtext{A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
+A preliminary analysis of performance on different types of plots is available in appendix \ref{app:plottypes}, but a larger study is needed for definitive results.
+The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
+Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.}
 

### 2014-11-17 Paper/Draft1/TestingVisualAptitude.Rnw
Move plot-type section to the appendix

@@
 
-\subsection{Lineup Plot Types}
-\comment{Susan, would you agree in the assessment that the results from this section are inconclusive at best?
-Unless I am overlooking anything, I don't see substantial findings.
-We could move this section to the appendix, and just refer to it as being inconclusive.}
 % \comment{What is the overall conclusion?}
 \newtext{
+While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
+Overall, there is an additional principal component gained from separating the lineup blocks by experiment type.
+As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
+}
+
+\newtext{The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
+}
+
+\newtext{A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
+A preliminary analysis of performance on different types of plots is available in appendix \ref{app:plottypes}, but a larger study is needed for definitive results.
+The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
+Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.}
+
+\section{Discussion and Conclusions}
+
+All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
+
+\bibliographystyle{IEEEtran}
+\bibliography{references}
+
+\appendices
+
+% \section{Examining potential testing biases}\label{app:ratests}
+% T-tests of results for Hillary and Stephanie: 
+% <<ttest, echo=F, include=T, results='asis'>>=
+% ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
+% card.rotation <- t.test(card_rot~tester, data=ans.summary)
+% paper.folding <- t.test(folding~tester, data=ans.summary)
+% fig.class <- t.test(fig_class~tester, data=ans.summary)
+% lineup <- t.test(lineup~tester, data=ans.summary)
+% 
+% res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
+%   tmp <- t.test(i~ans.summary$tester)
+%   data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
+%              t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
+% })
+% 
+% res$.id <- c("Card Rotation", "Paper Folding", "figure classification", "Lineups")
+% names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
+% 
+% print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
+% @
+
+\section{Scaling Scores}\label{app:ScoreAdj}
+To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
+The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
+
+This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
+To get a single unified score with standard deviation, we completed the following calculations: 
+\begin{align}
+\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
+\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
+\end{align}
+\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
+Substituting in the provided numbers, we get}
+\begin{align*}
+\mu_{\text{all}} &= \left(323\!\cdot\!114.9\! +\! 294\!\cdot\!120.0\right)/(323\!+\!294) \\
+& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
+\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
+& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
+\end{align*}
+
+Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
+We can achieve comparability by scaling the scores accordingly.
+For example, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
+
+\begin{eqnarray*}
+\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
+\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
+\end{eqnarray*}
+
+\section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
+Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
+Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).
+Due to sample size considerations, multiple testing corrections were not performed; in addition, the independent variables are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games.
+As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
+To better understand the effects of these variables, a much larger study would be required.
+<<ttests.categ, echo=F, eval=T, results='asis'>>=
+tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
+    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp)[1,]
+    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
+    rownames(coefs) <- NULL
+    coefs <- coefs[,c(6, 1:5)]
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "MeanSq", "F", "p.val")
+    coefs
+})
+
+tmp <- rbind.fill(tmp)
+tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
+tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
+print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Results of lineup score modeled by single demographic variables.
+\\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")
+@
+
+\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
+\subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
+
+<<pcasummary4,echo=F, results='asis'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+@
+Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
+
+\newtext{PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test, while}
+
+<<pcarotation4,echo=F, results='asis'>>=
+print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+@
+
+\newtext{
+PC3 differentiates these two from the paper folding test.
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.}
+
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
+PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
+\\label{fig:biplots4}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
+@
+
+\subsection{PCA of all Five Tests (Cognitive Tests and Lineups)}\label{app:pca5}
+PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
+Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}.
+
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+@
+
+Figure classification is strongly related to lineups (PC2, PC3).
+Performance on the visual search task  is also related to lineup performance (PC3).
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+
+The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+
+PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
+
+\section{Lineup Task Examples}\label{app:lineuptypes}
+\subsection{Lineup Set 1}
+
+\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
+\comment{Description of original purpose of the experiment}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
+\caption{Boxplots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
+\caption{Density plots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
+\caption{Dotplots used to compare the two distributions.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
+\caption{Histograms used to compare the two distributions.}
+\end{figure}
+\FloatBarrier
+\subsection{Lineup Set 2}
+\comment{Description of original purpose of the experiment. Note that this experiment appears to have more outliers.}
+
+\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
+Participants were much more accurate in this experiment ...
+}
+\comment{Add note about experimental performance, general accuracy rate, etc.}
+
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f94f6965d7}
+\caption{Boxplots.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
+\caption{Boxplots with jittered points.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9fc58d73}
+\caption{Bee swarm boxplots.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
+\caption{Violin plots.}
+\end{figure}
+\FloatBarrier
+\subsection{Lineup Set 3}
+\comment{Description of original purpose of the experiment.}
+\newtext{The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
+\comment{Add citations, more explanations, etc.}
+
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
+\caption{QQ Plot with guide line.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
+\caption{QQ Plot with acceptance bands.}
+\end{figure}
+\begin{figure}[htbp]
+\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
+\caption{QQ Plot rotated 45 degrees.}
+\end{figure}
+\section{Lineup Plot Types} \label{app:plottypes}
 We can also compare participants' performance on specific types of lineup plots compared with their scores on the visual aptitude tests, for instance, accuracy on lineups which require mental rotation may be related to performance on the card rotation task.
-}
 
@@
 @
-\newtext{
+
 Figure \ref{fig:plottypesdensity} compares performance on each different type of plot.
@@
 difference in medians), different tasks are shown as different colors, so that accuracy on tasks which are shown in blue can be compared to other blue density curves.
-}
-While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
-Overall, there is an additional principal component gained from separating the lineup blocks by experiment type.
-As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
-}
-
-\newtext{
+
 Figure \ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
@@
 In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots?
-}
+
 <<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores.\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
@@
                                      )
-\newtext{The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
-}
 
 cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-\newtext{A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
-A preliminary analysis of performance on different types of plots is available in appendix \ref{app:plottypes}, but a larger study is needed for definitive results.
-The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
-Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.}
 
@@
 
-\newtext{For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.}
-\newtext{While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
-}
-\newtext{Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in figure \ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
-}
-\section{Discussion and Conclusions}
-
-All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\appendices
-
-% \section{Examining potential testing biases}\label{app:ratests}
-% T-tests of results for Hillary and Stephanie: 
-% <<ttest, echo=F, include=T, results='asis'>>=
-% ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
-% card.rotation <- t.test(card_rot~tester, data=ans.summary)
-% paper.folding <- t.test(folding~tester, data=ans.summary)
-% fig.class <- t.test(fig_class~tester, data=ans.summary)
-% lineup <- t.test(lineup~tester, data=ans.summary)
-% 
-% res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
-%   tmp <- t.test(i~ans.summary$tester)
-%   data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
-%              t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
-% })
-% 
-% res$.id <- c("Card Rotation", "Paper Folding", "figure classification", "Lineups")
-% names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
-% 
-% print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
-% @
-
-\section{Scaling Scores}\label{app:ScoreAdj}
-To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
-The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
-
-This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
-To get a single unified score with standard deviation, we completed the following calculations: 
-\begin{align}
-\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
-\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
-\end{align}
-\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
-Substituting in the provided numbers, we get}
-\begin{align*}
-\mu_{\text{all}} &= \left(323\!\cdot\!114.9\! +\! 294\!\cdot\!120.0\right)/(323\!+\!294) \\
-& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
-\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
-& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
-\end{align*}
-
-Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
-We can achieve comparability by scaling the scores accordingly.
-For example, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
-
-\begin{eqnarray*}
-\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
-\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
-\end{eqnarray*}
-
-\section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
-Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).
-Due to sample size considerations, multiple testing corrections were not performed; in addition, the independent variables are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games.
-As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
-To better understand the effects of these variables, a much larger study would be required.
-<<ttests.categ, echo=F, eval=T, results='asis'>>=
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "MeanSq", "F", "p.val")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
-tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
-print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Results of lineup score modeled by single demographic variables.
-\\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")
-@
-
-\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
-\subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
-
-<<pcasummary4,echo=F, results='asis'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-@
-Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
-
-\newtext{PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test, while}
-
-<<pcarotation4,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
-
-\newtext{
-PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.}
-
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
-PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
-\\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
-@
-
-\subsection{PCA of all Five Tests (Cognitive Tests and Lineups)}\label{app:pca5}
-PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}.
-
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
-
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
-
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-
-\section{Lineup Task Examples}\label{app:lineuptypes}
-\subsection{Lineup Set 1}
-
-\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
-}
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
-\comment{Description of original purpose of the experiment}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
-\caption{Boxplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
-\caption{Density plots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
-\caption{Dotplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
-\caption{Histograms used to compare the two distributions.}
-\end{figure}
-\FloatBarrier
-\subsection{Lineup Set 2}
-\comment{Description of original purpose of the experiment. Note that this experiment appears to have more outliers.}
-
-\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment ...
-}
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f94f6965d7}
-\caption{Boxplots.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
-\caption{Boxplots with jittered points.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9fc58d73}
-\caption{Bee swarm boxplots.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
-\caption{Violin plots.}
-\end{figure}
-\FloatBarrier
-\subsection{Lineup Set 3}
-\comment{Description of original purpose of the experiment.}
-\newtext{The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
-\comment{Add citations, more explanations, etc.}
-
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
-\caption{QQ Plot with guide line.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
-\caption{QQ Plot with acceptance bands.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
-\caption{QQ Plot rotated 45 degrees.}
-\end{figure}
+For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
+Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in figure \ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
+
 \end{document}

### 2014-11-17 Paper/Draft1/TestingVisualAptitude.Rnw
Move new PC chunk up so it's included in the document.
Remove newtext for text that is no longer new

@@
 
-\newtext{Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants would be expected to determine that sub-plot 4 is the most different because the two groups have notably different medians.}
+Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants would be expected to determine that sub-plot 4 is the most different because the two groups have notably different medians.
 
@@
 \end{figure}
+
+<<>>=
+vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
+lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
+vmain <- rowSums(vistests)/4
+alltests <- data.frame(vmain, vistests-vmain, lineups)
+pctests <- prcomp(alltests)
+biplot(pctests)
+biplot(pctests, 3:4)
+
+@
+
 \section{Lineup Plot Types} \label{app:plottypes}
@@
 % \end{itemize}
-
-<<>>=
-vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
-lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
-vmain <- rowSums(vistests)/4
-alltests <- data.frame(vmain, vistests-vmain, lineups)
-pctests <- prcomp(alltests)
-biplot(pctests)
-biplot(pctests, 3:4)
-
-@

### 2014-11-18 Paper/Draft1/TestingVisualAptitude.Rnw
edits during the meeting

@@
 \textbf{A few sentences added, still need to tie into the existing literature}\\
-(4) I'll take care of the intro to lineups}
+(4) I'll take care of the intro to lineups\\
+(5) Include lineup results from other studies. For that, we need to compare percent correct attempted data. 
+}
 
@@
 
+\comment{ Just collecting items:
+lineups are classification tests, not necessarily visual tests; 
+individual tasks are somewhat more related to individual tests, but no overall general trend towards any of the other tests besides figure classification.
+
+mathematical training - could be interesting follow-up study.
+
+amazon turk study - how generalizable are they.
+
+use figure classification for pre-screening without biasing; i.e. as baseline criteria or for exclusion of participants from future amazon turk studies.
+
+}
+
 All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.

### 2014-11-19 Code/Analysis.R
calculating lineup percentage scores by section, with comment.

@@ -123,6 +123,8 @@ lineup.section.sum$testnum[is.na(lineup.section.sum$testnum)] <- ""
 lineup.section.sum$testnum[nchar(lineup.section.sum$testnum)>0] <- paste0("_", lineup.section.sum$testnum[nchar(lineup.section.sum$testnum)>0])
 lineup.section.sum$testtype <- paste0(lineup.section.sum$testtype, lineup.section.sum$testnum)
 
+lineup.section.pct <- dcast(subset(lineup.sec.sum, testtype=="lineup")%>%mutate(pct.score=pos.pts/total.pos), id~testtype+testnum, value.var="pct.score")
+
 lineup.section.summary <- merge(ans[,1:20], dcast(lineup.section.sum, id~testtype, value.var="score", na.rm=TRUE))
 rm(lineup.sec.sum2, lineup.sec.sum, lineup.section.sum)
 

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
calculating lineup percentage scores by section, with comment.
Lineup type scatterplot matrix, with visual and reasoning tests as well, plus some commentary.

@@
 (5) Include lineup results from other studies. For that, we need to compare percent correct attempted data. 
+\textbf{In \texttt{lineup.section.pct}.}
 }
@@
 
+\newtext{Figure \ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
+The visual search task was only slightly correlated with the three lineup tasks and has been omitted from this figure.
+Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
+Performance on lineup task 3, which dealt with the identification of nonnormal residual qq plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
+This suggests that certain lineup tasks may require more visual ability than others; in this case, the qq-plot lineups required participants to mentally rotate plots, compare vertical distances, and perform more mental manipulation than the first two lineup tasks.
+}
+
+<<LineupTypeSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=4, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
+All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
+\\label{fig:lineuptaskmatrix}">>=
+data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
+grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
+grid <- subset(grid, x != y & x<y)
+
+all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
+  xcol <- grid[i, "x"]
+  ycol <- grid[i, "y"]
+
+  rbind.fill(
+    data.frame(
+      xvar = names(data)[ycol], 
+      yvar = names(data)[xcol],
+      x = data[, xcol], 
+      y = data[, ycol], 
+      points = T,
+      data
+    ),
+    data.frame(
+      yvar = names(data)[ycol], 
+      xvar = names(data)[xcol],
+      x = mean(range(data[, ycol])), 
+      y = mean(range(data[, xcol])),
+      label = round(cor(data[,xcol], data[,ycol]), 3),
+      points = F
+    )
+  )
+}))
+all$xvar <- factor(all$xvar, levels=names(data))
+all$yvar <- factor(all$yvar, levels=names(data))
+densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
+    tmp <- density(data[,i], cut = 1, adjust=.75)
+    data.frame(
+      xvar = names(data)[i], 
+      yvar = names(data)[i],
+      x = tmp$x, 
+      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
+    )
+  }))
+fix.names <- function(values){
+  values %>% 
+    str_replace("lineup", "Lineups") %>% 
+    str_replace("vis_search", "Visual Search") %>%
+    str_replace("fig_class", "Figure Classification") %>% 
+    str_replace("card_rot", "Card Rotation") %>%
+    str_replace("folding", "Paper Folding")
+}
+
+all2 <- subset(all, grepl("lineup", yvar) & points & !grepl("vis_search", xvar))
+names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
+all2mean <- rbind(
+  ddply(subset(all2, !grepl("lineup", yvar)), .(xvar, yvar), summarize, label=paste0("Cor = ", round(cor(x,y), 3)), x=max(x), y=min(y)-1, hjust=1, vjust=0),
+  ddply(subset(all2, grepl("lineup", xvar) & grepl("lineup", yvar)), .(xvar, yvar), function(df){
+    with(df, data.frame(xvar=unique(yvar), yvar=unique(xvar), 
+                        x=mean(range(y)), y=mean(range(x)), 
+                        label=paste0("Correlation = \n", round(cor(x,y),3)),
+                        hjust=.5, vjust=.5))
+    })
+)
+par(mfrow=c(1,2),mar=c(4,4.5,.1,.1)
+
+ggplot()+
+  geom_point(data=subset(all2, points), aes(x=x, y=y)) + 
+  geom_smooth(data=subset(all2, points), aes(x=x, y=y), method="lm") + 
+  geom_text(data=all2mean, aes(x=x, y=y, label=label, hjust=hjust, vjust=vjust)) + 
+  geom_line(data=subset(densities, grepl("lineup", xvar)), aes(x=x, y=y)) + 
+  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
+  theme_bw() + 
+  theme(axis.title=element_blank())
+@
+
 In order to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of an experiment, we employ principal component analysis on participant scores averaged across each block of lineups.
@@
 }
-
-\comment{If PC6-7 aren't important, then we still only have 5 PCs, or one additional PC from the grouped analysis. }
+% \comment{If PC6-7 aren't important, then we still only have 5 PCs, or one additional PC from the grouped analysis. }
 %\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
Change number of columns in PCA plot

@@
 library(grid) # needed for arrow function
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
 @

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
Change interpretation of PCA of lineup types a bit for consistency with scatterplot matrix

@@
 PC6 and PC7 account for a joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.
-\newtext{PC2 emphasizes the effect of performance on lineup test 1, which is associated with all of the measures of visual aptitude.
-PC3 emphasizes the association between performance on lineup test 3 and the visual search test, while PC4 emphasizes the association between performance on lineup test 2 and performance on the paper folding test.
+\newtext{PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
+PC3 emphasizes the additional variation in performance on lineup test 3 and the visual search test, while PC4 emphasizes the extra variability in performance on lineup test 2 and the paper folding test.
 All three lineups, plus the figure classification, paper folding, and visual search tests contribute to PC5.

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
Remove newtext for older commentary on lineup type pca

@@
 % \comment{What is the overall conclusion?}
-\newtext{
+
 While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
@@
 As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
-}
-
-\newtext{The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
-}
-
-\newtext{A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
+
+The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
+
+A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
 A preliminary analysis of performance on different types of plots is available in appendix \ref{app:plottypes}, but a larger study is needed for definitive results.
 The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
-Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.}
+Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.
 
 \section{Discussion and Conclusions}
+Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by factors such as mathematical training, hours of video games played per week, and gender.
+In addition, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
+Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability. 
+This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
+
+In addition, when lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
 

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
Spacing of newtext boxes

@@
 The visual search task was only slightly correlated with the three lineup tasks and has been omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
-Performance on lineup task 3, which dealt with the identification of nonnormal residual qq plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in this case, the qq-plot lineups required participants to mentally rotate plots, compare vertical distances, and perform more mental manipulation than the first two lineup tasks.
-}
+Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.}
+\newtext{Performance on lineup task 3, which dealt with the identification of nonnormal residual qq plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
+This suggests that certain lineup tasks may require more visual ability than others; in this case, the qq-plot lineups required participants to mentally rotate plots, compare vertical distances, and perform more mental manipulation than the first two lineup tasks.}
 

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
remove comments

@@
 
-<<LineupTypeSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=4, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
+<<LineupTypeSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
 All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
@@
     str_replace("lineup", "Lineups") %>% 
+    str_replace("Lineups_", "Lineup ") %>%
     str_replace("vis_search", "Visual Search") %>%
-    str_replace("fig_class", "Figure Classification") %>% 
+    str_replace("fig_class", "Figure Class.") %>% 
     str_replace("card_rot", "Card Rotation") %>%
@@
 all2 <- subset(all, grepl("lineup", yvar) & points & !grepl("vis_search", xvar))
-names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
+# names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
 all2mean <- rbind(
-  ddply(subset(all2, !grepl("lineup", yvar)), .(xvar, yvar), summarize, label=paste0("Cor = ", round(cor(x,y), 3)), x=max(x), y=min(y)-1, hjust=1, vjust=0),
+  ddply(subset(all2, !grepl("lineup", xvar)), .(xvar, yvar), summarize, label=paste0("rho == ", round(cor(x,y), 3)), x=min(x)-diff(range(x))/5, y=max(y)+diff(range(y))/5, hjust=0, vjust=1, size=3),
   ddply(subset(all2, grepl("lineup", xvar) & grepl("lineup", yvar)), .(xvar, yvar), function(df){
@@
                         x=mean(range(y)), y=mean(range(x)), 
-                        label=paste0("Correlation = \n", round(cor(x,y),3)),
-                        hjust=.5, vjust=.5))
+                        label=paste0("rho == ", round(cor(x,y),3)),
+                        hjust=.5, vjust=.5, size=4))
     })
 )
-par(mfrow=c(1,2),mar=c(4,4.5,.1,.1)
 
@@
   geom_smooth(data=subset(all2, points), aes(x=x, y=y), method="lm") + 
-  geom_text(data=all2mean, aes(x=x, y=y, label=label, hjust=hjust, vjust=vjust)) + 
+  geom_text(data=all2mean, aes(x=x, y=y, label=label, hjust=hjust, vjust=vjust, size=size), parse=T) + 
+  scale_size_identity() + 
   geom_line(data=subset(densities, grepl("lineup", xvar)), aes(x=x, y=y)) + 
@@
 
-<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Lineup block PCA rotation matrix in graphical form.">>=
+<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Lineup block PCA rotation matrix in graphical form.", fig.width=5, fig.height=5>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 library(grid) # needed for arrow function
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=3) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
 @

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
Move mean-adjusted aptitude test PCA up

@@
 
+\paragraph{Removing the effect of overall aptitude}
+
+<<removeAptitude,echo=F,fig.width=8, fig.height=4>>=
+vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
+lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
+vmain <- rowSums(vistests)/4
+alltests <- data.frame(vmain, vistests-vmain, lineups)
+pctests <- prcomp(alltests)
+par(mfrow=c(1,2), mar=c(4,4.5,.1,.1), tcl=-0.01, cex.lab=.8,cex.axis=.7,mgp=c(2,.7,0),tcl=-.3)
+biplot(pctests)
+biplot(pctests, 3:4)
+@
+
 The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.

### 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
remove comments
Conclusion

@@
 }
-% \comment{If PC6-7 aren't important, then we still only have 5 PCs, or one additional PC from the grouped analysis. }
 %\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
-
-% \comment{What is the overall conclusion?}
 
@@
 \section{Discussion and Conclusions}
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by factors such as mathematical training, hours of video games played per week, and gender.
-In addition, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability. 
-This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
-
-In addition, when lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
+\newtext{Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by factors such as mathematical training, hours of video games played per week, and gender.}
+\newtext{In addition, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).}
+\newtext{Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.}
+\newtext{This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.}
+
+\newtext{In addition, when lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.}
+\newtext{Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).}
+\newtext{Future studies testing larger sets of lineups may be useful to understand which types of plots require additional visuospatial skills, as plots which appeal to a wider audience may be more successful when conveying important information.}
+
+\newtext{In addition to this theoretical information, the figure classification test may be useful for pre-screening participants in future online lineup studies.}
+\newtext{Such studies often suffer from participants who do not take the task seriously, and internal verification questions, as well as pre-qualification tasks are often used to reduce extraneous variability.}
+\newtext{While it would be impractical to require participants to score well on several different tests, it would be reasonable to ask participants to pre-qualify for a task by completing a figure classification test.}
+\newtext{As the figure classification test is different from the lineup task, this would not bias participants' scores on the domain of interest, but would ensure a participant pool that is sufficiently motivated to complete the lineup questions.}
+
+\newtext{The demographic results from this study indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
+}
 
@@
 use figure classification for pre-screening without biasing; i.e. as baseline criteria or for exclusion of participants from future amazon turk studies.
-
 }
@@
 \end{figure}
-
-<<>>=
-vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
-lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
-vmain <- rowSums(vistests)/4
-alltests <- data.frame(vmain, vistests-vmain, lineups)
-pctests <- prcomp(alltests)
-biplot(pctests)
-biplot(pctests, 3:4)
-
-@
 

### 2014-11-21 Paper/Draft1/TestingVisualAptitude.Rnw
read-through, small edits

--- 2014-11-20 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
+require(grid)
 ggplot()+
@@
   theme_bw() + 
-  theme(axis.title=element_blank())
+  theme(axis.title=element_blank(),
+        plot.margin=unit(c(0,0,0,0), unit="mm"))
 rm(list=c("data", "all", "densities"))
@@
 print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="ht", caption.placement="top")
 @
@@
 \subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous experiments which tested different types of plots to determine which plot type most effectively conveyed important characteristics of the underlying data set.
+Each of the three sets of 20 lineups was taken from previous studies on different designs to investigate which plot type most effectively conveyed important characteristics of the underlying data set.
 Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix \ref{app:lineuptypes}.
 
-\newtext{Figure \ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
-The visual search task was only slightly correlated with the three lineup tasks and has been omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.}
-\newtext{Performance on lineup task 3, which dealt with the identification of nonnormal residual qq plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in this case, the qq-plot lineups required participants to mentally rotate plots, compare vertical distances, and perform more mental manipulation than the first two lineup tasks.}
+Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
+The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
+Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
+Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
 
@@
 PC6 and PC7 account for a joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.
-\newtext{PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
+%
+PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
 PC3 emphasizes the additional variation in performance on lineup test 3 and the visual search test, while PC4 emphasizes the extra variability in performance on lineup test 2 and the paper folding test.
@@
 PC6 and PC7 jointly account for less than 10\% of the variance in the data and do not display any distinct patterns in the loadings.
-}
-% \comment{If PC6-7 aren't important, then we still only have 5 PCs, or one additional PC from the grouped analysis. }
 %\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
-
-% \comment{What is the overall conclusion?}
 
@@
 
-\paragraph{Removing the effect of overall aptitude}
-
+%\paragraph*{Removing the effect of overall aptitude}
 <<removeAptitude,echo=F,fig.width=8, fig.height=4>>=
@@
 
-A larger study could not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
-A preliminary analysis of performance on different types of plots is available in appendix \ref{app:plottypes}, but a larger study is needed for definitive results.
+A larger study might not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
+Preliminary results of performance on different types of plots are shown in appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
 The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
@@
 
-In addition, when lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
-
-\comment{ Just collecting items:
-lineups are classification tests, not necessarily visual tests; 
-individual tasks are somewhat more related to individual tests, but no overall general trend towards any of the other tests besides figure classification.
-
-mathematical training - could be interesting follow-up study.
-
-amazon turk study - how generalizable are they.
-
-use figure classification for pre-screening without biasing; i.e. as baseline criteria or for exclusion of participants from future amazon turk studies.
-
-}
+When lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
+Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).
+Future studies testing larger sets of lineups may be useful to understand which types of plots require additional visuospatial skills, as plots which appeal to a wider audience may be more successful when conveying important information.
+
+In addition to this theoretical information, the figure classification test may be useful for pre-screening participants in future online lineup studies.
+Such studies often suffer from participants who do not take the task seriously, and internal verification questions, as well as pre-qualification tasks are often used to reduce extraneous variability.
+While it would be impractical to require participants to score well on several different tests, it would be reasonable to ask participants to pre-qualify for a task by completing a figure classification test.
+As the figure classification test is different from the lineup task, this would not bias participants' scores on the domain of interest, but would ensure a participant pool that is sufficiently motivated to complete the lineup questions.
+
+The demographic results from this study indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
 
@@
 \end{align}
-\newtext{where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
-Substituting in the provided numbers, we get}
+where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
+Substituting in the provided numbers, we get
 \begin{align*}
@@
 As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
-To better understand the effects of these variables, a much larger study would be required.
+To better understand the effects of these variables, a larger study would be required.
 <<ttests.categ, echo=F, eval=T, results='asis'>>=
@@
 \subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
-
+%
 <<pcasummary4,echo=F, results='asis'>>=

### 2014-12-01 Paper/Draft1/TestingVisualAptitude.Rnw
Add text and reference for lineup section 1. Need performance stats

@@
 
-\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups with different means and outliers. 
+\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+The experiment was originally designed to explore the use of lineups to test plots of competing design\cite{hofmann2012graphical}.
+This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
 }
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
-\comment{Description of original purpose of the experiment}
+\comment{Describe performance - need to get data from original experiment.}
 \begin{figure}[htbp]

### 2014-12-08 Paper/Draft1/TestingVisualAptitude.Rnw
intro

@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-for IEEE Computer Society journal papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8a and later.
-
-\comment{Next few steps for the analysis: \\
-(3) in terms of why we would use lineups in a test comparison ...
-lineups give us a way to say 'this is correct' which regular graphics don't.
-So we should put a spin sentence on at the intro and the back that we are suggesting to use lineups as a tool to not only evaluate individuals (which we can see is possible) but also as a tool to evaluate charts ...
-which we've done before ...
-and tie it into the regular testing literature ...\\
-\textbf{A few sentences added, still need to tie into the existing literature}\\
-(4) I'll take care of the intro to lineups\\
-(5) Include lineup results from other studies. For that, we need to compare percent correct attempted data. 
-\textbf{In \texttt{lineup.section.pct}.}
-}
-
-\todo[inline]{Intro about statistical graphics, lineups, ...}
-\subsection{Spatial Reasoning and Statistical Graphics}
-Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
+\IEEEPARstart{S}{tatistical} graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
 Mathematical ability is important, even for the simplest graphs:  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph.
+
+%\todo[inline]{Intro about statistical graphics, lineups, ...}
+%\subsection{Spatial Reasoning and Statistical Graphics}
 Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
@@
 
-\subsection{The Lineup Protocol}
-\todo[inline]{Standard lineup protocol introduction here.}
-
-Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants would be expected to determine that sub-plot 4 is the most different because the two groups have notably different medians.
-
-Statistical lineups~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+%\subsection{The Lineup Protocol}
+%\todo[inline]{Standard lineup protocol introduction here.}
+
+\newtext{The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is introducing a testing framework that allows to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
+}
+
+Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
+
+Statistical lineups depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
 Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).

### 2014-12-15 Paper/Draft1/TestingVisualAptitude.Rnw
grammar

@@
 
-\newtext{The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is introducing a testing framework that allows to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
+\newtext{The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
 }

### 2014-12-15 Paper/Draft1/TestingVisualAptitude.Rnw
Format appendix, remove newtext, etc.

@@
 
-%\todo[inline]{Intro about statistical graphics, lineups, ...}
-%\subsection{Spatial Reasoning and Statistical Graphics}
 Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
@@
 
-%\subsection{The Lineup Protocol}
-%\todo[inline]{Standard lineup protocol introduction here.}
-
-\newtext{The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
-}
+The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
 
@@
 
-\newtext{PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test, while}
+PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.
 
@@
 
-\newtext{
-PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.}
-
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
 PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
@@
 
-\subsection{PCA of all Five Tests (Cognitive Tests and Lineups)}\label{app:pca5}
+\subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
@@
 
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.8\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
 \\label{fig:biplots5}", echo=F, fig.env="figure">>=
@@
 \subsection{Lineup Set 1}
-
-\newtext{The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
 The experiment was originally designed to explore the use of lineups to test plots of competing design\cite{hofmann2012graphical}.
 This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-}
+
 \comment{Describe performance - need to get data from original experiment.}
@@
 \subsection{Lineup Set 2}
-\comment{Description of original purpose of the experiment. Note that this experiment appears to have more outliers.}
-
-\newtext{The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment ...
-}
-\comment{Add note about experimental performance, general accuracy rate, etc.}
-
+The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
+Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
 \begin{figure}[htbp]
@@
 \subsection{Lineup Set 3}
-\comment{Description of original purpose of the experiment.}
-\newtext{The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
 Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.}
-\comment{Add citations, more explanations, etc.}
+As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.
+% \comment{Add citations, more explanations, etc.}
 
@@
 Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks.
-In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots?
+In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots.
 

### 2014-12-15 Paper/Draft1/TestingVisualAptitude.Rnw
Attempt at an abstract

@@
 \begin{abstract}
-The abstract goes here.
+Graphics convey statistical information very efficiently, but rely on a different set of mental processes than tabular displays.
+This study examines the demographic characteristics and visual skills associated with perception of statistical lineups.
+We conclude that statistical lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
+We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
 \end{abstract}

### 2015-01-05 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments for the intro

@@
 \begin{document}
-\title{Spatial Reasoning and Statistical Graphics}
+\title{Spatial Reasoning and Data Displays}
 
@@
 \begin{abstract}
-Graphics convey statistical information very efficiently, but rely on a different set of mental processes than tabular displays.
+\comment{Tone down the `statistical' part ...
+statistical, to the wrong ears sounds complicated.
+} Graphics convey statistical information very efficiently, but rely on a different set of mental processes than tabular displays.
+\newtext{}
 This study examines the demographic characteristics and visual skills associated with perception of statistical lineups.
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{S}{tatistical} graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
-Mathematical ability is important, even for the simplest graphs:  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph.
-
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-Graphics utilize higher-bandwidth visual pathways to encode information\cite{baddeley1974working}, allowing viewers to quickly and intuitively encode multiple dimensions of numerical relationships; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability.
+\comment{OK, slow down!
+The intro is still a bit disconnected - I've tried to add the sentences that I think are missing in the first paragraph, but I'm not sure what story you want to tell.
+What would help is to write a sentence each to summarise a paragraph, and then work on connections between the paragraph.
+}
+\IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but \newtext{not all displays are equally good, nor is any data display } %these displays may not be 
+equally useful to all viewers.
+
+Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively \newtext{relate} multiple dimensions of numerical \newtext{quantities.}%relationships
+%; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability. 
 Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
@@
 
+However,  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
+Spatial ability becomes more important, though, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+
+\comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
+In particular, we need to answer the questions:
+
+- why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and mak graphics a better tool for more people.
+
+- what are the limits of the studies that you have cited, ie is there something that we can do better?
+
+- what advantage do lineups bring to the table?
+}
+
+\comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
+
+In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
+
+\section{Methods}
+
+\subsection{The Lineup Protocol}
 The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
@@
 
-In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
-
-\section{Methods}
 \subsection{Measures of visuospatial ability}

### 2015-01-05 Paper/Draft1/TestingVisualAptitude.Rnw
notes during the meeting

@@
 % and "HIS" in caps to complete the first word.
-\comment{OK, slow down!
-The intro is still a bit disconnected - I've tried to add the sentences that I think are missing in the first paragraph, but I'm not sure what story you want to tell.
-What would help is to write a sentence each to summarise a paragraph, and then work on connections between the paragraph.
-}
 \IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but \newtext{not all displays are equally good, nor is any data display } %these displays may not be 
@@
 
-- why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and mak graphics a better tool for more people.
+- why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and make graphics a better tool for more people.
 
@@
 - what advantage do lineups bring to the table?
+}
+
+\comment{How do perceptual principles apply to statistical graphics (data displays)? - Lineups help us to isolate different parts of displays and allow in-built quantification of results.}
+
+\comment{Why are lineups variable?
+- Individual variability is big, we want to investigate which abilities individuals help the most in evaluating.
+- for this reason we need to motivate why lineups are useful.
 }

### 2015-01-07 Paper/Draft1/TestingVisualAptitude.Rnw
Paragraph linking lineups to graphical displays and justifying spatial reasoning tests.

--- 2014-12-15 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 \begin{abstract}
-Graphics convey statistical information very efficiently, but rely on a different set of mental processes than tabular displays.
-This study examines the demographic characteristics and visual skills associated with perception of statistical lineups.
-We conclude that statistical lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
+\comment{Tone down the `statistical' part ...
+statistical, to the wrong ears sounds complicated.
+} Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
+\newtext{}
+This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
+We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
 We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{S}{tatistical} graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers.
-Mathematical ability is important, even for the simplest graphs:  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph.
-
-Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-Graphics utilize higher-bandwidth visual pathways to encode information\cite{baddeley1974working}, allowing viewers to quickly and intuitively encode multiple dimensions of numerical relationships; however, this higher-bandwidth approach also depends on the viewer's visual and spatial reasoning ability.
+\IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
+Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
 Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
@@
 In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
-
+However,  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
+Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
+
+\newtext{
+"Lineups" are often used to evaluate the effectiveness of different graphical displays \cite{kosara2010mechanical,wickham2010graphical,majumder2013validation}; as the police counterpart, lineups consist of several distractor plots and one target.
+This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning. 
+}\newtext{
+Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
+This removes some of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require familiarity with data-based graphics. 
+}\newtext{
+While previous research has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills required by the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may differentially impact performance\cite{humanfactorslineups}.
+Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the data. 
+}
+
+% \comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
+% In particular, we need to answer the questions: 
+% 
+% - why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and make graphics a better tool for more people.
+% 
+% - what are the limits of the studies that you have cited, ie is there something that we can do better?
+% 
+% - what advantage do lineups bring to the table?
+% }
+% 
+% 
+% \comment{How do perceptual principles apply to statistical graphics (data displays)? - Lineups help us to isolate different parts of displays and allow in-built quantification of results.}
+% 
+% 
+% \comment{Why are lineups variable?
+- Individual variability is big, we want to investigate which abilities individuals help the most in evaluating.
+- for this reason we need to motivate why lineups are useful.
+}
+% 
+% 
+% \comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
+
+In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
+
+\section{Methods}
+
+\subsection{The Lineup Protocol}
 The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.

### 2015-01-07 Paper/Draft1/TestingVisualAptitude.Rnw
added and edited a few references; suggestion for intro

@@
 Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior\cite{zhang1997nature}.
-
-\newtext{
-"Lineups" are often used to evaluate the effectiveness of different graphical displays \cite{kosara2010mechanical,wickham2010graphical,majumder2013validation}; as the police counterpart, lineups consist of several distractor plots and one target.
-This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning. 
-}\newtext{
-Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
-This removes some of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require familiarity with data-based graphics. 
-}\newtext{
-While previous research has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills required by the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may differentially impact performance\cite{humanfactorslineups}.
-Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the data. 
-}
-
-% \comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
-% In particular, we need to answer the questions: 
-% 
-% - why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and make graphics a better tool for more people.
-% 
-% - what are the limits of the studies that you have cited, ie is there something that we can do better?
-% 
-% - what advantage do lineups bring to the table?
-% }
-% 
-% 
-% \comment{How do perceptual principles apply to statistical graphics (data displays)? - Lineups help us to isolate different parts of displays and allow in-built quantification of results.}
-% 
-% 
-% \comment{Why are lineups variable?
-- Individual variability is big, we want to investigate which abilities individuals help the most in evaluating.
-- for this reason we need to motivate why lineups are useful.
-}
-% 
-% 
-% \comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
-
-In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
-
-\section{Methods}
-
-\subsection{The Lineup Protocol}
-The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows us to quantify the significance of graphical findings with the same mathematical rigor as conventional hypothesis tests.
+It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~\cite{zhang1997nature}.
+
+\comment{I am not sure what the kosara reference has to do with the lineups - the only thing lineups and kosara have in common are the amazon studies.
+but you might want to keep the reference around for a discussion of the amazon results }
+``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
+Lineups are also useful in assessing the effectiveness of different graphical displays \cite{kosara2010mechanical, heike:2012, loy:2015}.
+Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
 
@@
 In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
-
-Statistical lineups depend on the ability to search for a signal amid distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
-Some lineups (polar coords) also depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
-By breaking the lineup task down into component parts, we may be able to determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups \cite{humanfactorslineups}.
-In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
 
@@
 
-One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively (providing an experimental protocol for comparing graphics based on the accuracy of user conclusions).
+%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
+
+Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
+This removes a lot of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
+
+\comment{`While previous research' - which research are you referring to? 
+Yifan's paper with the eyetracking study has looked at the spatial connection \cite{zhao2013mind}. But is that what you mean?
+
+`required by the lineup task' - that makes the lineup task strangely active. Could you rephrase that in terms of the participant?
+}
+\newtext{
+While previous research has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills required by the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
+Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the study results. 
+}
+
+\comments{This is going into the right direction, but a few pieces are still missing, 
+you need to still say what this paper is doing: we do a study to compare lineup performance with established tests, also go into a bit of discussion of the tests.}
+% \comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
+% In particular, we need to answer the questions: 
+% 
+% - why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and make graphics a better tool for more people.
+% 
+% - what are the limits of the studies that you have cited, ie is there something that we can do better?
+% 
+% - what advantage do lineups bring to the table?
+% }
+% 
+% 
+% \comment{How do perceptual principles apply to statistical graphics (data displays)? - Lineups help us to isolate different parts of displays and allow in-built quantification of results.}
+% 
+% 
+% \comment{Why are lineups variable?
+- Individual variability is big, we want to investigate which abilities individuals help the most in evaluating.
+- for this reason we need to motivate why lineups are useful.
+}
+% 
+% 
+% \comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
+
+In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
+
+\section{Methods}
+
+\subsection{The Lineup Protocol}
+The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
+
+\comment{you can dare to go into a bit of math here - introduce a bit of the lineup math here.  random variable and approximate distribution, explanation for choosing lineups of size 20. }
+
+Statistical lineups depend on the ability to search for a signal amid a set of distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
+\newtext{Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from   participants, e.g.~ polar coordinates} depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
+By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
+
+Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups \cite{humanfactorslineups}.
+In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
+
+One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively by providing an experimental protocol for comparing graphics based on the accuracy of user conclusions.
 In addition, lineups provide researchers with a rigorous framework for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
@@
 Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
-Tasks are times such  that participants are under  pressure to complete; participants are not expected to finish all of the problems in each section.
+Tasks are timed such  that participants are under  pressure to complete; participants are not expected to finish all of the problems in each section.
 This allows for a better discrimination between scores and prevents score compression at the top of the range.
@@
 
+\comment{Let's put figure \ref{fig:tests} into the appendix - that'll save us quite a bit of space. }
  \begin{figure*}[ht]
@@
 
-Between cognitive tasks, participants will also complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
 Participants have 5 minutes to complete each block of 20 lineups.
@@
 This is reasonably representative\footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}} of the university as a whole; in the fall 2012 semester, 23\% of students were associated with the college of engineering, 23\% were associated with the college of liberal arts and sciences, 23\% were associated with the college of human sciences, 12\% with the business school, and 14\% with the school of agriculture.
+
+\comment{here are the numbers for Fall 2014: http://www.registrar.iastate.edu/enrollment}
 
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + 
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey70")) + 
   facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 

### 2015-01-07 Paper/Draft1/TestingVisualAptitude.Rnw
comments throughout, but just minor ones

@@
 
-\comments{This is going into the right direction, but a few pieces are still missing, 
+\comment{This is going into the right direction, but a few pieces are still missing, 
 you need to still say what this paper is doing: we do a study to compare lineup performance with established tests, also go into a bit of discussion of the tests.}
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey70")) + 
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey")) + 
   facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 
@@
 
+\comment{First: what is shown in figure 8? THis is a non traditional plot, so you need to explain it first.}
 From figure~\ref{fig:lineupblockpca} we see that again, PC1 accounts for general visual aptitude. 
@@
 \section{Discussion and Conclusions}
+\comment{These read like the conclusions from earlier - video games aren't that prominently displayed anymore unless you are referring to figure 5.
+But if so, I would be cautious about summarising STEM major as 'mathematical training' - use the same names for the variables as in figure 5 and make sure to keep the order, i.e.
+gender is more impotant than video game hours.
+also make sure, that to mention that this is on a univariate level and doesn't take the correlation between the variables into account.
+}
+
 Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by factors such as mathematical training, hours of video games played per week, and gender.
@@
 \section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
+\comment{'Results' of lineup score - te caption in the table is too generic. give more details. }
 Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
@@
 \subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
+\comment{'Importance of principal components {\it in an analysis} of four tests' -- be more precise: you mean the PCA. Give some details as to how importance is measured. }
 %
@@
 @
-
+\comment{figure 9 and 10 are Biplots, say so in the caption - in the discussion of the figures and the tables in the text, please think again of adding the sentence summarising what is shown. }
 \subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}

### 2015-01-08 Paper/Draft1/TestingVisualAptitude.Rnw
Modified intro

@@
 
-\comment{I am not sure what the kosara reference has to do with the lineups - the only thing lineups and kosara have in common are the amazon studies.
+% \comment{I am not sure what the kosara reference has to do with the lineups - the only thing lineups and kosara have in common are the amazon studies.
 but you might want to keep the reference around for a discussion of the amazon results }
 ``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
-Lineups are also useful in assessing the effectiveness of different graphical displays \cite{kosara2010mechanical, heike:2012, loy:2015}.
+Lineups are also useful in assessing the effectiveness of different graphical displays \cite{heike:2012, loy:2015}.
 Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
@@
 Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
-This removes a lot of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-
-\comment{`While previous research' - which research are you referring to? 
-Yifan's paper with the eyetracking study has looked at the spatial connection \cite{zhao2013mind}. But is that what you mean?
-
-`required by the lineup task' - that makes the lineup task strangely active. Could you rephrase that in terms of the participant?
+This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
+
+% 
+% \comment{`While previous research' - which research are you referring to? 
+% Yifan's paper with the eyetracking study has looked at the spatial connection \cite{zhao2013mind}. But is that what you mean?
+% 
+% `required by the lineup task' - that makes the lineup task strangely active. Could you rephrase that in terms of the participant?
+% }
+\newtext{
+While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
+% Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the study results. 
+} \newtext{
+Our experiment compares lineup performance with established tests, such as the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
+The VST measures visual search speed, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning; 
+all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
+We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand human of the perception of statistical lineups.
+
 }
-\newtext{
-While previous research has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills required by the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
-Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the study results. 
-}
-
-\comment{This is going into the right direction, but a few pieces are still missing, 
-you need to still say what this paper is doing: we do a study to compare lineup performance with established tests, also go into a bit of discussion of the tests.}
+% 
+% \comment{This is going into the right direction, but a few pieces are still missing, 
+% you need to still say what this paper is doing: we do a study to compare lineup performance with established tests, also go into a bit of discussion of the tests.}
 % \comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
@@
 % \comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
-
-In this paper, we examine some demographic characteristics as well as characteristics such as  spatial and pattern recognition ability with the goal of understanding the skills necessary for accurate lineup performance as well as underlying variation in lineup performance.
 

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
few little things for the intro

@@
 ``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
-Lineups are also useful in assessing the effectiveness of different graphical displays \cite{heike:2012, loy:2015}.
+Lineups are also useful in assessing the effectiveness of different graphical displays \cite{hofmann2012graphical, loy:2015}.
 Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
@@
 % }
-\newtext{
+
 While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
 % Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the study results.
-} \newtext{
+
+\comment{good discussion, but this is the first time that you mention `our experiment' - include a sentence here, to say what this experiment is about.
+also: I think we should only say `study' or `experiment', and I'd prefer study, because experiment sounds like it could go BOOM.
+I don't want to give that impression.
+}
+
 Our experiment compares lineup performance with established tests, such as the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
+\comment{include citations for the tests}
 The VST measures visual search speed, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning; 
@@
 We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand human of the perception of statistical lineups. 
-
-}
+\comment {something went wrong with the last half sentence ... do you mean the human component in the statistical lineups?}
+
 % 
@@
 
-\section{Methods}
+\comment{and in the last paragraph of the intro: lay out the roadmap for the rest of the paper. In section \ref{sec:methods} we introduce ... }
+
+\section{Methods}\label{sec:methods}
 
@@
 Statistical lineups depend on the ability to search for a signal amid a set of distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
-\newtext{Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from   participants, e.g.~ polar coordinates} depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
+Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from   participants, e.g.~ polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
 By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
lineup math

@@
 
-\comment{you can dare to go into a bit of math here - introduce a bit of the lineup math here.  random variable and approximate distribution, explanation for choosing lineups of size 20. }
+\newtext{In a lineup test, the plot of the data is placed randomly among a set of, generally 19, distractor plots (or {\it null plots}) that are rendered from data generated by a model, without a signal (a null model).
+This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different".
+If observers identify the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
+Let $X$ be the number of observers (out of $n$) who identify the data plot from the lineup.
+Under the null hypothesis that the data plot is not different from the other plots, $X$ has approximately a Binomial distribution \cite{wickham2010graphical, majumder2013validation}.
+If $k$ of the observers identify the data plot from the lineup, the probability $P(X \ge k)$ is the $p$-value of the corresponding visual test.
+
+This aspect of lineups provides an important tool in the evaluation of graphical findings.
+However, the testing situation can be reversed, and lineups allow for an objective evaluation of an individual's visual ability based on the general's population performance on a lineup.
+
+}
+
+\comment{Include how an individual's performance is evaluated. }
 

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
Citations for tests

@@
 Our experiment compares lineup performance with established tests, such as the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
-\comment{include citations for the tests}
-The VST measures visual search speed, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning; 
+The VST measures visual search speed\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning\cite{ekstrom1976manual};
 all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
Fix brain fart

@@
 all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
-We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand human of the perception of statistical lineups. 
-\comment {something went wrong with the last half sentence ... do you mean the human component in the statistical lineups?}
+We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
 

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
2014 numbers

@@
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
-This is reasonably representative\footnote{\url{http://www.ir.iastate.edu/PDFfiles/2012-2013\%20Student\%20Profile.pdf}} of the university as a whole; in the fall 2012 semester, 23\% of students were associated with the college of engineering, 23\% were associated with the college of liberal arts and sciences, 23\% were associated with the college of human sciences, 12\% with the business school, and 14\% with the school of agriculture.
-
-\comment{here are the numbers for Fall 2014: http://www.registrar.iastate.edu/enrollment}
+This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the university as a whole; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
 

### 2015-01-09 Paper/Draft1/TestingVisualAptitude.Rnw
no reversal anymore

@@
 This aspect of lineups provides an important tool in the evaluation of graphical findings.
-However, the testing situation can be reversed, and lineups allow for an objective evaluation of an individual's visual ability based on the general's population performance on a lineup.
+Similarly, lineups allow for an objective evaluation of an individual's visual ability based on the general's population performance on a lineup.
 

### 2015-01-11 Paper/Draft1/TestingVisualAptitude.Rnw
Changes to main body of paper- intro, conclusion, moving sample tests to appendix, etc.

@@
 This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-
-% 
-% \comment{`While previous research' - which research are you referring to? 
-% Yifan's paper with the eyetracking study has looked at the spatial connection \cite{zhao2013mind}. But is that what you mean?
-% 
-% `required by the lineup task' - that makes the lineup task strangely active. Could you rephrase that in terms of the participant?
-% }
-
 While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
-% Exploring the participant characteristics which contribute to successful lineup evaluation allows us to better understand the lineup task and potential sources of variability in the study results.
-
-\comment{good discussion, but this is the first time that you mention `our experiment' - include a sentence here, to say what this experiment is about.
-also: I think we should only say `study' or `experiment', and I'd prefer study, because experiment sounds like it could go BOOM.
-I don't want to give that impression.
-}
-
-Our experiment compares lineup performance with established tests, such as the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
+
+\newtext{This study is designed to compare lineup performace with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. }
+We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
 The VST measures visual search speed\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning\cite{ekstrom1976manual};
@@
 
-% 
-% \comment{This is going into the right direction, but a few pieces are still missing, 
-% you need to still say what this paper is doing: we do a study to compare lineup performance with established tests, also go into a bit of discussion of the tests.}
-% \comment{Here is a paragraph missing, that connects the existing literature to what we are doing in this paper. 
-% In particular, we need to answer the questions: 
-% 
-% - why are we interested in evaluating which skills are necessary to evaluate graphical displays (this has been implicitly answered by saying that cognitive load was reduced, so we free up resources and make graphics a better tool for more people.
-% 
-% - what are the limits of the studies that you have cited, ie is there something that we can do better?
-% 
-% - what advantage do lineups bring to the table?
-% }
-% 
-% 
-% \comment{How do perceptual principles apply to statistical graphics (data displays)? - Lineups help us to isolate different parts of displays and allow in-built quantification of results.}
-% 
-% 
-% \comment{Why are lineups variable?
-- Individual variability is big, we want to investigate which abilities individuals help the most in evaluating.
-- for this reason we need to motivate why lineups are useful.
+\newtext{
+In section \ref{sec:methods}, we introduce the tests used in this study and describe how tests are scored.
+In section \ref{sec:results}, we discuss the test scores, comparing established tests with previously validated results, examining demographic characteristics associated with test scores.
+We discuss multicollinearity in the data, and use principal components analysis and linear regression to draw some conclusions about the similarity between lineups and aptitude tests.
+Finally, in section \ref{sec:discussion}, we discuss the implications of this study for the lineup protocol, and future work which could extend these results.
 }
-% 
-% 
-% \comment{We examine 'some demographic characteristics' is a bit lukewarm. I think once the missing paragraph is there, you can sell the rest of the paper a bit more passionately.}
-
-\comment{and in the last paragraph of the intro: lay out the roadmap for the rest of the paper. In section \ref{sec:methods} we introduce ... }
 
@@
 
-\newtext{In a lineup test, the plot of the data is placed randomly among a set of, generally 19, distractor plots (or {\it null plots}) that are rendered from data generated by a model, without a signal (a null model).
+In a lineup test, the plot of the data is placed randomly among a set of, generally 19, distractor plots (or {\it null plots}) that are rendered from data generated by a model, without a signal (a null model).
 This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different".
@@
 
-This aspect of lineups provides an important tool in the evaluation of graphical findings.
-Similarly, lineups allow for an objective evaluation of an individual's visual ability based on the general's population performance on a lineup.
-
+This aspect of lineups provides an important tool in the evaluation of graphical findings, in that we can aggregate responses from many individuals to evaluate a single graphic.
+In addition, however, we can aggregate an individual's score over several lineups to determine how well single individuals perform on the lineup task; data from previous lineup studies suggests that there is significant variation in individual performance as well.
+This study has the potential to explain some of that variance - if visual skills are important contributors, then perhaps we are simply measuring differences in spatial reasoning ability.
+\newtext{
+In order to evaluate an individual's overall lineup score, we would subtract a fraction of a point for each question answered incorrectly, and add a full point for each question answered correctly, as shown in
+\ref{eq.scoring1}.
+
+\begin{eqnarray}\label{eq.scoring1}
+\# \text{correct answers } - 1/(19) \cdot \# \text{wrong answers}.
+\end{eqnarray}
+
+This scoring scheme is chosen so that if participants are guessing, the expected score is 0. 
 }
-
-\comment{Include how an individual's performance is evaluated. }
+% 
+% \comment{Include how an individual's performance is evaluated. }
 
@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
-\begin{figure}[htp]\centering
-\includegraphics[width=\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task (VST).
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
-Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
-\end{figure}
 
@@
 As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
-
-\comment{Let's put figure \ref{fig:tests} into the appendix - that'll save us quite a bit of space. }
- \begin{figure*}[ht]
-  \centering
-   \subfloat[Figure Classification Task.
-Participants classify each figure in the second row as belonging to one of the groups above.
-Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.75\linewidth]{figureclassification}
-}
-\hfill
-    \subfloat[Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.75\linewidth]{cardrotation}
-    }
-    \hfill
-    \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
-Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.75\linewidth]{paperfolding}
-   }
-    \caption{Visuospatial tests}
-    \label{fig:tests}
-  \end{figure*}
 
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
@@
 
-\comment{First: what is shown in figure 8? THis is a non traditional plot, so you need to explain it first.}
-From figure~\ref{fig:lineupblockpca} we see that again, PC1 accounts for general visual aptitude. 
+\newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution each input variable makes to each principal component, weighted by the importance of each PC.
+This allows us to consider the rotation matrix visually; for instance, we see that again, PC1 accounts for most of the variance and seems to represent general visual aptitude.
+}
 % PC2 emphasizes the relationship between figure classification and lineup task 1 (and to a lesser extent, on lineup task 2).
@@
 
-\section{Discussion and Conclusions}
-\comment{These read like the conclusions from earlier - video games aren't that prominently displayed anymore unless you are referring to figure 5.
+\section{Discussion and Conclusions}\label{sec:discussion}
+% \comment{These read like the conclusions from earlier - video games aren't that prominently displayed anymore unless you are referring to figure 5.
 But if so, I would be cautious about summarising STEM major as 'mathematical training' - use the same names for the variables as in figure 5 and make sure to keep the order, i.e.
@@
 }
-
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by factors such as mathematical training, hours of video games played per week, and gender.
-In addition, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
+\newtext{
+Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
+In addition to these demographic factors, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
+}
 Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
@@
 
+\section{Visuospatial Aptitude Tests}
+The Visual Search Task (VST), shown in figure \ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
+\begin{figure}[htp]\centering
+\includegraphics[width=\linewidth]{VisualSearch}
+\caption[Visual Search Task]{Visual Search Task (VST).
+Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
+Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
+\end{figure}
+
+Figure \ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
+All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
+ \begin{figure*}[ht]
+  \centering
+   \subfloat[Figure Classification Task.
+Participants classify each figure in the second row as belonging to one of the groups above.
+Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
+  \includegraphics[width=.75\linewidth]{figureclassification}
+}
+\hfill
+    \subfloat[Card Rotation Task.
+Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
+Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
+\includegraphics[width=.75\linewidth]{cardrotation}
+    }
+    \hfill
+    \subfloat[Paper Folding Task.
+Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
+Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
+\includegraphics[width=.75\linewidth]{paperfolding}
+   }
+    \caption{Visuospatial tests}
+    \label{fig:tests}
+  \end{figure*}
 % \section{Examining potential testing biases}\label{app:ratests}

### 2015-01-11 Paper/Draft1/TestingVisualAptitude.Rnw

@@
 
+\newtext{
 This aspect of lineups provides an important tool in the evaluation of graphical findings, in that we can aggregate responses from many individuals to evaluate a single graphic.
-In addition, however, we can aggregate an individual's score over several lineups to determine how well single individuals perform on the lineup task; data from previous lineup studies suggests that there is significant variation in individual performance as well.
-This study has the potential to explain some of that variance - if visual skills are important contributors, then perhaps we are simply measuring differences in spatial reasoning ability.
-\newtext{
+In addition, however, we can aggregate an individual's score over several lineups to objectively compare individual performance on the lineup task.
+
 In order to evaluate an individual's overall lineup score, we would subtract a fraction of a point for each question answered incorrectly, and add a full point for each question answered correctly, as shown in

### 2015-01-13 Paper/Draft1/TestingVisualAptitude.Rnw
trying to fix the latex error

@@
 Participants are instructed to choose the plot which appears most different from the others.
-In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.
-}\label{fig:lineup}
+In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
 \end{figure}
@@
 
-\newtext{This study is designed to compare lineup performace with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. }
+\comment{`This study' - you still have not established that there is a study. Say it. In this paper we discuss the results of a study .... and now describe what this study tests.}
+\newtext{This study is designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. }
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
@@
 In section \ref{sec:results}, we discuss the test scores, comparing established tests with previously validated results, examining demographic characteristics associated with test scores.
-We discuss multicollinearity in the data, and use principal components analysis and linear regression to draw some conclusions about the similarity between lineups and aptitude tests.
+We discuss multi-collinearity in the data, and use principal components analysis and linear regression to draw some conclusions about the similarity between lineups and aptitude tests.
 Finally, in section \ref{sec:discussion}, we discuss the implications of this study for the lineup protocol, and future work which could extend these results.

### 2015-01-13 Paper/Draft1/TestingVisualAptitude.Rnw
few more comments, but almost there

@@
 
-Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
-In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
-
-\begin{figure}\centering
+\begin{figure}[htp]
+\centering
 \includegraphics[width=\linewidth]{lineup}
@@
 \end{figure}
+
+Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
 
@@
 
-\newtext{
-In section \ref{sec:methods}, we introduce the tests used in this study and describe how tests are scored.
-In section \ref{sec:results}, we discuss the test scores, comparing established tests with previously validated results, examining demographic characteristics associated with test scores.
-We discuss multi-collinearity in the data, and use principal components analysis and linear regression to draw some conclusions about the similarity between lineups and aptitude tests.
-Finally, in section \ref{sec:discussion}, we discuss the implications of this study for the lineup protocol, and future work which could extend these results.
-}
+In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
+In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
+We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
+Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
 
@@
 This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different".
-If observers identify the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
+If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
 Let $X$ be the number of observers (out of $n$) who identify the data plot from the lineup.
@@
 If $k$ of the observers identify the data plot from the lineup, the probability $P(X \ge k)$ is the $p$-value of the corresponding visual test.
-
-\newtext{
-This aspect of lineups provides an important tool in the evaluation of graphical findings, in that we can aggregate responses from many individuals to evaluate a single graphic.
-In addition, however, we can aggregate an individual's score over several lineups to objectively compare individual performance on the lineup task.
-
-In order to evaluate an individual's overall lineup score, we would subtract a fraction of a point for each question answered incorrectly, and add a full point for each question answered correctly, as shown in
-\ref{eq.scoring1}.
-
+%
+By aggregating responses from different individuals the lineup protocol therefore allows an objective evaluation of a graphical finding.
+
+Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task.
+
+For this approach, we derive a score for an individual as follows:
+
+Assume that an observer has evaluated $K$ lineups of size $m$ (consisting of $m-1$ decoy plots and 1 target), and successfully identified the target in $n_c$ of these plots, while missing the target in $n_w$ of them.
+The score for this individual is then given as:
 \begin{eqnarray}\label{eq.scoring1}
-\# \text{correct answers } - 1/(19) \cdot \# \text{wrong answers}.
+n_c - n_w/(m-1).
 \end{eqnarray}
-
-This scoring scheme is chosen so that if participants are guessing, the expected score is 0. 
-}
+Note that the sum of answers, $n_c + n_w$, is at most $K$, but could be less, if an observer chooses to not answer one of the lineup tasks or runs out of time.
+
+The scoring scheme as given in (\ref{eq.scoring1}) is chosen so that if participants are guessing, the expected score is 0.
+
 % 
@@
 
+\comment{add something about x and y axis in the description of the figure. }
 \newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution each input variable makes to each principal component, weighted by the importance of each PC.
@@
 par(mfrow=c(1,2), mar=c(4,4.5,.1,.1), tcl=-0.01, cex.lab=.8,cex.axis=.7,mgp=c(2,.7,0),tcl=-.3)
-biplot(pctests)
-biplot(pctests, 3:4)
+#biplot(pctests)
+#biplot(pctests, 3:4)
 @
@@
 }
-\newtext{
+
 Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
 In addition to these demographic factors, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-}
+
+\comment{reference is gone astray `those who score' ... the previous subject were facets of intelligence, and they do not score on tests - what you mean are people, so say that!}
+
 Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
@@
 
+\newpage
 \appendices

### 2015-01-19 Paper/Draft1/TestingVisualAptitude.Rnw
Fix abstract

@@
 \begin{abstract}
-\comment{Tone down the `statistical' part ...
-statistical, to the wrong ears sounds complicated.
-} Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
-\newtext{}
+Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
 This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.

### 2015-01-19 Paper/Draft1/TestingVisualAptitude.Rnw
Address "this study"

@@
 
-\comment{`This study' - you still have not established that there is a study. Say it. In this paper we discuss the results of a study .... and now describe what this study tests.}
-\newtext{This study is designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. }
+% \comment{`This study' - you still have not established that there is a study. Say it. In this paper we discuss the results of a study .... and now describe what this study tests.}
+\newtext{In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups.
+}
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 

### 2015-01-19 Paper/Draft1/TestingVisualAptitude.Rnw
PC discussion
Demographic factors re-phrasing

@@
 
-\comment{add something about x and y axis in the description of the figure. }
+% \comment{add something about x and y axis in the description of the figure. }
 \newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution each input variable makes to each principal component, weighted by the importance of each PC.
+The input variables are shown on the $y$ axis, with the importance to each PC shown on the $x$ axis.
 This allows us to consider the rotation matrix visually; for instance, we see that again, PC1 accounts for most of the variance and seems to represent general visual aptitude.
@@
 Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-In addition to these demographic factors, many facets of intelligence are highly correlated; those who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-
-\comment{reference is gone astray `those who score' ... the previous subject were facets of intelligence, and they do not score on tests - what you mean are people, so say that!}
+In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
+
+% \comment{reference is gone astray `those who score' ... the previous subject were facets of intelligence, and they do not score on tests - what you mean are people, so say that!}
 

### 2015-01-19 Paper/Draft1/TestingVisualAptitude.Rnw
Hide comment and change table caption

@@
 \section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-\comment{'Results' of lineup score - te caption in the table is too generic. give more details. }
+% \comment{'Results' of lineup score - the caption in the table is too generic. give more details. }
 Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
@@
 tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
-print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Results of lineup score modeled by single demographic variables.
+print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Participant demographics' impact on lineup score.
+The table below shows each single demographic variable's association with lineup score.
+STEM major, completion of Calculus I, time spent playing video games, and gender all show some association with score on statistical lineups.
 \\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")

### 2015-01-19 Paper/Draft1/TestingVisualAptitude.Rnw
PCA discussion

@@
 \section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
-\subsection{PCA of the Four Cognitive Tests}\label{app:pca4}
-\comment{'Importance of principal components {\it in an analysis} of four tests' -- be more precise: you mean the PCA. Give some details as to how importance is measured. }
+\subsection{Importance of principal components in an analysis of the four cognitive tests}\label{app:pca4}
+% \comment{'Importance of principal components {\it in an analysis} of four tests' -- be more precise: you mean the PCA. Give some details as to how importance is measured. }
 %
@@
 @
-Table~\ref{tab:PCAvariance4} contains the importance of each resultant PC and the proportion of the variance each PC represents.
-
+Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
 PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
@@
 <<pcarotation4,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="PCA Rotation matrix for the four cognitive tests.\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
-
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Plots of principal components 1-4 with observations.
-PCA was performed on the four cognitive tests used to understand the cognitive demands of the lineup protocol.
+print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+@
+
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
+Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
 \\label{fig:biplots4}", echo=F, fig.env="figure">>=
@@
 @
-\comment{figure 9 and 10 are Biplots, say so in the caption - in the discussion of the figures and the tables in the text, please think again of adding the sentence summarising what is shown. }
+% \comment{figure 9 and 10 are Biplots, say so in the caption - in the discussion of the figures and the tables in the text, please think again of adding the sentence summarising what is shown. }
+\newtext{Figure \ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
+PC2 begins to separate the four cognitive tests; PC3 and PC4 also differentiate the four tests.
+According to table \ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.}
+
 \subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}

### 2015-01-23 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments

@@
 
-% \comment{`This study' - you still have not established that there is a study. Say it. In this paper we discuss the results of a study .... and now describe what this study tests.}
-\newtext{In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups.
-}
+In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
@@
 A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
+
+\comment{You'll have to define the use of PCX still - a quick sentence on the order of the principal components being given by the amount of variance in the data that they explain should do it.}
+
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
@@
 the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
-Biplots of the remaining components are provided in Appendix \ref{app:pca5}.
+Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix \ref{app:pca5}.
 
@@
 % \comment{add something about x and y axis in the description of the figure. }
-\newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution each input variable makes to each principal component, weighted by the importance of each PC.
+\comment{Let's give this plots a name ...
+how about {\it PC importance plots}?
+Get rid of the * in the x axis title.
+Both `realtive contribution' and `importance' are a bit fuzzy terms ...
+I'm not sure what you are doing until I'm looking at the code.
+Define importance and each of the items, and yes, that might need a bit of math.}
+\newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution of each input variable  to each principal component (PC), weighted by the importance of each principal component.
 The input variables are shown on the $y$ axis, with the importance to each PC shown on the $x$ axis.
-This allows us to consider the rotation matrix visually; for instance, we see that again, PC1 accounts for most of the variance and seems to represent general visual aptitude.
+This allows us to consider the rotation matrix visually; for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
 }
@@
 @
-% \comment{figure 9 and 10 are Biplots, say so in the caption - in the discussion of the figures and the tables in the text, please think again of adding the sentence summarising what is shown. }
-\newtext{Figure \ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-PC2 begins to separate the four cognitive tests; PC3 and PC4 also differentiate the four tests.
-According to table \ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.}
+Figure \ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
+All of the following principal components separate between the cognitive tasks.
+\comment{How do the PCs exactly separate? Quick sentence for each.}
+
+According to table \ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
 

### 2015-01-23 Paper/Draft1/TestingVisualAptitude.Rnw
consistent referencing: all figures, tables and appendix references are now with upper case

@@
 
-Figure \ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+Figure~\ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
 In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
@@
 
-The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
+The \textbf{visual searching task} (VST), shown in Figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
 Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
@@
 This task is associated with inductive reasoning ablities (factor I  in~\cite{ekstrom1976manual}).
-An example is shown in figure~\ref{fig:figureclassification}.
+An example is shown in Figure~\ref{fig:figureclassification}.
 
@@
 It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in figure~\ref{fig:cardrotation}.
+An example is shown in Figure~\ref{fig:cardrotation}.
 The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-A sample question from the test is shown in figure~\ref{fig:paperfolding}.
+A sample question from the test is shown in Figure~\ref{fig:paperfolding}.
 It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
@@
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
-We compare Iowa State students' unscaled scores in table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
+We compare Iowa State students' unscaled scores in Table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
 
@@
 These demographic characteristics were chosen to account for life experience and personal skills which may have influenced the results.
-Statistical test results are available in appendix~\ref{app:categoricalresults}.
+Statistical test results are available in Appendix~\ref{app:categoricalresults}.
 
@@
 
-Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
+Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
 What is of more interest to us is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
@@
 PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
-More detailed results from the 4-test analysis are provided in Appendix \ref{app:pca}.
+More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
 
@@
 the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
-Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix \ref{app:pca5}.
+Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix~\ref{app:pca5}.
 
@@
 We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
-The result is shown in table~\ref{tab:m1}.
+The result is shown in Table~\ref{tab:m1}.
 Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores.
@@
 Each of the three sets of 20 lineups was taken from previous studies on different designs to investigate which plot type most effectively conveyed important characteristics of the underlying data set.
-Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix \ref{app:lineuptypes}.
+Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
 
@@
 Define importance and each of the items, and yes, that might need a bit of math.}
-\newtext{Figure \ref{fig:lineupblockpca} shows the relative contribution of each input variable  to each principal component (PC), weighted by the importance of each principal component.
+\newtext{Figure~\ref{fig:lineupblockpca} shows the relative contribution of each input variable  to each principal component (PC), weighted by the importance of each principal component.
 The input variables are shown on the $y$ axis, with the importance to each PC shown on the $x$ axis.
@@
 A larger study might not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
-Preliminary results of performance on different types of plots are shown in appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
+Preliminary results of performance on different types of plots are shown in Appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
 The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
@@
 \section{Visuospatial Aptitude Tests}
-The Visual Search Task (VST), shown in figure \ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
+The Visual Search Task (VST), shown in Figure~\ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
 \begin{figure}[htp]\centering
@@
 
-Figure \ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
+Figure~\ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
 All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
@@
 \section{Scaling Scores}\label{app:ScoreAdj}
-To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
+To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in Table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
 The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
@@
 Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
-PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; figure~\ref{fig:biplots4} and table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
 PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
@@
 @
-Figure \ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
+Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
 All of the following principal components separate between the cognitive tasks.
@@
 
-According to table \ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
+According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
 
@@
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-Biplots of the remaining principal components are shown in figure~\ref{fig:biplots5}.
+Biplots of the remaining principal components are shown in Figure~\ref{fig:biplots5}.
 
@@
 
-Figure \ref{fig:plottypesdensity} compares performance on each different type of plot.
+Figure~\ref{fig:plottypesdensity} compares performance on each different type of plot.
 As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs.
@@
 
-Figure \ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
+Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
 Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks.
@@
 For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
-Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in figure \ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
+Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in Figure~\ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
 

### 2015-01-27 Paper/Draft1/TestingVisualAptitude.Rnw
Principal component explanation

@@
 A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
-
-\comment{You'll have to define the use of PCX still - a quick sentence on the order of the principal components being given by the amount of variance in the data that they explain should do it.}
+Principal components (PC-) are ordered by importance (how much variability in the data they contain), and each principal component is uncorrelated with every other principal component.
 

### 2015-01-27 Paper/Draft1/TestingVisualAptitude.Rnw
Influence plot explanation

@@
 
-<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Lineup block PCA rotation matrix in graphical form.", fig.width=5, fig.height=5>>=
+\newtext{Principal component analysis separates multivariate data into orthogonal components using a rotation matrix, which contains information on the contribution of each input variable to the principal component.
+The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).}
+
+<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component importance plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 library(grid) # needed for arrow function
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=3) + xlab("Contribution to each PC * % Variance in PC") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=3) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
 @
@@
 % \comment{add something about x and y axis in the description of the figure. }
-\comment{Let's give this plots a name ...
+% \comment{Let's give this plots a name ...
 how about {\it PC importance plots}?
@@
 Define importance and each of the items, and yes, that might need a bit of math.}
-\newtext{Figure~\ref{fig:lineupblockpca} shows the relative contribution of each input variable  to each principal component (PC), weighted by the importance of each principal component.
-The input variables are shown on the $y$ axis, with the importance to each PC shown on the $x$ axis.
+Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component; the arrows are scaled in proportion to the importance of each principal component.
+The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
 This allows us to consider the rotation matrix visually; for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
-}
 % PC2 emphasizes the relationship between figure classification and lineup task 1 (and to a lesser extent, on lineup task 2).

### 2015-01-27 Paper/Draft1/TestingVisualAptitude.Rnw
Explanation of biplots

@@
 Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the following principal components separate between the cognitive tasks.
-\comment{How do the PCs exactly separate? Quick sentence for each.}
-
+All of the remaining principal components distinguish between the cognitive tasks;
+\newtext{PC2 and PC3 sseparate the paper folding, visual search, and figure classification tasks, while PC4 and PC5 separate lineups, card rotation, figure classification, and visual search.
+This separation allows us to compare the tasks which are similar from among the principal components.
+}
 According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.

### 2015-01-28 Paper/Draft1/TestingVisualAptitude.Rnw
Remove extraneous comment
still figure 6

--- 2015-01-27 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-\newtext{Principal component analysis separates multivariate data into orthogonal components using a rotation matrix, which contains information on the contribution of each input variable to the principal component.
-The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).}
-
+\comment{Susan, sorry, the description of figure 6 is still not working.
+You write impact at the x axis - so I'm wondering what a negative impact is supposed to be?
+You are mentioning contribution and importance in the paragraph below, but you don't really define either one of them.
+Be careful about mentioning Table 2 - that table doesn't have anything to do with this PCA - it might be better to include the two tables for this PCA in the appendix and work out there in detail, how the scaling is done.
+InfoVis does know about signular value composition, so you could start there ...
+}
+
+\newtext{
+Principal component analysis separates multivariate data into orthogonal components using a rotation matrix, which contains information on the contribution of each input variable to the principal component.
+The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
+}
 <<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component importance plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
@@
 & = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
-\sigma_{\text{all}} &= \sqrt{\left(323(27.8)^2 \!+\! 294(30)^2\right)/(323\!+\!294)} \\
+\sigma_{\text{all}} &= \sqrt{\left(323 \cdot 27.8^2 \!+\! 294 \cdot 30^2\right)/(323\!+\!294)} \\
 & = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
@@
 Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the following principal components separate between the cognitive tasks.
-\comment{How do the PCs exactly separate? Quick sentence for each.}
-
+All of the remaining principal components distinguish between the cognitive tasks;
+PC2 and PC3 separate  paper folding from visual search and from the lineup and figure classification tasks, while PC4 and PC5 mainly separate lineups from card rotation and figure classification.
+This separation allows us to compare the tasks which are similar from among the principal components.
 According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
@@
 
-\comment{Describe performance - need to get data from original experiment.}
+% \comment{Describe performance - need to get data from original experiment.}
 \begin{figure}[htbp]

### 2015-02-01 Paper/Draft1/TestingVisualAptitude.Rnw
last? couple of comments

@@
 % and "HIS" in caps to complete the first word.
+
+\comment{Susan, there's only a few things left to do, mostly to do with explanations for figures.
+x-axis, y-axis, ...
+Have a look at fig 6 again.
+It is a great chart, but it needs some explanation because it is a non-standard chart that you thought of.
+
+Fig 22 and 23 also need a bit more explanation.}
+
 \IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
@@
 \mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
-\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)},
+\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)}.
 \end{align}
-where $\mu_F$ and $\mu_M$ are the mean of female and male scores respectively, $N_F$ and $N_M$ are the number of participants in each group, and $\sigma_F^2$ and $\sigma_M^2$ are the variance of each group.
+Here $\mu_F$ and $\mu_M$ are the mean scores for females and males, respectively; $N_F$ and $N_M$ are the number of female and male participants, and $\sigma_F^2$ and $\sigma_M^2$ are the variances in scores for females and males.
+
 Substituting in the provided numbers, we get
@@
 
-<<lineup-types,echo=F, fig.cap="Plot of scaled scores for different types of lineups\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+\comment{For all of these figures: include first, what you are showing: x axis, y axis, ... before you go into an interpretation of the figure.}
+<<lineup-types,echo=F, fig.cap="Density plots of scaled scores for different types of lineups.
+XXX Main message?\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
 
-<<lineup-types-scores,echo=F, fig.cap="Plot of scaled lineup scores by aptitude test scores.\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
+<<lineup-types-scores,echo=F, fig.cap="Scatterplots of scaled lineup scores by aptitude test scores.
+XXX main message?\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
 lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)

### 2015-02-02 Paper/Draft1/TestingVisualAptitude.Rnw
Figure captions and influence plot explanation

@@
 
-\comment{Susan, sorry, the description of figure 6 is still not working.
+% 
+% \comment{Susan, sorry, the description of figure 6 is still not working.
 You write impact at the x axis - so I'm wondering what a negative impact is supposed to be?
 You are mentioning contribution and importance in the paragraph below, but you don't really define either one of them.
-Be careful about mentioning Table 2 - that table doesn't have anything to do with this PCA - it might be better to include the two tables for this PCA in the appendix and work out there in detail, how the scaling is done.
+% Be careful about mentioning Table 2 - that table doesn't have anything to do with this PCA - it might be better to include the two tables for this PCA in the appendix and work out there in detail, how the scaling is done.
 InfoVis does know about signular value composition, so you could start there ...
@@
 \newtext{
-Principal component analysis separates multivariate data into orthogonal components using a rotation matrix, which contains information on the contribution of each input variable to the principal component.
+Principal component analysis separates multivariate data into orthogonal components using a rotation matrix to transform correlated input data into an orthogonal space.
 The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
+For input variables $X_i$, $i=1, ..., K$ and output variables $Y_j$, $j=1, ..., K$, the rotation matrix, $M$ would have dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ would have dimension $1\times K$ (indexed by $j$).
+For each $X_i$ and $Y_j$, we could calculate the influence of variable $X_i$ on component $Y_j$ as follows:
+
+$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
+
+The influence is a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
+
+Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
+The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
+This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
 }
-<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component importance plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
+<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 Define importance and each of the items, and yes, that might need a bit of math.}
-Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component; the arrows are scaled in proportion to the importance of each principal component.
-The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
-This allows us to consider the rotation matrix visually; for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
+
 % PC2 emphasizes the relationship between figure classification and lineup task 1 (and to a lesser extent, on lineup task 2).
@@
 
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Plots of principal components 2-5 with observations.
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
 \\label{fig:biplots5}", echo=F, fig.env="figure">>=
@@
 <<lineup-types,echo=F, fig.cap="Density plots of scaled scores for different types of lineups.
-XXX Main message?\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+For the same experiment (shown by line color), certain types of plots are more difficult to read and are associated with lower participant scores.
+\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
 Figure~\ref{fig:plottypesdensity} compares performance on each different type of plot.
+The $x$ axis shows scaled score, the $y$ axis shows the density of participant scores.
 As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs.
@@
 <<lineup-types-scores,echo=F, fig.cap="Scatterplots of scaled lineup scores by aptitude test scores.
-XXX main message?\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
+There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ plots with confidence bands may require more skill at mental rotation than QQ plots without the bands.
+\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
 lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)

### 2015-02-03 Paper/Draft1/TestingVisualAptitude.Rnw
comment on figure 6 description
Fix would/could, and change PC_j notation

@@
 Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
-Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
+Both of these theories suggest spatial ability impacts a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
 In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
@@
 \end{eqnarray}
-Note that the sum of answers, $n_c + n_w$, is at most $K$, but could be less, if an observer chooses to not answer one of the lineup tasks or runs out of time.
+Note that the sum of answers, $n_c + n_w$, is at most $K$, but may be less, if an observer chooses to not answer one of the lineup tasks or runs out of time.
 
@@
 
+% \comment{PC$_j$ needs to be defined - I believe that you started out with the idea of calling it $Y_j$.
+The woulds and coulds need to go.
+Use the word 'define' so that it is absolutely clear that  it's not something you hypothetically 'could calculate'.}
 \newtext{
@@
 The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
-For input variables $X_i$, $i=1, ..., K$ and output variables $Y_j$, $j=1, ..., K$, the rotation matrix, $M$ would have dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ would have dimension $1\times K$ (indexed by $j$).
-For each $X_i$ and $Y_j$, we could calculate the influence of variable $X_i$ on component $Y_j$ as follows:
+For input variables $X_i$, $i=1, ..., K$ and output components $PC_j$, $j=1, ..., K$, the rotation matrix, $M$ has dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ has dimension $1\times K$ (indexed by $j$).
+For each $X_i$ and $PC_j$, we define the influence of variable $X_i$ on component $PC_j$ as follows:
 
@@
 
-A larger study might not only address the question of which lineup tasks require certain visual skills, but could also address the use of different types of plots from a perceptual perspective.
+A larger study might not only address the question of which lineup tasks require certain visual skills, but also the use of different types of plots from a perceptual perspective.
 Preliminary results of performance on different types of plots are shown in Appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
@@
 Such studies often suffer from participants who do not take the task seriously, and internal verification questions, as well as pre-qualification tasks are often used to reduce extraneous variability.
-While it would be impractical to require participants to score well on several different tests, it would be reasonable to ask participants to pre-qualify for a task by completing a figure classification test.
-As the figure classification test is different from the lineup task, this would not bias participants' scores on the domain of interest, but would ensure a participant pool that is sufficiently motivated to complete the lineup questions.
+While it is impractical to require participants to score well on several different tests, it is reasonable to ask participants to pre-qualify for a task by completing a figure classification test.
+As the figure classification test is different from the lineup task, this will not bias participants' scores on the domain of interest while ensuring that the participant pool is sufficiently motivated to complete the lineup questions.
 
@@
 As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
-To better understand the effects of these variables, a larger study would be required.
+To better understand the effects of these variables, a larger study is necessary.
 <<ttests.categ, echo=F, eval=T, results='asis'>>=
@@
 The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated qq plots showed lower performance because participants could more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants could more accurately compare the residuals to the reference bands.
+Rotated qq plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
 % \comment{Add citations, more explanations, etc.}

### 2015-02-03 Paper/Draft1/TestingVisualAptitude.Rnw
Remove old (commented out) comments

@@
 
-\comment{Susan, there's only a few things left to do, mostly to do with explanations for figures.
-x-axis, y-axis, ...
-Have a look at fig 6 again.
-It is a great chart, but it needs some explanation because it is a non-standard chart that you thought of.
-
-Fig 22 and 23 also need a bit more explanation.}
-
 \IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
@@
 
-% \comment{I am not sure what the kosara reference has to do with the lineups - the only thing lineups and kosara have in common are the amazon studies.
-but you might want to keep the reference around for a discussion of the amazon results }
 ``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
@@
 The scoring scheme as given in (\ref{eq.scoring1}) is chosen so that if participants are guessing, the expected score is 0.
-
-% 
-% \comment{Include how an individual's performance is evaluated. }
 
@@
 Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.
-% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class.
 Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
 Only one student is doing research who has not taken a statistics course.
-%\comment{The honors program puts freshman in research labs in their first year; Ryan worked in a corn biology lab as a physics/engineering major.
-If they are LBAR, they may still be placed in science labs, or they could be doing LBAR research - may not require stats.}
+
 A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
@@
 
-% 
-% \comment{Susan, sorry, the description of figure 6 is still not working.
-You write impact at the x axis - so I'm wondering what a negative impact is supposed to be?
-You are mentioning contribution and importance in the paragraph below, but you don't really define either one of them.
-% Be careful about mentioning Table 2 - that table doesn't have anything to do with this PCA - it might be better to include the two tables for this PCA in the appendix and work out there in detail, how the scaling is done.
-InfoVis does know about signular value composition, so you could start there ...
-}
-
-% \comment{PC$_j$ needs to be defined - I believe that you started out with the idea of calling it $Y_j$.
-The woulds and coulds need to go.
-Use the word 'define' so that it is absolutely clear that  it's not something you hypothetically 'could calculate'.}
-\newtext{
 Principal component analysis separates multivariate data into orthogonal components using a rotation matrix to transform correlated input data into an orthogonal space.
 The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
-For input variables $X_i$, $i=1, ..., K$ and output components $PC_j$, $j=1, ..., K$, the rotation matrix, $M$ has dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ has dimension $1\times K$ (indexed by $j$).
+
+\newtext{For input variables $X_i$, $i=1, ..., K$ and output components $PC_j$, $j=1, ..., K$, the rotation matrix, $M$ has dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ has dimension $1\times K$ (indexed by $j$).
 For each $X_i$ and $PC_j$, we define the influence of variable $X_i$ on component $PC_j$ as follows:
+}
 
@@
 This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
-}
 <<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
@@
 
-% \comment{add something about x and y axis in the description of the figure. }
-% \comment{Let's give this plots a name ...
-how about {\it PC importance plots}?
-Get rid of the * in the x axis title.
-Both `realtive contribution' and `importance' are a bit fuzzy terms ...
-I'm not sure what you are doing until I'm looking at the code.
-Define importance and each of the items, and yes, that might need a bit of math.}
-
-% PC2 emphasizes the relationship between figure classification and lineup task 1 (and to a lesser extent, on lineup task 2).
-PC3 differentiates the lineup task 3 from the visual search task; other tasks do not impact PC3 to a meaningful degree.
-PC4 separates lineup test 2 from the paper folding and card rotation tests and emphasizes the overlap with the visual search test.
-PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test.
-PC6 and PC7 account for a joint total of less than 10\% of the overall variability, and do not display any distinct patterns in the loadings.
-%
 PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
@@
 PC6 and PC7 jointly account for less than 10\% of the variance in the data and do not display any distinct patterns in the loadings.
-%\Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
 
@@
 
-%\paragraph*{Removing the effect of overall aptitude}
 <<removeAptitude,echo=F,fig.width=8, fig.height=4>>=
@@
 \section{Discussion and Conclusions}\label{sec:discussion}
-% \comment{These read like the conclusions from earlier - video games aren't that prominently displayed anymore unless you are referring to figure 5.
-But if so, I would be cautious about summarising STEM major as 'mathematical training' - use the same names for the variables as in figure 5 and make sure to keep the order, i.e.
-gender is more impotant than video game hours.
-also make sure, that to mention that this is on a univariate level and doesn't take the correlation between the variables into account.
-}
 
@@
 In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-
-% \comment{reference is gone astray `those who score' ... the previous subject were facets of intelligence, and they do not score on tests - what you mean are people, so say that!}
 
@@
 \section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-% \comment{'Results' of lineup score - the caption in the table is too generic. give more details. }
 Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
@@
 \subsection{Importance of principal components in an analysis of the four cognitive tests}\label{app:pca4}
-% \comment{'Importance of principal components {\it in an analysis} of four tests' -- be more precise: you mean the PCA. Give some details as to how importance is measured. }
-%
+
 <<pcasummary4,echo=F, results='asis'>>=
@@
 
-% \comment{Describe performance - need to get data from original experiment.}
 \begin{figure}[htbp]
@@
 As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
-% \comment{Add citations, more explanations, etc.}
 
@@
 
-\comment{For all of these figures: include first, what you are showing: x axis, y axis, ... before you go into an interpretation of the figure.}
 <<lineup-types,echo=F, fig.cap="Density plots of scaled scores for different types of lineups.

### 2015-02-06 Paper/Draft1/TestingVisualAptitude.Rnw
discussion of fig 6

@@
 
-\newtext{For input variables $X_i$, $i=1, ..., K$ and output components $PC_j$, $j=1, ..., K$, the rotation matrix, $M$ has dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ has dimension $1\times K$ (indexed by $j$).
-For each $X_i$ and $PC_j$, we define the influence of variable $X_i$ on component $PC_j$ as follows:
+\newtext{For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
+
+The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
+
+The influence of variable $X_i$ on component $PC_j$ is then defined as:
 }
@@
 
-The influence is a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
+\comment{Sign of influence? Influence is the smallest, if it is zero. The sign is kept to show separation between variables. You still need to comment on that.}
+Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
 

### 2015-02-08 Paper/Draft1/TestingVisualAptitude.Rnw
Sign of influence

--- 2015-02-03 Paper/Draft1/TestingVisualAptitude.Rnw
@@
 
-\newtext{For input variables $X_i$, $i=1, ..., K$ and output components $PC_j$, $j=1, ..., K$, the rotation matrix, $M$ has dimension $K\times K$ (indexed by $i$ and $j$ respectively), and the importance vector $I$ has dimension $1\times K$ (indexed by $j$).
-For each $X_i$ and $PC_j$, we define the influence of variable $X_i$ on component $PC_j$ as follows:
+For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
+
+The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
+
+The influence of variable $X_i$ on component $PC_j$ is then defined as:
+
+$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
+
+Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
+\newtext{Low values indicate that there is little influence of $X_i$ on $PC_j$; negative values show the direction of the influence (to maintain the separation of variables as part of principal components analysis).
 }
-
-$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
-
-The influence is a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
 

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
ready to go?

@@
 Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}. 
-
-Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
-The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
-Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
-
 <<LineupTypeSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
@@
 
+Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
+The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
+Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
+Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
+
 In order to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of an experiment, we employ principal component analysis on participant scores averaged across each block of lineups.
@@
 
-For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
-
-The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
-
-The influence of variable $X_i$ on component $PC_j$ is then defined as:
-
-$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
-
-Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
-\newtext{Low values indicate that there is little influence of $X_i$ on $PC_j$; negative values show the direction of the influence (to maintain the separation of variables as part of principal components analysis).
-}
-
-Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
-The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
-This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
 <<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
@@
 @
+
+For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
+
+The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
+
+The influence of variable $X_i$ on component $PC_j$ is then defined as:
+%
+$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
+%
+Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
+Small (absolute) values indicate that there is little influence of $X_i$ on $PC_j$; negative values show the direction of the influence (to maintain the separation of variables as part of principal components analysis).
+
+Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
+The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
+This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
 
@@
 The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated qq plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
 As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
@@
 \includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
-\caption{QQ Plot with guide line.}
+\caption{QQ-plot with guide line.}
 \end{figure}
@@
 \includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
-\caption{QQ Plot with acceptance bands.}
+\caption{QQ-plot with acceptance bands.}
 \end{figure}
@@
 \includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
-\caption{QQ Plot rotated 45 degrees.}
+\caption{QQ-plot rotated 45 degrees.}
 \end{figure}
@@
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
-                                              "Rotated QQ plot +\nAcceptance Band")
+                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
+                                              "Rotated QQ-plot +\nAcceptance Band")
                                      )
@@
 Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
-Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ plots tasks compared to the other tasks.
-In addition, rotated qq plots seem to be much more asssociated with the paper folding task scores than other qq plot tasks; this may be because they require more visual manipulation than other qq plots.
+Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
+In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
 
 <<lineup-types-scores,echo=F, fig.cap="Scatterplots of scaled lineup scores by aptitude test scores.
-There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ plots with confidence bands may require more skill at mental rotation than QQ plots without the bands.
+There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ-plots with confidence bands may require more skill at mental rotation than QQ-plots without the bands.
 \\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
@@
                                               "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ plot + Line", "QQ plot +\nAcceptance Band", 
-                                              "Rotated QQ plot +\nAcceptance Band"), 
+                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
+                                              "Rotated QQ-plot +\nAcceptance Band"), 
                                      labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
                                               "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
-                                              "QQ plot\n+ Line", "QQ plot\n+ Bands", 
+                                              "QQ-plot\n+ Line", "QQ-plot\n+ Bands", 
                                               "Rotated QQ\nplot + Bands")

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
Small edits to reduce paper size a bit.

@@
 \centering
-\includegraphics[width=\linewidth]{lineup}
+\includegraphics[width=\linewidth,trim=0cm .5cm 0cm .25cm]{lineup}
 \caption[Sample Lineup]{Sample lineup of boxplots.
@@
 \begin{figure}[ht]
-<<ResultsScaledScores, echo=F, include=T, fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=
+<<ResultsScaledScores, echo=F, include=T, fig.width=4.5, fig.height=4, out.width='.9\\linewidth', warning=F, message=F>>=
 
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=6, message=F, warning=F, fig.pos='ht'>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5.5, message=F, warning=F, fig.pos='ht'>>=
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
 \begin{figure*}[h!tb]\centering
-\includegraphics[width=.9\linewidth]{fig-VisReasoningCategorical-1}
+\includegraphics[width=.85\linewidth]{fig-VisReasoningCategorical-1}
 \caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
@@
 
-<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=5, fig.height=5>>=
+<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=6, fig.height=4>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 library(grid) # needed for arrow function
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=3) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
+qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
 @

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
Use latex figure environment explicitly... hopefully draft proof will show text in pictures this time?

@@
 
-<<VisReasoningSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
-Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
+\begin{figure}[ht]
+<<VisReasoningSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
 data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
@@
 Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}. 
-<<LineupTypeSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
-All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
-\\label{fig:lineuptaskmatrix}">>=
+\begin{figure}[ht]
+<<LineupTypeSPM,echo=FALSE,include=T, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
 data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
@@
 @
-
+\caption{Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
+All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
+\label{fig:lineuptaskmatrix}}
+\end{figure}
 Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
@@
 
-<<lineupblockpca, echo=F, warning=F, message=F, fig.cap="Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.", fig.width=6, fig.height=4>>=
+\begin{figure}[ht]
+<<lineupblockpca, echo=F, warning=F, message=F, fig.width=6, fig.height=4>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 @
+\caption{Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.\label{fig:influence}}
+\end{figure}
 
@@
 
-Figure~\ref{fig:lineupblockpca} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
+Figure~\ref{fig:influence} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
 The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
@@
 
-<<lineup-types,echo=F, fig.cap="Density plots of scaled scores for different types of lineups.
-For the same experiment (shown by line color), certain types of plots are more difficult to read and are associated with lower participant scores.
-\\label{fig:plottypesdensity}", fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+\begin{figure}[ht]
+<<lineup-types,echo=F, fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
 @
+\caption{Density plots of scaled scores for different types of lineups.
+For the same experiment (shown by line color), certain types of plots are more difficult to read and are associated with lower participant scores.
+\label{fig:plottypesdensity}}
+\end{figure}
 
@@
 
-<<lineup-types-scores,echo=F, fig.cap="Scatterplots of scaled lineup scores by aptitude test scores.
-There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ-plots with confidence bands may require more skill at mental rotation than QQ-plots without the bands.
-\\label{fig:scatterplottypes}", fig.width=6, fig.height=12, out.width="\\linewidth", fig.env="figure", warning=F, message=F>>=
+\begin{figure}[ht]
+
+<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, out.width="\\linewidth", warning=F, message=F>>=
 lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
@@
 @
+\caption{Scatterplots of scaled lineup scores by aptitude test scores.
+There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ-plots with confidence bands may require more skill at mental rotation than QQ-plots without the bands.
+\label{fig:scatterplottypes}}
+\end{figure}
 

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
embed ALL the fonts!

@@
 wd <- getwd()
-
+library(extrafont)
 library(knitr)
@@
 
-\author{Susan VanderPlas, Heike Hofmann} % , Stephanie De Graaf, Hillary Cheney ??
+\author{Susan VanderPlas, Heike Hofmann}
 
@@
 <<ResultsScaledScores, echo=F, include=T, fig.width=4.5, fig.height=4, out.width='.9\\linewidth', warning=F, message=F>>=
-
+loadfonts()
 score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
@@
   #coord_flip()
+embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
 @
@@
 <<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5.5, message=F, warning=F, fig.pos='ht'>>=
+loadfonts()
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
   theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
+embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
 @
@@
 }
-
+loadfonts()
 require(grid)
@@
 rm(list=c("data", "all", "densities"))
+embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
 @
@@
 )
-
+loadfonts()
 ggplot()+
@@
   theme(axis.title=element_blank())
+embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
 @
@@
 library(grid) # needed for arrow function
+loadfonts()
 qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
+embedFonts("./Figure/fig-lineupblockpca-1.pdf")
 @
@@
                                      )
+loadfonts()
 qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
@@
         axis.ticks.y=element_blank(), axis.text.y=element_blank())
+embedFonts("./Figure/fig-lineup-types-1.pdf")
 @
@@
 cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-
+loadfonts()
 qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
@@
         plot.margin=unit(c(0,0,0,0),"mm"))
+embedFonts("./Figure/fig-lineup-types-scores-1.pdf")
 @

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
external figure declarations

@@
 \centering
-\includegraphics[width=\linewidth,trim=0cm .5cm 0cm .25cm]{lineup}
+\includegraphics[width=\linewidth]{lineup}
 \caption[Sample Lineup]{Sample lineup of boxplots.
@@
 In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.
-
-\begin{figure}[ht]
-<<ResultsScaledScores, echo=F, include=T, fig.width=4.5, fig.height=4, out.width='.9\\linewidth', warning=F, message=F>>=
+<<ResultsScaledScores, echo=F, include=F, fig.width=4.5, fig.height=4, warning=F, message=F>>=
 loadfonts()
@@
 @
+
+\begin{figure}[ht]
+\includegraphics[width=.9\linewidth]{Figure/fig-ResultsScaledScores-1}
 \caption{Test scores for lineups and visuospatial tests.
@@
 
-\begin{figure}[ht]
-<<VisReasoningSPM,echo=FALSE,fig.pos="ht", include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
+<<VisReasoningSPM,echo=FALSE, include=F, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
 data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
@@
 @
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{Figure/fig-VisReasoningSPM-1}
 \caption{Pairwise scatterplots of test scores.
@@
 Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
-\begin{figure}[ht]
-<<LineupTypeSPM,echo=FALSE,include=T, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
+
+<<LineupTypeSPM,echo=FALSE,include=F, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
 data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
@@
 @
+
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{Figure/fig-LineupTypeSPM-1}
 \caption{Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
@@
 The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
-
-\begin{figure}[ht]
-<<lineupblockpca, echo=F, warning=F, message=F, fig.width=6, fig.height=4>>=
+<<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 @
+
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{Figure/fig-lineupblockpca-1}
 \caption{Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.\label{fig:influence}}
@@
 
-\begin{figure}[ht]
-<<lineup-types,echo=F, fig.width=4, fig.height=8, out.width=".9\\linewidth", warning=F, message=F>>=
+<<lineup-types,echo=F, fig.width=4, fig.height=8, include=F, warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
 @
+
+\begin{figure}[ht]
+\includegraphics[width=.9\linewidth]{Figure/fig-lineup-types-1}
 \caption{Density plots of scaled scores for different types of lineups.
@@
 
-\begin{figure}[ht]
-
-<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, out.width="\\linewidth", warning=F, message=F>>=
+<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, include=F, warning=F, message=F>>=
 lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
@@
 @
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}
 \caption{Scatterplots of scaled lineup scores by aptitude test scores.

### 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
change graphics device to CairoPDF?

@@
 library(knitr)
-opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|")
+opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
 datadir <- "../../Data/"

### 2015-03-16 Paper/InfoVis2015/TestingVisualAptitude.Rnw
InfoVis style file submission

--- 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
@@
-\documentclass[10pt,journal,compsoc]{IEEEtran}
-%---------------------------------------------------
-% TVCG packages and header
-% *** CITATION PACKAGES ***
-%
-\ifCLASSOPTIONcompsoc
-  % IEEE Computer Society needs nocompress option
-  % requires cite.sty v4.0 or later (November 2003)
-  \usepackage[nocompress]{cite}
-\else
-  % normal IEEE
-  \usepackage{cite}
-\fi
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
+\documentclass[journal]{vgtc}                % final (journal style)
+%\documentclass[review,journal]{vgtc}         % review (journal style)
+%\documentclass[widereview]{vgtc}             % wide-spaced review
+%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
+%\documentclass[electronic,journal]{vgtc}     % electronic version, journal
+
+%% Uncomment one of the lines above depending on where your paper is
+%% in the conference process. ``review'' and ``widereview'' are for review
+%% submission, ``preprint'' is for pre-publication, and the final version
+%% doesn't use a specific qualifier. Further, ``electronic'' includes
+%% hyperreferences for more convenient online viewing.
+
+%% Please use one of the ``review'' options in combination with the
+%% assigned online id (see below) ONLY if your paper uses a double blind
+%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
+%% in the past.
+
+%% Please note that the use of figures other than the optional teaser is not permitted on the first page
+%% of the journal version.  Figures should begin on the second page and be
+%% in CMYK or Grey scale format, otherwise, colour shifting may occur
+%% during the printing process.  Papers submitted with figures other than the optional teaser on the
+%% first page will be refused.
+
+%% These three lines bring in essential packages: ``mathptmx'' for Type 1
+%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
+%% for proper handling of the times font family.
+
+% \usepackage{mathptmx}
 \usepackage{graphicx}
+\usepackage{times}
+
+%% We encourage the use of mathptmx for consistent usage of times font
+%% throughout the proceedings. However, if you encounter conflicts
+%% with other math-related packages, you may want to disable it.
+
 % declare the path(s) where your graphic files are
@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% *** MATH PACKAGES ***
-%
+
+%---------------------------------------------------
 \usepackage[cmex10]{amsmath}
 \usepackage{amssymb}
-
-% *** SUBFigure~PACKAGES ***
-%\ifCLASSOPTIONcompsoc
+\usepackage{color}
+% \usepackage[dvipsnames,svgnames]{xcolor}
+% \usepackage{ulem}
+% \usepackage[section]{placeins}
 \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-
-% *** FLOAT PACKAGES ***
-%
-\usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
+% \usepackage{sidecap}
+\usepackage{multirow}
+\usepackage{bbm}
 \usepackage{url}
+\usepackage{xr}
+\usepackage{xr-hyper}
 
 %---------------------------------------------------
-\usepackage{color}
-\usepackage{fullpage}
-\usepackage[dvipsnames,svgnames]{xcolor}
-%\usepackage{wrapfig,float}
-%\usepackage{caption}
-%\usepackage{subcaption}
-%\usepackage{url}
-\usepackage{ulem}
-\usepackage[section]{placeins}
-\usepackage{sidecap}
-\usepackage{multirow}
-\usepackage{bbm}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage{rotating}
-\usepackage[referable]{threeparttablex}
-\usepackage{footnote} % table footnotes
-%\makesavenoteenv{tabular}
 
-% Bibliography without numbers or labels
-%\usepackage{natbib}
-%\bibliographystyle{apa}
+%% This turns references into clickable hyperlinks.
+\usepackage[bookmarks,backref=section,linkcolor=black]{hyperref} %,colorlinks
+\hypersetup{
+  pdfauthor = {},
+  pdftitle = {},
+  pdfsubject = {},
+  pdfkeywords = {},
+  colorlinks=true,
+  linkcolor= black,
+  citecolor= black,
+  pageanchor=true,
+  urlcolor = black,
+  plainpages = false,
+  linktocpage
+}
 
-%Optional Package to add PDF bookmarks and hypertext links
-\usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
-\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
-%---------------------------------------------------
+\externaldocument{appendix}
 
-%---------------------------------------------------
-%                 Editing Commands
-\newcommand{\done}[2][inline]{\todo[color=SpringGreen, #1]{#2}}  % for todos that have been seen and dealt with
-\newcommand{\meh}[2][inline]{\todo[color=White, #1]{#2}}   % for todos that may no longer be relevant 
-\newcommand{\comment}[2][inline]{\todo[color=SkyBlue, #1]{#2}} % for comments that may not be "to-do"s
-%\newcommand{\mcomment}[1]{\todo[color=SkyBlue]{#1}} % for margin comments
-\newcommand{\newtext}[1]{\todo[inline, color=White]{ \color{OliveGreen}{#1}}} % new text - not necessarily something to be done
-\newcommand{\move}[1]{\todo[inline, color=Lime]{#1}} % new to do item
-%
-%---------------------------------------------------
+%% If you are submitting a paper to a conference for review with a double
+%% blind reviewing process, please replace the value ``0'' below with your
+%% OnlineID. Otherwise, you may safely leave it at ``0''.
+\onlineid{0}
+
+%% declare the category of your paper, only shown in review mode
+\vgtccategory{Theory/Model}
+
+%% allow for this line if you want the electronic option to work properly
+\vgtcinsertpkg
+
+%% In preprint mode you may define your own headline.
+%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
+
+%% Paper title.
+
+\title{Spatial Reasoning and Data Displays}
+
+%% This is how authors are specified in the journal style
+
+%% indicate IEEE Member or Student Member in form indicated below
+\author{Susan VanderPlas, and Heike Hofmann, \textit{Member, IEEE}}
+\authorfooter{
+%% insert punctuation at end of each item
+\item Susan VanderPlas is a graduate student at Iowa State University. Email: skoons@iastate.edu.
+\item Heike Hofmann is a professor at Iowa State University in the Departments of Statistics and Human Computer Interaction. Email: hofmann@iastate.edu.
+}
+
+%other entries to be set up for journal
+\shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Data Displays}
+%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
+
+%% Abstract section.
+\abstract{
+Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
+This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
+We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
+We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
+} % end of abstract
+
+%% Keywords that describe your work. Will show as 'Index Terms' in journal
+%% please capitalize first letter and insert punctuation after last keyword
+\keywords{Data visualization, Perception, Statistical graphics, Statistical computing.}
+
+%% ACM Computing Classification System (CCS). 
+%% See <http://www.acm.org/class/1998/> for details.
+%% The ``\CCScat'' command takes four arguments.
+
+\CCScatlist{ % not used in journal version
+ \CCScat{H.2.8.c}{Data and knowledge visualization}%
+{Database applications}{Database management};
+ \CCScat{G.3.n}{Statistical Computing}{Probability and Statistics}{Mathematics of Computing}
+}
+
+%% Uncomment below to include a teaser figure.
+   \teaser{
+    \centering
+    \includegraphics[width=.4\linewidth]{lineup}
+    \caption[Sample Lineup]{Sample lineup of boxplots.
+Participants are instructed to choose the plot which appears most different from the others.
+In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
+  }
 
@@
 
+%% Uncomment below to disable the manuscript note
+%\renewcommand{\manuscriptnotetxt}{}
+
+%% Copyright space is enabled by default as required by guidelines.
+%% It is disabled by the 'review' option or via the following command:
+% \nocopyrightspace
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
-\title{Spatial Reasoning and Data Displays}
 
-\author{Susan VanderPlas, Heike Hofmann}
+%% The ``\maketitle'' command must be the first command after the
+%% ``\begin{document}'' command. It prepares and prints the title block.
 
-\author{Susan~VanderPlas,
-        and~Heike~Hofmann,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. VanderPlas and H. Hofmann are with the Department
-of Statistics and Statistical Laboratory, Iowa State University, Ames,
-IA, 50011.\protect\\
-% note need leading \protect in front of \\ to get a newline within \thanks as
-% \\ is fragile and will error, could use \hfil\break instead.
-E-mail: {skoons, hofmann}@iastate.edu}% <-this % stops an unwanted space
-\thanks{Manuscript received Month XX, 2015; revised Month XX, XXXX.}}
+%% the only exception to this rule is the \firstsection command
+\firstsection{Introduction}
 
-% The paper headers
-\markboth{IEEE TRANSACTIONS ON VISUALIZATION AND COMPUTER GRAPHICS, No.~X, Month~YYYY}%
-{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Statistical Graphics}
-
-\IEEEtitleabstractindextext{%
-\begin{abstract}
-Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
-This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
-We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
-We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
-\end{abstract}
-
-% Note that keywords are not normally used for peerreview papers.
-% \begin{IEEEkeywords}
-% Computer Society, IEEEtran, journal, \LaTeX, paper, template.
-% \end{IEEEkeywords}
-}
-
-% make the title area
 \maketitle
 
-\IEEEdisplaynontitleabstractindextext
-% \IEEEdisplaynontitleabstractindextext has no effect when using
-% compsoc or transmag under a non-conference mode.
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps (small caps for compsoc).
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-
-\IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
+\label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
 Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
@@
 Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=\linewidth]{lineup}
-\caption[Sample Lineup]{Sample lineup of boxplots.
-Participants are instructed to choose the plot which appears most different from the others.
-In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
-\end{figure}
+% 
+% \begin{figure}[htp]
+% 
+% \end{figure}
 
@@
 
-The \textbf{visual searching task} (VST), shown in Figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
+The \textbf{visual searching task} (VST), shown in appendix figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
 Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
@@
 The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning ablities (factor I  in~\cite{ekstrom1976manual}).
-An example is shown in Figure~\ref{fig:figureclassification}.
+This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
+An example is shown in appendix figure~\ref{fig:figureclassification}.
 
@@
 It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in Figure~\ref{fig:cardrotation}.
+An example is shown in appendix figure~\ref{fig:cardrotation}.
 The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
 
-% \begin{figure*}[hbt]\centering
-% \includegraphics[width=.75\linewidth]{cardrotation}
-% \caption[Card Rotation Task]{Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
-% \end{figure*}
 Two-dimensional comparisons are an important component of lineup performance.
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-A sample question from the test is shown in Figure~\ref{fig:paperfolding}.
+A sample question from the test is shown in appendix figure~\ref{fig:paperfolding}.
 It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
@@
 Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
-Participants have 5 minutes to complete each block of 20 lineups.
+Participants were given 5 minutes to complete each block of 20 lineups.
 Figure~\ref{fig:lineup} shows a sample lineup of box plots.
@@
 
-%\begin{savenotes}
-\begin{table*}[htb]
-\caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.
-The Visual Search task~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
-\label{tab:scorecomparison}}
-\centering
-\begin{threeparttable}  
+\begin{table*}[htb]\centering
 \begin{tabular}{rlllc}
@@
              & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
-             & \Sexpr{sprintf("%.1f", mean(ans.summary$fig_class))} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
+             & \Sexpr{sprintf("%.1f", mean(ans.summary$fig_class))} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnotemark[1]
              & \Sexpr{round(mean(ans.summary$vis_search), 1)} (\Sexpr{round(sd(ans.summary$vis_search), 1)})\\
@@
               & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}
+              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnotemark[2]
               & -- \\
-Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
+Unscaled Scores & 44.0 (24.6)\footnotemark[3]
                 & 13.8 (4.5)
-                & M:~120.0~(30.0), F:~114.9~(27.8)
-                & --\\
+                & \parbox[c]{.2\linewidth}{M:~120.0~(30.0)\\ F:~114.9~(27.8)}
+                & --\\\hline
 {\footnotesize Population}    
               & \parbox[t]{.15\linewidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
-              & \parbox[t]{.15\linewidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\linewidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
+              & \parbox[t]{.17\linewidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.2\linewidth}{\footnotesize suburban 11th \& 12th \\ grade students\\(288-300 M, 317-329 F)}
               & \\\hline
 \end{tabular}
-        \begin{tablenotes}
-            \item[2] ISU students took only Part I due to time constraints.
-            \item[3] Averages calculated assuming 294 males and 323 females.
-            \item[4] Data from Part I only.
-        \end{tablenotes}      
-    \end{threeparttable}
+\caption[Comparison of scores for cognitive tasks.]{Comparison of scores from Iowa State students and scores reported in~\protect\cite{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.
+The visual search task~\protect\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+\label{tab:scorecomparison}}
 \end{table*}
-%\end{savenotes}
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
+%%use following if all content of bibtex file should be shown
+%\nocite{*}
 \bibliography{references}
 
-\newpage
-\appendices
-
-\section{Visuospatial Aptitude Tests}
-The Visual Search Task (VST), shown in Figure~\ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
-\begin{figure}[htp]\centering
-\includegraphics[width=\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task (VST).
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
-Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
-\end{figure}
-
-Figure~\ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
-All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
- \begin{figure*}[ht]
-  \centering
-   \subfloat[Figure Classification Task.
-Participants classify each figure in the second row as belonging to one of the groups above.
-Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.75\linewidth]{figureclassification}
-}
-\hfill
-    \subfloat[Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.75\linewidth]{cardrotation}
-    }
-    \hfill
-    \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
-Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.75\linewidth]{paperfolding}
-   }
-    \caption{Visuospatial tests}
-    \label{fig:tests}
-  \end{figure*}
-% \section{Examining potential testing biases}\label{app:ratests}
-% T-tests of results for Hillary and Stephanie: 
-% <<ttest, echo=F, include=T, results='asis'>>=
-% ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
-% card.rotation <- t.test(card_rot~tester, data=ans.summary)
-% paper.folding <- t.test(folding~tester, data=ans.summary)
-% fig.class <- t.test(fig_class~tester, data=ans.summary)
-% lineup <- t.test(lineup~tester, data=ans.summary)
-% 
-% res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
-%   tmp <- t.test(i~ans.summary$tester)
-%   data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
-%              t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
-% })
-% 
-% res$.id <- c("Card Rotation", "Paper Folding", "figure classification", "Lineups")
-% names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
-% 
-% print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
-% @
-
-\section{Scaling Scores}\label{app:ScoreAdj}
-To calculate ``scaled" comparison scores between tests which included different numbers of test sections (as shown in Table~\ref{tab:scorecomparison}), we scaled the mean in direct proportion to the number of questions (thus, if there were two sections of equivalent size, and the reference score included only one of those sections, we multiplied the reported mean score by two).
-The variance calculation is a bit more complicated: In the case described in the main text, where the reference section contained half of the questions, the variance is multiplied by two, causing the standard deviation to be multiplied by approximately 1.41.
-
-This scaling gets slightly more complicated for scores which have two sub-groups, as with the figure classification test, which separately sumarizes male and female participants' scores. 
-To get a single unified score with standard deviation, we completed the following calculations: 
-\begin{align}
-\mu_{\text{all}} &= (N_F\mu_{F} + N_M\mu_{M})/(N_F + N_M)\\
-\sigma_{\text{all}} &= \sqrt{\left(N_F\sigma_F^2 + N_M\sigma_M^2\right)/(N_F + N_M)}.
-\end{align}
-Here $\mu_F$ and $\mu_M$ are the mean scores for females and males, respectively; $N_F$ and $N_M$ are the number of female and male participants, and $\sigma_F^2$ and $\sigma_M^2$ are the variances in scores for females and males.
-
-Substituting in the provided numbers, we get
-\begin{align*}
-\mu_{\text{all}} &= \left(323\!\cdot\!114.9\! +\! 294\!\cdot\!120.0\right)/(323\!+\!294) \\
-& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
-\sigma_{\text{all}} &= \sqrt{\left(323 \cdot 27.8^2 \!+\! 294 \cdot 30^2\right)/(323\!+\!294)} \\
-& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
-\end{align*}
-
-Whenever participants in two studies were not exposed to the same number of questions, the resulting scores are not comparable: both overall scores and their standard deviations are different.
-We can achieve comparability by scaling the scores accordingly.
-For example, in order to account for the fact that ISU students took only part I of two parts to the figure classification test (and thus completed half of the questions), we adjust the transformation as follows:
-
-\begin{eqnarray*}
-\mu_{\text{part I}} &= 1/2 \cdot \mu_{\text{all}}\\
-\sigma_{\text{part I}} &= 1/\sqrt{2} \cdot \sigma_{\text{all}}
-\end{eqnarray*}
-
-\section{Lineup Performance and Demographic Characteristics}\label{app:categoricalresults}
-Table~\ref{tab:ttest-demographics} provides the results of a sequence of linear models fit to the lineup data.
-Each row in the table represents a single model, with one predictor variable (a factor with two or more levels).
-Due to sample size considerations, multiple testing corrections were not performed; in addition, the independent variables are correlated: in our sample, males are more likely to have completed Calculus 1, but are also more likely to spend time playing video games.
-As such, a model including two or more of the significant predictor variables shows all included variables to be nonsignificant.
-To better understand the effects of these variables, a larger study is necessary.
-<<ttests.categ, echo=F, eval=T, results='asis'>>=
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "MeanSq", "F", "p.val")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
-tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
-print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Participant demographics' impact on lineup score.
-The table below shows each single demographic variable's association with lineup score.
-STEM major, completion of Calculus I, time spent playing video games, and gender all show some association with score on statistical lineups.
-\\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")
-@
-
-\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
-\subsection{Importance of principal components in an analysis of the four cognitive tests}\label{app:pca4}
-
-<<pcasummary4,echo=F, results='asis'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-@
-Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
-PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.
-
-<<pcarotation4,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
-
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
-Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
-@
-Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the remaining principal components distinguish between the cognitive tasks;
-PC2 and PC3 separate  paper folding from visual search and from the lineup and figure classification tasks, while PC4 and PC5 mainly separate lineups from card rotation and figure classification.
-This separation allows us to compare the tasks which are similar from among the principal components.
-According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
-
-\subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}
-PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-Biplots of the remaining principal components are shown in Figure~\ref{fig:biplots5}.
-
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
-
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
-
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-
-\section{Lineup Task Examples}\label{app:lineuptypes}
-\subsection{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-The experiment was originally designed to explore the use of lineups to test plots of competing design\cite{hofmann2012graphical}.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
-\caption{Boxplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
-\caption{Density plots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
-\caption{Dotplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
-\caption{Histograms used to compare the two distributions.}
-\end{figure}
-\FloatBarrier
-\subsection{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f94f6965d7}
-\caption{Boxplots.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
-\caption{Boxplots with jittered points.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9fc58d73}
-\caption{Bee swarm boxplots.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
-\caption{Violin plots.}
-\end{figure}
-\FloatBarrier
-\subsection{Lineup Set 3}
-The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
-
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
-\caption{QQ-plot with guide line.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
-\caption{QQ-plot with acceptance bands.}
-\end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
-\caption{QQ-plot rotated 45 degrees.}
-\end{figure}
-
-\section{Lineup Plot Types} \label{app:plottypes}
-We can also compare participants' performance on specific types of lineup plots compared with their scores on the visual aptitude tests, for instance, accuracy on lineups which require mental rotation may be related to performance on the card rotation task.
-
-<<lineup-types,echo=F, fig.width=4, fig.height=8, include=F, warning=F, message=F>>=
-lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
-lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
-lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
-                                     levels=c("violinplot", "jitter.boxplot", "bee.swarm.boxplot", "boxplot", 
-                                              "scatterplot", "density", "histogram", 
-                                              "qq.line", "qq.line.err", "qq.line.err.rot"), 
-                                     labels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
-                                              "Rotated QQ-plot +\nAcceptance Band")
-                                     )
-loadfonts()
-qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
-      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=2) + 
-  xlab("Proportion of Lineups Correct") + 
-  ylab("Density") + 
-  geom_vline(aes(xintercept=0), linetype=3) + 
-  scale_colour_discrete("Lineup Test Number") + 
-  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  scale_y_continuous(breaks=c(0, .5, 1)) +
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"), 
-        axis.ticks.y=element_blank(), axis.text.y=element_blank())
-embedFonts("./Figure/fig-lineup-types-1.pdf")
-@
-
-\begin{figure}[ht]
-\includegraphics[width=.9\linewidth]{Figure/fig-lineup-types-1}
-\caption{Density plots of scaled scores for different types of lineups.
-For the same experiment (shown by line color), certain types of plots are more difficult to read and are associated with lower participant scores.
-\label{fig:plottypesdensity}}
-\end{figure}
-
-Figure~\ref{fig:plottypesdensity} compares performance on each different type of plot.
-The $x$ axis shows scaled score, the $y$ axis shows the density of participant scores.
-As two different lineup tasks utilized boxplots to test different qualities of the distribution of data (outliers vs.
-difference in medians), different tasks are shown as different colors, so that accuracy on tasks which are shown in blue can be compared to other blue density curves.
-
-Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
-Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
-In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
-
-<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, include=F, warning=F, message=F>>=
-lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
-other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
-names(other.scores)[3] <- "testscore"
-lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
-names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
-lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
-lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
-                                     levels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
-                                              "Rotated QQ-plot +\nAcceptance Band"), 
-                                     labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
-                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
-                                              "QQ-plot\n+ Line", "QQ-plot\n+ Bands", 
-                                              "Rotated QQ\nplot + Bands")
-                                     )
-
-cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-loadfonts()
-qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
-      x=testscore, y=lineup.pct, geom=c("smooth", "jitter"), colour=I("grey30"), method="lm") + 
-  facet_grid(lineup.type~testtype, scales="free_x") + 
-  ylab("Scaled Lineup Score") + 
-  xlab("Test Score")+
-  geom_text(data=subset(cor.labels, testtype!="vis_search"), aes(x=x, y=1, label=label), hjust=0, vjust=1, size=3) + 
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"))
-embedFonts("./Figure/fig-lineup-types-scores-1.pdf")
-@
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}
-\caption{Scatterplots of scaled lineup scores by aptitude test scores.
-There is some indication that different types of lineup tasks may utilize different visual skills; for instance, QQ-plots with confidence bands may require more skill at mental rotation than QQ-plots without the bands.
-\label{fig:scatterplottypes}}
-\end{figure}
-
-For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
-Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in Figure~\ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.
-
 \end{document}
-
-% \section{Lit Review Rejects}
-% \begin{itemize}
-% \item authors in~\cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-% \item authors in~\cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
-They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
-Test administration differences may account for the change.
-They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
-\textit{In the current analysis, we have scored the paper folding test out of 20.
-Alternate scoring procedures still need to be investigated...
-they cite a poster, but it's not available anywhere online as far as I can tell.}
-% \item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
-They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
-% \item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
-The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
-that enable viewers to compare and observe relations among variables".
-% \begin{itemize}
-% \item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
-% Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
-In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
-The review also specifically discusses visuospatial ability, citing a speculation from authors in~\cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
-% \textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.
-That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}
-% \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-Displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
-Graphical displays have higher search and computational efficiency than text~\cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information~\cite{shah1999graphs}.
-% \item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
-Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
-This is pretty unlikely, and the literature is not so relevant to our cause.
-Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
-% \end{itemize}
-% \item authors in~\cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
-They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
-Their problem solving test was more practical (engineering-based), so isn't applicable to our study.
-They were testing instruction using visual + audio in sync, visual + audio out of sync, and then no instruction as a control.
-High-ability students could make more use of the in-sync presentation of visual and auditory information than low-ability students, but did not differ significantly in the out-of-sync information condition.
-Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-% \item authors in~\cite{hofmann2012graphical} for lineup stimuli and general lineup performance
-% \end{itemize}

### 2015-03-16 Paper/InfoVis2015/TestingVisualAptitude.tex
InfoVis style file submission

--- 2015-03-16 Paper/InfoVis2015/TestingVisualAptitude.Rnw
@@
-\documentclass[journal]{vgtc}                % final (journal style)
+\documentclass[journal]{vgtc}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}                % final (journal style)
 %\documentclass[review,journal]{vgtc}         % review (journal style)
@@
 
-<<VisualAptitude-setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
-rm(list=ls())
-wd <- getwd()
-library(extrafont)
-library(knitr)
-opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
-datadir <- "../../Data/"
-imgdir <- "Figure/"
-codedir <- "../../Code/"
-
-options(replace.assign=TRUE,scipen=3)
-
-library(reshape2)
-suppressMessages(library(ggplot2))
-library(plyr)
-library(grid)
-suppressMessages(library(gridExtra))
-library(xtable)
-library(stringr)
-library(magrittr)
-source(paste(codedir, "Analysis.R", sep=""))
-
-sanitize <- function(x){
-  str_replace(gsub("_", ".", x, fixed=T), "Pr\\((.{1,}))", "Pr$(\\1)$")
-}
-
-@
-
 %% Uncomment below to disable the manuscript note
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \section{Results}\label{sec:results}
-Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
-\Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
+Results are based on an evaluation of 38 undergraduate students at Iowa State University.
+61\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
@@
   & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
-             & \Sexpr{sprintf("%.1f", mean(ans.summary$fig_class))} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnotemark[1]
-             & \Sexpr{round(mean(ans.summary$vis_search), 1)} (\Sexpr{round(sd(ans.summary$vis_search), 1)})\\
-Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
+ISU Students & 83.4 (24.1) 
+             & 12.4 (3.7)
+             & 57.0 (23.8)\footnotemark[1]
+             & 21.9 (2.3)\\
+Scaled Scores & 88.0 (34.8)
               & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnotemark[2]
+              & 58.7 (14.4)\footnotemark[2]
               & -- \\
@@
 In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.
-<<ResultsScaledScores, echo=F, include=F, fig.width=4.5, fig.height=4, warning=F, message=F>>=
-loadfonts()
-score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
-score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
-qrange$sqrt.k <- sqrt(qrange$k-1)
-qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
-
-qr1 <- qrange
-qr1$Type <- "Theoretical Range"
-qr2 <- qrange
-qr2$Type <- "Scores indicating guessing"
-qr2$n <- with(qr2, -n/(k-1))
-
-qrange <- rbind(qr1, qr2)
-qrange$Type <- factor(qrange$Type)
-qrange$Type <- factor(qrange$Type, levels=levels(qrange$Type)[2:1])
-
-ggplot() + 
-  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, fill=Type), alpha=.35, stat="identity", position="identity", colour="red", inherit.aes=F) +
-  scale_fill_manual("", values=c(NA, "red")) + 
-  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
-  theme_bw()  + 
-  ylab("Test Scores") + 
-  xlab("") +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
-  theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.68,.125),  legend.box.just = "left", 
-        legend.key.height=unit(6, "mm")) 
-  #coord_flip()
-embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
-@
 
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5.5, message=F, warning=F, fig.pos='ht'>>=
-loadfonts()
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "STEM Major", "Video Game Hrs/Wk"))
-lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="m"] <- "Male"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="y"] <- "TRUE"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"
-
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
-lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
-
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey")) + 
-  facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 
-  ylab("Scaled Lineup Score") + 
-  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
-  theme_bw() + 
-  theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
-embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
-@
 \begin{figure*}[h!tb]\centering
@@
 \subsection{Understanding Visual Abilities and Lineup Performance}
-<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize", include=F>>=
-print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests.\\label{tab:corrmatrix5}"), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9, table.placement="htb", caption.placement="top")
-@
 
-<<VisReasoningSPM,echo=FALSE, include=F, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
-grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
-grid <- subset(grid, x != y & x<y)
-
-all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-  xcol <- grid[i, "x"]
-  ycol <- grid[i, "y"]
-
-  rbind.fill(
-    data.frame(
-      xvar = names(data)[ycol], 
-      yvar = names(data)[xcol],
-      x = data[, xcol], 
-      y = data[, ycol], 
-      points = T,
-      data
-    ),
-    data.frame(
-      yvar = names(data)[ycol], 
-      xvar = names(data)[xcol],
-      x = mean(range(data[, ycol])), 
-      y = mean(range(data[, xcol])),
-      label = round(cor(data[,xcol], data[,ycol]), 3),
-      points = F
-    )
-  )
-}))
-all$xvar <- factor(all$xvar, levels=names(data))
-all$yvar <- factor(all$yvar, levels=names(data))
-densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-    tmp <- density(data[,i], cut = 1, adjust=.75)
-    data.frame(
-      xvar = names(data)[i], 
-      yvar = names(data)[i],
-      x = tmp$x, 
-      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
-    )
-  }))
-fix.names <- function(values){
-  values %>% 
-    str_replace("lineup", "Lineups") %>% 
-    str_replace("vis_search", "Visual Search") %>%
-    str_replace("fig_class", "Figure Classification") %>% 
-    str_replace("card_rot", "Card Rotation") %>%
-    str_replace("folding", "Paper Folding")
-}
-loadfonts()
-require(grid)
-ggplot()+
-  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
-  geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
-  geom_text(data=subset(all, !points), aes(x=x, y=y, label=paste0("Correlation = \n", label))) + 
-  geom_line(data=densities, aes(x=x, y=y)) + 
-  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
-  theme_bw() + 
-  theme(axis.title=element_blank(),
-        plot.margin=unit(c(0,0,0,0), unit="mm"))
-rm(list=c("data", "all", "densities"))
-embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
-@
 \begin{figure}[ht]
@@
 
-<<pca,echo=F, warning=F, message=F>>=
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-# screeplot(pca)
-@
-
 Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
@@
 
-A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
+A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about 64\% of the total variability.
 Principal components (PC-) are ordered by importance (how much variability in the data they contain), and each principal component is uncorrelated with every other principal component.
@@
 PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-The other principal components span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
+The other principal components span another two dimensions, while the last dimension is weak (at 6\%).
 PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
@@
 From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.
-<<pcasummary5,echo=F, results='asis'>>=
-print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="ht", caption.placement="top")
-@
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 17:04:25 2015
+\begin{table}[ht]
+\centering
+\caption{Importance of principal components, analyzing all five tests.\label{tab:PCAvariance5}} 
+{\footnotesize
+\begin{tabular}{rrrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 & PC5 \\ 
+  \hline
+Standard deviation & 1.73 & 0.84 & 0.75 & 0.70 & 0.48 \\ 
+  Proportion of Variance & 0.60 & 0.14 & 0.11 & 0.10 & 0.05 \\ 
+  Cumulative Proportion & 0.60 & 0.74 & 0.85 & 0.95 & 1.00 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
 
-<<pcarotation5,echo=F, results='asis'>>=
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="ht", caption.placement="top")
-@
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 17:04:25 2015
+\begin{table}[ht]
+\centering
+\caption{PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
+             \label{tab:PCArotation5}} 
+\begin{tabular}{rrrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 & PC5 \\ 
+  \hline
+lineup & 0.42 & 0.49 & -0.46 & 0.60 & -0.10 \\ 
+  card.rot & 0.50 & -0.30 & 0.28 & 0.23 & 0.73 \\ 
+  fig.class & 0.43 & 0.45 & -0.15 & -0.75 & 0.18 \\ 
+  folding & 0.47 & 0.07 & 0.68 & 0.04 & -0.56 \\ 
+  vis.search & 0.41 & -0.69 & -0.48 & -0.15 & -0.33 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
 From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that  
-the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
+the first principal component, PC1, is again essentially an average across all tests and accounts for 60.1\% of the variance in the data. 
 Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix~\ref{app:pca5}.
@@
 % 
-% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
+% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for round(summary(pca)$importance[2,4]*100,1)\% of the variance.
 % PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
@@
 Note, that the correlation between the math and the stats score is almost zero.
-<<linearmodel, results='hide', echo=FALSE>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-ans.model <- data.frame(ans.summary, data.frame(pca2$x))
-# exclude all variables that are not supposed to go into the linear model
-# those are the unrotated tests; instead we have the PCs
-excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
 
-ans.model$math_science_research <- ans.model$math_science_research=="y"
-ans.model$stats_class <- ans.model$stats_class=="y"
-ans.model$calc_1 <- ans.model$calc_1=="y"
-pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills", "stem")], scale=T)
-pcmath$rotation
-# first component of math PCA is average of math_skills, calc_1 and stem
-# second component is stat_class and research
-
-#biplot(pcmath)
-#biplot(pcmath, 3:4)
-
-xtabs(~stem+sex, data=ans.model)
-# sad day!
-
-xtabs(~stem+calc_1, data=ans.model)
-# stem major is more associated with performance on lineups than calc 1...
-
-xtabs(~stats_class+math_science_research, data=ans.model)
-# only one kid is doing research without a stats class
-
-# recode to sum of them
-ans.model$STATS <- with(ans.model, stats_class + math_science_research)
-ans.model$MATH <- with(ans.model, calc_1+math_skills/5+stem)
-
-excl <- c(excl, c("math_science_research", "stats_class"))
-excl <- c(excl, c("calc_1", "math_skills", "stem"))
-incl <- setdiff(names(ans.model), excl)
-
-m0 <- lm(lineup~., data=ans.model[, incl])
-
-# use AIC in backward selection
-library(MASS)
-m1 <- stepAIC(m0, direction="both")
-@
 We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
@@
 
-<<modelresults, echo=FALSE, results='asis'>>=
-library(xtable)
-print(xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1"), caption.placement="top")
-@
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 17:04:25 2015
+\begin{table}[ht]
+\centering
+\caption{Estimates and significances of a linear model of lineup scores.} 
+\label{tab:m1}
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
+  \hline
+(Intercept) & 14.1192 & 1.9149 & 7.37 & 0.0000 \\ 
+  PC1 & 1.7672 & 0.5230 & 3.38 & 0.0018 \\ 
+  MATH & 2.1246 & 0.8732 & 2.43 & 0.0202 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
 \subsection{Lineup Types}
@@
 Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
-
-<<LineupTypeSPM,echo=FALSE,include=F, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
-data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
-grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
-grid <- subset(grid, x != y & x<y)
-
-all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-  xcol <- grid[i, "x"]
-  ycol <- grid[i, "y"]
-
-  rbind.fill(
-    data.frame(
-      xvar = names(data)[ycol], 
-      yvar = names(data)[xcol],
-      x = data[, xcol], 
-      y = data[, ycol], 
-      points = T,
-      data
-    ),
-    data.frame(
-      yvar = names(data)[ycol], 
-      xvar = names(data)[xcol],
-      x = mean(range(data[, ycol])), 
-      y = mean(range(data[, xcol])),
-      label = round(cor(data[,xcol], data[,ycol]), 3),
-      points = F
-    )
-  )
-}))
-all$xvar <- factor(all$xvar, levels=names(data))
-all$yvar <- factor(all$yvar, levels=names(data))
-densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-    tmp <- density(data[,i], cut = 1, adjust=.75)
-    data.frame(
-      xvar = names(data)[i], 
-      yvar = names(data)[i],
-      x = tmp$x, 
-      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
-    )
-  }))
-fix.names <- function(values){
-  values %>% 
-    str_replace("lineup", "Lineups") %>% 
-    str_replace("Lineups_", "Lineup ") %>%
-    str_replace("vis_search", "Visual Search") %>%
-    str_replace("fig_class", "Figure Class.") %>% 
-    str_replace("card_rot", "Card Rotation") %>%
-    str_replace("folding", "Paper Folding")
-}
-
-all2 <- subset(all, grepl("lineup", yvar) & points & !grepl("vis_search", xvar))
-# names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
-all2mean <- rbind(
-  ddply(subset(all2, !grepl("lineup", xvar)), .(xvar, yvar), summarize, label=paste0("rho == ", round(cor(x,y), 3)), x=min(x)-diff(range(x))/5, y=max(y)+diff(range(y))/5, hjust=0, vjust=1, size=3),
-  ddply(subset(all2, grepl("lineup", xvar) & grepl("lineup", yvar)), .(xvar, yvar), function(df){
-    with(df, data.frame(xvar=unique(yvar), yvar=unique(xvar), 
-                        x=mean(range(y)), y=mean(range(x)), 
-                        label=paste0("rho == ", round(cor(x,y),3)),
-                        hjust=.5, vjust=.5, size=4))
-    })
-)
-loadfonts()
-ggplot()+
-  geom_point(data=subset(all2, points), aes(x=x, y=y)) + 
-  geom_smooth(data=subset(all2, points), aes(x=x, y=y), method="lm") + 
-  geom_text(data=all2mean, aes(x=x, y=y, label=label, hjust=hjust, vjust=vjust, size=size), parse=T) + 
-  scale_size_identity() + 
-  geom_line(data=subset(densities, grepl("lineup", xvar)), aes(x=x, y=y)) + 
-  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
-  theme_bw() + 
-  theme(axis.title=element_blank())
-embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
-@
 
@@
 The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
-<<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
-lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-rotation <- melt(lineup.block.pca$rotation)
-importance <- as.data.frame(summary(lineup.block.pca)$importance)[2,]
-rotation <- merge(rotation, melt(importance, variable.name="Var2", value.name="importance"))
-rotation$Var1 <- factor(rotation$Var1, 
-                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_3", "lineup_2", "lineup_1"),
-                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
-library(grid) # needed for arrow function
-loadfonts()
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
-embedFonts("./Figure/fig-lineupblockpca-1.pdf")
-@
 
@@
 As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
-
-<<removeAptitude,echo=F,fig.width=8, fig.height=4>>=
-vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
-lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
-vmain <- rowSums(vistests)/4
-alltests <- data.frame(vmain, vistests-vmain, lineups)
-pctests <- prcomp(alltests)
-par(mfrow=c(1,2), mar=c(4,4.5,.1,.1), tcl=-0.01, cex.lab=.8,cex.axis=.7,mgp=c(2,.7,0),tcl=-.3)
-#biplot(pctests)
-#biplot(pctests, 3:4)
-@
 

### 2015-03-16 Paper/InfoVis2015/appendix.Rnw
InfoVis style file submission

--- 2015-02-09 Paper/Draft1/TestingVisualAptitude.Rnw
@@
-\documentclass[10pt,journal,compsoc]{IEEEtran}
-%---------------------------------------------------
-% TVCG packages and header
-% *** CITATION PACKAGES ***
-%
-\ifCLASSOPTIONcompsoc
-  % IEEE Computer Society needs nocompress option
-  % requires cite.sty v4.0 or later (November 2003)
-  \usepackage[nocompress]{cite}
-\else
-  % normal IEEE
-  \usepackage{cite}
-\fi
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
+\documentclass[journal]{vgtc}                % final (journal style)
+%\documentclass[review,journal]{vgtc}         % review (journal style)
+%\documentclass[widereview]{vgtc}             % wide-spaced review
+%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
+%\documentclass[electronic,journal]{vgtc}     % electronic version, journal
+
+%% Uncomment one of the lines above depending on where your paper is
+%% in the conference process. ``review'' and ``widereview'' are for review
+%% submission, ``preprint'' is for pre-publication, and the final version
+%% doesn't use a specific qualifier. Further, ``electronic'' includes
+%% hyperreferences for more convenient online viewing.
+
+%% Please use one of the ``review'' options in combination with the
+%% assigned online id (see below) ONLY if your paper uses a double blind
+%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
+%% in the past.
+
+%% Please note that the use of figures other than the optional teaser is not permitted on the first page
+%% of the journal version.  Figures should begin on the second page and be
+%% in CMYK or Grey scale format, otherwise, colour shifting may occur
+%% during the printing process.  Papers submitted with figures other than the optional teaser on the
+%% first page will be refused.
+
+%% These three lines bring in essential packages: ``mathptmx'' for Type 1
+%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
+%% for proper handling of the times font family.
+
+\usepackage{mathptmx}
 \usepackage{graphicx}
+\usepackage{times}
+
+%% We encourage the use of mathptmx for consistent usage of times font
+%% throughout the proceedings. However, if you encounter conflicts
+%% with other math-related packages, you may want to disable it.
+
 % declare the path(s) where your graphic files are
@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% *** MATH PACKAGES ***
-%
+
+%---------------------------------------------------
 \usepackage[cmex10]{amsmath}
 \usepackage{amssymb}
-
-% *** SUBFigure~PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-
-% *** FLOAT PACKAGES ***
-%
-\usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-
-%---------------------------------------------------
 \usepackage{color}
-\usepackage{fullpage}
 \usepackage[dvipsnames,svgnames]{xcolor}
-%\usepackage{wrapfig,float}
-%\usepackage{caption}
-%\usepackage{subcaption}
-%\usepackage{url}
 \usepackage{ulem}
 \usepackage[section]{placeins}
+\usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
 \usepackage{sidecap}
@@
 \usepackage{bbm}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage{rotating}
-\usepackage[referable]{threeparttablex}
-\usepackage{footnote} % table footnotes
-%\makesavenoteenv{tabular}
+\usepackage{url}
+\usepackage{xr}
+\usepackage{xr-hyper}
 
-% Bibliography without numbers or labels
-%\usepackage{natbib}
-%\bibliographystyle{apa}
-
-%Optional Package to add PDF bookmarks and hypertext links
-\usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
-\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}
 %---------------------------------------------------
 
-%---------------------------------------------------
-%                 Editing Commands
-\newcommand{\done}[2][inline]{\todo[color=SpringGreen, #1]{#2}}  % for todos that have been seen and dealt with
-\newcommand{\meh}[2][inline]{\todo[color=White, #1]{#2}}   % for todos that may no longer be relevant 
-\newcommand{\comment}[2][inline]{\todo[color=SkyBlue, #1]{#2}} % for comments that may not be "to-do"s
-%\newcommand{\mcomment}[1]{\todo[color=SkyBlue]{#1}} % for margin comments
-\newcommand{\newtext}[1]{\todo[inline, color=White]{ \color{OliveGreen}{#1}}} % new text - not necessarily something to be done
-\newcommand{\move}[1]{\todo[inline, color=Lime]{#1}} % new to do item
-%
-%---------------------------------------------------
+%% This turns references into clickable hyperlinks.
+\usepackage[bookmarks,backref=section,linkcolor=black]{hyperref} %,colorlinks
+\hypersetup{
+  pdfauthor = {},
+  pdftitle = {},
+  pdfsubject = {},
+  pdfkeywords = {},
+  colorlinks=true,
+  linkcolor= black,
+  citecolor= black,
+  pageanchor=true,
+  urlcolor = black,
+  plainpages = false,
+  linktocpage
+}
+
+\externaldocument{TestingVisualAptitude}
+
+%% If you are submitting a paper to a conference for review with a double
+%% blind reviewing process, please replace the value ``0'' below with your
+%% OnlineID. Otherwise, you may safely leave it at ``0''.
+\onlineid{0}
+
+%% declare the category of your paper, only shown in review mode
+\vgtccategory{Theory/Model}
+
+%% allow for this line if you want the electronic option to work properly
+\vgtcinsertpkg
+
+%% In preprint mode you may define your own headline.
+%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
+
+%% Paper title.
+
+\title{Spatial Reasoning and Data Displays}
+
+%% This is how authors are specified in the journal style
+
+%% indicate IEEE Member or Student Member in form indicated below
+\author{Susan VanderPlas, and Heike Hofmann, \textit{Member, IEEE}}
+\authorfooter{
+%% insert punctuation at end of each item
+\item
+ Susan VanderPlas is a graduate student at Iowa State University. %Email: skoons@iastate.edu.
+\item
+ Heike Hofmann is a professor at Iowa State University in the Departments of Statistics and Human Computer Interaction. %Email: hofmann@iastate.edu.
+}
+
+%other entries to be set up for journal
+\shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Data Displays}
+%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
+
+%% Abstract section.
+\abstract{
+Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
+This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
+We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
+We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
+} % end of abstract
+
+%% Keywords that describe your work. Will show as 'Index Terms' in journal
+%% please capitalize first letter and insert punctuation after last keyword
+\keywords{Data visualization, Perception, Statistical graphics, Statistical computing.}
+
+%% ACM Computing Classification System (CCS). 
+%% See <http://www.acm.org/class/1998/> for details.
+%% The ``\CCScat'' command takes four arguments.
+
+\CCScatlist{ % not used in journal version
+ \CCScat{H.2.8.c}{Data and knowledge visualization}%
+{Database applications}{Database management};
+ \CCScat{G.3.n}{Statistical Computing}{Probability and Statistics}{Mathematics of Computing}
+}
+
+%% Uncomment below to include a teaser figure.
+%    \teaser{
+%    \centering
+%    \includegraphics[width=16cm]{CypressView}
+%    \caption{In the Clouds: Vancouver from Cypress Mountain.}
+%   }
 
@@
 
+%% Uncomment below to disable the manuscript note
+%\renewcommand{\manuscriptnotetxt}{}
+
+%% Copyright space is enabled by default as required by guidelines.
+%% It is disabled by the 'review' option or via the following command:
+% \nocopyrightspace
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
-\title{Spatial Reasoning and Data Displays}
 
-\author{Susan VanderPlas, Heike Hofmann}
-
-\author{Susan~VanderPlas,
-        and~Heike~Hofmann,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. VanderPlas and H. Hofmann are with the Department
-of Statistics and Statistical Laboratory, Iowa State University, Ames,
-IA, 50011.\protect\\
-% note need leading \protect in front of \\ to get a newline within \thanks as
-% \\ is fragile and will error, could use \hfil\break instead.
-E-mail: {skoons, hofmann}@iastate.edu}% <-this % stops an unwanted space
-\thanks{Manuscript received Month XX, 2015; revised Month XX, XXXX.}}
-
-% The paper headers
-\markboth{IEEE TRANSACTIONS ON VISUALIZATION AND COMPUTER GRAPHICS, No.~X, Month~YYYY}%
-{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Statistical Graphics}
-
-\IEEEtitleabstractindextext{%
-\begin{abstract}
-Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
-This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
-We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
-We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
-\end{abstract}
-
-% Note that keywords are not normally used for peerreview papers.
-% \begin{IEEEkeywords}
-% Computer Society, IEEEtran, journal, \LaTeX, paper, template.
-% \end{IEEEkeywords}
-}
-
-% make the title area
-\maketitle
-
-\IEEEdisplaynontitleabstractindextext
-% \IEEEdisplaynontitleabstractindextext has no effect when using
-% compsoc or transmag under a non-conference mode.
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
-
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps (small caps for compsoc).
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-
-\IEEEPARstart{D}{ata} displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
-Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
-Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
-Under this framework, well-designed graphics reduce memory load and make more cognitive resources available for other tasks (such as drawing conclusions from the data), at the cost of depending on certain visuospatial reasoning abilities.
-
-Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
-Both of these theories suggest spatial ability impacts a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
-However,  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
-Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~\cite{zhang1997nature}.
-
-``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
-Lineups are also useful in assessing the effectiveness of different graphical displays \cite{hofmann2012graphical, loy:2015}.
-Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width=\linewidth]{lineup}
-\caption[Sample Lineup]{Sample lineup of boxplots.
-Participants are instructed to choose the plot which appears most different from the others.
-In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
-\end{figure}
-
-Figure~\ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
-In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
-
-%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
-
-Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
-This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
-
-In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
-We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
-The VST measures visual search speed\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning\cite{ekstrom1976manual};
-all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
-We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
-
-In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
-In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
-We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
-Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
-
-\section{Methods}\label{sec:methods}
-
-\subsection{The Lineup Protocol}
-The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
-
-In a lineup test, the plot of the data is placed randomly among a set of, generally 19, distractor plots (or {\it null plots}) that are rendered from data generated by a model, without a signal (a null model).
-This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different".
-If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
-Let $X$ be the number of observers (out of $n$) who identify the data plot from the lineup.
-Under the null hypothesis that the data plot is not different from the other plots, $X$ has approximately a Binomial distribution \cite{wickham2010graphical, majumder2013validation}.
-If $k$ of the observers identify the data plot from the lineup, the probability $P(X \ge k)$ is the $p$-value of the corresponding visual test.
-%
-By aggregating responses from different individuals the lineup protocol therefore allows an objective evaluation of a graphical finding.
-
-Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task.
-
-For this approach, we derive a score for an individual as follows:
-
-Assume that an observer has evaluated $K$ lineups of size $m$ (consisting of $m-1$ decoy plots and 1 target), and successfully identified the target in $n_c$ of these plots, while missing the target in $n_w$ of them.
-The score for this individual is then given as:
-\begin{eqnarray}\label{eq.scoring1}
-n_c - n_w/(m-1).
-\end{eqnarray}
-Note that the sum of answers, $n_c + n_w$, is at most $K$, but may be less, if an observer chooses to not answer one of the lineup tasks or runs out of time.
-
-The scoring scheme as given in (\ref{eq.scoring1}) is chosen so that if participants are guessing, the expected score is 0.
-
-Statistical lineups depend on the ability to search for a signal amid a set of distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
-Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from   participants, e.g.~ polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
-By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-
-Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups \cite{humanfactorslineups}.
-In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
-
-One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively by providing an experimental protocol for comparing graphics based on the accuracy of user conclusions.
-In addition, lineups provide researchers with a rigorous framework for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
-As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.
-
-\subsection{Measures of visuospatial ability}
-
-Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
-Tasks are timed such  that participants are under  pressure to complete; participants are not expected to finish all of the problems in each section.
-This allows for a better discrimination between scores and prevents score compression at the top of the range.
-
-The \textbf{visual searching task} (VST), shown in Figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
-Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
-The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
-
-The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning ablities (factor I  in~\cite{ekstrom1976manual}).
-An example is shown in Figure~\ref{fig:figureclassification}.
-
-The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
-In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
-As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
-
-The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in Figure~\ref{fig:cardrotation}.
-The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
-
-% \begin{figure*}[hbt]\centering
-% \includegraphics[width=.75\linewidth]{cardrotation}
-% \caption[Card Rotation Task]{Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants will complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.}\label{fig:cardrotation}
-% \end{figure*}
-Two-dimensional comparisons are an important component of lineup performance.
-In some lineup situations, these comparisons sometimes involve translation, but in other lineups, rotation is required.
-Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
-
-The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-A sample question from the test is shown in Figure~\ref{fig:paperfolding}.
-It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
-The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
-% \begin{figure*}[hbt]\centering
-% \includegraphics[width=.75\linewidth]{paperfolding}
-% \caption[Paper Folding Task]{Paper Folding Task.
-Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
-Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
-% \end{figure*}
-Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
-
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
-Participants have 5 minutes to complete each block of 20 lineups.
-Figure~\ref{fig:lineup} shows a sample lineup of box plots.
-
-In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
-These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
-
-\subsection{Test Scoring}\label{sec:scaling}
-
-All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
-Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
-\begin{eqnarray}\label{eq.scoring}
-\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
-\end{eqnarray}
-This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
-
-\section{Results}\label{sec:results}
-Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
-\Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
-Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
-This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the university as a whole; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
-
-\subsection{Comparison of Spatial Tests with Previously Validated Results}
-The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
-We compare Iowa State students' unscaled scores in Table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
-
-%\begin{savenotes}
-\begin{table*}[htb]
-\caption{Comparison of scores from Iowa State students and scores reported in~\cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.
-The Visual Search task~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
-\label{tab:scorecomparison}}
-\centering
-\begin{threeparttable}  
-\begin{tabular}{rlllc}
-\hline
-  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
-ISU Students & \Sexpr{round(mean(ans.summary$card_rot), 1)} (\Sexpr{round(sd(ans.summary$card_rot), 1)}) 
-             & \Sexpr{round(mean(ans.summary$folding), 1)} (\Sexpr{round(sd(ans.summary$folding), 1)})
-             & \Sexpr{sprintf("%.1f", mean(ans.summary$fig_class))} (\Sexpr{round(sd(ans.summary$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}
-             & \Sexpr{round(mean(ans.summary$vis_search), 1)} (\Sexpr{round(sd(ans.summary$vis_search), 1)})\\
-Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
-              & 13.8 (4.5)
-              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}
-              & -- \\
-Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
-                & 13.8 (4.5)
-                & M:~120.0~(30.0), F:~114.9~(27.8)
-                & --\\
-{\footnotesize Population}    
-              & \parbox[t]{.15\linewidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
-              & \parbox[t]{.15\linewidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.3\linewidth}{\footnotesize suburban 11th \& 12th grade students\\(288-300 males, 317-329 females)}
-              & \\\hline
-\end{tabular}
-        \begin{tablenotes}
-            \item[2] ISU students took only Part I due to time constraints.
-            \item[3] Averages calculated assuming 294 males and 323 females.
-            \item[4] Data from Part I only.
-        \end{tablenotes}      
-    \end{threeparttable}
-\end{table*}
-%\end{savenotes}
-
-Table~\ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
-Values have been adjusted to accommodate for differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix~\ref{app:ScoreAdj}).
-Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
-In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.
-
-Comparison population data was chosen to most closely match ISU undergraduate population demographics.
-Thus, if comparison data was available for 9th and 12th grade students, scores of Iowa State students were compared to scores of 12th grade students, who are closer in age to college students.
-When data was available from college students and Army enlistees, comparisons of scores were based on other college students, as college students are more likely to have a similar gender distribution to ISU students.
-
-Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
-Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
-Participants' scores on the VST indicate score compression; that is, both participants with medium and high visual search abilities scored at the extremely high end of the spectrum. 
-In future experiments, participants should be given less time (or more questions) to better differentiate participants with medium and high-ability.
-<<ResultsScaledScores, echo=F, include=F, fig.width=4.5, fig.height=4, warning=F, message=F>>=
-loadfonts()
-score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
-score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
-qrange$sqrt.k <- sqrt(qrange$k-1)
-qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure Class.", "Paper Folding", "Visual Search"))
-
-qr1 <- qrange
-qr1$Type <- "Theoretical Range"
-qr2 <- qrange
-qr2$Type <- "Scores indicating guessing"
-qr2$n <- with(qr2, -n/(k-1))
-
-qrange <- rbind(qr1, qr2)
-qrange$Type <- factor(qrange$Type)
-qrange$Type <- factor(qrange$Type, levels=levels(qrange$Type)[2:1])
-
-ggplot() + 
-  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, fill=Type), alpha=.35, stat="identity", position="identity", colour="red", inherit.aes=F) +
-  scale_fill_manual("", values=c(NA, "red")) + 
-  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
-  theme_bw()  + 
-  ylab("Test Scores") + 
-  xlab("") +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
-  theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
-  theme(legend.position=c(.68,.125),  legend.box.just = "left", 
-        legend.key.height=unit(6, "mm")) 
-  #coord_flip()
-embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
-@
-
-\begin{figure}[ht]
-\includegraphics[width=.9\linewidth]{Figure/fig-ResultsScaledScores-1}
-\caption{Test scores for lineups and visuospatial tests.
-As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing.
-We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability.
-\label{fig:Scores}}
-\end{figure}
-%
-
-% Applying the grading protocol discussed in section~\ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.
-Figure~\ref{fig:Scores} shows the range of possible scores and the observed score distribution.
-
-\subsection{Lineup Performance and Demographic Characteristics}
-
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5.5, message=F, warning=F, fig.pos='ht'>>=
-loadfonts()
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
-lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "STEM Major", "Video Game Hrs/Wk"))
-lineup.summary.categorical$value[lineup.summary.categorical$value=="f"] <- "Female"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="m"] <- "Male"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="y"] <- "TRUE"
-lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"
-
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
-lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
-
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey")) + 
-  facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 
-  ylab("Scaled Lineup Score") + 
-  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
-  theme_bw() + 
-  theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
-embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
-@
-\begin{figure*}[h!tb]\centering
-\includegraphics[width=.85\linewidth]{fig-VisReasoningCategorical-1}
-\caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
-Categories are ordered by effect size; majoring in a STEM field, calculus completion, hours spent playing video games per week, and sex are all associated with a significant difference in lineup score.
-}\label{fig:visualaptitudecat}
-\end{figure*}
-
-%Next, we examine the effect of demographic factors on lineup performance. 
-Previous work found a relationship between lineup performance and demographic factors such as education level, country of origin, and age \cite{humanfactorslineups}; our participant population is very homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
-
-Figure~\ref{fig:visualaptitudecat} shows participants' lineup scores in relationship to their responses in the questionnaire given at the beginning of the study; this allows us to explore effects of demographic characteristics (major, research experience, etc.) on test performance.
-
-Completion of Calculus I is associated with increased performance on lineups; this may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with findings in~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
-There is also a significant relationship between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
-There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
-These demographic characteristics were chosen to account for life experience and personal skills which may have influenced the results.
-Statistical test results are available in Appendix~\ref{app:categoricalresults}.
-
-\subsection{Understanding Visual Abilities and Lineup Performance}
-<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize", include=F>>=
-print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests.\\label{tab:corrmatrix5}"), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9, table.placement="htb", caption.placement="top")
-@
-
-<<VisReasoningSPM,echo=FALSE, include=F, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
-data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
-grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
-grid <- subset(grid, x != y & x<y)
-
-all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-  xcol <- grid[i, "x"]
-  ycol <- grid[i, "y"]
-
-  rbind.fill(
-    data.frame(
-      xvar = names(data)[ycol], 
-      yvar = names(data)[xcol],
-      x = data[, xcol], 
-      y = data[, ycol], 
-      points = T,
-      data
-    ),
-    data.frame(
-      yvar = names(data)[ycol], 
-      xvar = names(data)[xcol],
-      x = mean(range(data[, ycol])), 
-      y = mean(range(data[, xcol])),
-      label = round(cor(data[,xcol], data[,ycol]), 3),
-      points = F
-    )
-  )
-}))
-all$xvar <- factor(all$xvar, levels=names(data))
-all$yvar <- factor(all$yvar, levels=names(data))
-densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-    tmp <- density(data[,i], cut = 1, adjust=.75)
-    data.frame(
-      xvar = names(data)[i], 
-      yvar = names(data)[i],
-      x = tmp$x, 
-      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
-    )
-  }))
-fix.names <- function(values){
-  values %>% 
-    str_replace("lineup", "Lineups") %>% 
-    str_replace("vis_search", "Visual Search") %>%
-    str_replace("fig_class", "Figure Classification") %>% 
-    str_replace("card_rot", "Card Rotation") %>%
-    str_replace("folding", "Paper Folding")
-}
-loadfonts()
-require(grid)
-ggplot()+
-  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
-  geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
-  geom_text(data=subset(all, !points), aes(x=x, y=y, label=paste0("Correlation = \n", label))) + 
-  geom_line(data=densities, aes(x=x, y=y)) + 
-  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
-  theme_bw() + 
-  theme(axis.title=element_blank(),
-        plot.margin=unit(c(0,0,0,0), unit="mm"))
-rm(list=c("data", "all", "densities"))
-embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
-@
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-VisReasoningSPM-1}
-\caption{Pairwise scatterplots of test scores.
-Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores.
-Paper folding and card rotation scores are also highly correlated.\label{fig:scatterplotmatrix}}
-\end{figure}
-
-<<pca,echo=F, warning=F, message=F>>=
+<<pca,echo=F, warning=F, message=F,include=F>>=
 pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 @
-
-Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
-What is of more interest to us is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
-In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests using the same procedure.
-
-A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
-Principal components (PC-) are ordered by importance (how much variability in the data they contain), and each principal component is uncorrelated with every other principal component.
-
-PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-The other principal components span another two dimensions, while the last dimension is weak (at \Sexpr{round(summary(pca2)$importance[2,4],2)*100}\%).
-PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
-More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
-
-Incorporating the lineup task into the principal component analysis, we find the principal components to be fairly similar to the four-component analysis. 
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component.
-From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.
-<<pcasummary5,echo=F, results='asis'>>=
+<<pcasummary5,echo=F,include=F>>=
 print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="ht", caption.placement="top")
@@
 
-<<pcarotation5,echo=F, results='asis'>>=
+<<pcarotation5,echo=F,include=F>>=
 print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
@@
 @
-From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that  
-the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data. 
-Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix~\ref{app:pca5}.
 
-Figure classification is strongly related to lineups, and as in the four-component PCA, figure classification is strongly represented in the first two principal components.
-While lineups do span a separate dimension, the PCA suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
-
-% 
-% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
-% PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
-%
-
-This emphasizes the underpinnings of lineups: the test utilizes a visual medium, but is ultimately are a classification task presented in a graphical manner.
-Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.
-
-\subsection{Linear model of demographic factors}
-Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.
-Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
-Only one student is doing research who has not taken a statistics course.
-
-A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
-We therefore decided to use sums of these variables to come up with a separate math and a stats score.
-Note, that the correlation between the math and the stats score is almost zero.
-<<linearmodel, results='hide', echo=FALSE>>=
+<<linearmodel, include=FALSE, echo=FALSE>>=
 pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 @
-We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
-The result is shown in Table~\ref{tab:m1}.
-Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores.
-We can think of PC1 as a measure of innate visual or intellectual ability, while the MATH score is a matter of both ability and training.
-The remaining principal components were not sufficiently associated with lineup score to be included in the model.
-
-<<modelresults, echo=FALSE, results='asis'>>=
-library(xtable)
-print(xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1"), caption.placement="top")
-@
-\subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous studies on different designs to investigate which plot type most effectively conveyed important characteristics of the underlying data set.
-Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
-
-<<LineupTypeSPM,echo=FALSE,include=F, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
-data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
-grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
-grid <- subset(grid, x != y & x<y)
-
-all <- do.call("rbind", lapply(1:nrow(grid), function(i) {
-  xcol <- grid[i, "x"]
-  ycol <- grid[i, "y"]
-
-  rbind.fill(
-    data.frame(
-      xvar = names(data)[ycol], 
-      yvar = names(data)[xcol],
-      x = data[, xcol], 
-      y = data[, ycol], 
-      points = T,
-      data
-    ),
-    data.frame(
-      yvar = names(data)[ycol], 
-      xvar = names(data)[xcol],
-      x = mean(range(data[, ycol])), 
-      y = mean(range(data[, xcol])),
-      label = round(cor(data[,xcol], data[,ycol]), 3),
-      points = F
-    )
-  )
-}))
-all$xvar <- factor(all$xvar, levels=names(data))
-all$yvar <- factor(all$yvar, levels=names(data))
-densities <- do.call("rbind", lapply(1:ncol(data), function(i) {
-    tmp <- density(data[,i], cut = 1, adjust=.75)
-    data.frame(
-      xvar = names(data)[i], 
-      yvar = names(data)[i],
-      x = tmp$x, 
-      y = tmp$y/max(tmp$y)* diff(range(tmp$x)) + min(tmp$x)
-    )
-  }))
-fix.names <- function(values){
-  values %>% 
-    str_replace("lineup", "Lineups") %>% 
-    str_replace("Lineups_", "Lineup ") %>%
-    str_replace("vis_search", "Visual Search") %>%
-    str_replace("fig_class", "Figure Class.") %>% 
-    str_replace("card_rot", "Card Rotation") %>%
-    str_replace("folding", "Paper Folding")
-}
-
-all2 <- subset(all, grepl("lineup", yvar) & points & !grepl("vis_search", xvar))
-# names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
-all2mean <- rbind(
-  ddply(subset(all2, !grepl("lineup", xvar)), .(xvar, yvar), summarize, label=paste0("rho == ", round(cor(x,y), 3)), x=min(x)-diff(range(x))/5, y=max(y)+diff(range(y))/5, hjust=0, vjust=1, size=3),
-  ddply(subset(all2, grepl("lineup", xvar) & grepl("lineup", yvar)), .(xvar, yvar), function(df){
-    with(df, data.frame(xvar=unique(yvar), yvar=unique(xvar), 
-                        x=mean(range(y)), y=mean(range(x)), 
-                        label=paste0("rho == ", round(cor(x,y),3)),
-                        hjust=.5, vjust=.5, size=4))
-    })
-)
-loadfonts()
-ggplot()+
-  geom_point(data=subset(all2, points), aes(x=x, y=y)) + 
-  geom_smooth(data=subset(all2, points), aes(x=x, y=y), method="lm") + 
-  geom_text(data=all2mean, aes(x=x, y=y, label=label, hjust=hjust, vjust=vjust, size=size), parse=T) + 
-  scale_size_identity() + 
-  geom_line(data=subset(densities, grepl("lineup", xvar)), aes(x=x, y=y)) + 
-  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
-  theme_bw() + 
-  theme(axis.title=element_blank())
-embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
-@
-
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-LineupTypeSPM-1}
-\caption{Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
-All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
-\label{fig:lineuptaskmatrix}}
-\end{figure}
-Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
-The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
-Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
-
-In order to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of an experiment, we employ principal component analysis on participant scores averaged across each block of lineups.
-
-Principal component analysis separates multivariate data into orthogonal components using a rotation matrix to transform correlated input data into an orthogonal space.
-The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
 <<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
@@
                         labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
-library(grid) # needed for arrow function
-loadfonts()
-qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
-embedFonts("./Figure/fig-lineupblockpca-1.pdf")
+
 @
 
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-lineupblockpca-1}
-\caption{Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.\label{fig:influence}}
-\end{figure}
-
-For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
-
-The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
-
-The influence of variable $X_i$ on component $PC_j$ is then defined as:
-%
-$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
-%
-Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
-Small (absolute) values indicate that there is little influence of $X_i$ on $PC_j$; negative values show the direction of the influence (to maintain the separation of variables as part of principal components analysis).
-
-Figure~\ref{fig:influence} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
-The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
-This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
-
-PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
-PC3 emphasizes the additional variation in performance on lineup test 3 and the visual search test, while PC4 emphasizes the extra variability in performance on lineup test 2 and the paper folding test.
-All three lineups, plus the figure classification, paper folding, and visual search tests contribute to PC5.
-PC6 and PC7 jointly account for less than 10\% of the variance in the data and do not display any distinct patterns in the loadings.
-
-While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
-Overall, there is an additional principal component gained from separating the lineup blocks by experiment type.
-As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
-
-<<removeAptitude,echo=F,fig.width=8, fig.height=4>>=
-vistests <- scale(lineup.section.summary[,c("card_rot", "fig_class", "folding", "vis_search")])
-lineups <- scale(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3")])
-vmain <- rowSums(vistests)/4
-alltests <- data.frame(vmain, vistests-vmain, lineups)
-pctests <- prcomp(alltests)
-par(mfrow=c(1,2), mar=c(4,4.5,.1,.1), tcl=-0.01, cex.lab=.8,cex.axis=.7,mgp=c(2,.7,0),tcl=-.3)
-#biplot(pctests)
-#biplot(pctests, 3:4)
-@
-
-The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
-
-A larger study might not only address the question of which lineup tasks require certain visual skills, but also the use of different types of plots from a perceptual perspective.
-Preliminary results of performance on different types of plots are shown in Appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
-The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
-Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.
-
-\section{Discussion and Conclusions}\label{sec:discussion}
-
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-
-Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
-This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
-
-When lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
-Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).
-Future studies testing larger sets of lineups may be useful to understand which types of plots require additional visuospatial skills, as plots which appeal to a wider audience may be more successful when conveying important information.
-
-In addition to this theoretical information, the figure classification test may be useful for pre-screening participants in future online lineup studies.
-Such studies often suffer from participants who do not take the task seriously, and internal verification questions, as well as pre-qualification tasks are often used to reduce extraneous variability.
-While it is impractical to require participants to score well on several different tests, it is reasonable to ask participants to pre-qualify for a task by completing a figure classification test.
-As the figure classification test is different from the lineup task, this will not bias participants' scores on the domain of interest while ensuring that the participant pool is sufficiently motivated to complete the lineup questions.
-
-The demographic results from this study indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
-
-All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\newpage
-\appendices
-
+\appendix
 \section{Visuospatial Aptitude Tests}
@@
 \begin{figure}[htp]\centering
-\includegraphics[width=\linewidth]{VisualSearch}
+\includegraphics[width=.5\linewidth]{VisualSearch}
 \caption[Visual Search Task]{Visual Search Task (VST).
@@
 All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
- \begin{figure*}[ht]
+ \begin{figure}[ht]
   \centering
@@
     \label{fig:tests}
-  \end{figure*}
-% \section{Examining potential testing biases}\label{app:ratests}
-% T-tests of results for Hillary and Stephanie: 
-% <<ttest, echo=F, include=T, results='asis'>>=
-% ans.summary$tester <- c(rep("Hillary", 18), rep("Stephanie", nrow(ans.summary)-18))
-% card.rotation <- t.test(card_rot~tester, data=ans.summary)
-% paper.folding <- t.test(folding~tester, data=ans.summary)
-% fig.class <- t.test(fig_class~tester, data=ans.summary)
-% lineup <- t.test(lineup~tester, data=ans.summary)
-% 
-% res <- ldply(ans.summary[,c("card_rot", "folding", "fig_class", "lineup")], function(i){
-%   tmp <- t.test(i~ans.summary$tester)
-%   data.frame(Hillary = as.numeric(tmp$estimate)[1], Stephanie = as.numeric(tmp$estimate)[2],
-%              t = tmp$statistic, df=tmp$parameter, p=tmp$p.value, lb = tmp$conf.int[1], ub=tmp$conf.int[2])
-% })
-% 
-% res$.id <- c("Card Rotation", "Paper Folding", "figure classification", "Lineups")
-% names(res) <- c("Variable", "Mean.Hillary", "Mean.Stephanie", "t", "df", "p-value", "Diff.LB", "Diff.UB")
-% 
-% print(xtable(res), include.rownames = F, sanitize.text.function=sanitize)
-% @
+\end{figure}
 
@@
 <<ttests.categ, echo=F, eval=T, results='asis'>>=
+tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
+    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp)[1,]
+    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
+    rownames(coefs) <- NULL
+    coefs <- coefs[,c(6, 1:5)]
+    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
+    coefs
+})
+
+tmp <- rbind.fill(tmp)
+tmp <- tmp[order(tmp$p.value, decreasing=F),]
+
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
+lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
+
 tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
@@
 
+\bibliographystyle{abbrv}
+%%use following if all content of bibtex file should be shown
+%\nocite{*}
+\bibliography{references}
+
 \end{document}
-
-% \section{Lit Review Rejects}
-% \begin{itemize}
-% \item authors in~\cite{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same.
-% \item authors in~\cite{voyer1995magnitude} completed a meta-analysis of spatial tasks and gender differences.
-They concluded that ``rotation" tasks, such as the card rotation test, have robust sex differences across cohort, and while the sex differences seem to be declining for some tests, such as the card rotation test, the differences still exist for the time being.
-Test administration differences may account for the change.
-They also concluded that the paper-folding test showed no such sex difference, though this may be dependent on test scoring (a lower guessing penalty decreases the score difference).
-\textit{In the current analysis, we have scored the paper folding test out of 20.
-Alternate scoring procedures still need to be investigated...
-they cite a poster, but it's not available anywhere online as far as I can tell.}
-% \item authors in~\cite{lowrie2007solving} investigated students (9-10 yrs) ability to read mathematical graphs and their visual-spatial ability.
-They partition graphs into 6 ``graphical languages" (some of these languages aren't statistical graphs - networks, venn diagrams, maps), and use multivariate analysis to associate spatial ability with some of the types of graphs studied.
-% \item authors in~\cite{vekiri2002value} is lit review discussing the different theories about the utility of graphics in learning (dual coding theory, visual argument hypothesis, and conjoint retention hypothesis).
-The article contains a nice definition of graphics (though it might be too domain specific for statistical graphics, since it claims graphics have only one meaning and aren't prone to interpretation.) Depending on the context, ``maps", ``charts", and ``graphs" here could fall into a loose classification as ``statistical graphics", but in general ``graphs" is probably accurate for most of the lineups we're testing: ``[Referents are ] quantitative data ...
-that enable viewers to compare and observe relations among variables".
-% \begin{itemize}
-% \item \textbf{Dual Coding Theory}: people encode both words and pictures separately, so both are encoded for a graph. \\
-% Supports the use of graphics, and there is support for dual-coding in working memory research (pictures and words are processed separately but in parallel).
-In addition, since graphics depend on other abilities (spatial abilities), graphics that aren't properly designed create higher cognitive load and lower performance on graphical tasks.
-The review also specifically discusses visuospatial ability, citing a speculation from authors in~\cite{mayer1994whom} that students with low spatial ability perform worse with diagrams and graphs because more cognitive resources are required.
-% \textit{Lots of studies by Sweller cited throught the cognitive load discussion - I haven't gotten to that literature yet.
-That said, that research has some potential to justify the use of lineups from a psychological perspective - if you can claim that types of graphs that have poor performance do so because they create higher cognitive load (cue psych research into why that would be true, cleveland \& mcgill, etc.), you can justify the suckiness of those awful boxplot-jittered-things.}
-% \item \textbf{Visual Argument Hypothesis}: Graphics require less processing because their visuospatial properties encode some of the information through the arrangement of elements as well as the elements themselves.
-Displays that clearly present information without deep processing requirements are more effective\cite{zhang1994representations}.
-Graphical displays have higher search and computational efficiency than text~\cite{larkin1987diagram}, and when graphics are designed according to gestalt principles of organization, where grouping lines up with visual chunks, gestalt heuristics can be used to more quickly encode information~\cite{shah1999graphs}.
-% \item \textbf{Conjoint Retention Hypothesis} - based on dual coding theory, but claims that there are separate but interconnected memory codes for verbal and visual information.
-Assumes maps/graphs are encoded intact, with visuospatial properties preserved.
-This is pretty unlikely, and the literature is not so relevant to our cause.
-Spatial information is encoded, but words can actually interfere with this encoding; it is not ``intact"
-% \end{itemize}
-% \item authors in~\cite{mayer1994whom} used the card rotation test (10 figures, 80 questions total) and the paper folding test (10 items) to evaluate spatial performance, along with a problem solving test.
-They used the same penalty structure on the card rotation test, and don't seem to have used any penalty on the paper folding test; they then scaled the two tests.
-Their problem solving test was more practical (engineering-based), so isn't applicable to our study.
-They were testing instruction using visual + audio in sync, visual + audio out of sync, and then no instruction as a control.
-High-ability students could make more use of the in-sync presentation of visual and auditory information than low-ability students, but did not differ significantly in the out-of-sync information condition.
-Thus, visualization ability moderates the results of the information presented by allowing students to process information concurrently more efficiently, but doesn't help students process the information separately or help students come up with solutions to the problem-solving task.
-% \item authors in~\cite{hofmann2012graphical} for lineup stimuli and general lineup performance
-% \end{itemize}

### 2015-03-16 Paper/InfoVis2015/appendix.tex
InfoVis style file submission

--- 2015-03-16 Paper/InfoVis2015/appendix.Rnw
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
-
-<<pca,echo=F, warning=F, message=F,include=F>>=
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-# screeplot(pca)
-@
-<<pcasummary5,echo=F,include=F>>=
-print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="ht", caption.placement="top")
-@
-
-<<pcarotation5,echo=F,include=F>>=
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="ht", caption.placement="top")
-@
-
-<<linearmodel, include=FALSE, echo=FALSE>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-ans.model <- data.frame(ans.summary, data.frame(pca2$x))
-# exclude all variables that are not supposed to go into the linear model
-# those are the unrotated tests; instead we have the PCs
-excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")
-
-ans.model$math_science_research <- ans.model$math_science_research=="y"
-ans.model$stats_class <- ans.model$stats_class=="y"
-ans.model$calc_1 <- ans.model$calc_1=="y"
-pcmath <- prcomp(ans.model[,c("math_science_research", "stats_class", "calc_1", "math_skills", "stem")], scale=T)
-pcmath$rotation
-# first component of math PCA is average of math_skills, calc_1 and stem
-# second component is stat_class and research
-
-#biplot(pcmath)
-#biplot(pcmath, 3:4)
-
-xtabs(~stem+sex, data=ans.model)
-# sad day!
-
-xtabs(~stem+calc_1, data=ans.model)
-# stem major is more associated with performance on lineups than calc 1...
-
-xtabs(~stats_class+math_science_research, data=ans.model)
-# only one kid is doing research without a stats class
-
-# recode to sum of them
-ans.model$STATS <- with(ans.model, stats_class + math_science_research)
-ans.model$MATH <- with(ans.model, calc_1+math_skills/5+stem)
-
-excl <- c(excl, c("math_science_research", "stats_class"))
-excl <- c(excl, c("calc_1", "math_skills", "stem"))
-incl <- setdiff(names(ans.model), excl)
-
-m0 <- lm(lineup~., data=ans.model[, incl])
-
-# use AIC in backward selection
-library(MASS)
-m1 <- stepAIC(m0, direction="both")
-@
-<<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
-lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-rotation <- melt(lineup.block.pca$rotation)
-importance <- as.data.frame(summary(lineup.block.pca)$importance)[2,]
-rotation <- merge(rotation, melt(importance, variable.name="Var2", value.name="importance"))
-rotation$Var1 <- factor(rotation$Var1, 
-                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_3", "lineup_2", "lineup_1"),
-                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
-
-@
 
@@
 \mu_{\text{all}} &= \left(323\!\cdot\!114.9\! +\! 294\!\cdot\!120.0\right)/(323\!+\!294) \\
-& = \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)}\\
+& = 58.7\\
 \sigma_{\text{all}} &= \sqrt{\left(323 \cdot 27.8^2 \!+\! 294 \cdot 30^2\right)/(323\!+\!294)} \\
-& = \Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)}.
+& = 14.4.
 \end{align*}
@@
 To better understand the effects of these variables, a larger study is necessary.
-<<ttests.categ, echo=F, eval=T, results='asis'>>=
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
-    coefs
-})
+\begin{kframe}
 
-tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
-lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
-
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "MeanSq", "F", "p.val")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
-tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
-print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Participant demographics' impact on lineup score.
+{\ttfamily\noindent\bfseries\color{errorcolor}{| Error in eval(expr, envir, enclos): contrasts can be applied only to factors with 2 or more levels}}\end{kframe}% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 16:58:33 2015
+\begin{table}[ht]
+\centering
+\caption{Participant demographics' impact on lineup score.
 The table below shows each single demographic variable's association with lineup score.
 STEM major, completion of Calculus I, time spent playing video games, and gender all show some association with score on statistical lineups.
-\\label{tab:ttest-demographics}"), include.rownames=F, floating.environment="table", caption.placement="top")
-@
+\label{tab:ttest-demographics}}
+\begin{tabular}{lrrrr}
+  \hline
+Variable & DF & Mean.Squared & F.value & p.value \\ 
+  \hline
+stem & 1 & 401.517 & 14.44 & 0.001 \\ 
+  calc\_1 & 1 & 204.569 & 6.15 & 0.018 \\ 
+  vidgame\_hrs\_factor & 3 & 108.847 & 3.44 & 0.028 \\ 
+  sex & 1 & 140.844 & 4.02 & 0.053 \\ 
+  art\_skills & 4 & 75.891 & 2.28 & 0.082 \\ 
+  verbal\_skills & 3 & 60.220 & 1.68 & 0.191 \\ 
+  math\_science\_research & 1 & 59.670 & 1.60 & 0.214 \\ 
+  AutoCAD & 1 & 50.893 & 1.36 & 0.252 \\ 
+  age & 1 & 34.434 & 0.91 & 0.348 \\ 
+  math\_skills & 3 & 37.039 & 0.98 & 0.416 \\ 
+  stats\_class & 1 & 9.062 & 0.23 & 0.631 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
@@
 
-<<pcasummary4,echo=F, results='asis'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-@
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 16:58:33 2015
+\begin{table}[htb]
+\centering
+\caption{Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\label{tab:PCAvariance4}} 
+{\footnotesize
+\begin{tabular}{rrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 \\ 
+  \hline
+Standard deviation & 1.61 & 0.81 & 0.73 & 0.49 \\ 
+  Proportion of Variance & 0.64 & 0.16 & 0.13 & 0.06 \\ 
+  Cumulative Proportion & 0.64 & 0.81 & 0.94 & 1.00 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
+
 Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
-PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC1 accounts for about 60\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
 PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.
+PC4 is not particularly significant (it accounts for 5.9\% of the variance), but it differentiates the card rotation task from the paper folding task.
 
-<<pcarotation4,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 16:58:33 2015
+\begin{table}[htb]
+\centering
+\caption{Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\label{tab:PCArotation4}} 
+\begin{tabular}{rrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 \\ 
+  \hline
+card.rot & 0.55 & -0.19 & -0.38 & 0.72 \\ 
+  fig.class & 0.46 & 0.58 & 0.66 & 0.14 \\ 
+  folding & 0.52 & 0.33 & -0.53 & -0.59 \\ 
+  vis.search & 0.46 & -0.72 & 0.38 & -0.34 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=.7\linewidth]{Figure/fig-biplot-pca4-1} 
+\includegraphics[width=.7\linewidth]{Figure/fig-biplot-pca4-2}
+
+}
+
+\caption{Biplots of principal components 1-4 with observations.
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
-@
+\label{fig:biplots4}}\label{fig:biplot-pca4}
+\end{figure}
+
+\end{knitrout}
 Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
@@
 This separation allows us to compare the tasks which are similar from among the principal components.
-According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
+According to Table~\ref{tab:PCArotation4}, the first three principal components account for 94.1\% of the variance.
 
@@
 
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=.7\linewidth]{Figure/fig-biplots-pca5-1} 
+\includegraphics[width=.7\linewidth]{Figure/fig-biplots-pca5-2}
+
+}
+
+\caption{Biplots of principal components 2-5 with observations.
 The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
 PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
+\label{fig:biplots5}}\label{fig:biplots-pca5}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<lineup-types,echo=F, fig.width=4, fig.height=8, include=F, warning=F, message=F>>=
-lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
-lineup.type.sum2$plot.type <- str_replace(as.character(lineup.type.sum2$plot.type), "error", "err")
-lineup.type.sum2$plot.type <- factor(lineup.type.sum2$plot.type, 
-                                     levels=c("violinplot", "jitter.boxplot", "bee.swarm.boxplot", "boxplot", 
-                                              "scatterplot", "density", "histogram", 
-                                              "qq.line", "qq.line.err", "qq.line.err.rot"), 
-                                     labels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
-                                              "Rotated QQ-plot +\nAcceptance Band")
-                                     )
-loadfonts()
-qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
-      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=2) + 
-  xlab("Proportion of Lineups Correct") + 
-  ylab("Density") + 
-  geom_vline(aes(xintercept=0), linetype=3) + 
-  scale_colour_discrete("Lineup Test Number") + 
-  scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
-  scale_y_continuous(breaks=c(0, .5, 1)) +
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"), 
-        axis.ticks.y=element_blank(), axis.text.y=element_blank())
-embedFonts("./Figure/fig-lineup-types-1.pdf")
-@
-
 \begin{figure}[ht]
@@
 
-<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, include=F, warning=F, message=F>>=
-lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
-other.scores <- subset(lineup.type.sum3, testtype!="lineup")[,c(1,2,13)]
-names(other.scores)[3] <- "testscore"
-lineup.scores <- melt(subset(lineup.type.sum3, testtype=="lineup")[,c(1,3:12)], id.vars = 1)
-names(lineup.scores)[2:3] <- c("lineup.type", "lineup.pct")
-lineup.type.sum3 <- merge(other.scores, lineup.scores, all.x=T, all.y=T)
-lineup.type.sum3$lineup.type <- factor(lineup.type.sum3$lineup.type, 
-                                     levels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
-                                              "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
-                                              "Rotated QQ-plot +\nAcceptance Band"), 
-                                     labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
-                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
-                                              "QQ-plot\n+ Line", "QQ-plot\n+ Bands", 
-                                              "Rotated QQ\nplot + Bands")
-                                     )
-
-cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-loadfonts()
-qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
-      x=testscore, y=lineup.pct, geom=c("smooth", "jitter"), colour=I("grey30"), method="lm") + 
-  facet_grid(lineup.type~testtype, scales="free_x") + 
-  ylab("Scaled Lineup Score") + 
-  xlab("Test Score")+
-  geom_text(data=subset(cor.labels, testtype!="vis_search"), aes(x=x, y=1, label=label), hjust=0, vjust=1, size=3) + 
-  theme_bw() + 
-  theme(legend.position="bottom", legend.direction="horizontal", 
-        axis.text.x=element_text(angle=30, hjust=1),
-        plot.margin=unit(c(0,0,0,0),"mm"))
-embedFonts("./Figure/fig-lineup-types-scores-1.pdf")
-@
 \begin{figure}[ht]
@@
 
-For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
+For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was 0.505, the correlation between general lineup score and the figure classification test was 0.512, and the correlation between lineup score and the paper folding test was 0.471.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
 Additional data is imperative to understand the reasoning required for specific types of plots - it is likely that the 5-10 trials per participant presented in each chart in Figure~\ref{fig:scatterplottypes} are simply not sufficient to uncover any specific relationship between reasoning ability and lineup task.

### 2015-03-16 Paper/InfoVis2015/appendix.Rnw
Appendix fixes

@@
 opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
-datadir <- "../../Data/"
+
+setwd("../../")
+datadir <- "Data/"
 imgdir <- "Figure/"
-codedir <- "../../Code/"
+codedir <- "Code/"
 
@@
 library(magrittr)
-source(paste(codedir, "Analysis.R", sep=""))
+
+source(paste0(codedir, "Analysis.R"))
 
@@
 \begin{document}
-
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{A\arabic{figure}}
 <<pca,echo=F, warning=F, message=F,include=F>>=
@@
 Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.75\linewidth]{figureclassification}
+  \includegraphics[width=.8\linewidth]{figureclassification}
 }
@@
 Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.75\linewidth]{cardrotation}
+\includegraphics[width=.8\linewidth]{cardrotation}
     }
@@
 Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.75\linewidth]{paperfolding}
+\includegraphics[width=.8\linewidth]{paperfolding}
    }
@@
 <<ttests.categ, echo=F, eval=T, results='asis'>>=
-tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
-    coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
-    rownames(coefs) <- NULL
-    coefs <- coefs[,c(6, 1:5)]
-    names(coefs) <- c("Variable", "DF", "Sum.of.Squares", "Mean.Squared", "F.value", "p.value")
-    coefs
-})
-
-tmp <- rbind.fill(tmp)
-tmp <- tmp[order(tmp$p.value, decreasing=F),]
-
-lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, levels=unique(tmp$Variable))
-lineup.summary.categorical$value <- factor(lineup.summary.categorical$value, levels=c("0","1", "2","3","4","5","[1, 2)", "[2, 5)", "5+", "18-20", "21+", "TRUE", "FALSE", "Female", "Male"))
+lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
+lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="math_skills", value=1))
+lineup.summary.categorical$variable <- factor(lineup.summary.categorical$variable, labels=c("Age", "Sex", "Math/Science Research", "Statistics Class", "Calculus 1", "Verbal Skills", "Math Skills", "Art Skills", "AutoCAD Experience", "STEM Major", "Video Game Hrs/Wk"))
 
 tmp <- dlply(lineup.summary.categorical, .(variable), function(df){
-    tmp <- anova(lm(lineup~factor(value, ordered=F), data=df))
-    coefs <- as.data.frame(tmp)[1,]
+    tmp2 <- anova(lm(lineup~factor(value, ordered=F), data=df))
+    coefs <- as.data.frame(tmp2)[1,]
     coefs$level <- gsub("factor(value, ordered = F)", unique(df$variable), rownames(coefs), fixed=T)
@@
 
+library(stringr)
 tmp <- rbind.fill(tmp)
 tmp2 <- tmp[order(tmp$p.val, decreasing=F),-3]
-tmp2$Variable <- gsub("Math/Science", "STEM", gsub(" Experience", "", gsub(" Hrs/Wk", " hrs", tmp2$Variable)), fixed=T)
+tmp2$Variable <- tmp2$Variable %>% 
+  str_replace(fixed("Math/Science"), "STEM") %>%
+  str_replace(" Experience", "") %>%
+  str_replace(" Hrs/Wk", " Hours")
 print(xtable(tmp2, digits=c(1, 1, 0, 3, 2, 3), caption="Participant demographics' impact on lineup score.
@@
 
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
@@
 
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.7\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
 The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
@@
 
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
 \caption{Boxplots used to compare the two distributions.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
 \caption{Density plots used to compare the two distributions.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
 \caption{Dotplots used to compare the two distributions.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
 \caption{Histograms used to compare the two distributions.}
 \end{figure}
-\FloatBarrier
+
 \subsection{Lineup Set 2}
@@
 Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f94f6965d7}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f94f6965d7}
 \caption{Boxplots.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
 \caption{Boxplots with jittered points.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f9fc58d73}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9fc58d73}
 \caption{Bee swarm boxplots.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
 \caption{Violin plots.}
 \end{figure}
-\FloatBarrier
+
 \subsection{Lineup Set 3}
@@
 
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13876a225d34-single}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13876a225d34-single}
 \caption{QQ-plot with guide line.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
 \caption{QQ-plot with acceptance bands.}
 \end{figure}
-\begin{figure}[htbp]
-\includegraphics[keepaspectratio=T,width=.9\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
+\begin{figure}[htbp]\centering
+\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
 \caption{QQ-plot rotated 45 degrees.}
@@
 
-\begin{figure}[ht]
-\includegraphics[width=.9\linewidth]{Figure/fig-lineup-types-1}
+\begin{figure}[ht]\centering
+\includegraphics[width=.8\linewidth]{Figure/fig-lineup-types-1}
 \caption{Density plots of scaled scores for different types of lineups.
@@
 
-<<lineup-types-scores,echo=F,fig.width=6, fig.height=12, include=F, warning=F, message=F>>=
+<<lineup-types-scores,echo=F,fig.width=12, fig.height=5, include=F, warning=F, message=F>>=
 lineup.type.sum3 <- dcast(lineup.type.sum2, id + testtype ~ plot.type, value.var="score", fun.aggregate = mean)
@@
                                      )
+lineup.type.sum3 <- ddply(lineup.type.sum3, .(testtype), transform, scaled.testscore=testscore/max(testscore))
 
-cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
+cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(scaled.testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
 loadfonts()
 qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
-      x=testscore, y=lineup.pct, geom=c("smooth", "jitter"), colour=I("grey30"), method="lm") + 
-  facet_grid(lineup.type~testtype, scales="free_x") + 
+      x=scaled.testscore, y=lineup.pct, geom=c("smooth", "jitter"), colour=I("grey30"), method="lm") + 
+  facet_grid(testtype~lineup.type, scales="free") + 
   ylab("Scaled Lineup Score") + 
-  xlab("Test Score")+
+  xlab("Scaled Test Score")+
   geom_text(data=subset(cor.labels, testtype!="vis_search"), aes(x=x, y=1, label=label), hjust=0, vjust=1, size=3) + 
@@
         plot.margin=unit(c(0,0,0,0),"mm"))
-embedFonts("./Figure/fig-lineup-types-scores-1.pdf")
+# embedFonts("./Figure/fig-lineup-types-scores-1.pdf")
 @
-\begin{figure}[ht]
+\begin{figure*}[ht]\centering
 \includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}
@@
 \label{fig:scatterplottypes}}
-\end{figure}
+\end{figure*}
 

### 2015-03-16 Paper/InfoVis2015/appendix.tex
Appendix fixes

@@
 \begin{document}
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{A\arabic{figure}}
 
@@
 To better understand the effects of these variables, a larger study is necessary.
-\begin{kframe}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{| Error in eval(expr, envir, enclos): contrasts can be applied only to factors with 2 or more levels}}\end{kframe}% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 16:58:33 2015
+% latex table generated in R 3.1.2 by xtable 1.7-1 package
+% Mon Mar 16 17:53:53 2015
 \begin{table}[ht]
@@
   \hline
-Variable & DF & Mean.Squared & F.value & p.value \\ 
+Variable & DF & MeanSq & F & p.val \\ 
   \hline
-stem & 1 & 401.517 & 14.44 & 0.001 \\ 
-  calc\_1 & 1 & 204.569 & 6.15 & 0.018 \\ 
-  vidgame\_hrs\_factor & 3 & 108.847 & 3.44 & 0.028 \\ 
-  sex & 1 & 140.844 & 4.02 & 0.053 \\ 
-  art\_skills & 4 & 75.891 & 2.28 & 0.082 \\ 
-  verbal\_skills & 3 & 60.220 & 1.68 & 0.191 \\ 
-  math\_science\_research & 1 & 59.670 & 1.60 & 0.214 \\ 
+STEM Major & 1 & 401.517 & 14.44 & 0.001 \\ 
+  Calculus 1 & 1 & 204.569 & 6.15 & 0.018 \\ 
+  Video Game Hours & 3 & 108.847 & 3.44 & 0.028 \\ 
+  Sex & 1 & 140.844 & 4.02 & 0.053 \\ 
+  Art Skills & 4 & 75.891 & 2.28 & 0.082 \\ 
+  Verbal Skills & 3 & 60.220 & 1.68 & 0.191 \\ 
+  STEM Research & 1 & 59.670 & 1.60 & 0.214 \\ 
   AutoCAD & 1 & 50.893 & 1.36 & 0.252 \\ 
-  age & 1 & 34.434 & 0.91 & 0.348 \\ 
-  math\_skills & 3 & 37.039 & 0.98 & 0.416 \\ 
-  stats\_class & 1 & 9.062 & 0.23 & 0.631 \\ 
+  Age & 1 & 34.434 & 0.91 & 0.348 \\ 
+  Math Skills & 3 & 37.039 & 0.98 & 0.416 \\ 
+  Statistics Class & 1 & 9.062 & 0.23 & 0.631 \\ 
    \hline
@@
 % latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 16:58:33 2015
+% Mon Mar 16 17:53:53 2015
 \begin{table}[htb]
@@
 
-{\centering \includegraphics[width=.7\linewidth]{Figure/fig-biplot-pca4-1} 
-\includegraphics[width=.7\linewidth]{Figure/fig-biplot-pca4-2}
+{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-1} 
+\includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-2}
 
@@
 
-{\centering \includegraphics[width=.7\linewidth]{Figure/fig-biplots-pca5-1} 
-\includegraphics[width=.7\linewidth]{Figure/fig-biplots-pca5-2}
+{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-1} 
+\includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-2}
 
@@
 
-\begin{figure}[ht]
+\begin{figure*}[ht]\centering
 \includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}

### 2015-03-22 Paper/InfoVis2015/TestingVisualAptitude.Rnw
Moved appendix sections to paper

@@
 
-The \textbf{visual searching task} (VST), shown in appendix figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
+The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
 Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
@@
 
+\begin{figure}[htp]\centering
+\includegraphics[width=.8\linewidth]{VisualSearch}
+\caption[Visual Search Task]{Visual Search Task (VST).
+Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
+Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
+\end{figure}
+
 The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
 This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
-An example is shown in appendix figure~\ref{fig:figureclassification}.
+An example is shown in figure~\ref{fig:figureclassification}.
 
@@
 It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in appendix figure~\ref{fig:cardrotation}.
+An example is shown in figure~\ref{fig:cardrotation}.
 The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
@@
 The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-A sample question from the test is shown in appendix figure~\ref{fig:paperfolding}.
+A sample question from the test is shown in figure~\ref{fig:paperfolding}.
 It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
-% \begin{figure*}[hbt]\centering
-% \includegraphics[width=.75\linewidth]{paperfolding}
-% \caption[Paper Folding Task]{Paper Folding Task.
-Participants are instructed to pick the figure~matching the sequence of steps shown in the left-hand figure.
-Participants will complete up to 20 of these tasks in 6 minutes.}\label{fig:paperfolding}
-% \end{figure*}
+
+\begin{figure}[ht]
+  \centering
+   \subfloat[Figure Classification Task.
+Participants classify each figure in the second row as belonging to one of the groups above.
+Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
+  \includegraphics[width=.8\linewidth]{figureclassification}
+}
+\hfill
+    \subfloat[Card Rotation Task.
+Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
+Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
+\includegraphics[width=.8\linewidth]{cardrotation}
+    }
+    \hfill
+    \subfloat[Paper Folding Task.
+Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
+Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
+\includegraphics[width=.8\linewidth]{paperfolding}
+   }
+    \caption{Visuospatial tests}
+    \label{fig:tests}
+\end{figure}
+
 Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
@@
 ggplot()+
-  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
+  geom_point(data=subset(all, points), aes(x=x, y=y), shape=1) + 
   geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
@@
 PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
-More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
+% More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
 
-Incorporating the lineup task into the principal component analysis, we find the principal components to be fairly similar to the four-component analysis. 
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component.
-From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.
+<<pcasummary4,echo=F, results='asis'>>=
+pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
+print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
+@
+Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
+PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
+PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.
+
+<<pcarotation4,echo=F, results='asis'>>=
+print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+@
+
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
+Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
+\\label{fig:biplots4}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
+@
+Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
+All of the remaining principal components distinguish between the cognitive tasks; we can utilize this separation by incorporating the lineup task and examining changes in the principal components.
+
+Next, we add the lineup results to our principal components analysis.
+The results of the principal component analysis of the four cognitive tests and lineup results are similar to the results of the cognitive tests alone; this suggests that lineups require at least some of the skills measured by the other cognitive tests.
+Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
 <<pcasummary5,echo=F, results='asis'>>=
@@
 the first principal component, PC1, is again essentially an average across all tests and accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data.
-Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix~\ref{app:pca5}.
+
+<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+@
+
+Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
+Figure classification is strongly related to lineups (PC2, PC3).
+Performance on the visual search task  is also related to lineup performance (PC3).
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+
+The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+
+PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
 
@@
 While lineups do span a separate dimension, the PCA suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
-
-% 
-% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for \Sexpr{round(summary(pca)$importance[2,4]*100,1)}\% of the variance.
-% PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
-%
 
@@
 \subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous studies on different designs to investigate which plot type most effectively conveyed important characteristics of the underlying data set.
-Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
+Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
+\begin{figure*}[htbp]\centering
+\subfloat[Set 1: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}}
+\subfloat[Set 1: Density plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}}
+\subfloat[Set 1: Dotplot ]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}}
+\subfloat[Set 1: Histogram]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}}\\
+\subfloat[Set 2: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f94f6965d7}}
+\subfloat[Set 2: Boxplot with jittered points]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9e65f0e4}}
+\subfloat[Set 2: Bee swarm boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9fc58d73}}
+\subfloat[Set 2: Violin plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f93acbb7e4}}\\
+\subfloat[Set 3: QQ-plot with guide line]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13876a225d34-single}}
+\subfloat[Set 3: QQ-plot with acceptance bands]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13877c93dc78-single}}
+\subfloat[Set 3: QQ-plot rotated 45 degrees.]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
+\caption{Lineup plot types used in the experiment. \label{fig:lineups}}
+\end{figure*}
 
-<<LineupTypeSPM,echo=FALSE,include=F, eval=T, fig.width=4,fig.height=6.5, message=F, warning=F>>=
+\paragraph{Lineup Set 1}
+The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
+Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
+
+\paragraph{Lineup Set 2}
+The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
+Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
+
+\paragraph{Lineup Set 3}
+The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
+
+<<LineupTypeSPM,echo=FALSE,include=F, eval=T, fig.width=6,fig.height=9.75, message=F, warning=F>>=
 data <- lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "vis_search", "fig_class", "card_rot", "folding")]
@@
 ggplot()+
-  geom_point(data=subset(all2, points), aes(x=x, y=y)) + 
+  geom_point(data=subset(all2, points), aes(x=x, y=y), shape=1) + 
   geom_smooth(data=subset(all2, points), aes(x=x, y=y), method="lm") + 

### 2015-03-22 Paper/InfoVis2015/TestingVisualAptitude.tex
Moved appendix sections to paper

@@
 PC2 differentiates the figure classification test from the visual searching test, whereas PC3 differentiates these two tests from the paper folding test. 
-More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
+% More detailed results from the 4-test analysis are provided in Appendix~\ref{app:pca}.
 
-Incorporating the lineup task into the principal component analysis, we find the principal components to be fairly similar to the four-component analysis. 
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component.
-From the distribution of the variance components, we see that the lineup test  spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 17:04:25 2015
+% latex table generated in R 3.1.3 by xtable 1.7-1 package
+% Sun Mar 22 15:27:33 2015
+\begin{table}[htb]
+\centering
+\caption{Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\label{tab:PCAvariance4}} 
+{\footnotesize
+\begin{tabular}{rrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 \\ 
+  \hline
+Standard deviation & 1.61 & 0.81 & 0.73 & 0.49 \\ 
+  Proportion of Variance & 0.64 & 0.16 & 0.13 & 0.06 \\ 
+  Cumulative Proportion & 0.64 & 0.81 & 0.94 & 1.00 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
+
+Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
+PC1 accounts for about 60\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
+PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
+PC4 is not particularly significant (it accounts for 5.9\% of the variance), but it differentiates the card rotation task from the paper folding task.
+
+% latex table generated in R 3.1.3 by xtable 1.7-1 package
+% Sun Mar 22 15:27:33 2015
+\begin{table}[htb]
+\centering
+\caption{Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\label{tab:PCArotation4}} 
+\begin{tabular}{rrrrr}
+  \hline
+ & PC1 & PC2 & PC3 & PC4 \\ 
+  \hline
+card.rot & 0.55 & -0.19 & -0.38 & 0.72 \\ 
+  fig.class & 0.46 & 0.58 & 0.66 & 0.14 \\ 
+  folding & 0.52 & 0.33 & -0.53 & -0.59 \\ 
+  vis.search & 0.46 & -0.72 & 0.38 & -0.34 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-1} 
+\includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-2}
+
+}
+
+\caption{Biplots of principal components 1-4 with observations.
+Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
+\label{fig:biplots4}}\label{fig:biplot-pca4}
+\end{figure}
+
+\end{knitrout}
+Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
+All of the remaining principal components distinguish between the cognitive tasks; we can utilize this separation by incorporating the lineup task and examining changes in the principal components.
+
+Next, we add the lineup results to our principal components analysis.
+The results of the principal component analysis of the four cognitive tests and lineup results are similar to the results of the cognitive tests alone; this suggests that lineups require at least some of the skills measured by the other cognitive tests.
+Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
+% latex table generated in R 3.1.3 by xtable 1.7-1 package
+% Sun Mar 22 15:27:33 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 17:04:25 2015
+% latex table generated in R 3.1.3 by xtable 1.7-1 package
+% Sun Mar 22 15:27:33 2015
 \begin{table}[ht]
@@
 the first principal component, PC1, is again essentially an average across all tests and accounts for 60.1\% of the variance in the data.
-Biplots \cite{biplots:gabriel} of the remaining components are provided in Appendix~\ref{app:pca5}.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-1} 
+\includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-2}
+
+}
+
+\caption{Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\label{fig:biplots5}}\label{fig:biplots-pca5}
+\end{figure}
+
+\end{knitrout}
+
+Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
+Figure classification is strongly related to lineups (PC2, PC3).
+Performance on the visual search task  is also related to lineup performance (PC3).
+These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
+
+The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
+Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
+
+PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
+PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
 
@@
 While lineups do span a separate dimension, the PCA suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
-
-% 
-% The additional dimension, PC4 in the five-component PCA, separates lineups from the figure classification task and incorporates the relatively weak correlation between performance on the card rotation task and lineup performance; this dimension accounts for round(summary(pca)$importance[2,4]*100,1)\% of the variance.
-% PC5 is almost identical to PC4 in the four-component PCA; as before, it does not account for a significant portion of the variance in the data.
-%
 
@@
 \subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous studies on different designs to investigate which plot type most effectively conveyed important characteristics of the underlying data set.
-Examples lineups for each experiment and more detailed explanations for each of these experiments are provided in Appendix~\ref{app:lineuptypes}.
+Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
+\begin{figure*}[htbp]\centering
+\subfloat[Set 1: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}}
+\subfloat[Set 1: Density plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}}
+\subfloat[Set 1: Dotplot ]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}}
+\subfloat[Set 1: Histogram]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}}\\
+\subfloat[Set 2: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f94f6965d7}}
+\subfloat[Set 2: Boxplot with jittered points]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9e65f0e4}}
+\subfloat[Set 2: Bee swarm boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9fc58d73}}
+\subfloat[Set 2: Violin plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f93acbb7e4}}\\
+\subfloat[Set 3: QQ-plot with guide line]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13876a225d34-single}}
+\subfloat[Set 3: QQ-plot with acceptance bands]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13877c93dc78-single}}
+\subfloat[Set 3: QQ-plot rotated 45 degrees.]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
+\caption{Lineup plot types used in the experiment. \label{fig:lineups}}
+\end{figure*}
+
+\paragraph{Lineup Set 1}
+The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
+Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
+
+\paragraph{Lineup Set 2}
+The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
+Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
+
+\paragraph{Lineup Set 3}
+The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
 

### 2015-03-22 Paper/InfoVis2015/appendix.Rnw
Moved appendix sections to paper

@@
 \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
-\usepackage{sidecap}
 \usepackage{multirow}
@@
 
-\title{Spatial Reasoning and Data Displays}
+\title{Appendix: Spatial Reasoning and Data Displays}
 
@@
 \author{Susan VanderPlas, and Heike Hofmann, \textit{Member, IEEE}}
-\authorfooter{
-%% insert punctuation at end of each item
-\item
- Susan VanderPlas is a graduate student at Iowa State University. %Email: skoons@iastate.edu.
-\item
- Heike Hofmann is a professor at Iowa State University in the Departments of Statistics and Human Computer Interaction. %Email: hofmann@iastate.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Data Displays}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{
-Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
+% \authorfooter{
+% %% insert punctuation at end of each item
+% \item
+%  Susan VanderPlas is a graduate student at Iowa State University. Email: skoons@iastate.edu.
+% \item
+%  Heike Hofmann is a professor at Iowa State University in the Departments of Statistics and Human Computer Interaction. Email: hofmann@iastate.edu.
+% }
+% 
+% 
+% %other entries to be set up for journal
+% \shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Spatial Reasoning and Data Displays}
+% %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
+% 
+% 
+% %% Abstract section.
+% \abstract{
+% Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
 This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
@@
 We also examine the possibility that specific graphical tasks may be associated with certain visual skills and conclude that more research is necessary to understand which visual skills are required in order to understand certain plot types.
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Data visualization, Perception, Statistical graphics, Statistical computing.}
+% } % end of abstract
+% 
+% % Keywords that describe your work. Will show as 'Index Terms' in journal
+% % please capitalize first letter and insert punctuation after last keyword
+% \keywords{Data visualization, Perception, Statistical graphics, Statistical computing.}
 
@@
 \begin{document}
+\maketitle
+
+\appendix
+
 \setcounter{table}{0}
@@
 \renewcommand{\thetable}{A\arabic{figure}}
-<<pca,echo=F, warning=F, message=F,include=F>>=
-pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-# screeplot(pca)
-@
-<<pcasummary5,echo=F,include=F>>=
-print(xtable(summary(pca), caption="Importance of principal components, analyzing all five tests.\\label{tab:PCAvariance5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="ht", caption.placement="top")
-@
-
-<<pcarotation5,echo=F,include=F>>=
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="ht", caption.placement="top")
-@
 
@@
 @
-
-\appendix
-\section{Visuospatial Aptitude Tests}
-The Visual Search Task (VST), shown in Figure~\ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
-\begin{figure}[htp]\centering
-\includegraphics[width=.5\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task (VST).
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
-Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
-\end{figure}
-
-Figure~\ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
-All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
- \begin{figure}[ht]
-  \centering
-   \subfloat[Figure Classification Task.
-Participants classify each figure in the second row as belonging to one of the groups above.
-Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.8\linewidth]{figureclassification}
-}
-\hfill
-    \subfloat[Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.8\linewidth]{cardrotation}
-    }
-    \hfill
-    \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
-Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.8\linewidth]{paperfolding}
-   }
-    \caption{Visuospatial tests}
-    \label{fig:tests}
-\end{figure}
 
@@
 
-\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
-\subsection{Importance of principal components in an analysis of the four cognitive tests}\label{app:pca4}
-
-<<pcasummary4,echo=F, results='asis'>>=
-pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
-print(xtable(summary(pca2), caption="Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\\label{tab:PCAvariance4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
-@
-Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
-PC1 accounts for about \Sexpr{round(summary(pca2)$importance[3,1],1)*100}\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for \Sexpr{round(summary(pca2)$importance[2,4]*100,1)}\% of the variance), but it differentiates the card rotation task from the paper folding task.
-
-<<pcarotation4,echo=F, results='asis'>>=
-print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
-@
-
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
-Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
-@
-Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the remaining principal components distinguish between the cognitive tasks;
-PC2 and PC3 separate  paper folding from visual search and from the lineup and figure classification tasks, while PC4 and PC5 mainly separate lineups from card rotation and figure classification.
-This separation allows us to compare the tasks which are similar from among the principal components.
-According to Table~\ref{tab:PCArotation4}, the first three principal components account for \Sexpr{round(summary(pca2)$importance[3,3]*100,1)}\% of the variance.
-
-\subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}
-PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-Biplots of the remaining principal components are shown in Figure~\ref{fig:biplots5}.
-
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-@
-
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
-
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-
-\section{Lineup Task Examples}\label{app:lineuptypes}
-\subsection{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-The experiment was originally designed to explore the use of lineups to test plots of competing design\cite{hofmann2012graphical}.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
-\caption{Boxplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
-\caption{Density plots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
-\caption{Dotplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
-\caption{Histograms used to compare the two distributions.}
-\end{figure}
-
-\subsection{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f94f6965d7}
-\caption{Boxplots.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
-\caption{Boxplots with jittered points.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9fc58d73}
-\caption{Bee swarm boxplots.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
-\caption{Violin plots.}
-\end{figure}
-
-\subsection{Lineup Set 3}
-The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
-
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13876a225d34-single}
-\caption{QQ-plot with guide line.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
-\caption{QQ-plot with acceptance bands.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
-\caption{QQ-plot rotated 45 degrees.}
-\end{figure}
-
 \section{Lineup Plot Types} \label{app:plottypes}
@@
 
-<<lineup-types,echo=F, fig.width=4, fig.height=8, include=F, warning=F, message=F>>=
+<<lineup-types,echo=F, fig.width=8, fig.height=6, include=F, warning=F, message=F>>=
 lineup.type.sum2$plot.type <- str_replace(lineup.type.sum2$plot.type, "rotated", "rot")
@@
                                      labels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
+                                              "Scatterplot", "Density Plot", "Histogram", 
                                               "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
@@
                                      )
+lineup.type.sum2$testnum <- paste0("Set ", lineup.type.sum2$testnum)
 loadfonts()
 qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
-      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=4) + 
-  facet_wrap(~plot.type, ncol=2) + 
+      x=score, y=..scaled.., colour=factor(testnum), geom="density", cut=2, adjust=1.25) + 
+  facet_wrap(~testnum+plot.type, ncol=4) + 
   xlab("Proportion of Lineups Correct") + 
@@
   geom_vline(aes(xintercept=0), linetype=3) + 
-  scale_colour_discrete("Lineup Test Number") + 
+  scale_colour_discrete(guide="none") + 
   scale_x_continuous(breaks=c(-.33, 0, .33, .67, 1)) +
@@
 \begin{figure}[ht]\centering
-\includegraphics[width=.8\linewidth]{Figure/fig-lineup-types-1}
+\includegraphics[width=\linewidth]{Figure/fig-lineup-types-1}
 \caption{Density plots of scaled scores for different types of lineups.
@@
 
-Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
-Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
-In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
-
 <<lineup-types-scores,echo=F,fig.width=12, fig.height=5, include=F, warning=F, message=F>>=
@@
                                      levels=c("Violin Plot", "Boxplot + Points", "Bee Swarm Boxplot", "Boxplot", 
-                                              "Scatterplot", "Density Plot", "Overlaid Histogram", 
+                                              "Scatterplot", "Density Plot", "Histogram", 
                                               "QQ-plot + Line", "QQ-plot +\nAcceptance Band", 
@@
                                      labels=c("Violin Plot", "Boxplot\n + Points", "Bee Swarm \nBoxplot", "Boxplot", 
-                                              "Scatter\nPlot", "Density\nPlot", "Overlaid\nHistogram", 
+                                              "Scatter\nPlot", "Density\nPlot", "Histogram", 
                                               "QQ-plot\n+ Line", "QQ-plot\n+ Bands", 
@@
 lineup.type.sum3 <- ddply(lineup.type.sum3, .(testtype), transform, scaled.testscore=testscore/max(testscore))
-
+lineup.type.sum3$testtype <- factor(lineup.type.sum3$testtype, levels=c("vis_search", "card_rot", "fig_class", "folding"),
+                                    labels=c("vis_search", "Card Rot.", "Fig. Class.", "Paper Folding"))
 cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(scaled.testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
@@
 @
-\begin{figure*}[ht]\centering
+\begin{figure*}[htbp]\centering
 \includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}
@@
 
+Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
+Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
+In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
+
 For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$card_rot))}, the correlation between general lineup score and the figure classification test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$fig_class))}, and the correlation between lineup score and the paper folding test was \Sexpr{sprintf("%.3f", cor(ans.summary$lineup, ans.summary$folding))}.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.
@@
 
-\bibliographystyle{abbrv}
-%%use following if all content of bibtex file should be shown
-%\nocite{*}
-\bibliography{references}
+% \bibliographystyle{abbrv}
+% %%use following if all content of bibtex file should be shown
+% %\nocite{*}
+% \bibliography{references}
 

### 2015-03-22 Paper/InfoVis2015/appendix.tex
Moved appendix sections to paper

@@
 \renewcommand{\thetable}{A\arabic{figure}}
-
-\appendix
-\section{Visuospatial Aptitude Tests}
-The Visual Search Task (VST), shown in Figure~\ref{fig:VST}, is used to measure a person's ability to locate a target amid a field of distractors. 
-\begin{figure}[htp]\centering
-\includegraphics[width=.5\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task (VST).
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
-Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
-\end{figure}
-
-Figure~\ref{fig:tests} shows samples from the figure classification task, the card rotation task, and the paper folding task.
-All three tasks are part of the Kit of Factor-Referenced Cognitive Tests \cite{ekstrom1976manual}.
- \begin{figure}[ht]
-  \centering
-   \subfloat[Figure Classification Task.
-Participants classify each figure in the second row as belonging to one of the groups above.
-Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.8\linewidth]{figureclassification}
-}
-\hfill
-    \subfloat[Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.8\linewidth]{cardrotation}
-    }
-    \hfill
-    \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
-Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.8\linewidth]{paperfolding}
-   }
-    \caption{Visuospatial tests}
-    \label{fig:tests}
-\end{figure}
 
@@
 To better understand the effects of these variables, a larger study is necessary.
-% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 17:53:53 2015
+% latex table generated in R 3.1.3 by xtable 1.7-1 package
+% Sun Mar 22 15:22:39 2015
 \begin{table}[ht]
@@
 
-\section{Principal Component Analysis of Visuo-Spatial Tests}\label{app:pca}
-\subsection{Importance of principal components in an analysis of the four cognitive tests}\label{app:pca4}
-
-% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 17:53:53 2015
-\begin{table}[htb]
-\centering
-\caption{Importance of principal components in an analysis of four tests of spatial ability: figure classification, paper folding, card rotation, and visual search.\label{tab:PCAvariance4}} 
-{\footnotesize
-\begin{tabular}{rrrrr}
-  \hline
- & PC1 & PC2 & PC3 & PC4 \\ 
-  \hline
-Standard deviation & 1.61 & 0.81 & 0.73 & 0.49 \\ 
-  Proportion of Variance & 0.64 & 0.16 & 0.13 & 0.06 \\ 
-  Cumulative Proportion & 0.64 & 0.81 & 0.94 & 1.00 \\ 
-   \hline
-\end{tabular}
-}
-\end{table}
-
-Table~\ref{tab:PCAvariance4} contains the proportion of the variance in the four cognitive tasks represented by each principal component.
-PC1 accounts for about 60\% of the variance; Figure~\ref{fig:biplots4} and Table~\ref{tab:PCArotation4} confirm that PC1 is a measure of the similarity between all 4 tests; that is, a participant's general (or visual) aptitude.
-PC2 differentiates the figure classification test from the visual searching test, while PC3 differentiates these two from the paper folding test.
-PC4 is not particularly significant (it accounts for 5.9\% of the variance), but it differentiates the card rotation task from the paper folding task.
-
-% latex table generated in R 3.1.2 by xtable 1.7-1 package
-% Mon Mar 16 17:53:53 2015
-\begin{table}[htb]
-\centering
-\caption{Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\label{tab:PCArotation4}} 
-\begin{tabular}{rrrrr}
-  \hline
- & PC1 & PC2 & PC3 & PC4 \\ 
-  \hline
-card.rot & 0.55 & -0.19 & -0.38 & 0.72 \\ 
-  fig.class & 0.46 & 0.58 & 0.66 & 0.14 \\ 
-  folding & 0.52 & 0.33 & -0.53 & -0.59 \\ 
-  vis.search & 0.46 & -0.72 & 0.38 & -0.34 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-
-{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-1} 
-\includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-2}
-
-}
-
-\caption{Biplots of principal components 1-4 with observations.
-Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\label{fig:biplots4}}\label{fig:biplot-pca4}
-\end{figure}
-
-\end{knitrout}
-Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the remaining principal components distinguish between the cognitive tasks;
-PC2 and PC3 separate  paper folding from visual search and from the lineup and figure classification tasks, while PC4 and PC5 mainly separate lineups from card rotation and figure classification.
-This separation allows us to compare the tasks which are similar from among the principal components.
-According to Table~\ref{tab:PCArotation4}, the first three principal components account for 94.1\% of the variance.
-
-\subsection{PCA of Cognitive Tests and Lineups}\label{app:pca5}
-PC1 is essentially an average across all tests representing a general ``visual intelligence" factor.
-Biplots of the remaining principal components are shown in Figure~\ref{fig:biplots5}.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-
-{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-1} 
-\includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-2}
-
-}
-
-\caption{Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\label{fig:biplots5}}\label{fig:biplots-pca5}
-\end{figure}
-
-\end{knitrout}
-
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
-
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-
-\section{Lineup Task Examples}\label{app:lineuptypes}
-\subsection{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-The experiment was originally designed to explore the use of lineups to test plots of competing design\cite{hofmann2012graphical}.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}
-\caption{Boxplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}
-\caption{Density plots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}
-\caption{Dotplots used to compare the two distributions.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}
-\caption{Histograms used to compare the two distributions.}
-\end{figure}
-
-\subsection{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots.
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f94f6965d7}
-\caption{Boxplots.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9e65f0e4}
-\caption{Boxplots with jittered points.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f9fc58d73}
-\caption{Bee swarm boxplots.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups2/file147f93acbb7e4}
-\caption{Violin plots.}
-\end{figure}
-
-\subsection{Lineup Set 3}
-The final lineup section explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
-
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13876a225d34-single}
-\caption{QQ-plot with guide line.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13877c93dc78-single}
-\caption{QQ-plot with acceptance bands.}
-\end{figure}
-\begin{figure}[htbp]\centering
-\includegraphics[keepaspectratio=T,width=.8\linewidth]{../../Images/Lineups3/file13878e5c92f-single}
-\caption{QQ-plot rotated 45 degrees.}
-\end{figure}
-
 \section{Lineup Plot Types} \label{app:plottypes}
@@
 
-Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
-Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
-In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
-
-\begin{figure*}[ht]\centering
+\begin{figure*}[htbp]\centering
 \includegraphics[width=\linewidth]{Figure/fig-lineup-types-scores-1}
@@
 
+Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
+Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
+In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
+
 For comparison, the correlation between general lineup score (non-subdivided) and the card rotation test score was 0.505, the correlation between general lineup score and the figure classification test was 0.512, and the correlation between lineup score and the paper folding test was 0.471.While we can compare the correlation strength between tasks, it is clear that the correlation between the score on any single lineup type and a particular visual aptitude score is lower than the overall relationship that we attribute to visual ability.

### 2015-03-28 Paper/InfoVis2015/TestingVisualAptitude.Rnw
intro

@@
 
-% \usepackage{mathptmx}
+\usepackage{mathptmx}
 \usepackage{graphicx}
@@
 \usepackage{xr-hyper}
+\usepackage{afterpage}
+
+\usepackage{todonotes}
+\newcommand{\hhnote}[1]{\todo[inline,color=magenta!40]{#1}}
 
@@
 \item Susan VanderPlas is a graduate student at Iowa State University. Email: skoons@iastate.edu.
-\item Heike Hofmann is a professor at Iowa State University in the Departments of Statistics and Human Computer Interaction. Email: hofmann@iastate.edu.
+\item Heike Hofmann is Professor of Statistics at Iowa State University and member of the faculty in the Human Computer Interaction program. Email: hofmann@iastate.edu.
 }
@@
     \includegraphics[width=.4\linewidth]{lineup}
-    \caption[Sample Lineup]{Sample lineup of boxplots.
-Participants are instructed to choose the plot which appears most different from the others.
-In this lineup, plot 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
+    \caption[Sample Lineup]{``Which of these plots is the most different from the others?"\newline
+    This is the question participants are asked to answer for each of the lineups in the study.
+In the figure, a sample lineup of boxplots is shown.
+%Participants are instructed to choose the plot which appears most different from the others.
+    Panel 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
   }
@@
 
-``Lineups" have recently been introduced \cite{buja2009statistical, wickham2010graphical, majumder2013validation} as a tool to evaluate the statistical significance of a graphical finding.
+In this paper we investigate the extent to which observers make use of spatial ability in their evaluation of and reasoning about statistical graphics.
+As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol  \cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
+
+``Lineups" have recently been introduced as a tool to evaluate the statistical significance of a graphical finding.
 Lineups are also useful in assessing the effectiveness of different graphical displays \cite{hofmann2012graphical, loy:2015}.
@@
 \begin{eqnarray}\label{eq.scoring1}
-n_c - n_w/(m-1).
+n_c - \frac{n_w}{m-1}.
 \end{eqnarray}
@@
 % 
+\afterpage{\clearpage}
 
@@
 \end{figure*}
+\afterpage{\clearpage}
 
@@
 \end{figure}
+\afterpage{\clearpage}
 
@@
 @
+\afterpage{\clearpage}
+%
 \subsection{Lineup Types}

### 2015-03-28 Paper/InfoVis2015/TestingVisualAptitude.tex
intro

@@
 
-% latex table generated in R 3.1.3 by xtable 1.7-1 package
-% Sun Mar 22 15:27:33 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Mar 28 11:14:52 2015
 \begin{table}[htb]
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
 
@@
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\label{fig:biplots4}}\label{fig:biplot-pca4}
+\label{fig:biplots4}\label{fig:biplot-pca4}}
 \end{figure}
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.3 by xtable 1.7-1 package
-% Sun Mar 22 15:27:33 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Mar 28 11:14:52 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.3 by xtable 1.7-1 package
-% Sun Mar 22 15:27:33 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Mar 28 11:14:52 2015
 \begin{table}[ht]
@@
 PC3.
-\label{fig:biplots5}}\label{fig:biplots-pca5}
+\label{fig:biplots5}\label{fig:biplots-pca5}}
 \end{figure}
@@
 
+\afterpage{\clearpage}
+%
 \subsection{Lineup Types}

### 2015-03-29 Paper/InfoVis2015/TestingVisualAptitude.Rnw
idea for the teaser

@@
     \centering
-    \includegraphics[width=.4\linewidth]{lineup}
+\begin{minipage}[c]{.5\textwidth}
+\includegraphics[width=\textwidth]{lineup}
+\end{minipage}
+\begin{minipage}[c]{.3\textwidth}
+    \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}
+    \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-2}
+\end{minipage}
     \caption[Sample Lineup]{``Which of these plots is the most different from the others?"\newline
@@
 %Participants are instructed to choose the plot which appears most different from the others.
-    Panel 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}}
+    Panel 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}
+    The plots on the right show biplots of principal components 2-5 with observations. The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs. PC3.}
   }
@@
 
-In this paper we investigate the extent to which observers make use of spatial ability in their evaluation of and reasoning about statistical graphics.
+Here, we want to investigate the extent to which observers make use of spatial ability in their evaluation of and reasoning about statistical graphics.
 As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol  \cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=.9\linewidth]{Figure/fig-ResultsScaledScores-1}
@@
 % 
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=9,fig.height=5.5, message=F, warning=F, fig.pos='ht'>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8.75,fig.height=6, message=F, warning=F, fig.pos='ht'>>=
 loadfonts()
@@
 
-qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot", fill=I("grey")) + 
+qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="blank") +
+  geom_boxplot(fill="grey", outlier.size=1.5) + 
   facet_wrap(~variable, scales="free_x", ncol=4) + xlab("") + 
   ylab("Scaled Lineup Score") + 
-  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
+  geom_point(aes(x=value, y=lineup), shape=1, size=2.5) + 
   theme_bw() + 
@@
 \end{figure*}
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
@@
 \begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-VisReasoningSPM-1}
+\centering
+\includegraphics[width=.9\linewidth]{Figure/fig-VisReasoningSPM-1}
 \caption{Pairwise scatterplots of test scores.
@@
 \end{figure}
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
@@
 @
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 %
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=\linewidth]{Figure/fig-LineupTypeSPM-1}

### 2015-03-29 Paper/InfoVis2015/TestingVisualAptitude.tex
idea for the teaser

@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 11:14:52 2015
+% Sat Mar 28 22:35:16 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 11:14:52 2015
+% Sat Mar 28 22:35:16 2015
 \begin{table}[ht]
@@
 
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 %

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
ggbiplot attempt

@@
 
+<<ggbiplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+ggbiplot <- function(pca, choices) {
+  u <- pca$x[,choices]
+  v <- pca$rotation[,choices]
+  u <- data.frame(sweep(u, MARGIN=2, 1/pca$sdev[choices], FUN="*"))
+  v <- data.frame(sweep(v, MARGIN=2, 1/sqrt(rowSums(v^2))[choices], FUN="*"))
+  v$label <- row.names(v)
+
+  qplot(u[,1], u[,2], data=u, colour=I("grey50"), shape=I(1), size=I(3)) + theme_bw() + theme(aspect.ratio=1) + coord_equal() + 
+  geom_segment(aes(x=0, y=0, xend=v[,1], yend=v[,2]), data=v, arrow=arrow(length = unit(0.3, "cm"))) + 
+    xlab(sprintf("PC%d",choices[1])) + ylab(sprintf("PC%d",choices[2])) +
+    geom_text(aes(x=v[,1], y=v[,2], label=label, hjust=0.5*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v)
+}
+
+ggbiplot(pca, choices=2:3) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
+ggbiplot(pca, choices=3:4)
+@
 Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
color for segments

--- 2015-03-29 Paper/InfoVis2015/TestingVisualAptitude.Rnw
@@
 
+<<ggbiplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+ggbiplot <- function(pca, choices, segColor="black") {
+  u <- pca$x[,choices]
+  v <- pca$rotation[,choices]
+  u <- data.frame(sweep(u, MARGIN=2, 1/pca$sdev[choices], FUN="*"))
+  v <- data.frame(sweep(v, MARGIN=2, 1/sqrt(rowSums(v^2))[choices], FUN="*"))
+  v$label <- row.names(v)
+
+  segColor <- rep(segColor, length=nrow(v))
+
+  qplot(u[,1], u[,2], data=u, colour=I("grey50"), shape=I(1), size=I(3)) + theme_bw() + 
+    theme(aspect.ratio=1) + coord_equal() + 
+  geom_segment(aes(x=0, y=0, xend=v[,1], yend=v[,2], colour=segColor), data=v, arrow=arrow(length = unit(0.3, "cm"))) + 
+    xlab(sprintf("PC%d",choices[1])) + ylab(sprintf("PC%d",choices[2])) +
+    geom_text(aes(x=v[,1], y=v[,2], label=label, colour=segColor, 
+                  hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
+    scale_colour_identity()
+}
+
+dimnames(pca$rotation)[[1]] <- c("lineup", "card rotation", "figure classification", "folding", "visual search")
+ggbiplot(pca, choices=2:3, segColor=c("darkred", rep("black", 4))) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
+ggbiplot(pca, choices=3:4, segColor=c("darkred", rep("black", 4)))
+@
 Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
switched out biplot by ggbiplot

--- 2015-03-29 Paper/InfoVis2015/TestingVisualAptitude.Rnw
@@
 
+ggbiplot <- function(pca, choices, segColor="black") {
+  u <- pca$x[,choices]
+  v <- pca$rotation[,choices]
+  u <- data.frame(sweep(u, MARGIN=2, 1/pca$sdev[choices], FUN="*"))
+  v <- data.frame(sweep(v, MARGIN=2, 1/sqrt(rowSums(v^2))[choices], FUN="*"))
+  v$label <- row.names(v)
+
+  segColor <- rep(segColor, length=nrow(v))
+
+  qplot(u[,1], u[,2], data=u, colour=I("grey50"), shape=I(1), size=I(3)) + theme_bw() + 
+    theme(aspect.ratio=1) + coord_equal() + 
+  geom_segment(aes(x=0, y=0, xend=v[,1], yend=v[,2], colour=segColor), data=v, arrow=arrow(length = unit(0.3, "cm"))) + 
+    xlab(sprintf("PC%d",choices[1])) + ylab(sprintf("PC%d",choices[2])) +
+    geom_text(aes(x=v[,1], y=v[,2], label=label, colour=segColor, 
+                  hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
+    scale_colour_identity()
+}
 @
@@
 \\label{fig:biplots4}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca2, choices=1:2, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca2, choices=3:4, pc.biplot=T, cex=c(.5, 1), xlim=c(-3,3), ylim=c(-2.5,2.5))
+#par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+dimnames(pca2$rotation)[[1]] <- c("card rotation", "figure classification", "paper folding", "visual search")
+
+ggbiplot(pca2, choices=1:2) + xlim(c(-2.5,2.5)) +  ylim(c(-2.5,2.5))
+ggbiplot(pca2, choices=3:4) + xlim(c(-3,3)) + ylim(c(-2.5,2.5))
 @
@@
 
+% <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+% par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+% biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+% biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+% @
+
 <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
@@
 \\label{fig:biplots5}", echo=F, fig.env="figure">>=
-par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+
+dimnames(pca$rotation)[[1]] <- c("lineup", "card rotation", "figure classification", "folding", "visual search")
+ggbiplot(pca, choices=2:3, segColor=c("red", rep("black", 4))) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
+ggbiplot(pca, choices=4:5, segColor=c("#FF3333", rep("black", 4)))
 @
-
 Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.tex
switched out biplot by ggbiplot

@@
-\documentclass[journal]{vgtc}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}                % final (journal style)
+\documentclass[journal]{vgtc}                % final (journal style)
 %\documentclass[review,journal]{vgtc}         % review (journal style)
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
+% Mon Mar 30 14:32:39 2015
 \begin{table}[htb]
@@
 \end{table}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-1} 
-\includegraphics[width=.45\linewidth]{Figure/fig-biplot-pca4-2}
-
-}
-
-\caption{Biplots of principal components 1-4 with observations.
-Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\label{fig:biplots4}\label{fig:biplot-pca4}}
-\end{figure}
-
-\end{knitrout}
-Figure~\ref{fig:biplots4} shows that the first PC does not differentiate between any of the tasks; it might be best understood as a general aptitude factor.
-All of the remaining principal components distinguish between the cognitive tasks; we can utilize this separation by incorporating the lineup task and examining changes in the principal components.
-
-Next, we add the lineup results to our principal components analysis.
-The results of the principal component analysis of the four cognitive tests and lineup results are similar to the results of the cognitive tests alone; this suggests that lineups require at least some of the skills measured by the other cognitive tests.
-Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
-\begin{table}[ht]
-\centering
-\caption{Importance of principal components, analyzing all five tests.\label{tab:PCAvariance5}} 
-{\footnotesize
-\begin{tabular}{rrrrrr}
-  \hline
- & PC1 & PC2 & PC3 & PC4 & PC5 \\ 
-  \hline
-Standard deviation & 1.73 & 0.84 & 0.75 & 0.70 & 0.48 \\ 
-  Proportion of Variance & 0.60 & 0.14 & 0.11 & 0.10 & 0.05 \\ 
-  Cumulative Proportion & 0.60 & 0.74 & 0.85 & 0.95 & 1.00 \\ 
-   \hline
-\end{tabular}
-}
-\end{table}
-
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
-\begin{table}[ht]
-\centering
-\caption{PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \label{tab:PCArotation5}} 
-\begin{tabular}{rrrrrr}
-  \hline
- & PC1 & PC2 & PC3 & PC4 & PC5 \\ 
-  \hline
-lineup & 0.42 & 0.49 & -0.46 & 0.60 & -0.10 \\ 
-  card.rot & 0.50 & -0.30 & 0.28 & 0.23 & 0.73 \\ 
-  fig.class & 0.43 & 0.45 & -0.15 & -0.75 & 0.18 \\ 
-  folding & 0.47 & 0.07 & 0.68 & 0.04 & -0.56 \\ 
-  vis.search & 0.41 & -0.69 & -0.48 & -0.15 & -0.33 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-From the rotation matrix (see Table~\ref{tab:PCArotation5}) we see that  
-the first principal component, PC1, is again essentially an average across all tests and accounts for 60.1\% of the variance in the data.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-{\centering \includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-1} 
-\includegraphics[width=.45\linewidth]{Figure/fig-biplots-pca5-2}
-
-}
-
-\caption{Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\label{fig:biplots5}\label{fig:biplots-pca5}}
-\end{figure}
-
-\end{knitrout}
-
-Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
-Figure classification is strongly related to lineups (PC2, PC3).
-Performance on the visual search task  is also related to lineup performance (PC3).
-These two components highlight the shared demands of the lineup task and the figure classification task: participants must establish categories from provided stimuli and then classify the stimuli accordingly.
-
-The visual search task is also clearly important to lineup performance: PC3 captures the similarity between the visual search and lineup performance, and aspects of these tasks are negatively correlated with aspects of the paper folding and card rotation tasks within PC3.
-Paper folding does not seem to be strongly associated with lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.
-
-PC4 captures the similarity between lineups and the card rotation task and separates this similarity from the figure classification task; this similarity does not account for much extra variance (10\%), but it may be that only some lineups require spatial rotation skills.
-PC5 contains only 5\% of the remaining variance, and is thus not of much interest, however, it seems to capture the relationship between the card rotation task and the paper folding and visual search tasks.
-
-Figure classification is strongly related to lineups, and as in the four-component PCA, figure classification is strongly represented in the first two principal components.
-While lineups do span a separate dimension, the PCA suggests that they are most closely related to the figure classification task, and least related to the visual searching task.
-
-This emphasizes the underpinnings of lineups: the test utilizes a visual medium, but is ultimately are a classification task presented in a graphical manner.
-Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.
-
-\subsection{Linear model of demographic factors}
-Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.
-Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
-Only one student is doing research who has not taken a statistics course.
-
-A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research.
-We therefore decided to use sums of these variables to come up with a separate math and a stats score.
-Note, that the correlation between the math and the stats score is almost zero.
-
-We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests, selecting the best model using AIC and stepwise backwards selection.
-The result is shown in Table~\ref{tab:m1}.
-Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores.
-We can think of PC1 as a measure of innate visual or intellectual ability, while the MATH score is a matter of both ability and training.
-The remaining principal components were not sufficiently associated with lineup score to be included in the model.
-
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
-\begin{table}[ht]
-\centering
-\caption{Estimates and significances of a linear model of lineup scores.} 
-\label{tab:m1}
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
-  \hline
-(Intercept) & 14.1192 & 1.9149 & 7.37 & 0.0000 \\ 
-  PC1 & 1.7672 & 0.5230 & 3.38 & 0.0018 \\ 
-  MATH & 2.1246 & 0.8732 & 2.43 & 0.0202 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-%\afterpage{\clearpage}
-%
-\subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
-\begin{figure*}[htbp]\centering
-\subfloat[Set 1: Boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}}
-\subfloat[Set 1: Density plot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}}
-\subfloat[Set 1: Dotplot ]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}}
-\subfloat[Set 1: Histogram]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}}\\
-\subfloat[Set 2: Boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f94f6965d7}}
-\subfloat[Set 2: Boxplot with jittered points]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9e65f0e4}}
-\subfloat[Set 2: Bee swarm boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9fc58d73}}
-\subfloat[Set 2: Violin plot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f93acbb7e4}}\\
-\subfloat[Set 3: QQ-plot with guide line]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13876a225d34-single}}
-\subfloat[Set 3: QQ-plot with acceptance bands]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13877c93dc78-single}}
-\subfloat[Set 3: QQ-plot rotated 45 degrees.]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
-\caption{Lineup plot types used in the experiment. \label{fig:lineups}}
-\end{figure*}
-
-\paragraph{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
-
-\paragraph{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-
-\paragraph{Lineup Set 3}
-The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\linewidth]{Figure/fig-LineupTypeSPM-1}
-\caption{Pairwise scatterplots of test scores, with lineups separated into the three lineup tasks.
-All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
-\label{fig:lineuptaskmatrix}}
-\end{figure}
-Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks.
-The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
-Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
-
-In order to examine which lineup tasks are most closely associated with visual abilities tested in the aptitude portions of an experiment, we employ principal component analysis on participant scores averaged across each block of lineups.
-
-Principal component analysis separates multivariate data into orthogonal components using a rotation matrix to transform correlated input data into an orthogonal space.
-The importance of a principal component is also evaluated to assess the proportion of the overall variance in the data contained within the component (Table \ref{tab:PCAvariance5} shows the importance of each PC for the analysis of the aggregate lineup score and the four aptitude tests).
-
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{Figure/fig-lineupblockpca-1}
-\caption{Principal Component influence plot, showing each input variable's contribution to the principal component, scaled by the proportion of variability in the data contained in the principal component.\label{fig:influence}}
-\end{figure}
-
-For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
-
-The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
-
-The influence of variable $X_i$ on component $PC_j$ is then defined as:
-%
-$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
-%
-Influence is therefore a measure of the contribution of an input variable to the principal component, scaled by the importance of that principal component to the overall variability in the data.
-Small (absolute) values indicate that there is little influence of $X_i$ on $PC_j$; negative values show the direction of the influence (to maintain the separation of variables as part of principal components analysis).
-
-Figure~\ref{fig:influence} shows the influence of each input variable on each principal component for a principal component analysis  including each lineup task as a separate input variable.
-The input variables are shown on the $y$ axis, with the influence of the variable on each PC shown on the $x$ axis.
-This allows us to consider the rotation matrix visually (while accounting for the importance of each principal component); for instance, we see that again, the first PC accounts for most of the variance and seems to represent general visual aptitude.
-
-PC2 emphasizes the overlapping variation in performance on lineup test 1 and the figure classification test.
-PC3 emphasizes the additional variation in performance on lineup test 3 and the visual search test, while PC4 emphasizes the extra variability in performance on lineup test 2 and the paper folding test.
-All three lineups, plus the figure classification, paper folding, and visual search tests contribute to PC5.
-PC6 and PC7 jointly account for less than 10\% of the variance in the data and do not display any distinct patterns in the loadings.
-
-While lineups constitute a distinct principal component when aggregated into a single score, this PCA of the separate lineup types and the cognitive tests indicates that different lineup experiments exist in different principal component loadings.
-Overall, there is an additional principal component gained from separating the lineup blocks by experiment type.
-As lineup tasks 1 and 2 contained similar plot types, it is possible that those two tasks overlap in the component space while lineup task 3 is distinct.
-
-The relationship between participant performance on different types of lineups (and different types of plots) and performance on tests of spatial ability bears further investigation; this study suggests that there may be an effect, but there is simply not enough variation to make definitive conclusions about the relationship between different measures of visual aptitude and performance on specific lineup tasks.
-
-A larger study might not only address the question of which lineup tasks require certain visual skills, but also the use of different types of plots from a perceptual perspective.
-Preliminary results of performance on different types of plots are shown in Appendix~\ref{app:plottypes}, but a larger study is needed for definitive results.
-The advantage of the lineup protocol is that it allows us to not only consider individual performance but also to compare aggregate performance on different types of plots.
-Integrating information about the visual skills required for each type of plot provides information about the underlying perceptual skills and experience required to read different types of plots.
-
-\section{Discussion and Conclusions}\label{sec:discussion}
-
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-
-Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
-This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
-
-When lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
-Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).
-Future studies testing larger sets of lineups may be useful to understand which types of plots require additional visuospatial skills, as plots which appeal to a wider audience may be more successful when conveying important information.
-
-In addition to this theoretical information, the figure classification test may be useful for pre-screening participants in future online lineup studies.
-Such studies often suffer from participants who do not take the task seriously, and internal verification questions, as well as pre-qualification tasks are often used to reduce extraneous variability.
-While it is impractical to require participants to score well on several different tests, it is reasonable to ask participants to pre-qualify for a task by completing a figure classification test.
-As the figure classification test is different from the lineup task, this will not bias participants' scores on the domain of interest while ensuring that the participant pool is sufficiently motivated to complete the lineup questions.
-
-The demographic results from this study indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
-
-All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
-
-\bibliographystyle{abbrv}
-%%use following if all content of bibtex file should be shown
-%\nocite{*}
-\bibliography{references}
-
-\end{document}

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
Caption fixes

@@
 
-Figure~\ref{fig:lineup} shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
 In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
@@
 
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations.
+<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations (same as figure 1(b)).
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.Rnw
scaling in biplot

@@
   v <- pca$rotation[,choices]
-  u <- data.frame(sweep(u, MARGIN=2, 1/pca$sdev[choices], FUN="*"))
-  v <- data.frame(sweep(v, MARGIN=2, 1/sqrt(rowSums(v^2))[choices], FUN="*"))
+  n <- nrow(u)
+  u <- data.frame(sweep(u, MARGIN=2, 1/(pca$sdev[choices]), FUN="*"))
+  v <- data.frame(sweep(v, MARGIN=2, 3 * (pca$sdev[choices]), FUN="*"))
   v$label <- row.names(v)
@@
                   hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
-    scale_colour_identity()
+    scale_colour_identity() + theme(aspect.ratio=1)
 }

### 2015-03-30 Paper/InfoVis2015/TestingVisualAptitude.tex
scaling in biplot

--- 2015-03-29 Paper/InfoVis2015/TestingVisualAptitude.tex
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
+% Mon Mar 30 14:37:38 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
+% Mon Mar 30 14:37:38 2015
 \begin{table}[ht]
@@
 
+% <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
+PC3.
+\\label{fig:biplots5}", echo=F, fig.env="figure">>=
+% par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
+% biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
+% biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
+% @
+
 \begin{knitrout}
@@
 \end{knitrout}
-
 Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Mar 28 22:35:16 2015
+% Mon Mar 30 14:37:39 2015
 \begin{table}[ht]

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.Rnw
fiddling with the teaser image

@@
 % declare the path(s) where your graphic files are
-\graphicspath{{Figure/}{Images/}{../../Images/}}
+\graphicspath{{Figure/}{Images/}}
 % and their extensions so you won't have to specify these with
@@
       \centering
-\begin{minipage}[c]{.5\textwidth}
-\includegraphics[width=\textwidth]{lineup}
+      \begin{minipage}[c]{0.4\textwidth}
+         \subfloat[][``Which of these plots is the most different from the others?"]{\label{fig:lineup} \includegraphics[height=\textwidth]{lineup.png}}
+        %
         \end{minipage}
-\begin{minipage}[c]{.3\textwidth}
-    \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}
+        \begin{minipage}[c]{.4\textwidth}
+         \subfloat[][Biplots]{\label{fig:biplots}
+          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
           \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-2}
+         }
           \end{minipage}
-    \caption[Sample Lineup]{``Which of these plots is the most different from the others?"\newline
-    This is the question participants are asked to answer for each of the lineups in the study.
-In the figure, a sample lineup of boxplots is shown.
+    \caption{On the left a sample lineup of boxplots is shown.
+Participants are asked to identify the most different plot.
 %Participants are instructed to choose the plot which appears most different from the others.
-    Panel 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}
+    Here, panel 4 is the target plot, because the two groups have a large difference in medians.\\
     The plots on the right show biplots of principal components 2-5 with observations. The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs. PC3.}
@@
 dimnames(pca$rotation)[[1]] <- c("lineup", "card rotation", "figure classification", "folding", "visual search")
-ggbiplot(pca, choices=2:3, segColor=c("red", rep("black", 4))) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
-ggbiplot(pca, choices=4:5, segColor=c("#FF3333", rep("black", 4)))
+ggbiplot(pca, choices=2:3, segColor=c("#CC3333", rep("black", 4))) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
+ggbiplot(pca, choices=4:5, segColor=c("#CC3333", rep("black", 4)))
 @

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.tex
fiddling with the teaser image

@@
       \centering
-\begin{minipage}[c]{.5\textwidth}
-\includegraphics[width=\textwidth]{lineup}
+      \begin{minipage}[c]{0.4\textwidth}
+         \subfloat[][``Which of these plots is the most different from the others?"]{\label{fig:lineup} \includegraphics[height=\textwidth]{lineup.png}}
+        %
         \end{minipage}
-\begin{minipage}[c]{.3\textwidth}
-    \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}
+        \begin{minipage}[c]{.1\textwidth}
+         \subfloat[][Biplots]{\label{fig:biplots}
+          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
           \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-2}
+         }
           \end{minipage}
-    \caption[Sample Lineup]{``Which of these plots is the most different from the others?"\newline
-    This is the question participants are asked to answer for each of the lineups in the study.
-In the figure, a sample lineup of boxplots is shown.
+    \caption{On the left a sample lineup of boxplots is shown.
+Participants are asked to identify the most different plot.
 %Participants are instructed to choose the plot which appears most different from the others.
-    Panel 4 is the target plot, because the two groups have a large difference in medians.\label{fig:lineup}
+    Here, panel 4 is the target plot, because the two groups have a large difference in medians.\\
     The plots on the right show biplots of principal components 2-5 with observations. The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs. PC3.}
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 14:37:38 2015
+% Mon Mar 30 19:13:11 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 14:37:38 2015
+% Mon Mar 30 19:13:11 2015
 \begin{table}[ht]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 14:37:39 2015
+% Mon Mar 30 19:13:12 2015
 \begin{table}[ht]

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.Rnw
labels fixed

@@
       \centering
+      \hfill
       \begin{minipage}[c]{0.4\textwidth}
-         \subfloat[][``Which of these plots is the most different from the others?"]{\label{fig:lineup} \includegraphics[height=\textwidth]{lineup.png}}
+          \includegraphics[height=\textwidth]{lineup.png}
+          (a) ``Which of these plots is the most different from the others?"
         %
         \end{minipage}
+      %  \hfill
         \begin{minipage}[c]{.4\textwidth}
-         \subfloat[][Biplots]{\label{fig:biplots}
-          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
-          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-2}
-         }
+        \centering
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-2}\hfill\newline
+
+        (b) Biplots of visuo-spatial tests and lineup task.
 \end{minipage}
-    \caption{On the left a sample lineup of boxplots is shown.
+\hfill
+    \caption{\label{fig:lineup}On the left a sample lineup of boxplots is shown.
 Participants are asked to identify the most different plot.
 %Participants are instructed to choose the plot which appears most different from the others.
-    Here, panel 4 is the target plot, because the two groups have a large difference in medians.\\
-    The plots on the right show biplots of principal components 2-5 with observations. The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs. PC3.}
+    Panel 4 is the target plot, because the two groups have a large difference in medians.\\
+    The biplots on the right show principal components 2-5 with observations from the user study. The lineup task appears to be most similar to the figure classification task.}
   }
@@
     geom_text(aes(x=v[,1], y=v[,2], label=label, colour=segColor, 
-                  hjust=0.15*(0.5-sign(v[,1])), vjust=0.5-sign(v[,2])), data=v) +
+                  hjust=(0.5-sign(v[,1])/2), vjust=(0.5-sign(v[,2])/2) ), data=v) +
     scale_colour_identity() + theme(aspect.ratio=1)
@@
 Participants were given 5 minutes to complete each block of 20 lineups.
-Figure~\ref{fig:lineup} shows a sample lineup of box plots.
+Figure~\ref{fig:lineup}(a) shows a sample lineup of box plots.
 
@@
 
-<<biplot-pca4, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations (same as figure 1(b)).
+<<biplot-pca4, fig.cap="", fig.width=5.5, fig.height=5.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 1-4 with observations (same as figure 1(b)).
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
@@
 #par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-dimnames(pca2$rotation)[[1]] <- c("card rotation", "figure classification", "paper folding", "visual search")
+dimnames(pca2$rotation)[[1]] <- c("card\nrotation", "figure\nclassification", "paper\nfolding", "visual\nsearch")
 
-ggbiplot(pca2, choices=1:2) + xlim(c(-2.5,2.5)) +  ylim(c(-2.5,2.5))
-ggbiplot(pca2, choices=3:4) + xlim(c(-3,3)) + ylim(c(-2.5,2.5))
+ggbiplot(pca2, choices=1:2) + xlim(c(-2.5,4.1)) +  ylim(c(-2.5,2.5))
+ggbiplot(pca2, choices=3:4) + xlim(c(-3,3.5)) + ylim(c(-3,3))
 @
@@
 
-% <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
+<<biplots-pca5, fig.cap="", fig.width=5.5, fig.height=5.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
 The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
 PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-% par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-% biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-% biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-% @
+(These are the same biplots as shown in figure \\ref{fig:lineup}b)  \\label{fig:biplots5}", echo=F, fig.env="figure">>=
 
-<<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-
-dimnames(pca$rotation)[[1]] <- c("lineup", "card rotation", "figure classification", "folding", "visual search")
-ggbiplot(pca, choices=2:3, segColor=c("#CC3333", rep("black", 4))) + xlim(c(-2.6,2.6)) +  ylim(c(-2.6,2.6))
-ggbiplot(pca, choices=4:5, segColor=c("#CC3333", rep("black", 4)))
+dimnames(pca$rotation)[[1]] <- c("lineup", "card\nrotation", "figure\nclassification", "paper\nfolding", "visual\nsearch")
+ggbiplot(pca, choices=2:3, segColor=c("#CC3333", rep("black", 4))) + xlim(c(-2.4,2.8)) +  ylim(c(-2.6,2.6))
+ggbiplot(pca, choices=4:5, segColor=c("#CC3333", rep("black", 4))) + xlim(c(-2.8,2.8)) + ylim(c(-2.8,2.8))
 @

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.tex
labels fixed

@@
       \centering
+      \hfill
       \begin{minipage}[c]{0.4\textwidth}
-         \subfloat[][``Which of these plots is the most different from the others?"]{\label{fig:lineup} \includegraphics[height=\textwidth]{lineup.png}}
+          \includegraphics[height=\textwidth]{lineup.png}
+          (a) ``Which of these plots is the most different from the others?"
         %
         \end{minipage}
-        \begin{minipage}[c]{.1\textwidth}
-         \subfloat[][Biplots]{\label{fig:biplots}
-          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
-          \includegraphics[height=.8\textwidth]{Figure/fig-biplots-pca5-2}
-         }
+      %  \hfill
+        \begin{minipage}[c]{.4\textwidth}
+        \centering
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-2}\hfill\newline
+
+        (b) Biplots of visuo-spatial tests and lineup task.
 \end{minipage}
-    \caption{On the left a sample lineup of boxplots is shown.
+\hfill
+    \caption{\label{fig:lineup}On the left a sample lineup of boxplots is shown.
 Participants are asked to identify the most different plot.
 %Participants are instructed to choose the plot which appears most different from the others.
-    Here, panel 4 is the target plot, because the two groups have a large difference in medians.\\
-    The plots on the right show biplots of principal components 2-5 with observations. The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs. PC3.}
+    Panel 4 is the target plot, because the two groups have a large difference in medians.\\
+    The biplots on the right show principal components 2-5 with observations from the user study. The lineup task appears to be most similar to the figure classification task.}
   }
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 19:13:11 2015
+% Mon Mar 30 23:08:34 2015
 \begin{table}[htb]
@@
 
-\caption{Biplots of principal components 1-4 with observations.
+\caption{Biplots of principal components 1-4 with observations (same as figure 1(b)).
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 19:13:11 2015
+% Mon Mar 30 23:08:35 2015
 \begin{table}[ht]
@@
 
-% <<biplots-pca5, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Biplots of principal components 2-5 with observations.
-The lineup task appears to be most similar to the figure classification task, based on the plot of PC2 vs.
-PC3.
-\\label{fig:biplots5}", echo=F, fig.env="figure">>=
-% par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
-% biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), xlim=c(-2.5,2.5), ylim=c(-2.5,2.5))
-% biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), xlim=c(-2,2), ylim=c(-2,2))
-% @
-
 \begin{knitrout}
@@
 PC3.
-\label{fig:biplots5}\label{fig:biplots-pca5}}
+(These are the same biplots as shown in figure \ref{fig:lineup}b)  \label{fig:biplots5}\label{fig:biplots-pca5}}
 \end{figure}
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 19:13:12 2015
+% Mon Mar 30 23:08:36 2015
 \begin{table}[ht]

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.Rnw
Updated compiled version

@@
       \centering
-      \hfill
       \begin{minipage}[c]{0.4\textwidth}
@@
         \centering
-          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-1}\hfill\newline
-          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-2}\hfill\newline
-
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-1}\\
+          \includegraphics[height=.5\textwidth]{Figure/fig-biplots-pca5-2}\\
         (b) Biplots of visuo-spatial tests and lineup task.
         \end{minipage}
-\hfill
-    \caption{\label{fig:lineup}On the left a sample lineup of boxplots is shown.
+    \caption{\protect\label{fig:lineup}On the left a sample lineup of boxplots is shown.
 Participants are asked to identify the most different plot.
 %Participants are instructed to choose the plot which appears most different from the others.
-    Panel 4 is the target plot, because the two groups have a large difference in medians.\\
+    Panel 4 is the target plot, because the two groups have a large difference in medians.
     The biplots on the right show principal components 2-5 with observations from the user study. The lineup task appears to be most similar to the figure classification task.}

### 2015-03-31 Paper/InfoVis2015/TestingVisualAptitude.tex
Updated compiled version

@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 23:08:34 2015
+% latex table generated in R 3.1.3 by xtable 1.7-3 package
+% Tue Mar 31 10:05:43 2015
 \begin{table}[htb]
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
 
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 23:08:35 2015
+% latex table generated in R 3.1.3 by xtable 1.7-3 package
+% Tue Mar 31 10:05:43 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 23:08:35 2015
+% latex table generated in R 3.1.3 by xtable 1.7-3 package
+% Tue Mar 31 10:05:43 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Mar 30 23:08:36 2015
+% latex table generated in R 3.1.3 by xtable 1.7-3 package
+% Tue Mar 31 10:05:44 2015
 \begin{table}[ht]

### 2015-04-03 Paper/InfoVis2015/TestingVisualAptitude.tex
merge

@@
 \newcommand{\hhnote}[1]{\todo[inline,color=magenta!40]{#1}}
-
+\newcommand{\newtext}[1]{\todo[inline,color=green!40]{#1}}
 %---------------------------------------------------
@@
 In a lineup test, the plot of the data is placed randomly among a set of, generally 19, distractor plots (or {\it null plots}) that are rendered from data generated by a model, without a signal (a null model).
-This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different".
+This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different"
+\newtext{(this phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies \cite{wickham2010graphical, majumder2013validation}).}
 If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
@@
 
-One of the most useful features of the lineup protocol is that it allows researchers to conclusively determine which graphics show certain features more conclusively by providing an experimental protocol for comparing graphics based on the accuracy of user conclusions.
+One of the most useful features of the lineup protocol is that it allows researchers to determine which visualizations show certain features more conclusively by providing an experimental protocol for comparing different visualization methods based on the accuracy of user conclusions.
 In addition, lineups provide researchers with a rigorous framework for determining whether a specific graph shows a real, statistically significant effect by comparing a target plot with plots formed using permutations of the same data, providing a randomization test protocol for graphics.
 As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.
+\newtext{This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization. }
 
@@
 Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
-What is of more interest to us is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
-In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests using the same procedure.
+\newtext{While test scores are correlated, the relationships between test scores appear to be linear, with sufficient variability around the trend line to merit further investigation.
+The correlations are to be expected, as all of the tests in this experiment are conducted in a visual domain, and require general reasoning, intelligence, and other factors.
+}
+What is of more interest to us is 
+\newtext{whether lineups recruit specific cognitive abilities, such as visual search, reasoning and categorization ability, mental rotation.} 
+In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests (four cognitive tests, and additionally lineups) using the same procedure.
 
-A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about 64\% of the total variability.
+\newtext{
+Principal component analysis (PCA) is a statistical technique to transform the coordinate system of correlated data, producing orthogonal components in a new coordinate space (see \cite{pcaReference} for more information).
+The most stringent assumption required for PCA is that of linearity, which is reasonable given Figure \ref{fig:scatterplotmatrix}.
+Additionally, PCA assumes that the interesting portion of the data is the variance between any two variables (a perfectly linear relationship between two variables would be uninteresting, as the variables contain the same information).
+As a result, PCA can be used to examine which variables contain similar information (i.e.
+which variables are represented in each principal component).
+It is this feature of principal components which we utilize in this paper.
+}
+
+A principal component analysis of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about 64\% of the total variability.
 Principal components (PC-) are ordered by importance (how much variability in the data they contain), and each principal component is uncorrelated with every other principal component.
@@
 
-% latex table generated in R 3.1.3 by xtable 1.7-3 package
-% Tue Mar 31 10:05:43 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Tue Jun  9 10:29:29 2015
 \begin{table}[htb]
@@
 Principal component analysis was performed on the four cognitive tests used to understand the association between the cognitive skills required for these tests and the skills required for the lineup protocol.
-\label{fig:biplots4}\label{fig:biplot-pca4}}
+\label{fig:biplots4}}\label{fig:biplot-pca4}
 \end{figure}
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.3 by xtable 1.7-3 package
-% Tue Mar 31 10:05:43 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Tue Jun  9 10:29:30 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.3 by xtable 1.7-3 package
-% Tue Mar 31 10:05:43 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Tue Jun  9 10:29:30 2015
 \begin{table}[ht]
@@
 PC3.
-(These are the same biplots as shown in figure \ref{fig:lineup}b)  \label{fig:biplots5}\label{fig:biplots-pca5}}
+(These are the same biplots as shown in figure \ref{fig:lineup}b)  \label{fig:biplots5}}\label{fig:biplots-pca5}
 \end{figure}
@@
 
-% latex table generated in R 3.1.3 by xtable 1.7-3 package
-% Tue Mar 31 10:05:44 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Tue Jun  9 10:29:30 2015
 \begin{table}[ht]
@@
 All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
-\label{fig:lineuptaskmatrix}}
+\label{fig:lineuptaskmatrix} \newtext{While the correlations shown in this graph are all less than 0.60, there are some surprisingly strong relationships for human subjects data.}}
 \end{figure}

### 2015-06-08 Paper/InfoVis2015/ReviewerResponses.Rmd
Reviewer responses and slightly modified regex in paper

---
title: "ReviewerResponses"
author: "Susan VanderPlas"
date: "June 8, 2015"
output: html_document
---

# Summary Rating

   4  (Accept) The paper should be accepted with some minor revisions. Once these have been completed it will meet the quality standard.

## The Summary Review

### Overview

   The paper attempts to solve a quite interesting problem, which is
   understanding the relationship between demographic characteristics /
   visual skills and perception of graphical lineups. Reviewers agree that
   this topic is quite interesting, and the authors did conduct a quite
   extensive study. R4, particularly, is strongly supportive for this paper,
   saying "The work represented here may be the most important recent
   advance in methodology for evaluating the effectiveness of visual
   displays." However, other reviewers listed several concerns in
   experimental design, data analysis, and conclusions.

### Strengths

   - The topic of this study is very interesting. (R3, R4)
   - The paper is well presented. (R1, R2, R3, R4)
   - The study is extensively evaluated. (R1)
   - Supplementary materials provide additional details on methodologies
   used. (R2)
   - A groundbreaking methodological contribution is made. (R4)
   - Deal with important issues in spatial reasoning on graphics. (R1)

### Weaknesses

   - A clearer motivation of certain experimental design choices is
   necessary. (R1)
   - Experiment procedures should have more proper control and be more
   detailed. (R1, R3)
   - The statistical analysis is of questionable validity. (R2, R3)
   - Results are not conclusive and open to interpretation. (R1, R2)
   - A lack of discussion on visualization design in general. (R3)
   - A separate related work section is necessary. (R3)

   However, these concerns, especially regarding statistical analysis, are
   rather critical. The reviewers recommended to accept this paper in the
   condition that these weaknesses are properly handled within the revision
   cycle. Thus, as the primary reviewer of this paper, I will review the
   revision of the paper and ask the following questions, and if the answers
   to all the questions are "yes", I will recommend to accept this paper.

   Q1. Are all the experimental design choices properly motivated? (refer to
   the first concern of R1).
   Q2. Are the experimental procedure sufficiently detailed? (refer to the
   second concern of R1 and )
   Q3. The results of statistical analysis do not seem to be conclusive, are
   the interpretations of these results properly revised? (refer to the
   third concern of R1)
   Q4. Does the revision contain additional discussion of why PCA is a
   proper approach to interpret the data for the VIS community members who
   do not have sufficient statistics background? (refer to the concern of
   R2)
   Q5. Does the revision has the clear description of "most different"?
   (refer to the first concern of R3)
   Q6. Does the revision has the proper discussion of visualization design
   (refer to the second concern of R3)
   Q7. Does the revision resolve a potential concern of confounding factors
   (refer to the conern of R3)
   Q8. Does the revision have an explicit related works section? (refer to
   the last concern of R3)
   Q9. Does the revision handle all the minor issues raised by reviewers?

Second round comments (public)

Second round supplementary materials check

Second round supplementary materials comments

# Review 2
Title: Spatial Reasoning and Data Displays

_Reviewer_: primary

_Paper type_   Evaluation

_Expertise_ 2  (Knowledgeable)

_Overall Rating_ 3.5 - Between Possible Accept and Accept

_Supplemental Materials_ Acceptable

## Justification

   Overall, the study appears to be meticulously conducted, and the research
   question of "what kinds of factors really influence on visual tasks" is a
   very needed to be scientifically answered. However, I have a concern
   whether PCA is a proper inferencing tool (I think it as a more
   dimensionality reduction tool) and the resulting take-home messages are
   rather unclear.

## The Review

   Overall, various statistical methods (e.g., scaling scores) appeared to
   be meticulously used, and the authors even provided supplementary
   materials to detail such processes, which shows a great practice. Using a
   previously tested approach, line-ups, to investigate which factors will
   impact on visualization task performances is also an interesting idea. I
   believe that our community needs more work like this.

   However, I have a concern. The authors largely rely on principal
   component analysis (PCA) to infer what kinds of cognitive abilities are
   relevant with the line-up tasks. However, I wonder if that is a
   statistically sound approach. I am not a statistician, so I cannot make a
   definitive claim that this is a right or wrong approach (I hope that
   other reviewers have a stronger background on this approach).

   However, I found the following page
   (http://www.alglib.net/dataanalysis/principalcomponentsanalysis.php)
   describing the underlying assumptions of PCA:

   "It may be noted that the PCA is based on the following assumptions:

   The assumption that the dimensionality of data can be efficiently reduced
   by linear transformation
   The assumption that most information is contained in those directions
   where input data variance is maximum.
   As it is evident, these conditions are by no means always met. For
   example, if points of an input set are positioned on the surface of a
   hypersphere, no linear transformation can reduce dimension (nonlinear
   transformation, however, can easily cope with this task). This
   disadvantage is equally attributable to all linear algorithms, and it can
   be eliminated due to the use of complementary dummy variables that are
   nonlinear functions of the input data set elements (the so-called "kernel
   trick").

   The second disadvantage of the PCA consists in the fact that the
   directions maximizing variance do not always maximize information. The
   page of the LDA subroutines gives an example of such a task, wherein the
   maximum-variance variable affords almost no information, whilst the
   minimum-variance variable permits classes to be wholly separated. In this
   case, the PCA will give preference to the first (less informative)
   variable. This drawback is closely connected to the fact that the PCA
   does not perform linear separation of classes, linear regression or other
   similar operations, but it merely permits the input vector to be best
   restored on the basis of the partial information about it. All additional
   information pertaining to the vector (such as the identification of an
   image with one of the classes) is ignored."

   Based on this description, and if this description is correct, I believe
   that the authors should provide more concrete evidence showing that such
   assumptions are met in their analysis, so that the authors can make some
   inferences from the PCA results. However, it is unclear to me based on
   the current manuscript.

   In addition, I found that it is unclear what are the take-home messages
   from the current analysis. It sounds that more research should be done.
   If that is the case, this study sounds like a quite exploratory at this
   stage.

   * page 2
   “speed[9]”

   Put a space right before a citation throughout this paper.

   * page 3
   "Lineups require similar manipulations in two-dimensional space, and also
   require the ability to perform complex spatial manipulations mentally;
   for instance, comparing the interquartile range of two box- plots as well
   as their relative alignment to a similar set of two boxplots in another
   panel."

   I understand that the experiment follows the line-up protocols, but it is
   unclear how it was conducted in detail. What are the visualizations used
   for the three blocks of 20 lineups. What are tasks? (Task examples are
   shown in Figure 1(a), but not sure if the same task was used for all
   lineups). I later found that more examples of line-up tasks are described
   in Section 3.6, but the detailed descriptions of tasks are still
   necessary.

   * page 6
   "3.5 Lineup Types"

   This should be go to the Methods section, not the result section.

   ---

   Regarding the concern regarding PCA, a reviewer who had the sufficient
   statistics background mentioned that my concern is irrelevant to the
   application in the paper. I hope that the authors clarify potential
   questions from readers by clarifying why PCA would be an appropriate
   approach for the people who do not have sufficient statistical
   background.

# Review 1

_Title_: Spatial Reasoning and Data Displays

_Reviewer_:           external

_Paper type_   Evaluation

_Expertise_ 2  (Knowledgeable)

_Overall Rating_ 3.5 - Between Possible Accept and Accept

_Supplemental Materials_  Not applicable (no supplemental materials were submitted with the paper)

## Justification

The paper addresses an important issue in spatial reasoning on graphics.
   While the paper is well presented and the study is extensively evaluated,
   I find that the paper could benefit from stronger motivation of certain
   design choices, better explanation of experimental procedures, and more
   conclusive evaluation of the results.

## The Review

   The paper addresses an important issue in spatial reasoning on graphics
   that I believe is highly valuable to the InfoVis community. The paper is
   well presented and the study is extensively evaluated.

   My main concerns about the paper are threefold.

   Firstly, the paper could benefit from stronger motivation of certain
   design choices. For instance, it is not fully clear how the measures of
   visuospatial ability were chosen from a much larger set of tests in the
   given reference. How are these related to spatial and reasoning ability
   and what were the hypotheses about their relation to the graphics
   presented in the lineup tasks?

   Secondly, the descriptions of the experimental procedures is lacking
   detail. In 2.2 the different tests are explained but it is not clear in
   which order these were presented. Was it a randomised or counterbalanced
   design? Also, what was contained in the 20 line ups presented between
   each session in relation to the different plot types presented in Fig 9?
   Why were the lineups interleaved with the other four tests? What was the
   experimental setup: screen used, viewing distance, lab environment? What
   were the instructions for each of the experimental sessions? What was the
   total experiment time and were breaks included? Beware that such studies
   can be influenced by many factors, such as order effects, fatigue, and
   learning effects, for which reason it is really important to provide
   sufficient details on the experimental design.

   Lastly, while the evaluation appears to be very extensive, I find the
   results not as conclusive as they are interpreted by the authors. Some
   rather low correlations around 0.3-0.4 are interpreted by the authors as
   being high, which I would not agree with. For instance, in Fig 10
   differences in correlation between 0.2 and 0.3 are interpreted to be
   significant while the respective scatter plots do not suggest for one or
   the other to be significantly better. Such small differences in low
   correlation regimes can often be misleading and should not be
   overinterpreted. I appreciate the authors' effort to also draw strong
   conclusions on the results of the PCA, but also these appear to be not
   overly convincing. For instance, at the end of 3.5 the author state that
   lineup 3 is different from lineup 1 and 2. I agree with that from visual
   observation in Fig 9, but I would be hesitant to infer that from the PCA
   results.

   Some more detailed and minor comments:
   - In 2.2, Wolfe's work on visual search should be mentioned in relation
   to the VST.
   - Equation (2) states the same as equation (1).
   - In section 3, you mention one exception in relation to the
   participants' age, please be more specific.
   - In section 3, I suppose you mean that the demographics are
   representative for the student population and not for the whole
   university incl staff.
   - With regard to your assessment of the correlation in Fig 6. I would
   discriminate more between the individual results rather than stating that
   they are all highly correlated.
   - It would be good to move Table 1 to the previous page.
   - Concerning Fig 5: I am not surprised that there is no difference for
   age, as the difference is minimal and the groups would probably be put
   into the same group would there be a wider spread of age in the
   participant population.
   - In the last paragraph of section 3.3: There is a typo "... is
   ultimately are a classification... ".
   - Beginning of 3.4: you mention that all demographic variables are highly
   correlated. Really all? Please be more specific.
   - Task for lineup 3 is given. What were the tasks given for the
   evaluation of lineups 1 and 2?
   - The VST task is excluded from Fig 10 due to its low correlation. Please
   include it for completeness.
   - PCA is explained in detail in 3.5 although it has been used extensively
   earlier. Please put the explanation before the first use.
   - Section 4: I don't agree with the initial statement that lineups are
   strongly related to the performance on tests of visual ability. There is
   certainly evidence provided for that in the paper, but I would not refer
   to it as strong evidence.
   - I propose to put Appendix C into the main body of the paper.
   - In Appendix C: I don't agree with the statement that the card rotation
   task is much more associated with the QQ plots. The correlations are
   actually higher for the first four plot types on the left in the figure.

# Review 3

_Title_: Spatial Reasoning and Data Displays

_Reviewer_:           external

_Paper type_  Evaluation

_Expertise_ 2  (Knowledgeable)

_Overall Rating_ 2.5 - Between Reject and Possible Accept

_Supplemental Materials_ Acceptable

## Justification

   This paper pairs visualization judgments with cognitive tests to try to
   disentangle cognitive strategies for asserting the "most different" plot
   in a set. The paper asks an interesting question and is very well
   written. However, it is unclear from the manuscript what the experimental
   task is and how it's controlled for, and the analysis is of questionable
   validity as it aggregates results across a wide breadth of individual
   designs without consideration for how different visualization designs
   might lead to different cognitive strategies. As a result, the paper is a
   near-miss: it asks an interesting question, but makes a few very critical
   oversights that make its results both questionable and of little utility
   for the visualization community.

## The Review

   The goal of this paper is to measure the role of different forms of
   spatial reasoning in interpreting visualizations, with the ultimate goal
   of understanding the visual strategies underlying judgments made using a
   visualization. Turning to cognition to try to illuminate different
   strategies for interpreting visualizations is an extremely interesting
   idea. However, there are several elements of the experimental design and
   analysis that are concerning. Of particular concern are the ambiguity of
   the lineup task and the lack of consideration of different design
   elements within the analysis. While I believe that the paper in its
   current form is unsuitable for InfoVis, the manuscript is headed in the
   right direction and could be a significant contribution to the community
   in the future.

   It is unclear in the experimental task what “most different” means.
   From the text, it seems as if its intended to measure distributional
   variation. However, “most different” can (and likely will) be
   interpreted by participants in different ways (e.g. largest difference in
   variance, largest difference in mean or median). Was this communicated to
   the participants? Were they given a tutorial? How was the difficulty of
   this judgment controlled? How were other potentially related statistics
   controlled for (e.g. how can we verify that participants were actually
   performing the requested task and not a related proxy task)? What
   information was available to participants for each visualization design
   to make this judgment? In a box plot, for example, we’re only getting
   distributional characteristics, it’s possible that the box plot is
   insufficient to characterize the distribution effectively (e.g.
   Anscombe’s quartet).

   A larger issue related to this point is that the manuscript lacks a
   discussion of visualization design in general. For example, the
   manuscript clearly indicates that  there is a difference in the reported
   correlations across different sets of visualizations (e.g. lineup one,
   two, and three). This directly suggests that the different visualization
   design components , which is well-explored in the graphical perception
   literature (e.g. Cleveland & McGill, 1984, Heer, Kong, & Agrawala, 2009,
   Fuchs et al, 2013, Albers, Correll, & Gleicher, 2014 among many others).
   This would also imply that the findings as reported are biased by the
   sampling of visualization types used in the experiment. Further, the
   comparison within lineups seems somewhat unprincipled. For example,
   it’s unclear how a dot-plot is more similar to set one than set two.

   This will likely be a significant confound in the comparisons with the
   more traditional  cognitive tests. Assuming the tests use exactly the
   same form of stimuli founding figures 2 and 3, these stimuli are uniform
   in design elements across all of the visualization designs. As a result,
   it’s difficult to tell if the results are a by-product of similarity in
   visual features or in visual strategy. This analysis feels like it might
   be comparing apples to oranges--I am not convinced that it makes sense to
   compare the lineup results to performance on these tests in aggregate due
   to complications in visual design. Given this confound, it is unclear how
   a visualization designer could use these results in practice.

   For example, the lineup test is an example of a more free-form visual
   search task: participants are actively searching for the “most
   different” element. As a result, deviations from these two tasks might
   be a function of difference in strategy (as suggested in the paper) or
   due to difference in stimuli (e.g. detecting configurations vs. box
   plots) or both. The figure classification test may provide the most
   similar results for many visualization designs simply because the visual
   hull of the data (the “cloud” in a scatterplot, the box in a boxplot)
   is the most salient global feature for making similarity comparisons,
   which aligns well with the "form" of a figure. This would lead to
   correlation between the tests, but the actual strategy may not be
   rule-based.

   To make these comparisons more compelling, these same experimental
   procedures can be conducted with the stimuli in each of these tests
   adapted to more closely mirror the tested visualization type. I would
   encourage the authors to consider restructuring these tests to use
   stimuli more relevant to each respective visualization lineup condition
   or, at a minimum, to at least break down the results with respect to each
   individual visualization design and compare strategies across visual
   feature. Aggregating across all visualizations is an oversimplification
   of the task—different visualization may require different strategies,
   and, in fact, graphical perception suggests that they do.

   I would also recommend that a revised manuscript include an explicit
   related works section. Traditionally in this community, the related work
   is separate from the introduction. While I think this paper reads well
   without that structure, the authors may wish to consider revising their
   structure to accommodate a more thorough discussion of related work. Part
   of this recommendation is that there is a good breadth of relevant work
   that is not considered in this manuscript. For example, “The lineup
   protocol [12, 25, 3] is a testing framework that allows researchers to
   quantify the statistical significance of a graphical finding with the
   same mathematical rigor as conventional hypothesis tests” is a very
   contentious claim. In particular, recent work in visualization, such as
   (Pak, Hutchinson, & Turk-Brown, 2014, Talbot, Setlur, & Anand, 2014 (more
   distant comparison is harder), Correll & Gleicher, 2014, and a myriad of
   papers on limitations in interpreting boxplots), suggest that people are
   not good at performing visual significance testing. Visual search is
   among the most commonly used tasks in graphical perception experiments.
   For example, Fuchs et al, 2014 includes experiments that explicitly
   involve visual search tasks in small multiple (lineup) conditions.

   Minor Issues:
   Additional experimental details about the type of visualizations tested
   (e.g. Figure 9) should be presented in the Methods section. A pointer to
   another set of papers with large numbers of potential designs is
   insufficient to understand the experiment. Additionally, explanations of
   the baseline cognitive tests should be presented up front to ground the
   analysis: while many in the visualization community may have seen the
   compared tests (visual search task (VST), paper folding test, card
   rotation test, and figure classification test), especially the visual
   search task, many other have not. Give the necessary grounding and
   explain why these particular tasks were chosen as a baseline.

## Spelling & Grammar:

   Overall this paper is very well written. However, throughout the paper,
   the term “graphics” is used instead of “visualization”.
   “Graphics” has a much broader possible meaning and
   “visualization” is conventional for the InfoVis audience.

   speed[9]  -> speed [9]
   reasoning[6] -> reasoning [6]

# Review 4

_Title_: Spatial Reasoning and Data Displays

_Reviewer_:           secondary

_Paper type_   Theory / Model

_Expertise_   3  (Expert)

_Overall Rating_ 5 - Strong Accept    
The paper is an excellent contribution and should certainly be accepted, possibly with some minor revisions.    
It more than meets the quality standard.

_Supplemental Materials_ Acceptable

## Justification

   Nothing that I haven't listed in the review. I think this is an
   extraordinary paper because the authors have taken a fundamentally
   important contribution and extended it to an area where nobody has
   thought about it so far.

## The Review

   The work represented here may be the most important recent advance in
   methodology for evaluating the effectiveness of visual displays. It
   appears to be, as well, the most powerful method for evaluating
   individual differences in graph reading ability. What is remarkable is
   that such a simple paradigm can be coupled with a formal statistical
   distribution (the binomial) for quantifying these results. Instead of
   separating graph reading into subtasks that may be highly correlated (and
   therefore not highly informative taken as a whole set) this team has
   devised a method for dealing with highly complex displays (probability
   plots, billets, etc.) as well as simple ones.

   Moreover, this team has implemented a methodology that fits perfectly
   into a tradition of careful psychological measurement. Instead of
   requesting subjects to rate things numerically (as Cleveland did, for
   example), these researchers simply ask for the point of a finger, so to
   speak. The response is similar in nature to that of classical perceptual
   direct scaling studies going back almost a century.

   Sometimes authors or teams milk a simple idea by generating minor
   variations and submitting them to a bunch of journals. What's remarkable
   here is that the collection of their publications given in the references
   ranges from a formal statistical article with proofs etc. to several
   visualization articles having very different focuses. The same is true
   for this article. I have seen presentations of this research and I never
   thought of the possibility that it could be used in the context of the
   present article.

   This paper is very well organized and written. I have few considerations
   for revision and plan to nominate it for best paper.

### Details:

   I'd suggest a mention of the groundbreaking mental rotation research of
   Shepard & Metzler. This was 3D rotation, but it did stimulate thinking by
   later psychologists on the rotation problem. What's more, their results
   were confirmed by subsequent neurological research in Science magazine.

### 2015-06-08 Paper/InfoVis2015/TestingVisualAptitude.Rnw
Reviewer responses and slightly modified regex in paper
Modifications - "most different" = feature, not bug; PCA discussion + additional reference

@@
 sanitize <- function(x){
-  str_replace(gsub("_", ".", x, fixed=T), "Pr\\((.{1,}))", "Pr$(\\1)$")
+  str_replace(gsub("_", ".", x, fixed=T), "Pr\\((.{1,})\\)", "Pr$(\\1)$")
 }
@@
 Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
-What is of more interest to us is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
-In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests using the same procedure.
+\newtext{While test scores are correlated, the relationships between test scores appear to be linear, with sufficient variability around the trend line to merit further investigation.
+The correlations are to be expected, as all of the tests in this experiment are conducted in a visual domain, and require general reasoning, intelligence, and other factors.
+}
+What is of more interest to us is 
+\newtext{whether lineups recruit specific cognitive abilities, such as visual search, reasoning and categorization ability, mental rotation.} 
+In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests (four cognitive tests, and additionally lineups) using the same procedure.
 
-A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
+\newtext{
+Principal component analysis (PCA) is a statistical technique to transform the coordinate system of correlated data, producing orthogonal components in a new coordinate space (see \cite{pcaReference} for more information).
+The most stringent assumption required for PCA is that of linearity, which is reasonable given Figure \ref{fig:scatterplotmatrix}.
+Additionally, PCA assumes that the interesting portion of the data is the variance between any two variables (a perfectly linear relationship between two variables would be uninteresting, as the variables contain the same information).
+As a result, PCA can be used to examine which variables contain similar information (i.e.
+which variables are represented in each principal component).
+It is this feature of principal components which we utilize in this paper.
+}
+
+A principal component analysis of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about \Sexpr{round(summary(pca2)$importance[3,1],2)*100}\% of the total variability.
 Principal components (PC-) are ordered by importance (how much variability in the data they contain), and each principal component is uncorrelated with every other principal component.

### 2015-06-09 Paper/InfoVis2015Revision/ReviewerResponses.Rmd
InfoVis Revision

--- 2015-06-08 Paper/InfoVis2015/ReviewerResponses.Rmd
@@
 
-   Q1. Are all the experimental design choices properly motivated? (refer to
+1. Are all the experimental design choices properly motivated? (refer to
    the first concern of R1).
-   Q2. Are the experimental procedure sufficiently detailed? (refer to the
+2. Are the experimental procedure sufficiently detailed? (refer to the
    second concern of R1 and )
-   Q3. The results of statistical analysis do not seem to be conclusive, are
+3. The results of statistical analysis do not seem to be conclusive, are
    the interpretations of these results properly revised? (refer to the
    third concern of R1)
-   Q4. Does the revision contain additional discussion of why PCA is a
+4. Does the revision contain additional discussion of why PCA is a
    proper approach to interpret the data for the VIS community members who
@@
    R2)
-   Q5. Does the revision has the clear description of "most different"?
+5. Does the revision has the clear description of "most different"?
    (refer to the first concern of R3)
-   Q6. Does the revision has the proper discussion of visualization design
+6. Does the revision has the proper discussion of visualization design
    (refer to the second concern of R3)
-   Q7. Does the revision resolve a potential concern of confounding factors
+7. Does the revision resolve a potential concern of confounding factors
    (refer to the conern of R3)
-   Q8. Does the revision have an explicit related works section? (refer to
+8. Does the revision have an explicit related works section? (refer to
    the last concern of R3)
-   Q9. Does the revision handle all the minor issues raised by reviewers?
+9. Does the revision handle all the minor issues raised by reviewers?
 

### 2015-06-15 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
merge

--- 2015-06-08 Paper/InfoVis2015/TestingVisualAptitude.Rnw
@@
   #coord_flip()
+if (file.exists("./Figure/fig-ResultsScaledScores-1.pdf")) 
   embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
@@
   theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
+if (file.exists("./Figure/fig-VisReasoningCategorical-1.pdf"))
   embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
@@
 rm(list=c("data", "all", "densities"))
-embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
+if (file.exists("./Figure/fig-VisReasoningSPM-1.pdf")) embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
 @
@@
   theme(axis.title=element_blank())
-embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
+if (file.exists("./Figure/fig-LineupTypeSPM-1.pdf")) embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
 @
@@
 qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
-embedFonts("./Figure/fig-lineupblockpca-1.pdf")
+if (file.exists("./Figure/fig-lineupblockpca-1.pdf")) embedFonts("./Figure/fig-lineupblockpca-1.pdf")
 @

### 2015-06-15 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
changes look good, but we need to document them in the Rmd file

@@
 %% insert punctuation at end of each item
-\item Susan VanderPlas is a graduate student at Iowa State University. Email: skoons@iastate.edu.
+\item Susan VanderPlas is a Post Doc in the Department of Statistics at Iowa State University. Email: skoons@iastate.edu.
 \item Heike Hofmann is Professor of Statistics at Iowa State University and member of the faculty in the Human Computer Interaction program. Email: hofmann@iastate.edu.
@@
 This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different"
-\newtext{(this phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies \cite{wickham2010graphical, majumder2013validation}).}
+\newtext{This phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies employing lineups \cite{wickham2010graphical, majumder2013validation}.}
 If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.

### 2015-06-15 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Fixing reviewer complaints, etc. and expanding other information

@@
 \newcommand{\hhnote}[1]{\todo[inline,color=magenta!40]{#1}}
-\newcommand{\newtext}[1]{\todo[inline,color=green!40]{#1}}
+\newcommand{\comment}[1]{\todo[inline,color=green!40]{#1}}
+\newcommand{\newtext}[1]{\textcolor{blue}{#1}}
 %---------------------------------------------------
@@
 As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol  \cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
-
-``Lineups" have recently been introduced as a tool to evaluate the statistical significance of a graphical finding.
+\comment{Not sure where to add the Related Work section; the introduction is really the most similar thing, but adding another sub-section in would really kill the flow of the paragraph, in my opinion.}
+``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding\cite{wickham2010graphical}.
 Lineups are also useful in assessing the effectiveness of different graphical displays \cite{hofmann2012graphical, loy:2015}.
@@
 In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
+\newtext{
+Our goal in this study is not to evaluate the statistical significance of particular plots, but rather, to consider the cognitive demands that we require of participants in lineup studies. 
+}
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
@@
 By aggregating responses from different individuals the lineup protocol therefore allows an objective evaluation of a graphical finding. 
-
 Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task. 
-
+\newtext{In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score. 
+}
 For this approach, we derive a score for an individual as follows:
@@
 
+\newtext{Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
+This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic.
+This study uses three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots.
+Each block of lineups are a subset of lineups designed to compare the visual significance of different plot types (examples of each plot type are shown in figure \ref{fig:lineups}).
+Our purpose in this study is to compare the cognitive demands required to successfully evaluate these datasets, which contain variance difference, mean difference, or violations of normality, as displayed in various common statistical visualizations.}
+
 \subsection{Measures of visuospatial ability}
@@
 
+\newtext{All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in \cite{ekstrom1976manual}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; pilot tests did not indicate an effect of task order.
+}
+
 Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
@@
 \begin{eqnarray}\label{eq.scoring}
-\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}.
+\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}
 \end{eqnarray}
+\newtext{Equation \ref{eq.scoring} is a restatement of equation \ref{eq.scoring1} provided for convenience.}
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
@@
 Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
-This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the university as a whole; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
+This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the student population of Iowa State; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
 
@@
 Values have been adjusted to accommodate for differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix~\ref{app:ScoreAdj}).
-Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
+Once adjusted, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics.
 In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.
@@
 
-This emphasizes the underpinnings of lineups: the test utilizes a visual medium, but is ultimately are a classification task presented in a graphical manner.
+This emphasizes the underpinnings of lineups: the test utilizes a visual medium, but ultimately lineups are a classification task presented in a graphical manner.
 Using lineups as a proxy for statistical significance tests is similar to using a classifier on pictoral data: while the data is presented ``graphically", the participant is actually classifying the data based on underlying summary statistics.
@@
 
-all2 <- subset(all, grepl("lineup", yvar) & points & !grepl("vis_search", xvar))
+all2 <- subset(all, grepl("lineup", yvar) & points )#& !grepl("vis_search", xvar)
 # names(all2)[1:4] <- c("yvar", "xvar", "y", "x")
@@
 Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks. 
-The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
+%The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
 Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
@@
 %
-$$\text{Influence of variable} X_i \text{ on PC}_j = M_{ij}\times I_j$$
+$$\text{Influence of variable } X_i \text{ on PC}_j = M_{ij}\times I_j$$
 %

### 2015-06-15 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
InfoVis Revision
merge
Fixing reviewer complaints, etc. and expanding other information

--- 2015-04-03 Paper/InfoVis2015/TestingVisualAptitude.tex
@@
 \newtext{This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization. }
+
+\newtext{Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
+This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic.
+This study uses three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots.
+Each block of lineups are a subset of lineups designed to compare the visual significance of different plot types (examples of each plot type are shown in figure \ref{fig:lineups}).
+Our purpose in this study is to compare the cognitive demands required to successfully evaluate these datasets, which contain variance difference, mean difference, or violations of normality, as displayed in various common statistical visualizations.}
 
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Tue Jun  9 10:29:29 2015
+% latex table generated in R 3.2.0 by xtable 1.7-1 package
+% Mon Jun 15 15:31:31 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Tue Jun  9 10:29:30 2015
+% latex table generated in R 3.2.0 by xtable 1.7-1 package
+% Mon Jun 15 15:31:31 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Tue Jun  9 10:29:30 2015
+% latex table generated in R 3.2.0 by xtable 1.7-1 package
+% Mon Jun 15 15:31:31 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Tue Jun  9 10:29:30 2015
+% latex table generated in R 3.2.0 by xtable 1.7-1 package
+% Mon Jun 15 15:31:32 2015
 \begin{table}[ht]

### 2015-06-16 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Addressing reviewer comments

@@
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
-Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
+Graphics utilize higher-bandwidth visual pathways to encode information~~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
 Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
@@
 
-Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
+Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~~\cite{vekiri2002value}.
 Both of these theories suggest spatial ability impacts a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
-However,  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
-Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~\cite{zhang1997nature}.
+In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~~\cite{lowrie2007solving}.
+However,  mathematical ability, not spatial ability, was shown~~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
+Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+It is theorized that graphics are a form of ``external cognition"~~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~~\cite{zhang1997nature}.
 
 Here, we want to investigate the extent to which observers make use of spatial ability in their evaluation of and reasoning about statistical graphics.
-As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol  \cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
+As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol ~~\cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
 \comment{Not sure where to add the Related Work section; the introduction is really the most similar thing, but adding another sub-section in would really kill the flow of the paragraph, in my opinion.}
-``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding\cite{wickham2010graphical}.
-Lineups are also useful in assessing the effectiveness of different graphical displays \cite{hofmann2012graphical, loy:2015}.
+``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
+Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
 Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
@@
 This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance \cite{humanfactorslineups}.
+While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
 
@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
-The VST measures visual search speed\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning\cite{ekstrom1976manual};
+The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
 all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
@@
 \subsection{The Lineup Protocol}
-The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
+The lineup protocol~~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
 
@@
 This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different"
-\newtext{This phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies employing lineups \cite{wickham2010graphical, majumder2013validation}.}
+\newtext{This phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies employing lineups~\cite{wickham2010graphical, majumder2013validation}.}
 If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
 Let $X$ be the number of observers (out of $n$) who identify the data plot from the lineup.
-Under the null hypothesis that the data plot is not different from the other plots, $X$ has approximately a Binomial distribution \cite{wickham2010graphical, majumder2013validation}.
+Under the null hypothesis that the data plot is not different from the other plots, $X$ has approximately a Binomial distribution~\cite{wickham2010graphical, majumder2013validation}.
 If $k$ of the observers identify the data plot from the lineup, the probability $P(X \ge k)$ is the $p$-value of the corresponding visual test.
@@
 
-Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups \cite{humanfactorslineups}.
-In addition, lineup performance can be partially explained using statistical distance metrics \cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
+Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups~\cite{humanfactorslineups}.
+In addition, lineup performance can be partially explained using statistical distance metrics~\cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
 
@@
 This study uses three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots.
-Each block of lineups are a subset of lineups designed to compare the visual significance of different plot types (examples of each plot type are shown in figure \ref{fig:lineups}).
+Each block of lineups are a subset of lineups designed to compare the visual significance of different plot types (examples of the visualization types presented in each block of lineups are shown in figure \ref{fig:lineups}).
 Our purpose in this study is to compare the cognitive demands required to successfully evaluate these datasets, which contain variance difference, mean difference, or violations of normality, as displayed in various common statistical visualizations.}
+
+Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
+\begin{figure*}[htbp]\centering
+\subfloat[Set 1: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}}
+\subfloat[Set 1: Density plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}}
+\subfloat[Set 1: Dotplot ]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}}
+\subfloat[Set 1: Histogram]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}}\\
+\subfloat[Set 2: Boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f94f6965d7}}
+\subfloat[Set 2: Boxplot with jittered points]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9e65f0e4}}
+\subfloat[Set 2: Bee swarm boxplot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9fc58d73}}
+\subfloat[Set 2: Violin plot]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f93acbb7e4}}\\
+\subfloat[Set 3: QQ-plot with guide line]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13876a225d34-single}}
+\subfloat[Set 3: QQ-plot with acceptance bands]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13877c93dc78-single}}
+\subfloat[Set 3: QQ-plot rotated 45 degrees.]{
+  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
+\caption{Lineup plot types used in the experiment.
+\label{fig:lineups} \newtext{Each set of plots is representative of the plot types shown in a block of lineups in the experiment, and each block represents a subset of plots from a traditional lineup experiment to determine which visualization method facilitates identification of the plot with the strongest visual signal.
+Set 1 manipulates group means and variance, set 2 manipulates group means, and set 3 examines violations of normality.
+These tasks were chosen as common statistical visualization tasks, and while not exhaustive, should provide a broad basis for interpreting the results of this experiment.}}
+\end{figure*}
+
+\paragraph{Lineup Set 1}
+The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
+Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
+
+\paragraph{Lineup Set 2}
+The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
+Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
+
+\paragraph{Lineup Set 3}
+The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
 
@@
 The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
-Historically, visual search has been used as a measure of brain damage \cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~\cite{anderson1983interactive}.
-The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations.
+Historically, this visual search task has been used as a measure of brain damage~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~~\cite{anderson1983interactive}.
+The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations, such as those in~\cite{wolfe1990limitations,treisman1980feature}.
 
@@
 The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
+This task is associated with inductive reasoning abilities (factor I  in~~\cite{ekstrom1976manual}).
 An example is shown in figure~\ref{fig:figureclassification}.
@@
 The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+It tests mental rotation skills, and is classified as a test of spatial orientation in~~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
 An example is shown in figure~\ref{fig:cardrotation}.
-The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids \cite{mayer1994whom} and statistical graphs \cite{lowrie2007solving} in education.
+The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids~\cite{mayer1994whom} and statistical graphs~\cite{lowrie2007solving} in education.
 
@@
 A sample question from the test is shown in figure~\ref{fig:paperfolding}.
-It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
+It is classified as part of the visualization factor in~~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
@@
 
-\newtext{All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in \cite{ekstrom1976manual}.
+\newtext{All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
 The order of tasks was the same for each participant due to the constraints of in-person testing; pilot tests did not indicate an effect of task order.
@@
 
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~~\cite{hofmann2012graphical,majumder2013validation}.
 Participants were given 5 minutes to complete each block of 20 lineups.
@@
 These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+
+\newtext{After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
+The duration of testing for each participant was approximately 1 hour.
+}
 
@@
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests~~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~~\cite{educational1963kit}.
 
@@
 \end{tabular}
-\caption[Comparison of scores for cognitive tasks.]{Comparison of scores from Iowa State students and scores reported in~\protect\cite{ekstrom1976manual}.
+\caption[Comparison of scores for cognitive tasks.]{Comparison of scores from Iowa State students and scores reported in~\protect~\cite{ekstrom1976manual}.
 Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
 Data shown are from the population most similar to ISU students, out of the data available.
-The visual search task~\protect\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+The visual search task~\protect~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
 \label{tab:scorecomparison}}
@@
 %Next, we examine the effect of demographic factors on lineup performance. 
-Previous work found a relationship between lineup performance and demographic factors such as education level, country of origin, and age \cite{humanfactorslineups}; our participant population is very homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
+Previous work found a relationship between lineup performance and demographic factors such as education level, country of origin, and age~\cite{humanfactorslineups}; our participant population is very homogeneous, which allows us to explore factors such as educational background and skills on performance in lineup tests.
 
@@
 Completion of Calculus I is associated with increased performance on lineups; this may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with findings in~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
+This association is consistent with findings in~~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
 There is also a significant relationship between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~~\cite{hausmann2000sex}.
 There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
@@
 \newtext{
-Principal component analysis (PCA) is a statistical technique to transform the coordinate system of correlated data, producing orthogonal components in a new coordinate space (see \cite{pcaReference} for more information).
+Principal component analysis (PCA) is a statistical technique to transform the coordinate system of correlated data, producing orthogonal components in a new coordinate space (see~\cite{pcaReference} for more information).
 The most stringent assumption required for PCA is that of linearity, which is reasonable given Figure \ref{fig:scatterplotmatrix}.
@@
 @
-Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
+Biplots~\cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
 Figure classification is strongly related to lineups (PC2, PC3).
@@
 \subsection{Linear model of demographic factors}
-Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.
+Note that many of the demographic variables in the survey show dependencies, for example there is a high correlation between STEM majors and taking calculus.
 Similarly, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high.
 Only one student is doing research who has not taken a statistics course.
+Students' ratings of their math skills are also related to whether they are majoring in a STEM field or whether they have completed calculus.
 
@@
 \subsection{Lineup Types}
-Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
-\begin{figure*}[htbp]\centering
-\subfloat[Set 1: Boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/boxplot_1_15_5_4_13}}
-\subfloat[Set 1: Density plot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/density_1_15_5_4_13}}
-\subfloat[Set 1: Dotplot ]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/dotplot_1_15_5_4_13}}
-\subfloat[Set 1: Histogram]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups1/histogram_1_15_5_4_13}}\\
-\subfloat[Set 2: Boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f94f6965d7}}
-\subfloat[Set 2: Boxplot with jittered points]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9e65f0e4}}
-\subfloat[Set 2: Bee swarm boxplot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f9fc58d73}}
-\subfloat[Set 2: Violin plot]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups2/file147f93acbb7e4}}\\
-\subfloat[Set 3: QQ-plot with guide line]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13876a225d34-single}}
-\subfloat[Set 3: QQ-plot with acceptance bands]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13877c93dc78-single}}
-\subfloat[Set 3: QQ-plot rotated 45 degrees.]{
-  \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
-\caption{Lineup plot types used in the experiment. \label{fig:lineups}}
-\end{figure*}
-
-\paragraph{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
-
-\paragraph{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-
-\paragraph{Lineup Set 3}
-The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
+Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks. 
+%The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
+Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
+Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
 
@@
 \end{figure}
-Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks. 
-%The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
-Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
-Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
 

### 2015-06-16 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Addressing reviewer comments

@@
 % latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Mon Jun 15 15:31:31 2015
+% Tue Jun 16 12:39:22 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Mon Jun 15 15:31:31 2015
+% Tue Jun 16 12:39:22 2015
 \begin{table}[ht]
@@
 \end{knitrout}
-Biplots \cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
+Biplots~\cite{biplots:gabriel} of the remaining components are provided in figure~\ref{fig:biplots5}.
 Figure classification is strongly related to lineups (PC2, PC3).
@@
 % latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Mon Jun 15 15:31:32 2015
+% Tue Jun 16 12:39:23 2015
 \begin{table}[ht]

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
attempt at discussing related work

@@
 \comment{Not sure where to add the Related Work section; the introduction is really the most similar thing, but adding another sub-section in would really kill the flow of the paragraph, in my opinion.}
+\hhnote{I agree, it'll disrupt the flow, but we'll have to add a related work section nevertheless. We could add a bit of discussion on separating lineups from plots judging significance.}
+
+\hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
+first attempt: It has been shown~\citep{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\citep{correll2013}.
+Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
+Statistical properties are then derived from this assessment.
+This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\citet{majumder2013validation}.
+The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
+a correlation coefficient is based on only a few summary statistics, whereas e.g.
+a corresponding  scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
+This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \citet{anscombe:1973}.
+}
 ``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
attempt at discussing related work

@@
 
-% latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Tue Jun 16 12:39:22 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Mon Jun 22 14:48:56 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Tue Jun 16 12:39:22 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Mon Jun 22 14:48:57 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Tue Jun 16 12:39:22 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Mon Jun 22 14:48:57 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.0 by xtable 1.7-1 package
-% Tue Jun 16 12:39:23 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Mon Jun 22 14:48:57 2015
 \begin{table}[ht]

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
with correct citations now

@@
 \hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\citep{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\citep{correll2013}.
+first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
 Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
@@
 a corresponding  scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
-This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \citet{anscombe:1973}.
+This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
 }

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
ah, and now the citations are actually in the right file ... grrr

@@
 Statistical properties are then derived from this assessment.
-This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\citet{majumder2013validation}.
+This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
 The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
ah, and now the citations are actually in the right file ... grrr

@@
 \hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\citep{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\citep{correll2013}.
+first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
 Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
 Statistical properties are then derived from this assessment.
-This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\citet{majumder2013validation}.
+This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
 The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 14:48:56 2015
+% Mon Jun 22 15:06:35 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 14:48:57 2015
+% Mon Jun 22 15:06:36 2015
 \begin{table}[ht]

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
what's still missing

@@
 
+\hhnote{XXX Still missing: we need to extend the discussion a little bit that we want to extract the common core from lineups rather than investigate the relationship between individual designs and the  tasks.
+I believe, that the lineup commonalities are stronger than the design differences - but we might need to get some statistical backup for that.
+...
+didn't we run those PCAs as well?
+}
+
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
@@
 
-\hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
+``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
+Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
+Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
+% 
+% \begin{figure}[htp]
+% 
+% \end{figure}
+
+Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
+
+%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
+
+Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
+This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
+While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
+
+In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
+\newtext{
+Our goal in this study is not to evaluate the statistical significance of particular plots, but rather, to consider the cognitive demands that we require of participants in lineup studies. 
+}
+We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
+The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
+We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
+
+In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
+In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
+We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
+Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+
+\paragraph{Related Work}
+
+\hhnote{First attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
 Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
@@
 }
-``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
-Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
-Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-% 
-% \begin{figure}[htp]
-% 
-% \end{figure}
-
-Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
-In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
-
-%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
-
-Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
-This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
-
-In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
-\newtext{
-Our goal in this study is not to evaluate the statistical significance of particular plots, but rather, to consider the cognitive demands that we require of participants in lineup studies. 
-}
-We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
-The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
-We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
-
-In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
-In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
-We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
-Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+\hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
 

### 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
still fiddling with the references
what's still missing

@@
 
-\hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
-Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
-Statistical properties are then derived from this assessment.
-This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
-The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
-a correlation coefficient is based on only a few summary statistics, whereas e.g.
-a corresponding  scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
-This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
-}
 ``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
@@
 Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+
+\paragraph{Related Work}
+\hhnote{First attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
+Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
+Statistical properties are then derived from this assessment.
+This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
+The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
+a correlation coefficient is based on only a few summary statistics, whereas e.g.
+a corresponding  scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
+This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
+}
+\hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
 
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:35 2015
+% Mon Jun 22 15:20:39 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:36 2015
+% Mon Jun 22 15:20:39 2015
 \begin{table}[ht]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:36 2015
+% Mon Jun 22 15:20:40 2015
 \begin{table}[ht]

### 2015-06-23 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
significance of first PC between three lineups?
little comment

@@
 ...
-didn't we run those PCAs as well?
+one way to do that would be to do a PCA on the three lineups and investigate whether anything beyond the first dimension is significant.
+I don't think so.
+}
+
+\hhnote{XXX we should add: the lineups come from three different lineup experiments that discuss and compare the design choices (cite papers).
+This is NOT what we are doing here.
+Regardless of design choices, we want to find out, what visual skills are necessary for dealing with lineups.
 }

### 2015-06-23 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Changes during meeting

--- 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
@@
 
+% \hhnote{XXX Still missing: we need to extend the discussion a little bit that we want to extract the common core from lineups rather than investigate the relationship between individual designs and the  tasks.
+I believe, that the lineup commonalities are stronger than the design differences - but we might need to get some statistical backup for that.
+...
+one way to do that would be to do a PCA on the three lineups and investigate whether anything beyond the first dimension is significant.
+I don't think so.
+}
+
+% \hhnote{XXX we should add: the lineups come from three different lineup experiments that discuss and compare the design choices (cite papers).
+This is NOT what we are doing here.
+Regardless of design choices, we want to find out, what visual skills are necessary for dealing with lineups.
+}
+
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
@@
 As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol ~~\cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
-\comment{Not sure where to add the Related Work section; the introduction is really the most similar thing, but adding another sub-section in would really kill the flow of the paragraph, in my opinion.}
-\hhnote{I agree, it'll disrupt the flow, but we'll have to add a related work section nevertheless. We could add a bit of discussion on separating lineups from plots judging significance.}
-
-\hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
+
+``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
+Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
+Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
+% 
+% \begin{figure}[htp]
+% 
+% \end{figure}
+
+Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
+In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
+
+%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
+
+Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
+This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
+While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
+
+In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
+\newtext{
+Our goal in this study is not to evaluate the statistical significance of particular plots or visualization designs, but rather, to consider the cognitive demands that we require of participants in lineup studies.
+}
+We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
+The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
+\newtext{
+The cognitive tests from \cite{ekstrom1976manual} are highly validated tests that have been constructed to measure specific cognitive factors.
+We have additionally added the VST in order to additionally account for the basic visual search component of the lineup task.
+}
+We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
+
+In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
+In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
+We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
+Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+
+\paragraph{Related Work}
+\newtext{It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
 Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
@@
 This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
-}
-``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
-Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
-Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-% 
-% \begin{figure}[htp]
-% 
-% \end{figure}
-
-Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
-In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
-
-%This method is described in more detail in the next section; here we motivate the importance of understanding the connection between lineups and spatial reasoning.
-
-Lineups provide a quantitative measurement of the effectiveness of a particular plot: if participants consistently identify the target plot rather than the randomly-generated distractors, the plot effectively shows the difference between real data and random noise.
-This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
-
-In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
-\newtext{
-Our goal in this study is not to evaluate the statistical significance of particular plots, but rather, to consider the cognitive demands that we require of participants in lineup studies. 
-}
-We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
-The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task. 
-We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
-
-In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
-In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
-We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
-Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+In addition, the lineup task does not require participants to make numerical judgments, as in \cite{talbot}, but rather to indicate (though not identify) a visual difference; this leads to a much more natural testing procedure.}
+% \hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
 
@@
 It is classified as part of the visualization factor in~~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
-The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
+The paper folding test is associated with the ability to extrapolate symmetry and reflection.
 
@@
 
-\subsection{Comparison of Spatial Tests with Previously Validated Results}
+\subsection{Comparison of Spatial Tests with Validated Results}
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
@@
 
-<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=8.75,fig.height=6, message=F, warning=F, fig.pos='ht'>>=
+<<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=6,fig.height=5, message=F, warning=F, fig.pos='ht'>>=
 loadfonts()
@@
 lineup.summary.categorical$value[lineup.summary.categorical$value=="n"] <- "FALSE"
+
+lineup.summary.categorical$variable <- gsub("Math/Science Research", "Math/Sci Research", lineup.summary.categorical$variable)
 
@@
 @
-\begin{figure*}[h!tb]\centering
-\includegraphics[width=.85\linewidth]{fig-VisReasoningCategorical-1}
+\begin{figure}[h!tb]\centering
+\includegraphics[width=.95\linewidth]{fig-VisReasoningCategorical-1}
 \caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
@@
 }\label{fig:visualaptitudecat}
-\end{figure*}
+\end{figure}
 %\afterpage{\clearpage}
@@
 
-\subsection{Understanding Visual Abilities and Lineup Performance}
+\subsection{Visual Abilities and Lineup Performance}
 <<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize", include=F>>=
@@
 <<pcarotation4,echo=F, results='asis'>>=
+row.names(pca2$rotation) <- c("Card Rot", "Fig Class", "Folding", "VST")
 print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")

### 2015-06-23 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Additional meeting changes - additional references, etc.

@@
 Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-% 
-% \begin{figure}[htp]
-% 
-% \end{figure}
-
 Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
@@
 all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task.
-\newtext{
-The cognitive tests from \cite{ekstrom1976manual} are highly validated tests that have been constructed to measure specific cognitive factors.
-We have additionally added the VST in order to additionally account for the basic visual search component of the lineup task.
-}
+
 We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
@@
 In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
-In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
+In section \ref{sec:results} we discuss the study results and compare them with scores on previously established tests, accounting for demographic characteristics associated with test scores.
 We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
@@
 \paragraph{Related Work}
-\newtext{It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
-Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
-Statistical properties are then derived from this assessment.
+\newtext{
+It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well and may even mislead viewers~\cite{correll2013}.
+Lineup tests approach the evaluation of statistical visualizations differently; they do not ask participants to judge significances or estimate numerical differences directly (as in \cite{cleveland:1985, fuchs, talbot}) but are only based on comparisons.
+Statistical properties are then derived from this assessment using results aggregated across individual participants.
 This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
@@
 a correlation coefficient is based on only a few summary statistics, whereas e.g.
-a corresponding scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
+a corresponding scatterplot shows the data in more detail, perhaps leading participants to a different evaluation.
 This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
-In addition, the lineup task does not require participants to make numerical judgments, as in \cite{talbot}, but rather to indicate (though not identify) a visual difference; this leads to a much more natural testing procedure.}
+}
 % \hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
@@
 \newtext{Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
-This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic.
-This study uses three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots.
-Each block of lineups are a subset of lineups designed to compare the visual significance of different plot types (examples of the visualization types presented in each block of lineups are shown in figure \ref{fig:lineups}).
-Our purpose in this study is to compare the cognitive demands required to successfully evaluate these datasets, which contain variance difference, mean difference, or violations of normality, as displayed in various common statistical visualizations.}
+This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic; this examination utilizes data aggregated over all participants who evaluated a specific lineup image.
+
+In this study, we utilize three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots (examples of the visualization types presented in each block of lineups are shown in figure \ref{fig:lineups}).
+For example, in lineup set 1, the same five datasets are displayed as a dot plot, density plot, box plot, and histogram.
+Each block of lineups are a subset of stimuli used in experiments designed to compare the visual significance of different plot types, however, our purpose in this study is not to determine the efficacy of different types of plots.
+Rather, our goal is to compare the cognitive demands required to successfully evaluate these lineups, which display variance difference, mean difference, or violations of normality using common statistical plots.}
 
@@
   \includegraphics[keepaspectratio=T,width=.245\linewidth]{../../Images/Lineups3/file13878e5c92f-single}}
-\caption{Lineup plot types used in the experiment.
+\protect\caption{Lineup plot types used in the experiment.
 \label{fig:lineups} \newtext{Each set of plots is representative of the plot types shown in a block of lineups in the experiment, and each block represents a subset of plots from a traditional lineup experiment to determine which visualization method facilitates identification of the plot with the strongest visual signal.
@@
 \paragraph{Lineup Set 1}
-The experiment in the first lineup section examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+The lineups in this set are taken from an experiment which examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
 This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
@@
 \paragraph{Lineup Set 2}
-The second lineup section also explored two groups of data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
+The lineups in the second set are taken from another experiment which examined visualizations of categorical data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
 Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
@@
 \paragraph{Lineup Set 3}
-The final lineup set explored QQ-plots from various model simulations, using reference lines, acceptance bands, and rotation to determine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were too liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands.
+Lineups in the final set are taken from an experiment which examined modifications of QQ-plots, using data from various model simulations.
+Reference lines, acceptance bands, and rotation were utilized to examine which plots allowed participants to most effectively identify violations of normality.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were overly liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands \cite{loy:2015}.
 
@@
 Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
-Tasks are timed such  that participants are under  pressure to complete; participants are not expected to finish all of the problems in each section.
-This allows for a better discrimination between scores and prevents score compression at the top of the range.
+Tasks are timed such  that participants are under pressure; participants are not expected to finish all of the problems in each section.
+This allows for better discrimination between scores and reduces score compression at the top of the range.
 
@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations, such as those in~\cite{wolfe1990limitations,treisman1980feature}.
+\newtext{We have included visual search in the cognitive testing battery for this experiment because it is the most simple of the tasks required to successfully evaluate a lineup; if participants cannot complete visual search tasks successfully, they will not be able to accurately identify which lineup plot is different.
+}
 
@@
 
+\newtext{
+The remaining cognitive tests are adapted from \cite{ekstrom1976manual} and are highly validated tests that have been constructed to measure specific cognitive factors, such as inductive reasoning and spatial rotation.
+}
+
 The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
@@
 Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
+\newtext{Three-dimensional rotation is somewhat less relevant to most statistical graphics, which are typically two-dimensional in nature.
+While \cite{ekstrom1976manual} includes three-dimensional rotation tasks similar to \cite{shepard1971mental}, these three-dimensional tasks were not included in this experiment due to time constraints and limited face validity.}
 
@@
 The paper folding test is associated with the ability to extrapolate symmetry and reflection.
+
+\newtext{
+Tests in the Kit of Factor-Referenced Cognitive tests are classified according to the reasoning required to complete the test; in this way, we can be reasonably sure that the paper folding test, card rotation test, and figure classification test have construct validity.
+}
 
@@
 \Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
-Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
+Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception in the 25-28 bracket.
 This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the student population of Iowa State; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
@@
 %
-\subsection{Lineup Types}
+\subsection{Visual Aptitude for Specific Lineup Tasks}
 Figure~\ref{fig:lineuptaskmatrix} shows the correlations between the three lineup tasks and the figure classification, card rotation, and paper folding tasks. 
@@
 Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks.
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks, \newtext{which required participants to examine categorical data.}
 
@@
 
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
-
-Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
+Performance on lineups is related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
+\newtext{Additionally, due to the nature of observational studies, our results cannot be interpreted causally; students with a higher overall aptitude likely have strong visual abilities, but these same students may be more likely to choose STEM majors and score well on lineup tests.}
+
+Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than to tests of spatial ability.
 This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
 
-When lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
+When lineup tasks designed to examine different statistical features are viewed separately, there is some indication that different tasks are associated with different visual abilities.
 Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).
@@
 
-The demographic results from this study indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
+The demographic results from this study also indicate that in future lineup studies, it may be important to record information about participants' mathematical training, so that studies can be compared across participant pools with more reliability.
 

### 2015-06-23 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
significance of first PC between three lineups?
Additional meeting changes - additional references, etc.

--- 2015-06-22 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
@@
 As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol ~~\cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
-\comment{Not sure where to add the Related Work section; the introduction is really the most similar thing, but adding another sub-section in would really kill the flow of the paragraph, in my opinion.}
-\hhnote{I agree, it'll disrupt the flow, but we'll have to add a related work section nevertheless. We could add a bit of discussion on separating lineups from plots judging significance.}
 
-\hhnote{ these are the references we are supposed to discuss (not sure, how to add the talbot one - that is a study extending cleveland and then heer's).
-first attempt: It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well or are even misleading viewers~\cite{correll2013}.
-Lineup tests are different as they do not ask participants to judge significances directly but are only based on comparisons.
-Statistical properties are then derived from this assessment.
-This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
-The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
-a correlation coefficient is based on only a few summary statistics, whereas e.g.
-a corresponding  scatterplot is showing a lot more details of the data which might lead participants to a different evaluation.
-This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
-}
 ``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
@@
 \newtext{
-Our goal in this study is not to evaluate the statistical significance of particular plots, but rather, to consider the cognitive demands that we require of participants in lineup studies. 
+Our goal in this study is not to evaluate the statistical significance of particular plots or visualization designs, but rather, to consider the cognitive demands that we require of participants in lineup studies.
 }
@@
 all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task.
+
 We hope to facilitate comparison of the lineup task to known cognitive tests, inform the design of future studies, and better understand the perception of statistical lineups.
@@
 In section \ref{sec:methods} we introduce the tests used in the study and describe how the tests are scored.
-In section \ref{sec:results} we discuss the study results and compare them with scores previously established test, that also take  demographic characteristics associated with test scores into account.
+In section \ref{sec:results} we discuss the study results and compare them with scores on previously established tests, accounting for demographic characteristics associated with test scores.
 We discuss multi-collinearity in the study results, and use principal components analysis and linear models to draw some conclusions about the similarity between lineups and aptitude tests.
 Finally, in section \ref{sec:discussion}, we discuss the implications of the study results for the lineup protocol and  possible extensions.
+
+\paragraph{Related Work}
+\newtext{
+It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well and may even mislead viewers~\cite{correll2013}.
+Lineup tests approach the evaluation of statistical visualizations differently; they do not ask participants to judge significances or estimate numerical differences directly (as in \cite{cleveland:1985, fuchs, talbot}) but are only based on comparisons.
+Statistical properties are then derived from this assessment using results aggregated across individual participants.
+This has been shown to be similarly powerful to standard statistical tests in situations where these classical tests hold, and more powerful in situations where conditions of classical tests are not met~\cite{majumder2013validation}.
+The difference between statistical properties and visualizations is very easy to explain, as statistical significance of, e.g.
+a correlation coefficient is based on only a few summary statistics, whereas e.g.
+a corresponding scatterplot shows the data in more detail, perhaps leading participants to a different evaluation.
+This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
+}
+% \hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
 
@@
 \end{figure}
+
+\newtext{
+The remaining cognitive tests are adapted from \cite{ekstrom1976manual} and are highly validated tests that have been constructed to measure specific cognitive factors, such as inductive reasoning and spatial rotation.
+}
 
@@
 It is classified as part of the visualization factor in~~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
-The paper folding test is associated with the ability to extrapolate symmetry and reflection over multiple steps.
+The paper folding test is associated with the ability to extrapolate symmetry and reflection.
+
+\newtext{
+Tests in the Kit of Factor-Referenced Cognitive tests are classified according to the reasoning required to complete the test; in this way, we can be reasonably sure that the paper folding test, card rotation test, and figure classification test have construct validity.
+}
 
@@
 61\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
-Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception.
+Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception in the 25-28 bracket.
 This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the student population of Iowa State; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
 
-\subsection{Comparison of Spatial Tests with Previously Validated Results}
+\subsection{Comparison of Spatial Tests with Validated Results}
 The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
@@
 
-\begin{figure*}[h!tb]\centering
-\includegraphics[width=.85\linewidth]{fig-VisReasoningCategorical-1}
+\begin{figure}[h!tb]\centering
+\includegraphics[width=.95\linewidth]{fig-VisReasoningCategorical-1}
 \caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score.
@@
 
-\subsection{Understanding Visual Abilities and Lineup Performance}
+\subsection{Visual Abilities and Lineup Performance}
 
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:35 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Tue Jun 23 15:35:56 2015
 \begin{table}[htb]
@@
   \hline
-card.rot & 0.55 & -0.19 & -0.38 & 0.72 \\ 
-  fig.class & 0.46 & 0.58 & 0.66 & 0.14 \\ 
-  folding & 0.52 & 0.33 & -0.53 & -0.59 \\ 
-  vis.search & 0.46 & -0.72 & 0.38 & -0.34 \\ 
+Card Rot & 0.55 & -0.19 & -0.38 & 0.72 \\ 
+  Fig Class & 0.46 & 0.58 & 0.66 & 0.14 \\ 
+  Folding & 0.52 & 0.33 & -0.53 & -0.59 \\ 
+  VST & 0.46 & -0.72 & 0.38 & -0.34 \\ 
    \hline
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:36 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Tue Jun 23 15:35:57 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:36 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Tue Jun 23 15:35:57 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Mon Jun 22 15:06:36 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Tue Jun 23 15:35:58 2015
 \begin{table}[ht]
@@
 
-Performance on lineups is strongly related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-In addition to these demographic factors, many facets of intelligence are highly correlated; participants who score higher on general aptitude tests may score higher on tests of visual ability (and may also score higher on lineup tests).
+Performance on lineups is related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
+\newtext{Additionally, due to the nature of observational studies, our results cannot be interpreted causally; students with a higher overall aptitude likely have strong visual abilities, but these same students may be more likely to choose STEM majors and score well on lineup tests.}
 
-Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than tests of spatial ability.
+Despite these caveats, we have demonstrated that the general lineup task is most closely related to a classification task, rather than to tests of spatial ability.
 This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
 
-When lineup tasks with different goals are viewed separately, there is some indication that different tasks are associated with different visual abilities.
+When lineup tasks designed to examine different statistical features are viewed separately, there is some indication that different tasks are associated with different visual abilities.
 Lineup tasks 1 and 2 are quite similar, and are more associated with the figure classification task; lineup task 3, while still moderately correlated with the figure classification task, is also moderately correlated with the visuospatial ability tests (paper folding, card rotation).

### 2015-06-23 Paper/InfoVis2015Revision/appendix.Rnw
Additional meeting changes - additional references, etc.

--- 2015-03-22 Paper/InfoVis2015/appendix.Rnw
@@
 Figure~\ref{fig:scatterplottypes} shows the association between scaled score on each type of lineup and score on the visual reasoning tests.
-Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the card rotation task is much more significantly associated with the QQ-plots tasks compared to the other tasks.
-In addition, rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
+Sample size for each plot type is fairly small - between 5 and 10 plots per individual, so there is low power for systematic inference, but we can establish that the QQ plots are more strongly associated with the card rotation task than with the figure classification task, particularly when confidence bands are included.
+Rotated QQ-plots seem to be much more asssociated with the paper folding task scores than other QQ-plot tasks; this may be because they require more visual manipulation than other QQ-plots.
 

### 2015-06-23 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
Responses to reviewers

--- 2015-06-09 Paper/InfoVis2015Revision/ReviewerResponses.Rmd
@@
 ---
-title: "ReviewerResponses"
+title: "Responses To Reviewers"
 author: "Susan VanderPlas"
-date: "June 8, 2015"
-output: html_document
+date: "June 23, 2015"
+output: word_document
 ---
 
-# Summary Rating
-
-   4  (Accept) The paper should be accepted with some minor revisions. Once these have been completed it will meet the quality standard.
-
-## The Summary Review
-
-### Overview
-
-   The paper attempts to solve a quite interesting problem, which is
-   understanding the relationship between demographic characteristics /
-   visual skills and perception of graphical lineups. Reviewers agree that
-   this topic is quite interesting, and the authors did conduct a quite
-   extensive study. R4, particularly, is strongly supportive for this paper,
-   saying "The work represented here may be the most important recent
-   advance in methodology for evaluating the effectiveness of visual
-   displays." However, other reviewers listed several concerns in
-   experimental design, data analysis, and conclusions.
-
-### Strengths
-
-   - The topic of this study is very interesting. (R3, R4)
-   - The paper is well presented. (R1, R2, R3, R4)
-   - The study is extensively evaluated. (R1)
-   - Supplementary materials provide additional details on methodologies
-   used. (R2)
-   - A groundbreaking methodological contribution is made. (R4)
-   - Deal with important issues in spatial reasoning on graphics. (R1)
-
-### Weaknesses
-
-   - A clearer motivation of certain experimental design choices is
-   necessary. (R1)
-   - Experiment procedures should have more proper control and be more
-   detailed. (R1, R3)
-   - The statistical analysis is of questionable validity. (R2, R3)
-   - Results are not conclusive and open to interpretation. (R1, R2)
-   - A lack of discussion on visualization design in general. (R3)
-   - A separate related work section is necessary. (R3)
-
-   However, these concerns, especially regarding statistical analysis, are
-   rather critical. The reviewers recommended to accept this paper in the
-   condition that these weaknesses are properly handled within the revision
-   cycle. Thus, as the primary reviewer of this paper, I will review the
-   revision of the paper and ask the following questions, and if the answers
-   to all the questions are "yes", I will recommend to accept this paper.
-
+# Major Concerns
 1. Are all the experimental design choices properly motivated? (refer to
    the first concern of R1).    
+*The purpose of each of the cognitive tests is described in section 2.2.
+In summary, the paper-folding, card rotation, and figure classification tasks are part of the kit of factor-referenced cognitive tests (and are thus extremely well-studied); additionally, the authors chose the specific visual searching task (VST) because of its similarity of form to statistical lineups and its' compatibility with the written form of the other cognitive tests and lineup tests.*
+*As mentioned in the paper, the paper folding and card rotation tests have been used in many previous studies of statistical graphics and data displays.
+The figure-classification task examines inductive reasoning, which is another component of the lineup task (determining the dimension on which the lineup components differ, then applying that to answer the question successfully).*
 2. Are the experimental procedure sufficiently detailed? (refer to the
@@
 9. Does the revision handle all the minor issues raised by reviewers?
-
-Second round comments (public)
-
-Second round supplementary materials check
-
-Second round supplementary materials comments
-
-# Review 2
-Title: Spatial Reasoning and Data Displays
-
-_Reviewer_: primary
-
-_Paper type_   Evaluation
-
-_Expertise_ 2  (Knowledgeable)
-
-_Overall Rating_ 3.5 - Between Possible Accept and Accept
-
-_Supplemental Materials_ Acceptable
-
-## Justification
-
-   Overall, the study appears to be meticulously conducted, and the research
-   question of "what kinds of factors really influence on visual tasks" is a
-   very needed to be scientifically answered. However, I have a concern
-   whether PCA is a proper inferencing tool (I think it as a more
-   dimensionality reduction tool) and the resulting take-home messages are
-   rather unclear.
-
-## The Review
-
-   Overall, various statistical methods (e.g., scaling scores) appeared to
-   be meticulously used, and the authors even provided supplementary
-   materials to detail such processes, which shows a great practice. Using a
-   previously tested approach, line-ups, to investigate which factors will
-   impact on visualization task performances is also an interesting idea. I
-   believe that our community needs more work like this.
-
-   However, I have a concern. The authors largely rely on principal
-   component analysis (PCA) to infer what kinds of cognitive abilities are
-   relevant with the line-up tasks. However, I wonder if that is a
-   statistically sound approach. I am not a statistician, so I cannot make a
-   definitive claim that this is a right or wrong approach (I hope that
-   other reviewers have a stronger background on this approach).
-
-   However, I found the following page
-   (http://www.alglib.net/dataanalysis/principalcomponentsanalysis.php)
-   describing the underlying assumptions of PCA:
-
-   "It may be noted that the PCA is based on the following assumptions:
-
-   The assumption that the dimensionality of data can be efficiently reduced
-   by linear transformation
-   The assumption that most information is contained in those directions
-   where input data variance is maximum.
-   As it is evident, these conditions are by no means always met. For
-   example, if points of an input set are positioned on the surface of a
-   hypersphere, no linear transformation can reduce dimension (nonlinear
-   transformation, however, can easily cope with this task). This
-   disadvantage is equally attributable to all linear algorithms, and it can
-   be eliminated due to the use of complementary dummy variables that are
-   nonlinear functions of the input data set elements (the so-called "kernel
-   trick").
-
-   The second disadvantage of the PCA consists in the fact that the
-   directions maximizing variance do not always maximize information. The
-   page of the LDA subroutines gives an example of such a task, wherein the
-   maximum-variance variable affords almost no information, whilst the
-   minimum-variance variable permits classes to be wholly separated. In this
-   case, the PCA will give preference to the first (less informative)
-   variable. This drawback is closely connected to the fact that the PCA
-   does not perform linear separation of classes, linear regression or other
-   similar operations, but it merely permits the input vector to be best
-   restored on the basis of the partial information about it. All additional
-   information pertaining to the vector (such as the identification of an
-   image with one of the classes) is ignored."
-
-   Based on this description, and if this description is correct, I believe
-   that the authors should provide more concrete evidence showing that such
-   assumptions are met in their analysis, so that the authors can make some
-   inferences from the PCA results. However, it is unclear to me based on
-   the current manuscript.
-
-   In addition, I found that it is unclear what are the take-home messages
-   from the current analysis. It sounds that more research should be done.
-   If that is the case, this study sounds like a quite exploratory at this
-   stage.
-
-   * page 2
-   “speed[9]”
-
-   Put a space right before a citation throughout this paper.
-
-   * page 3
-   "Lineups require similar manipulations in two-dimensional space, and also
-   require the ability to perform complex spatial manipulations mentally;
-   for instance, comparing the interquartile range of two box- plots as well
-   as their relative alignment to a similar set of two boxplots in another
-   panel."
-
-   I understand that the experiment follows the line-up protocols, but it is
-   unclear how it was conducted in detail. What are the visualizations used
-   for the three blocks of 20 lineups. What are tasks? (Task examples are
-   shown in Figure 1(a), but not sure if the same task was used for all
-   lineups). I later found that more examples of line-up tasks are described
-   in Section 3.6, but the detailed descriptions of tasks are still
-   necessary.
-
-   * page 6
-   "3.5 Lineup Types"
-
-   This should be go to the Methods section, not the result section.
-
-   ---
-
-   Regarding the concern regarding PCA, a reviewer who had the sufficient
-   statistics background mentioned that my concern is irrelevant to the
-   application in the paper. I hope that the authors clarify potential
-   questions from readers by clarifying why PCA would be an appropriate
-   approach for the people who do not have sufficient statistical
-   background.
-
-# Review 1
-
-_Title_: Spatial Reasoning and Data Displays
-
-_Reviewer_:           external
-
-_Paper type_   Evaluation
-
-_Expertise_ 2  (Knowledgeable)
-
-_Overall Rating_ 3.5 - Between Possible Accept and Accept
-
-_Supplemental Materials_  Not applicable (no supplemental materials were submitted with the paper)
-
-## Justification
-
-The paper addresses an important issue in spatial reasoning on graphics.
-   While the paper is well presented and the study is extensively evaluated,
-   I find that the paper could benefit from stronger motivation of certain
-   design choices, better explanation of experimental procedures, and more
-   conclusive evaluation of the results.
-
-## The Review
-
-   The paper addresses an important issue in spatial reasoning on graphics
-   that I believe is highly valuable to the InfoVis community. The paper is
-   well presented and the study is extensively evaluated.
-
-   My main concerns about the paper are threefold.
-
-   Firstly, the paper could benefit from stronger motivation of certain
-   design choices. For instance, it is not fully clear how the measures of
-   visuospatial ability were chosen from a much larger set of tests in the
-   given reference. How are these related to spatial and reasoning ability
-   and what were the hypotheses about their relation to the graphics
-   presented in the lineup tasks?
-
-   Secondly, the descriptions of the experimental procedures is lacking
-   detail. In 2.2 the different tests are explained but it is not clear in
-   which order these were presented. Was it a randomised or counterbalanced
-   design? Also, what was contained in the 20 line ups presented between
-   each session in relation to the different plot types presented in Fig 9?
-   Why were the lineups interleaved with the other four tests? What was the
-   experimental setup: screen used, viewing distance, lab environment? What
-   were the instructions for each of the experimental sessions? What was the
-   total experiment time and were breaks included? Beware that such studies
-   can be influenced by many factors, such as order effects, fatigue, and
-   learning effects, for which reason it is really important to provide
-   sufficient details on the experimental design.
-
-   Lastly, while the evaluation appears to be very extensive, I find the
-   results not as conclusive as they are interpreted by the authors. Some
-   rather low correlations around 0.3-0.4 are interpreted by the authors as
-   being high, which I would not agree with. For instance, in Fig 10
-   differences in correlation between 0.2 and 0.3 are interpreted to be
-   significant while the respective scatter plots do not suggest for one or
-   the other to be significantly better. Such small differences in low
-   correlation regimes can often be misleading and should not be
-   overinterpreted. I appreciate the authors' effort to also draw strong
-   conclusions on the results of the PCA, but also these appear to be not
-   overly convincing. For instance, at the end of 3.5 the author state that
-   lineup 3 is different from lineup 1 and 2. I agree with that from visual
-   observation in Fig 9, but I would be hesitant to infer that from the PCA
-   results.
-
-   Some more detailed and minor comments:
-   - In 2.2, Wolfe's work on visual search should be mentioned in relation
-   to the VST.
-   - Equation (2) states the same as equation (1).
-   - In section 3, you mention one exception in relation to the
-   participants' age, please be more specific.
-   - In section 3, I suppose you mean that the demographics are
-   representative for the student population and not for the whole
-   university incl staff.
-   - With regard to your assessment of the correlation in Fig 6. I would
-   discriminate more between the individual results rather than stating that
-   they are all highly correlated.
-   - It would be good to move Table 1 to the previous page.
-   - Concerning Fig 5: I am not surprised that there is no difference for
-   age, as the difference is minimal and the groups would probably be put
-   into the same group would there be a wider spread of age in the
-   participant population.
-   - In the last paragraph of section 3.3: There is a typo "... is
-   ultimately are a classification... ".
-   - Beginning of 3.4: you mention that all demographic variables are highly
-   correlated. Really all? Please be more specific.
-   - Task for lineup 3 is given. What were the tasks given for the
-   evaluation of lineups 1 and 2?
-   - The VST task is excluded from Fig 10 due to its low correlation. Please
-   include it for completeness.
-   - PCA is explained in detail in 3.5 although it has been used extensively
-   earlier. Please put the explanation before the first use.
-   - Section 4: I don't agree with the initial statement that lineups are
-   strongly related to the performance on tests of visual ability. There is
-   certainly evidence provided for that in the paper, but I would not refer
-   to it as strong evidence.
-   - I propose to put Appendix C into the main body of the paper.
-   - In Appendix C: I don't agree with the statement that the card rotation
-   task is much more associated with the QQ plots. The correlations are
-   actually higher for the first four plot types on the left in the figure.
-
-# Review 3
-
-_Title_: Spatial Reasoning and Data Displays
-
-_Reviewer_:           external
-
-_Paper type_  Evaluation
-
-_Expertise_ 2  (Knowledgeable)
-
-_Overall Rating_ 2.5 - Between Reject and Possible Accept
-
-_Supplemental Materials_ Acceptable
-
-## Justification
-
-   This paper pairs visualization judgments with cognitive tests to try to
-   disentangle cognitive strategies for asserting the "most different" plot
-   in a set. The paper asks an interesting question and is very well
-   written. However, it is unclear from the manuscript what the experimental
-   task is and how it's controlled for, and the analysis is of questionable
-   validity as it aggregates results across a wide breadth of individual
-   designs without consideration for how different visualization designs
-   might lead to different cognitive strategies. As a result, the paper is a
-   near-miss: it asks an interesting question, but makes a few very critical
-   oversights that make its results both questionable and of little utility
-   for the visualization community.
-
-## The Review
-
-   The goal of this paper is to measure the role of different forms of
-   spatial reasoning in interpreting visualizations, with the ultimate goal
-   of understanding the visual strategies underlying judgments made using a
-   visualization. Turning to cognition to try to illuminate different
-   strategies for interpreting visualizations is an extremely interesting
-   idea. However, there are several elements of the experimental design and
-   analysis that are concerning. Of particular concern are the ambiguity of
-   the lineup task and the lack of consideration of different design
-   elements within the analysis. While I believe that the paper in its
-   current form is unsuitable for InfoVis, the manuscript is headed in the
-   right direction and could be a significant contribution to the community
-   in the future.
-
-   It is unclear in the experimental task what “most different” means.
-   From the text, it seems as if its intended to measure distributional
-   variation. However, “most different” can (and likely will) be
-   interpreted by participants in different ways (e.g. largest difference in
-   variance, largest difference in mean or median). Was this communicated to
-   the participants? Were they given a tutorial? How was the difficulty of
-   this judgment controlled? How were other potentially related statistics
-   controlled for (e.g. how can we verify that participants were actually
-   performing the requested task and not a related proxy task)? What
-   information was available to participants for each visualization design
-   to make this judgment? In a box plot, for example, we’re only getting
-   distributional characteristics, it’s possible that the box plot is
-   insufficient to characterize the distribution effectively (e.g.
-   Anscombe’s quartet).
-
-   A larger issue related to this point is that the manuscript lacks a
-   discussion of visualization design in general. For example, the
-   manuscript clearly indicates that  there is a difference in the reported
-   correlations across different sets of visualizations (e.g. lineup one,
-   two, and three). This directly suggests that the different visualization
-   design components , which is well-explored in the graphical perception
-   literature (e.g. Cleveland & McGill, 1984, Heer, Kong, & Agrawala, 2009,
-   Fuchs et al, 2013, Albers, Correll, & Gleicher, 2014 among many others).
-   This would also imply that the findings as reported are biased by the
-   sampling of visualization types used in the experiment. Further, the
-   comparison within lineups seems somewhat unprincipled. For example,
-   it’s unclear how a dot-plot is more similar to set one than set two.
-
-   This will likely be a significant confound in the comparisons with the
-   more traditional  cognitive tests. Assuming the tests use exactly the
-   same form of stimuli founding figures 2 and 3, these stimuli are uniform
-   in design elements across all of the visualization designs. As a result,
-   it’s difficult to tell if the results are a by-product of similarity in
-   visual features or in visual strategy. This analysis feels like it might
-   be comparing apples to oranges--I am not convinced that it makes sense to
-   compare the lineup results to performance on these tests in aggregate due
-   to complications in visual design. Given this confound, it is unclear how
-   a visualization designer could use these results in practice.
-
-   For example, the lineup test is an example of a more free-form visual
-   search task: participants are actively searching for the “most
-   different” element. As a result, deviations from these two tasks might
-   be a function of difference in strategy (as suggested in the paper) or
-   due to difference in stimuli (e.g. detecting configurations vs. box
-   plots) or both. The figure classification test may provide the most
-   similar results for many visualization designs simply because the visual
-   hull of the data (the “cloud” in a scatterplot, the box in a boxplot)
-   is the most salient global feature for making similarity comparisons,
-   which aligns well with the "form" of a figure. This would lead to
-   correlation between the tests, but the actual strategy may not be
-   rule-based.
-
-   To make these comparisons more compelling, these same experimental
-   procedures can be conducted with the stimuli in each of these tests
-   adapted to more closely mirror the tested visualization type. I would
-   encourage the authors to consider restructuring these tests to use
-   stimuli more relevant to each respective visualization lineup condition
-   or, at a minimum, to at least break down the results with respect to each
-   individual visualization design and compare strategies across visual
-   feature. Aggregating across all visualizations is an oversimplification
-   of the task—different visualization may require different strategies,
-   and, in fact, graphical perception suggests that they do.
-
-   I would also recommend that a revised manuscript include an explicit
-   related works section. Traditionally in this community, the related work
-   is separate from the introduction. While I think this paper reads well
-   without that structure, the authors may wish to consider revising their
-   structure to accommodate a more thorough discussion of related work. Part
-   of this recommendation is that there is a good breadth of relevant work
-   that is not considered in this manuscript. For example, “The lineup
-   protocol [12, 25, 3] is a testing framework that allows researchers to
-   quantify the statistical significance of a graphical finding with the
-   same mathematical rigor as conventional hypothesis tests” is a very
-   contentious claim. In particular, recent work in visualization, such as
-   (Pak, Hutchinson, & Turk-Brown, 2014, Talbot, Setlur, & Anand, 2014 (more
-   distant comparison is harder), Correll & Gleicher, 2014, and a myriad of
-   papers on limitations in interpreting boxplots), suggest that people are
-   not good at performing visual significance testing. Visual search is
-   among the most commonly used tasks in graphical perception experiments.
-   For example, Fuchs et al, 2014 includes experiments that explicitly
-   involve visual search tasks in small multiple (lineup) conditions.
-
-   Minor Issues:
-   Additional experimental details about the type of visualizations tested
-   (e.g. Figure 9) should be presented in the Methods section. A pointer to
-   another set of papers with large numbers of potential designs is
-   insufficient to understand the experiment. Additionally, explanations of
-   the baseline cognitive tests should be presented up front to ground the
-   analysis: while many in the visualization community may have seen the
-   compared tests (visual search task (VST), paper folding test, card
-   rotation test, and figure classification test), especially the visual
-   search task, many other have not. Give the necessary grounding and
-   explain why these particular tasks were chosen as a baseline.
-
-## Spelling & Grammar:
-
-   Overall this paper is very well written. However, throughout the paper,
-   the term “graphics” is used instead of “visualization”.
-   “Graphics” has a much broader possible meaning and
-   “visualization” is conventional for the InfoVis audience.
-
-   speed[9]  -> speed [9]
-   reasoning[6] -> reasoning [6]
-
-# Review 4
-
-_Title_: Spatial Reasoning and Data Displays
-
-_Reviewer_:           secondary
-
-_Paper type_   Theory / Model
-
-_Expertise_   3  (Expert)
-
-_Overall Rating_ 5 - Strong Accept    
-The paper is an excellent contribution and should certainly be accepted, possibly with some minor revisions.    
-It more than meets the quality standard.
-
-_Supplemental Materials_ Acceptable
-
-## Justification
-
-   Nothing that I haven't listed in the review. I think this is an
-   extraordinary paper because the authors have taken a fundamentally
-   important contribution and extended it to an area where nobody has
-   thought about it so far.
-
-## The Review
-
-   The work represented here may be the most important recent advance in
-   methodology for evaluating the effectiveness of visual displays. It
-   appears to be, as well, the most powerful method for evaluating
-   individual differences in graph reading ability. What is remarkable is
-   that such a simple paradigm can be coupled with a formal statistical
-   distribution (the binomial) for quantifying these results. Instead of
-   separating graph reading into subtasks that may be highly correlated (and
-   therefore not highly informative taken as a whole set) this team has
-   devised a method for dealing with highly complex displays (probability
-   plots, billets, etc.) as well as simple ones.
-
-   Moreover, this team has implemented a methodology that fits perfectly
-   into a tradition of careful psychological measurement. Instead of
-   requesting subjects to rate things numerically (as Cleveland did, for
-   example), these researchers simply ask for the point of a finger, so to
-   speak. The response is similar in nature to that of classical perceptual
-   direct scaling studies going back almost a century.
-
-   Sometimes authors or teams milk a simple idea by generating minor
-   variations and submitting them to a bunch of journals. What's remarkable
-   here is that the collection of their publications given in the references
-   ranges from a formal statistical article with proofs etc. to several
-   visualization articles having very different focuses. The same is true
-   for this article. I have seen presentations of this research and I never
-   thought of the possibility that it could be used in the context of the
-   present article.
-
-   This paper is very well organized and written. I have few considerations
-   for revision and plan to nominate it for best paper.
-
-### Details:
-
-   I'd suggest a mention of the groundbreaking mental rotation research of
-   Shepard & Metzler. This was 3D rotation, but it did stimulate thinking by
-   later psychologists on the rotation problem. What's more, their results
-   were confirmed by subsequent neurological research in Science magazine.

### 2015-06-25 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
Responses to reviewers, removing \newtext{} from new text for the resubmit.

@@
 ---
-title: "Responses To Reviewers"
-author: "Susan VanderPlas"
-date: "June 23, 2015"
+title: "Summary of Changes"
+author: "Susan VanderPlas & Heike Hofmann"
 output: word_document
@@
 # Major Concerns
-1. Are all the experimental design choices properly motivated? (refer to
-   the first concern of R1).    
+1. **Are all the experimental design choices properly motivated?**    
+**(refer to the first concern of R1).**    
 *The purpose of each of the cognitive tests is described in section 2.2.
@@
 The figure-classification task examines inductive reasoning, which is another component of the lineup task (determining the dimension on which the lineup components differ, then applying that to answer the question successfully).*
-2. Are the experimental procedure sufficiently detailed? (refer to the
-   second concern of R1 and )
-3. The results of statistical analysis do not seem to be conclusive, are
-   the interpretations of these results properly revised? (refer to the
-   third concern of R1)
-4. Does the revision contain additional discussion of why PCA is a
-   proper approach to interpret the data for the VIS community members who
-   do not have sufficient statistics background? (refer to the concern of
-   R2)
-5. Does the revision has the clear description of "most different"?
-   (refer to the first concern of R3)
-6. Does the revision has the proper discussion of visualization design
-   (refer to the second concern of R3)
-7. Does the revision resolve a potential concern of confounding factors
-   (refer to the conern of R3)
-8. Does the revision have an explicit related works section? (refer to
-   the last concern of R3)
-9. Does the revision handle all the minor issues raised by reviewers?
+2. **Are the experimental procedure sufficiently detailed?**    
+**(refer to the second concern of R1 and )**    
+*We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown).
+The testing documents are comprehensive, and oral instructions mirror the instructions on the introduction page of each task (students are familiar with this format as it is commonly used on standardized tests, such as the SAT and ACT).*
+3. **The results of statistical analysis do not seem to be conclusive, are the interpretations of these results properly revised?**    
+**(refer to the third concern of R1)**    
+*The results of the statistical analysis are as conclusive as the nature of the study allows.
+While correlations of 0.3-0.4 are low compared to the absolute range of 0 - 1, they are strong for human subjects data (which is how we have interpreted them here).
+Because so many different factors influence human behavior, we generally do not see correlations above 0.5 on distinct tasks.*
+*We have provided both graphics and PCA results in order to demonstrate our results in more than one way; while these results are preliminary, the cost involved in designing and executing a study which would enable multivariate hypothesis testing (given the multiple correlations between dependent variables) is prohibitive.
+This study is the first to examine the visual and cognitive skills necessary to evaluate statistical lineups; follow-up studies may provide more statistically precise results.*
+4. **Does the revision contain additional discussion of why PCA is a proper approach to interpret the data for the VIS community members who do not have sufficient statistics background?**    
+**(refer to the concern of R2)**    
+*The revision includes a reference to a PCA tutorial (for those unfamiliar with the procedure) and explains the rationale for utilizing PCA in some additional detail.
+Section 3.3 - Visual Abilities and Lineup Performance has been modified extensively to provide additional details about PCA.*
+5. **Does the revision has the clear description of "most different"?**    
+**(refer to the first concern of R3)**    
+*The use of 'most different' is a feature of the lineup protocol, and does not bias the participants to examine specific distributional characteristics.
+While participants may choose different distributional features to identify the 'most different' plot, this mirrors how we perceive graphics (we cue in to different features).
+In addition, many statistical tests react to different features of the data - a t-test may not produce significant results even with a truly significant difference in means if the two distributions have wildly different variances.
+Finally, the use of 'most different' allows the lineup protocol to be used even in populations which do not have mathematical training and might not understand the use of terms such as mean and variance.*
+*We have added additional language in Section 2.1 to emphasize that the phrasing is a standard part of the lineup protocol.*
+6. **Does the revision has the proper discussion of visualization design**    
+**(refer to the second concern of R3)**    
+*This study was not designed to directly impact visualization design recommendations.
+While the lineup protocol can easily be used to compare many different visualization designs, the goal of this experiment was to examine the cognitive requirements of lineup evaluation at the participant level.
+As such, it would not be appropriate to discuss visualization design in detail.*
+*While there are differences in the correlation of scores on lineup sets 1, 2, and 3 with the cognitive tests, this does not provide any additional information about whether designs in lineup set 1 are superior to those in lineup set 2 and lineup set 3, and in fact, the sets of lineups are not comparable in this way.
+Each set of lineups was taken from an experiment which did aim to rank designs by effectiveness, but in this study we are instead examining individual differences in cognitive abilities and lineup performance.*
+7. **Does the revision resolve a potential concern of confounding factors**    
+**(refer to the conern of R3)**    
+*The issue of confounding factors is related to the issue raised in #6, that is, that we have not discussed visualization design in this paper.
+As explained above, our goal was not to compare types of visualizations, but rather to examine individual differences in cognitive abilities and performance on lineup tasks.
+The tasks which underlie each set of lineups (as described in the caption of Fig.
+2 in the revision) are common tasks in statistical graphics.
+In addition, the similarity in test protocol between the VST, the tests from the factor-referenced kit of cognitive tests, and the lineup protocol was intentional; by controlling the experimental protocol, we could reduce participant confusion and the corresponding variability.
+As our goal was not to examine the elements of visualization design experimentally, controlling those elements is one element of appropriate experimental design.*
+8. **Does the revision have an explicit related works section?**    
+**(refer to the last concern of R3)**    
+*An explicit related works section has been added at the conclusion of Sec. 1.*
+9. **Does the revision handle all the minor issues raised by reviewers?**
+
+# Minor Issues
+**Minor issues raised by Reviewer 1:**
+
+- In 2.2, Wolfe’s work on visual search should be mentioned in relation to the VST.    
+*A citation has been added, but Wolfe's work on visual search differs from the VST used in this study; while the underlying cognitive mechanisms may be the same, a detailed explanation of the feature integration component of visual search seems extraneous to the study described in this paper.*
+- Equation (2) states the same as equation (1).    
+*The repetition has been removed due to space constraints.*
+- In section 3, you mention one exception in relation to the participants’ age, please be more specific.    
+*This wording has been modified to provide more detail*
+- In section 3, I suppose you mean that the demographics are representative for the student population and not for the whole university incl staff.    
+*Indeed, though the wording has been modified for clarity in the revised draft.*
+- With regard to your assessment of the correlation in Fig 6. I would discriminate more between the individual results rather than stating that they are all highly correlated.    
+*Additional wording has been added to this effect.* 
+- It would be good to move Table 1 to the previous page.    
+*Due to LaTeX's float placement, it is difficult to place a figure/table which spans two columns on the same page as a lone single-column figure.
+Table 1 is as close to Section 3.1 as is possible given these constraints.*
+- Concerning Fig 5: I am not surprised that there is no difference for age, as the difference is minimal and the groups would probably be put into the same group would there be a wider spread of age in the participant population.
+*Absolutely; however, even in pilot studies (in populations with more age variability) there did not seem to be a difference due to age in lineup evaluations.
+Age was included as a demographic variable, but we did not expect there to be significant differences due to age.*
+- In the last paragraph of section 3.3: There is a typo “… is ultimately are a classification…”.    
+*This has been resolved in the revised draft.*
+- Beginning of 3.4: you mention that all demographic variables are highly correlated. Really all? Please be more specific.    
+*This has been re-phrased.*
+- Task for lineup 3 is given. What were the tasks given for the evaluation of lineups 1 and 2?    
+*We are not sure what part of the paper you're referencing here; we have added additional information about lineup set 1 and 2 at the beginning of Section 3.5 for clarity.*
+- The VST task is excluded from Fig 10 due to its low correlation. Please include it for completeness.    
+*Fig. 10 has been modified accordingly.*
+- PCA is explained in detail in 3.5 although it has been used extensively earlier. Please put the explanation before the first use.    
+*The description of PCA in Sec.
+3.5 is specifically focused on the development of Fig.
+11, which is not a traditional plot used to display PCA data.
+Additional detail about PCA has been added earlier in the paper, but the development of the influence plot remains in Sec.
+3.5 as well.*
+- Section 4: I don’t agree with the initial statement that lineups are strongly related to the performance on tests of visual ability.
+There is certainly evidence provided for that in the paper, but I would not refer to it as strong evidence.
+*We have added additional language to clarify that 'strong' here is relative to other human subjects data, which typically is more variable and thus has lower correlations than data from other scientific pursuits.*
+- I propose to put Appendix C into the main body of the paper.    
+*Due to space constraints, Appendix C does not fit in the main body of the paper.
+Additionally, due to small sample sizes, more work is required to draw any strong conclusions from the analysis in Appendix C.
+Analyzing the correlations on the individual plot design level would be an interesting follow up study (and thus, we have provided an initial analysis as an appendix).*
+- In Appendix C: I don’t agree with the statement that the card rotation task is much more associated with the QQ plots.
+The correlations are actually higher for the first four plot types on the left in the figure.
+*This statement has been modified for clarity.*
+
+**Minor issues raised by Reviewer 2:**
+
+- page 2 “speed[9]”: Put a space right before a citation throughout this paper.    
+*This change has been made throughout.*
+- page 3 “Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two box- plots as well as their relative alignment to a similar set of two boxplots in another panel.”
+I understand that the experiment follows the line-up protocols, but it is unclear how it was conducted in detail.
+What are the visualizations used for the three blocks of 20 lineups.
+What are tasks?
+(Task examples are shown in Figure 1(a), but not sure if the same task was used for all lineups).
+I later found that more examples of line-up tasks are described in Section 3.6, but the detailed descriptions of tasks are still necessary.
+*The additional detail provided in the caption of Fig.
+2 addresses this; we have also included the test documents to provide more detail and facilitate replication of this experiment.
+In addition, we have moved the figure in Section 3.6 to the methods section so that this part of the experimental design is clear from the beginning.*
+- page 6 “3.5 Lineup Types”: This should be go to the Methods section, not the result section.    
+*The analysis has remained in section 3.5, but the title has been modified for clarity. The figure with all of the lineup types has also been moved to the methods section.*
+
+**Minor issues raised by Reviewer 3:**
+
+- Additional experimental details about the type of visualizations tested (e.g.
+Figure 9) should be presented in the Methods section.
+A pointer to another set of papers with large numbers of potential designs is insufficient to understand the experiment.
+*Some additional detail has been added, but we have also included the test documents as supplementary material; they are designed to provide all of the instructions a participant would need to complete the experiment (much as standardized tests are designed to require relatively little instruction outside of the test booklet).*
+- Additionally, explanations of the baseline cognitive tests should be presented up front to ground the analysis: while many in the visualization community may have seen the compared tests (visual search task (VST), paper folding test, card rotation test, and figure classification test), especially the visual search task, many other have not.
+Give the necessary grounding and explain why these particular tasks were chosen as a baseline.
+*Some additional justification of the cognitive tests has been added, but many of these tests are commonly used in studies of statistical graphics and cognition (and those studies have been cited in the test description section).
+The visual search task (VST) has not been previously used in statistical graphics, but the justification for using this form of visual search task is provided in Section 2.2.*
+- Throughout the paper, the term “graphics” is used instead of “visualization”. “Graphics” has a much broader possible meaning and “visualization” is conventional for the InfoVis audience.    
+*Visualization is a much broader term than ``statistical graphics'', but we have attempted to clarify the use of the term graphics as much as possible.*
+- speed[9] -> speed [9] reasoning[6] -> reasoning [6]    
+*This change has been made throughout the paper.*
+
+**Minor issues raised by Reviewer 4:**
+
+- I’d suggest a mention of the groundbreaking mental rotation research of Shepard & Metzler.
+This was 3D rotation, but it did stimulate thinking by later psychologists on the rotation problem.
+What’s more, their results were confirmed by subsequent neurological research in Science magazine.
+*We have cited the canonical Shepard & Metzler paper, but as graphics are typically two-dimensional we did not utilize the three-dimensional rotation task from the Kit of Factor Referenced Cognitive Tests.
+This would be an interesting addition to future studies.*

### 2015-06-25 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Responses to reviewers, removing \newtext{} from new text for the resubmit.

@@
 In this paper, we present the results of a study designed to compare lineup performance with visual aptitude and reasoning tests, examining the skills necessary to successfully evaluate lineups. 
-\newtext{
 Our goal in this study is not to evaluate the statistical significance of particular plots or visualization designs, but rather, to consider the cognitive demands that we require of participants in lineup studies.
-}
+
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test. 
@@
 \paragraph{Related Work}
-\newtext{
 It has been shown~\cite{pak2014} that visualizations do not necessarily reflect statistical properties well and may even mislead viewers~\cite{correll2013}.
@@
 This, however, is not a drawback of the visualization but only shows that summary statistics might not be enough to come to an appropriate conclusion \cite{anscombe:1973}.
-}
-% \hhnote{XXX still missing: talbot reference; not sure, how to add the talbot one - it is a really nice study  extending cleveland's and then heer's, but not directly related.}
 
@@
 This sheet of charts is then shown to human observers, who are asked to identify the display that is ``the most different"
-\newtext{This phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies employing lineups~\cite{wickham2010graphical, majumder2013validation}.}
+This phrasing is intentionally ambiguous to prevent biasing participants and is commonly used in other studies employing lineups~\cite{wickham2010graphical, majumder2013validation}.
 If an observer identifies the plot drawn from the actual data, this can be reasonably taken as evidence that the data it shows is different from the data of other plots.
@@
 Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task. 
-\newtext{In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score. 
-}
+In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score.
+
 For this approach, we derive a score for an individual as follows:
@@
 \begin{eqnarray}\label{eq.scoring1}
-n_c - \frac{n_w}{m-1}.
+n_c - \textstyle\frac{n_w}{m-1}.
 \end{eqnarray}
@@
 As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.
-\newtext{This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization. }
-
-\newtext{Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
+This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization.
+
+Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
 This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic; this examination utilizes data aggregated over all participants who evaluated a specific lineup image.
@@
 Each block of lineups are a subset of stimuli used in experiments designed to compare the visual significance of different plot types, however, our purpose in this study is not to determine the efficacy of different types of plots.
-Rather, our goal is to compare the cognitive demands required to successfully evaluate these lineups, which display variance difference, mean difference, or violations of normality using common statistical plots.}
+Rather, our goal is to compare the cognitive demands required to successfully evaluate these lineups, which display variance difference, mean difference, or violations of normality using common statistical plots.
 
@@
 \protect\caption{Lineup plot types used in the experiment.
-\label{fig:lineups} \newtext{Each set of plots is representative of the plot types shown in a block of lineups in the experiment, and each block represents a subset of plots from a traditional lineup experiment to determine which visualization method facilitates identification of the plot with the strongest visual signal.
+\label{fig:lineups} Each set of plots is representative of the plot types shown in a block of lineups in the experiment, and each block represents a subset of plots from a traditional lineup experiment to determine which visualization method facilitates identification of the plot with the strongest visual signal.
 Set 1 manipulates group means and variance, set 2 manipulates group means, and set 3 examines violations of normality.
-These tasks were chosen as common statistical visualization tasks, and while not exhaustive, should provide a broad basis for interpreting the results of this experiment.}}
+These tasks were chosen as common statistical visualization tasks, and while not exhaustive, should provide a broad basis for interpreting the results of this experiment.}
 \end{figure*}
@@
 The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations, such as those in~\cite{wolfe1990limitations,treisman1980feature}.
-\newtext{We have included visual search in the cognitive testing battery for this experiment because it is the most simple of the tasks required to successfully evaluate a lineup; if participants cannot complete visual search tasks successfully, they will not be able to accurately identify which lineup plot is different.
-}
+We have included visual search in the cognitive testing battery for this experiment because it is the most simple of the tasks required to successfully evaluate a lineup; if participants cannot complete visual search tasks successfully, they will not be able to accurately identify which lineup plot is different.
 
@@
 
-\newtext{
 The remaining cognitive tests are adapted from \cite{ekstrom1976manual} and are highly validated tests that have been constructed to measure specific cognitive factors, such as inductive reasoning and spatial rotation.
-}
 
@@
 Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
-\newtext{Three-dimensional rotation is somewhat less relevant to most statistical graphics, which are typically two-dimensional in nature.
-While \cite{ekstrom1976manual} includes three-dimensional rotation tasks similar to \cite{shepard1971mental}, these three-dimensional tasks were not included in this experiment due to time constraints and limited face validity.}
+Three-dimensional rotation is somewhat less relevant to most statistical graphics, which are typically two-dimensional in nature.
+While \cite{ekstrom1976manual} includes three-dimensional rotation tasks similar to \cite{shepard1971mental}, these three-dimensional tasks were not included in this experiment due to time constraints and limited face validity.
 
@@
 
-\newtext{
 Tests in the Kit of Factor-Referenced Cognitive tests are classified according to the reasoning required to complete the test; in this way, we can be reasonably sure that the paper folding test, card rotation test, and figure classification test have construct validity.
-}
 
@@
 
-\newtext{All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; pilot tests did not indicate an effect of task order.
-}
+All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; internal pilot tests did not indicate an effect of task order.
 
@@
 
-\newtext{After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
+After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
 The duration of testing for each participant was approximately 1 hour.
-}
 
@@
 
-All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
-Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
-\begin{eqnarray}\label{eq.scoring}
-\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}
-\end{eqnarray}
-\newtext{Equation \ref{eq.scoring} is a restatement of equation \ref{eq.scoring1} provided for convenience.}
+All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(m-1)$, where $m$ is the total number of possible answers to the question.
+Thus, for a test consisting of multiple choice questions with $m$ distinct answers with a single correct answer each, the score is calculated as in equation \ref{eq.scoring1}.
+
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
@@
 Results from the visuospatial tests used in this experiment are highly correlated, as shown in Figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are in some way measuring individuals' visual ability.
-\newtext{While test scores are correlated, the relationships between test scores appear to be linear, with sufficient variability around the trend line to merit further investigation.
+While test scores are correlated, the relationships between test scores appear to be linear, with sufficient variability around the trend line to merit further investigation.
 The correlations are to be expected, as all of the tests in this experiment are conducted in a visual domain, and require general reasoning, intelligence, and other factors.
-}
-What is of more interest to us is 
-\newtext{whether lineups recruit specific cognitive abilities, such as visual search, reasoning and categorization ability, mental rotation.} 
+What is of more interest to us is whether lineups recruit specific cognitive abilities, such as visual search, reasoning and categorization ability, mental rotation.
 In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the battery of cognitive tests (other than lineups) using principal components analysis on the scaled test scores, then we examine all five tests (four cognitive tests, and additionally lineups) using the same procedure.
 
-\newtext{
 Principal component analysis (PCA) is a statistical technique to transform the coordinate system of correlated data, producing orthogonal components in a new coordinate space (see~\cite{pcaReference} for more information).
@@
 It is this feature of principal components which we utilize in this paper.
-}
 
@@
 Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks, \newtext{which required participants to examine categorical data.}
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks, which required participants to examine categorical data.
 
@@
 All lineup tasks are moderately correlated with the figure classification task, and while tasks 1 and 2 are most strongly correlated with figure classification, lineup task 3 is most strongly correlated with the card rotation and paper folding tasks.
-\label{fig:lineuptaskmatrix} \newtext{While the correlations shown in this graph are all less than 0.60, there are some surprisingly strong relationships for human subjects data.}}
+\label{fig:lineuptaskmatrix} While the correlations shown in this graph are all less than 0.60, there are some surprisingly strong relationships for human subjects data.}
 \end{figure}
@@
 For variables $X_i$, $i=1, ..., K$ a principal component analysis results in a set of $K$ principal  components $PC_j$, $j=1, ..., K$, given as the rows of the rotation matrix, $M$ (which has dimension $K\times K$, indexed by $i$ and $j$ respectively).
-
 The importance $I_j$ of each component is determined by the amount of variability the data exhibits along each of the principal axis.
-
 The influence of variable $X_i$ on component $PC_j$ is then defined as:
@@
 \section{Discussion and Conclusions}\label{sec:discussion}
-
 Performance on lineups is related to performance on tests of visual ability; however, this relationship is mediated by demographic factors such as major (STEM or not) and completion of calculus I.
-\newtext{Additionally, due to the nature of observational studies, our results cannot be interpreted causally; students with a higher overall aptitude likely have strong visual abilities, but these same students may be more likely to choose STEM majors and score well on lineup tests.}
+Additionally, due to the nature of observational studies, our results cannot be interpreted causally; students with a higher overall aptitude likely have strong visual abilities, but these same students may be more likely to choose STEM majors and score well on lineup tests.
 
@@
 This is an important verification of a tool that is useful for examining statistical graphics, as it emphasizes the idea that while the testing medium is graphical in nature, the task is in fact a classification task, where the viewer must determine the most important features of each plot and then identify which plot is different.
-
 When lineup tasks designed to examine different statistical features are viewed separately, there is some indication that different tasks are associated with different visual abilities.
@@
 All results and data shown here were collected and analyzed in accordance with IRB \# 13-581.
-
 \bibliographystyle{abbrv}

### 2015-06-25 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Responses to reviewers, removing \newtext{} from new text for the resubmit.

@@
 As a result, lineups are a useful and innovative tool for evaluating charts; on an individual level, they can also be used to evaluate a specific participant's perceptual reasoning ability in the context of statistical graphics.
-\newtext{This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization. }
+This study uses the lineup protocol to examine differences in individual performance, rather than using aggregate scores to determine the effectiveness of a single visualization.
 
-\newtext{Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
+Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
 This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic; this examination utilizes data aggregated over all participants who evaluated a specific lineup image.
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Tue Jun 23 15:35:56 2015
+% Thu Jun 25 09:34:25 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Tue Jun 23 15:35:57 2015
+% Thu Jun 25 09:34:26 2015
 \begin{table}[ht]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Tue Jun 23 15:35:58 2015
+% Thu Jun 25 09:34:26 2015
 \begin{table}[ht]

### 2015-06-25 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
couple of comments

@@
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
-Graphics utilize higher-bandwidth visual pathways to encode information~~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
+Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
 Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
@@
 
-Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~~\cite{vekiri2002value}.
+Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner~\cite{vekiri2002value}.
 Both of these theories suggest spatial ability impacts a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements.
-In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~~\cite{lowrie2007solving}.
-However,  mathematical ability, not spatial ability, was shown~~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
-Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed ~~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
-It is theorized that graphics are a form of ``external cognition"~~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~~\cite{zhang1997nature}.
+In addition, previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs~\cite{lowrie2007solving}.
+However,  mathematical ability, not spatial ability, was shown~\cite{shah1995conceptual} to be associated with accuracy on a simple two-dimensional line graph. 
+Spatial ability becomes more important when more complicated graphical displays are used in comparison tasks: the lower performance of individuals with low spatial ability on tests utilizing diagrams and graphs is attributed~\cite{mayer1994whom} to the fact that more cognitive resources are required to process the visual stimuli, which leaves fewer resources to make connections and draw conclusions from those stimuli.
+It is theorized that graphics are a form of ``external cognition"~\cite{scaife1996external} that guide, constrain, and facilitate cognitive behavior~\cite{zhang1997nature}.
 
 Here, we want to investigate the extent to which observers make use of spatial ability in their evaluation of and reasoning about statistical graphics.
-As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol ~~\cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
+As a quantitative measure of how successful an evaluation is, we are making use of the lineup protocol~\cite{buja2009statistical, wickham2010graphical, majumder2013validation}.
 
 ``Lineups" were introduced as a tool to evaluate the statistical significance of a graphical finding~\cite{wickham2010graphical}.
-Lineups are also useful in assessing the effectiveness of different graphical displays~\cite{hofmann2012graphical, loy:2015}.
-Like their police counterpart, lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
-Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; participants are expected to identify the most different among the plots shown.
-In this example, sub-plot 4 is the target  because of the noticeably different locations of the two boxplot medians.
+Lineups are also useful in assessing the effectiveness of alternative graphical displays~\cite{hofmann2012graphical, loy:2015}.
+Like their police counterpart, graphical lineups consist of several distractor plots (of randomly generated data) and one target (the data plot).
+Figure~\ref{fig:lineup}(a) shows a sample lineup of boxplots; in this example, sub-plot 4 shows the target data (recognizable by the shift between the two boxplot medians).
 
@@
 This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-While previous research~~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
+\hhnote{should we say something here about using charts only for comparisons?
+We almost do that already.
+Maybe just drive that message home?
+}
+While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
 
@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test.
+
 The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
@@
 \subsection{The Lineup Protocol}
-The lineup protocol~~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
+The lineup protocol~\cite{hofmann2012graphical,wickham2010graphical, buja2009statistical} is a testing framework that allows researchers to quantify the statistical significance of a graphical finding with the same mathematical rigor as conventional hypothesis tests.
 
@@
 Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task.
-In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score.
-
-For this approach, we derive a score for an individual as follows:
+
+In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score, derived as follows:
+
+%For this approach, we derive a score for an individual as follows:
 
@@
 Statistical lineups depend on the ability to search for a signal amid a set of distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
-Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from participants, e.g.~ polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
-By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
-
-Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups~\cite{humanfactorslineups}.
+Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from participants, e.g.~polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
+By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen\newtext{, externally validated,} cognitive tests to assess these aspects of visuospatial ability.
+
+Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors as well as gender had an effect on the amount of time spent on lineups~\cite{humanfactorslineups}.
 In addition, lineup performance can be partially explained using statistical distance metrics~\cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
@@
 The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning abilities (factor I  in~~\cite{ekstrom1976manual}).
+This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
 An example is shown in figure~\ref{fig:figureclassification}.
@@
 The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in~~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
 An example is shown in figure~\ref{fig:cardrotation}.
@@
 A sample question from the test is shown in figure~\ref{fig:paperfolding}.
-It is classified as part of the visualization factor in~~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
+It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
 The paper folding test is associated with the ability to extrapolate symmetry and reflection.
@@
 
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~~\cite{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
 Participants were given 5 minutes to complete each block of 20 lineups.
@@
 \subsection{Test Scoring}\label{sec:scaling}
-
-All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(m-1)$, where $m$ is the total number of possible answers to the question.
-Thus, for a test consisting of multiple choice questions with $m$ distinct answers with a single correct answer each, the score is calculated as in equation \ref{eq.scoring1}.
-
+%
+All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
+Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
+\begin{eqnarray}\label{eq.scoring}
+\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}
+\end{eqnarray}
+Equation \ref{eq.scoring} is a restatement of equation \ref{eq.scoring1} provided for convenience.
 This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests~~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~~\cite{educational1963kit}.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
 
@@
 Completion of Calculus I is associated with increased performance on lineups; this may be related to general math education level, or it may be that success in both lineups and calculus requires certain visual skills.
-This association is consistent with findings in~~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
+This association is consistent with findings in~\cite{shah1995conceptual}, which associated  mathematical ability to performance on simple graph description tasks.
 There is also a significant relationship between hours of video games played per week and score on lineups, however, this association is not monotonic and the groups do not have equal sample size, so the conclusion may be suspect.
-There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~~\cite{hausmann2000sex}.
+There is a (nearly) significant difference between male and female performance on lineups; this is not particularly surprising, as men perform better on many spatial tests~\cite{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women~\cite{hausmann2000sex}.
 There is no significant difference in lineup performance for participants of different age, self-assessed skills in various domains, previous participation in math or science research, completion of a statistics class, or experience with AutoCAD.
@@
 %The visual search task is only slightly correlated with the three lineup tasks and is therefore omitted from this figure.
+
 Performance on lineup tasks 1 and 2, which dealt with the distributuion of two groups of numerical data, is most strongly correlated with the performance on the figure classification task, which measures general reasoning ability.
 Performance on lineup task 3, which investigated the potential to visually identify nonnormality in residual QQ-plots, is more associated with the card rotation and paper folding tests, which measure visuospatial ability.
-This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks, which required participants to examine categorical data.
+This suggests that certain lineup tasks may require more visual ability than others; in the case of the QQ-lineups a successful evaluation  needed  participants to mentally rotate plots to compare vertical distances, requiring more mental manipulation than the first two lineup tasks, which asked from participants to examine categorical data.
 
@@
 <<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
+
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 rotation$Var1 <- factor(rotation$Var1, 
-                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_3", "lineup_2", "lineup_1"),
-                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
+                        levels=c("vis_search", "folding", "card_rot", "fig_class", "Lineup-Core"),
+                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Core"))
+
+#rotation$Var1 <- factor(rotation$Var1, 
+#                        levels=c("vis_search", "folding", "card_rot", "fig_class", #"lineup_3", "lineup_2", "lineup_1"),
+#                        labels=c("Visual Search", "Paper Folding", "Card Rotation", #"Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
 library(grid) # needed for arrow function

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
couple of changes ...

@@
 
+We would like to thank the reviewers for the detailed comments. We hope to have addressed all of the questions.
+
+We are detailing our response below:
+
 # Major Concerns
@@
 **(refer to the first concern of R1).**
-*The purpose of each of the cognitive tests is described in section 2.2.
-In summary, the paper-folding, card rotation, and figure classification tasks are part of the kit of factor-referenced cognitive tests (and are thus extremely well-studied); additionally, the authors chose the specific visual searching task (VST) because of its similarity of form to statistical lineups and its' compatibility with the written form of the other cognitive tests and lineup tests.*
+
+*We have extended the discussion of the choice of cognitive tests; the purpose of each of the cognitive tests is described in section 2.2.
+In summary, the paper-folding, card rotation, and figure classification tasks are part of the kit of factor-referenced cognitive tests (and are thus extremely well-studied); additionally, we included a visual searching task (VST) because of its similarity of form to statistical lineups and its' compatibility with the written form of the other cognitive tests and lineup tests.*
 *As mentioned in the paper, the paper folding and card rotation tests have been used in many previous studies of statistical graphics and data displays.
 The figure-classification task examines inductive reasoning, which is another component of the lineup task (determining the dimension on which the lineup components differ, then applying that to answer the question successfully).*
+*XXX hypotheses about the relationship with lineups? XXX*
+
 2. **Are the experimental procedure sufficiently detailed?**    
 **(refer to the second concern of R1 and )**    
-*We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown).
+*We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown; however, there are various sources online, that -illegally- provide access to parts of the actual test sets, e.g.
+http://spatiallearning.org/resource-info/Spatial_Ability_Tests/Paper_Folding_Test.pdf ).
 The testing documents are comprehensive, and oral instructions mirror the instructions on the introduction page of each task (students are familiar with this format as it is commonly used on standardized tests, such as the SAT and ACT).*

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
couple of changes ...

@@
 Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
-This study examines the demographic characteristics and visual skills associated with perception of graphical lineups.
+here, we present a study relating demographic characteristics and visual skills to perception of graphical lineups.
 We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
@@
 
-% \hhnote{XXX Still missing: we need to extend the discussion a little bit that we want to extract the common core from lineups rather than investigate the relationship between individual designs and the  tasks.
-I believe, that the lineup commonalities are stronger than the design differences - but we might need to get some statistical backup for that.
-...
-one way to do that would be to do a PCA on the three lineups and investigate whether anything beyond the first dimension is significant.
-I don't think so.
-}
-
-% \hhnote{XXX we should add: the lineups come from three different lineup experiments that discuss and compare the design choices (cite papers).
-This is NOT what we are doing here.
-Regardless of design choices, we want to find out, what visual skills are necessary for dealing with lineups.
-}
-
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test.
-
-The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task.
+\newtext{The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~}\cite{educational1963kit, ekstrom1976manual}.
+\newtext{These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
+}
+
+The VST measures visual search speed~\cite{goldstein1973validity}; \newtext{we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.}
+Both  card rotation and paper folding tests measure spatial manipulation ability in \newtext{two- and three dimensions}, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+\newtext{we hypothesized that} all of these skills are %at least peripherally
+recruited during the lineup task, but \newtext{are interested which of these dominate} in predicting performance on the lineup task.
 

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
couple of changes ...

@@
 
-% \hhnote{XXX Still missing: we need to extend the discussion a little bit that we want to extract the common core from lineups rather than investigate the relationship between individual designs and the  tasks.
-I believe, that the lineup commonalities are stronger than the design differences - but we might need to get some statistical backup for that.
-...
-one way to do that would be to do a PCA on the three lineups and investigate whether anything beyond the first dimension is significant.
-I don't think so.
-}
-
-% \hhnote{XXX we should add: the lineups come from three different lineup experiments that discuss and compare the design choices (cite papers).
-This is NOT what we are doing here.
-Regardless of design choices, we want to find out, what visual skills are necessary for dealing with lineups.
-}
-
 \label{sec:introduction}Data displays provide quick summaries of data, models, and results, but not all displays are equally good, nor is any data display equally useful to all viewers. 
-Graphics utilize higher-bandwidth visual pathways to encode information~~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
+Graphics utilize higher-bandwidth visual pathways to encode information~\cite{baddeley1974working}, allowing viewers to quickly and intuitively relate multiple dimensions of numerical quantities.
 Well-designed graphics emphasize and present important features of the data while minimizing  features of lesser importance, guiding the viewer towards conclusions that are meaningful in context and supported by the data while maximizing the information encoded in working memory.
@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test.
-The VST measures visual search speed~\cite{goldstein1973validity}, the paper folding and card rotation tests measures spatial manipulation ability, and the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-all of these skills are at least peripherally recruited during the lineup task, but some may dominate in predicting performance on the lineup task.
+\newtext{The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~}\cite{educational1963kit, ekstrom1976manual}.
+\newtext{These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
+}
+
+The VST measures visual search speed~\cite{goldstein1973validity}; \newtext{we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.}
+Both  card rotation and paper folding tests measure spatial manipulation ability in \newtext{two- and three dimensions}, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+\newtext{we hypothesized that} all of these skills are %at least peripherally
+recruited during the lineup task, but \newtext{are interested which of these dominate} in predicting performance on the lineup task.
 
@@
 Additionally, however, we can aggregate the scores from the same individual on several lineups to objectively assess an individual's performance on the lineup task.
-In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score.
 
-For this approach, we derive a score for an individual as follows:
+In this experiment, we are interested in individual skill at evaluating lineups rather than the data displayed in a single lineup, so we will consider an individual's aggregate lineup score, derived as follows:
+
+%For this approach, we derive a score for an individual as follows:
 
@@
 Statistical lineups depend on the ability to search for a signal amid a set of distractors (visual search) and the ability to infer patterns from stimuli (pattern recognition).
-Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from participants, e.g.~ polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
-By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen cognitive tests to assess these aspects of visuospatial ability.
+Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from participants, e.g.~polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
+By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen\newtext{, externally validated,} cognitive tests to assess these aspects of visuospatial ability.
 
-Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors plus gender had an effect on the amount of time spent on lineups~\cite{humanfactorslineups}.
+Demographic factors are known to impact lineup performance: country, education, and age affected score on lineup tests, and all of those factors as well as gender had an effect on the amount of time spent on lineups~\cite{humanfactorslineups}.
 In addition, lineup performance can be partially explained using statistical distance metrics~\cite{distancemetriclineups}, but these metrics do not completely succeed in predicting human performance, in part due to the difficulty of representing human visual ability algorithmically.
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Thu Jun 25 09:34:25 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Fri Jun 26 14:46:58 2015
 \begin{table}[htb]
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Thu Jun 25 09:34:25 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Fri Jun 26 14:46:59 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Thu Jun 25 09:34:26 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Fri Jun 26 14:46:59 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Thu Jun 25 09:34:26 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Fri Jun 26 14:46:59 2015
 \begin{table}[ht]

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
response, better formulated for hypothesis of relationship

@@
 The figure-classification task examines inductive reasoning, which is another component of the lineup task (determining the dimension on which the lineup components differ, then applying that to answer the question successfully).*
-*XXX hypotheses about the relationship with lineups? XXX*
+*Regarding our hypotheses about the relationship of the visuospatial tests and lineups: we extended the discussion in the introduction; in summary we hypothesize that visual search is the basis for any lineup evaluation.
+We are further intrested in whether the spatial component (measured in 2d by the card rotation test or in 3d by the paper folding test) dominates lineup performance or inductive reasoning (as measured by the figure classification test).*
 

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
Spacing

@@
 **(refer to the first concern of R1).**    
-
 *We have extended the discussion of the choice of cognitive tests; the purpose of each of the cognitive tests is described in section 2.2.

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Saving space

@@
 The score for this individual is then given as:
-\begin{eqnarray}\label{eq.scoring1}
-n_c - \textstyle\frac{n_w}{m-1}.
-\end{eqnarray}
+\begin{equation}\label{eq.scoring1}
+n_c - n_w/(m-1)
+\end{equation}
 Note that the sum of answers, $n_c + n_w$, is at most $K$, but may be less, if an observer chooses to not answer one of the lineup tasks or runs out of time.
-
 The scoring scheme as given in (\ref{eq.scoring1}) is chosen so that if participants are guessing, the expected score is 0.
@@
 
-\paragraph{Lineup Set 1}
-The lineups in this set are taken from an experiment which examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
-This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
-
-\paragraph{Lineup Set 2}
-The lineups in the second set are taken from another experiment which examined visualizations of categorical data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
-Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
-
-\paragraph{Lineup Set 3}
-Lineups in the final set are taken from an experiment which examined modifications of QQ-plots, using data from various model simulations.
-Reference lines, acceptance bands, and rotation were utilized to examine which plots allowed participants to most effectively identify violations of normality.
-Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were overly liberal.
-As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands \cite{loy:2015}.
-
-\subsection{Measures of visuospatial ability}
-
-Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
-Tasks are timed such  that participants are under pressure; participants are not expected to finish all of the problems in each section.
-This allows for better discrimination between scores and reduces score compression at the top of the range.
-
-The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
-Historically, this visual search task has been used as a measure of brain damage~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~~\cite{anderson1983interactive}.
-The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations, such as those in~\cite{wolfe1990limitations,treisman1980feature}.
-We have included visual search in the cognitive testing battery for this experiment because it is the most simple of the tasks required to successfully evaluate a lineup; if participants cannot complete visual search tasks successfully, they will not be able to accurately identify which lineup plot is different.
-
-\begin{figure}[htp]\centering
-\includegraphics[width=.8\linewidth]{VisualSearch}
-\caption[Visual Search Task]{Visual Search Task (VST).
-Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
-Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
-\end{figure}
-
-The remaining cognitive tests are adapted from \cite{ekstrom1976manual} and are highly validated tests that have been constructed to measure specific cognitive factors, such as inductive reasoning and spatial rotation.
-
-The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
-This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
-An example is shown in figure~\ref{fig:figureclassification}.
-
-The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
-In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
-As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
-
-The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
-It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
-An example is shown in figure~\ref{fig:cardrotation}.
-The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids~\cite{mayer1994whom} and statistical graphs~\cite{lowrie2007solving} in education.
-
-Two-dimensional comparisons are an important component of lineup performance.
-In some lineup situations, these comparisons sometimes involve translation, but in other lineups, rotation is required.
-Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
-Three-dimensional rotation is somewhat less relevant to most statistical graphics, which are typically two-dimensional in nature.
-While \cite{ekstrom1976manual} includes three-dimensional rotation tasks similar to \cite{shepard1971mental}, these three-dimensional tasks were not included in this experiment due to time constraints and limited face validity.
-
-The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
-A sample question from the test is shown in figure~\ref{fig:paperfolding}.
-It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
-The paper folding test is associated with the ability to extrapolate symmetry and reflection.
-
-Tests in the Kit of Factor-Referenced Cognitive tests are classified according to the reasoning required to complete the test; in this way, we can be reasonably sure that the paper folding test, card rotation test, and figure classification test have construct validity.
-
-\begin{figure}[ht]
-  \centering
-   \subfloat[Figure Classification Task.
-Participants classify each figure in the second row as belonging to one of the groups above.
-Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
-  \includegraphics[width=.8\linewidth]{figureclassification}
-}
-\hfill
-    \subfloat[Card Rotation Task.
-Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
-Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
-\includegraphics[width=.8\linewidth]{cardrotation}
-    }
-    \hfill
-    \subfloat[Paper Folding Task.
-Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
-Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
-\includegraphics[width=.8\linewidth]{paperfolding}
-   }
-    \caption{Visuospatial tests}
-    \label{fig:tests}
-\end{figure}
-
-Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
-
-All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; internal pilot tests did not indicate an effect of task order.
-
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
-Participants were given 5 minutes to complete each block of 20 lineups.
-Figure~\ref{fig:lineup}(a) shows a sample lineup of box plots.
-
-In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
-These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
-
-After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
-The duration of testing for each participant was approximately 1 hour.
-
-\subsection{Test Scoring}\label{sec:scaling}
-%
-All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
-Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
-\begin{eqnarray}\label{eq.scoring}
-\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}
-\end{eqnarray}
-Equation \ref{eq.scoring} is a restatement of equation \ref{eq.scoring1} provided for convenience.
-This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
-Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
-The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
-
-\section{Results}\label{sec:results}
-Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
-\Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
-Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception in the 25-28 bracket.
-This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the student population of Iowa State; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
-
-\subsection{Comparison of Spatial Tests with Validated Results}
-The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
-We compare Iowa State students' unscaled scores in Table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
-
-\begin{table*}[htb]\centering
+\begin{table*}[htbp]\centering
 \begin{tabular}{rlllc}
@@
 \end{table*}
+
+\paragraph{Lineup Set 1}
+The lineups in this set are taken from an experiment which examined the use of boxplots, density plots, histograms, and dotplots to compare two groups which vary in mean and sample size.
+This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
+Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.\hphantom{\ref{tab:scorecomparison}}
+
+\paragraph{Lineup Set 2}
+The lineups in the second set are taken from another experiment which examined visualizations of categorical data, this time comparing boxplots, bee swarm boxplots, boxplots with overlaid jittered data, and violin plots, as shown in figure \ref{fig:lineups} (e)-(h).
+Participants were much more accurate in this experiment than in the experiment described previously, because of the types of plots compared as well as the underlying data distributions.
+
+\paragraph{Lineup Set 3}
+Lineups in the final set are taken from an experiment which examined modifications of QQ-plots, using data from various model simulations.
+Reference lines, acceptance bands, and rotation were utilized to examine which plots allowed participants to most effectively identify violations of normality.
+Rotated QQ-plots showed lower performance because participants were able to more accurately compare acceptance bands to residuals, and thus could identify that the reference bands were overly liberal.
+As a result, performance was somewhat lower for rotated plots, even though participants were more accurate when comparing the residuals to the reference bands \cite{loy:2015}.
+
+\subsection{Measures of visuospatial ability}
+
+Participants are asked to complete several cognitive tests designed to measure spatial and reasoning ability.
+Tasks are timed such  that participants are under pressure; participants are not expected to finish all of the problems in each section.
+This allows for better discrimination between scores and reduces score compression at the top of the range.
+
+The \textbf{visual searching task} (VST), shown in figure~\ref{fig:VST}, is designed to test a participant's ability to find a target stimulus in a field of distractors, thus making the visual search task similar in concept to lineups.
+Historically, this visual search task has been used as a measure of brain damage~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological}; however, similar tasks have been used to measure cognitive performance in a variety of situations, for example under the influence of drugs in~~\cite{anderson1983interactive}.
+The similarity to the lineup protocol as well as the simplicity of the test and its' lack of color justify the slight deviation from forms of visual search tasks typically used in normal populations, such as those in~\cite{wolfe1990limitations,treisman1980feature}.
+We have included visual search in the cognitive testing battery for this experiment because it is the most simple of the tasks required to successfully evaluate a lineup; if participants cannot complete visual search tasks successfully, they will not be able to accurately identify which lineup plot is different.
+
+\begin{figure}[htp]\centering
+\includegraphics[width=.8\linewidth]{VisualSearch}
+\caption[Visual Search Task]{Visual Search Task (VST).
+Participants are instructed to find the plot numbered 1-24 which matches the plot labeled ``Target".
+Participants will complete up to 25 of these tasks in 5 minutes.}\label{fig:VST}
+\end{figure}
+
+The remaining cognitive tests are adapted from \cite{ekstrom1976manual} and are highly validated tests that have been constructed to measure specific cognitive factors, such as inductive reasoning and spatial rotation.
+
+The \textbf{figure classification task} tests a participant's ability to extrapolate rules from provided figures.
+This task is associated with inductive reasoning abilities (factor I  in~\cite{ekstrom1976manual}).
+An example is shown in figure~\ref{fig:figureclassification}.
+
+The figure classification test requires the same type of reasoning as the lineups: participants must determine the rules from the provided classes, and extrapolate from those rules to classify new figures.
+In lineups, participants must determine the rules based on the panels appearing in the lineup; they must then identify the plot which does not conform.
+As such, the figure classification test has content validity in relation to lineup performance: it is measuring similar underlying criteria.
+
+The \textbf{card rotation test} measures a participant's ability to rotate objects in two dimensions in order to distinguish between left-hand and right-hand versions of the same figure.
+It tests mental rotation skills, and is classified as a test of spatial orientation in~\cite{ekstrom1976manual}, though it does require that participants have both mental rotation ability and short-term visual memory.
+An example is shown in figure~\ref{fig:cardrotation}.
+The card rotation test is often used in studies investigating the effect of visual ability on the use of visual aids~\cite{mayer1994whom} and statistical graphs~\cite{lowrie2007solving} in education.
+
+Two-dimensional comparisons are an important component of lineup performance.
+In some lineup situations, these comparisons sometimes involve translation, but in other lineups, rotation is required.
+Lineups also require visual short-term memory, so the additional factor measured implicitly by this test does not reduce its potential relevance to lineup performance.
+Three-dimensional rotation is somewhat less relevant to most statistical graphics, which are typically two-dimensional in nature.
+While \cite{ekstrom1976manual} includes three-dimensional rotation tasks similar to \cite{shepard1971mental}, these three-dimensional tasks were not included in this experiment due to time constraints and limited face validity.
+
+The \textbf{paper folding test} measures participants' ability to visualize and mentally manipulate figures in three dimensions.
+A sample question from the test is shown in figure~\ref{fig:paperfolding}.
+It is classified as part of the visualization factor in~\cite{ekstrom1976manual}, which differs from the spatial orientation factor because it requires participants to visualize, manipulate, and transform the figure~mentally, which makes it a more complex and demanding task than simple rotation.
+The paper folding test is associated with the ability to extrapolate symmetry and reflection.
+
+Tests in the Kit of Factor-Referenced Cognitive tests are classified according to the reasoning required to complete the test; in this way, we can be reasonably sure that the paper folding test, card rotation test, and figure classification test have construct validity.
+
+\begin{figure}[ht]
+  \centering
+   \subfloat[Figure Classification Task.
+Participants classify each figure in the second row as belonging to one of the groups above.
+Participants complete up to 14 of these tasks (each consisting of 8 figures to classify) in 8 minutes.\label{fig:figureclassification}]{%
+  \includegraphics[width=.8\linewidth]{figureclassification}
+}
+\hfill
+    \subfloat[Card Rotation Task.
+Participants mark each figure on the right hand side as either the same as or different than the figure on the left hand side of the dividing line.
+Participants complete up to 20 of these tasks (each consisting of 8 figures) in 6 minutes.\label{fig:cardrotation}]{%
+\includegraphics[width=.8\linewidth]{cardrotation}
+    }
+    \hfill
+    \subfloat[Paper Folding Task.
+Participants are instructed to pick the figure matching the sequence of steps shown on the left-hand side.
+Participants  complete up to 20 of these tasks in 6 minutes.\label{fig:paperfolding}]{%
+\includegraphics[width=.8\linewidth]{paperfolding}
+   }
+    \caption{Visuospatial tests}
+    \label{fig:tests}
+\end{figure}
+
+Lineups require similar manipulations in two-dimensional space, and also require the ability to perform complex spatial manipulations mentally; for instance, comparing the interquartile range of two boxplots as well as their relative alignment to a similar set of two boxplots in another panel.
+
+All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; internal pilot tests did not indicate an effect of task order.
+
+Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
+Participants were given 5 minutes to complete each block of 20 lineups.
+Figure~\ref{fig:lineup}(a) shows a sample lineup of box plots.
+
+In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
+These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
+
+After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
+The duration of testing for each participant was approximately 1 hour.
+
+\subsection{Test Scoring}\label{sec:scaling}
+%
+All test results were scored so that random guessing produces an expected value of 0; therefore each question answered correctly contributes to the score by 1, while a wrong answer is scored by $-1/(k-1)$, where $k$ is the total number of possible answers to the question.
+Thus, for a test consisting of multiple choice questions with $k$ suggested answers with a single correct answer each, the score is calculated as
+\begin{eqnarray}\label{eq.scoring}
+\# \text{correct answers } - 1/(k-1) \cdot \# \text{wrong answers}
+\end{eqnarray}
+Equation \ref{eq.scoring} is a restatement of equation \ref{eq.scoring1} provided for convenience.
+This allows us to compare each participant's score in light of how many problems were attempted as well as the number of correct responses.
+Combining accuracy and speed into a single number does not only make a comparison of test scores easier,  this scoring mechanism is also used on many standardized tests, such as the SAT and the battery of psychological tests~\cite{diamond1973correction, ekstrom1976manual} from which parts of this test are drawn.
+The advantage of using tests from the Kit of Factor Referenced Cognitive tests~\cite{ekstrom1976manual} is that the tests are extremely well studied (including an extensive meta-analysis in~\cite{voyer1995magnitude} of the spatial tests we are using in this study) and comparison data are available from the validation of these factors~\cite{schaie1998longitudinal,hampson1990variations,mayer1994whom} and previous versions of the kit~\cite{educational1963kit}.
+
+\section{Results}\label{sec:results}
+Results are based on an evaluation of \Sexpr{nrow(ans)} undergraduate students at Iowa State University.
+\Sexpr{round(sum(ans.summary$stem)/nrow(ans.summary)*100)}\% of the participants were in STEM fields, the others were distributed relatively evenly between agriculture, business, and the social sciences.
+Students were evenly distributed by gender, and were between 18 and 24 years of age with only one exception in the 25-28 bracket.
+This is reasonably representative\footnote{\url{http://www.registrar.iastate.edu/sites/default/files/uploads/stats/university/F14summary.pdf}} of the student population of Iowa State; in the fall 2014 semester, 26\% of students were associated with the college of engineering, 24\% were associated with the college of liberal arts and sciences, 15\% were associated with the college of human sciences, 7\% with the college of design, 13\% with the business school, and 15\% with the school of agriculture.
+
+\subsection{Comparison of Spatial Tests with Validated Results}
+The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).
+We compare Iowa State students' unscaled scores in Table~\ref{tab:scorecomparison}, adjusting data from other populations to account for subpopulation structure and test length.
 
@@
 row.names(pca2$rotation) <- c("Card Rot", "Fig Class", "Folding", "VST")
-print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="htb", caption.placement="top")
+print(xtable(pca2$rotation, caption="Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\\label{tab:PCArotation4}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", table.placement="htb", caption.placement="top")
 @
@@
 <<pcarotation5,echo=F, results='asis'>>=
-print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \\label{tab:PCArotation5}",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, table.placement="ht", caption.placement="top")
+row.names(pca$rotation) <- c("Lineup", "Card Rot", "Fig Class", "Folding", "VST")
+print(
+  xtable(
+    pca$rotation, 
+    caption="PCA Rotation matrix for all five tests. The first principal component is essentially an average of performance on all five tests. \\label{tab:PCArotation5}", 
+    digits=2), 
+  sanitize.text.function=sanitize, 
+  sanitize.rownames.function=sanitize, 
+  sanitize.colnames.function=sanitize, 
+  table.placement="ht", caption.placement="top")
 @

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
response, better formulated for hypothesis of relationship
Saving space

@@
 
+\begin{table*}[htbp]\centering
+\begin{tabular}{rlllc}
+\hline
+  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
+ISU Students & 83.4 (24.1) 
+             & 12.4 (3.7)
+             & 57.0 (23.8)\footnotemark[1]
+             & 21.9 (2.3)\\
+Scaled Scores & 88.0 (34.8)
+              & 13.8 (4.5)
+              & 58.7 (14.4)\footnotemark[2]
+              & -- \\
+Unscaled Scores & 44.0 (24.6)\footnotemark[3]
+                & 13.8 (4.5)
+                & \parbox[c]{.2\linewidth}{M:~120.0~(30.0)\\ F:~114.9~(27.8)}
+                & --\\\hline
+{\footnotesize Population}    
+              & \parbox[t]{.15\linewidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
+              & \parbox[t]{.17\linewidth}{\footnotesize 46 college students\\(1963~version)}
+              & \parbox[t]{.2\linewidth}{\footnotesize suburban 11th \& 12th \\ grade students\\(288-300 M, 317-329 F)}
+              & \\\hline
+\end{tabular}
+\caption[Comparison of scores for cognitive tasks.]{Comparison of scores from Iowa State students and scores reported in~\protect~\cite{ekstrom1976manual}.
+Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
+Data shown are from the population most similar to ISU students, out of the data available.
+The visual search task~\protect~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
+\label{tab:scorecomparison}}
+\end{table*}
+
 \paragraph{Lineup Set 1}
@@
 This set of lineups consists of 20 plots selected from the plots used in the full experiment; each set of data is displayed with each of the four plot types.
-Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.
+Figure \ref{fig:lineups} (a)-(d) provides an example of the four plot types displaying the same data.\hphantom{\ref{tab:scorecomparison}}
 
@@
 
-\begin{table*}[htb]\centering
-\begin{tabular}{rlllc}
-\hline
-  & Card Rotation & Paper Folding & Figure Classification & Visual Search  \\\hline
-ISU Students & 83.4 (24.1) 
-             & 12.4 (3.7)
-             & 57.0 (23.8)\footnotemark[1]
-             & 21.9 (2.3)\\
-Scaled Scores & 88.0 (34.8)
-              & 13.8 (4.5)
-              & 58.7 (14.4)\footnotemark[2]
-              & -- \\
-Unscaled Scores & 44.0 (24.6)\footnotemark[3]
-                & 13.8 (4.5)
-                & \parbox[c]{.2\linewidth}{M:~120.0~(30.0)\\ F:~114.9~(27.8)}
-                & --\\\hline
-{\footnotesize Population}    
-              & \parbox[t]{.15\linewidth}{\footnotesize approx.\ 550 male\\ naval recruits} 
-              & \parbox[t]{.17\linewidth}{\footnotesize 46 college students\\(1963~version)}
-              & \parbox[t]{.2\linewidth}{\footnotesize suburban 11th \& 12th \\ grade students\\(288-300 M, 317-329 F)}
-              & \\\hline
-\end{tabular}
-\caption[Comparison of scores for cognitive tasks.]{Comparison of scores from Iowa State students and scores reported in~\protect~\cite{ekstrom1976manual}.
-Scaled scores are calculated based on information reported in the manual, scaled to account for differences in the number of questions answered during this experiment.
-Data shown are from the population most similar to ISU students, out of the data available.
-The visual search task~\protect~\cite{goldstein1973validity,demita1981validity,moerland1986neuropsychological} is not part of the Kit of Factor Referenced Cognitive Test data, and thus we do not have comparison data for the form used in this experiment.
-\label{tab:scorecomparison}}
-\end{table*}
-
 Table~\ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations.
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Fri Jun 26 14:46:58 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 12:42:40 2015
 \begin{table}[htb]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Fri Jun 26 14:46:59 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 12:42:40 2015
 \begin{table}[htb]
@@
 \caption{Rotation matrix for principal component analysis of the four cognitive tests (visual search, paper folding, card rotation, figure classification).\label{tab:PCArotation4}} 
+{\footnotesize
 \begin{tabular}{rrrrr}
@@
 \end{tabular}
+}
 \end{table}
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Fri Jun 26 14:46:59 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 12:42:40 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Fri Jun 26 14:46:59 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 12:42:40 2015
 \begin{table}[ht]
 \centering
-\caption{PCA Rotation matrix for all five tests. The first principal component is essentially an average of all five tests. 
-             \label{tab:PCArotation5}} 
+\caption{PCA Rotation matrix for all five tests. The first principal component is essentially an average of performance on all five tests. \label{tab:PCArotation5}} 
 \begin{tabular}{rrrrrr}
@@
   \hline
-lineup & 0.42 & 0.49 & -0.46 & 0.60 & -0.10 \\ 
-  card.rot & 0.50 & -0.30 & 0.28 & 0.23 & 0.73 \\ 
-  fig.class & 0.43 & 0.45 & -0.15 & -0.75 & 0.18 \\ 
-  folding & 0.47 & 0.07 & 0.68 & 0.04 & -0.56 \\ 
-  vis.search & 0.41 & -0.69 & -0.48 & -0.15 & -0.33 \\ 
+Lineup & 0.42 & 0.49 & -0.46 & 0.60 & -0.10 \\ 
+  Card Rot & 0.50 & -0.30 & 0.28 & 0.23 & 0.73 \\ 
+  Fig Class & 0.43 & 0.45 & -0.15 & -0.75 & 0.18 \\ 
+  Folding & 0.47 & 0.07 & 0.68 & 0.04 & -0.56 \\ 
+  VST & 0.41 & -0.69 & -0.48 & -0.15 & -0.33 \\ 
    \hline
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Fri Jun 26 14:46:59 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 12:42:41 2015
 \begin{table}[ht]

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Updated appendix, regenerated pdfs, abstract typo fix

@@
 Graphics convey numerical information very efficiently, but rely on a different set of mental processes than tabular displays.
-here, we present a study relating demographic characteristics and visual skills to perception of graphical lineups.
+Here, we present a study relating demographic characteristics and visual skills to perception of graphical lineups.
 We conclude that lineups are essentially a classification test in a visual domain, and that performance on the lineup protocol is associated with general aptitude, rather than specific tasks such as card rotation and spatial manipulation.
@@
 The duration of testing for each participant was approximately 1 hour.
+All test documents (with redacted copyrighted material) are available as supplemental materials.
 

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Updated appendix, regenerated pdfs, abstract typo fix

@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:42:40 2015
+% Fri Jun 26 12:56:40 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:42:40 2015
+% Fri Jun 26 12:56:40 2015
 \begin{table}[ht]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:42:41 2015
+% Fri Jun 26 12:56:41 2015
 \begin{table}[ht]

### 2015-06-26 Paper/InfoVis2015Revision/appendix.tex
Addressing reviewer comments
Updated appendix, regenerated pdfs, abstract typo fix

--- 2015-03-22 Paper/InfoVis2015/appendix.tex
@@
 To better understand the effects of these variables, a larger study is necessary.
-% latex table generated in R 3.1.3 by xtable 1.7-1 package
-% Sun Mar 22 15:22:39 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Fri Jun 26 13:01:47 2015
 \begin{table}[ht]

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
another comma

@@
 **(refer to the second concern of R1 and )**
+
 *We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown; however, there are various sources online, that -illegally- provide access to parts of the actual test sets, e.g.
@@
 The testing documents are comprehensive, and oral instructions mirror the instructions on the introduction page of each task (students are familiar with this format as it is commonly used on standardized tests, such as the SAT and ACT).*
+
 3. **The results of statistical analysis do not seem to be conclusive, are the interpretations of these results properly revised?**    
@@
 In addition, the similarity in test protocol between the VST, the tests from the factor-referenced kit of cognitive tests, and the lineup protocol was intentional; by controlling the experimental protocol, we could reduce participant confusion and the corresponding variability.
-As our goal was not to examine the elements of visualization design experimentally, controlling those elements is one element of appropriate experimental design.*
+As our goal was not to examine the elements of visualization design experimentally, controlling those elements adheres to good experimental design principles.*
 8. **Does the revision have an explicit related works section?**    

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Remove note, slight wording change

@@
 This removes much of the subjectivity from user evaluations of display effectiveness, and the procedure is simple enough that it does not generally require participants to be very familiar with data-based graphics.
-\hhnote{should we say something here about using charts only for comparisons?
-We almost do that already.
-Maybe just drive that message home?
-}
 While previous research~\cite{lowrie2007solving,mayer1994whom} has examined the link between certain types of graphical perception and spatial skills, it is important to identify any additional visual skills participants utilize to complete the lineup task, as well as better understand demographic characteristics (math education, research experience, age, gender) which may impact performance~\cite{humanfactorslineups}.
@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test.
-\newtext{The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~}\cite{educational1963kit, ekstrom1976manual}.
-\newtext{These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
-}
-
-The VST measures visual search speed~\cite{goldstein1973validity}; \newtext{we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.}
-Both  card rotation and paper folding tests measure spatial manipulation ability in \newtext{two- and three dimensions}, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-\newtext{we hypothesized that} all of these skills are %at least peripherally
-recruited during the lineup task, but \newtext{are interested which of these dominate} in predicting performance on the lineup task.
+The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~\cite{educational1963kit, ekstrom1976manual}.
+These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
+
+The VST measures visual search speed~\cite{goldstein1973validity}; we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.
+Both  card rotation and paper folding tests measure spatial manipulation ability in two- and three dimensions, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+we hypothesized that all of these skills are recruited during the lineup task, but are interested which of these dominate in predicting performance on the lineup task.
 
@@
 Depending on the choice of plot shown in the lineup, the task of identifying the most different plot might require additional abilities from participants, e.g.~polar coordinates depend on the ability to mentally rotate stimuli (spatial rotation) and mentally manipulate graphs (spatial rotation and manipulation).
-By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen\newtext{, externally validated,} cognitive tests to assess these aspects of visuospatial ability.
+By breaking the lineup task down into its components, we  determine which visuospatial factors most strongly correlate with lineup performance, using carefully chosen, externally validated, cognitive tests to assess these aspects of visuospatial ability.
 

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Remove note, slight wording change

@@
 We compare lineup performace to the visual search task (VST), paper folding test, card rotation test, and figure classification test.
-\newtext{The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~}\cite{educational1963kit, ekstrom1976manual}.
-\newtext{These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
-}
+The last three tests are part of the Kit of Factor-Referenced Cognitive Tests~\cite{educational1963kit, ekstrom1976manual}.
+These tests measure three separate dimensions of visuospatial ability and have been virtually unchanged in use  since the 1960s, and are thus extremely well-studied.
 
-The VST measures visual search speed~\cite{goldstein1973validity}; \newtext{we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.}
-Both  card rotation and paper folding tests measure spatial manipulation ability in \newtext{two- and three dimensions}, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
-\newtext{we hypothesized that} all of these skills are %at least peripherally
-recruited during the lineup task, but \newtext{are interested which of these dominate} in predicting performance on the lineup task.
+The VST measures visual search speed~\cite{goldstein1973validity}; we considered visual search as the basis of a lineup evaluation and included this test mainly to identify potential problems in the results, if participants were unable to perform this task.
+Both  card rotation and paper folding tests measure spatial manipulation ability in two- and three dimensions, while the figure classification test measures inductive reasoning~\cite{ekstrom1976manual};
+we hypothesized that all of these skills are recruited during the lineup task, but are interested which of these dominate in predicting performance on the lineup task.
 
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:56:40 2015
+% Fri Jun 26 14:47:21 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:56:40 2015
+% Fri Jun 26 14:47:21 2015
 \begin{table}[ht]
@@
 % latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 12:56:41 2015
+% Fri Jun 26 14:47:22 2015
 \begin{table}[ht]

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
Remove note, slight wording change
just edits to the response

@@
 **(refer to the first concern of R1).**
+
 *We have extended the discussion of the choice of cognitive tests; the purpose of each of the cognitive tests is described in section 2.2.
@@
 
-*We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown; however, there are various sources online, that -illegally- provide access to parts of the actual test sets, e.g.
+*We have added additional detail about the experimental procedures, and have included the testing materials as supplementary documents (with the factor-referenced kit tests redacted due to copyright issues; in these cases only the instruction page is shown; however, there are various sources online, that  provide access to parts of the actual test sets, e.g.
 http://spatiallearning.org/resource-info/Spatial_Ability_Tests/Paper_Folding_Test.pdf ).
@@
 **(refer to the third concern of R1)**
-*The results of the statistical analysis are as conclusive as the nature of the study allows.
-While correlations of 0.3-0.4 are low compared to the absolute range of 0 - 1, they are strong for human subjects data (which is how we have interpreted them here).
+
+*While correlations of 0.3-0.4 are low compared to the absolute range of 0 - 1, they are fairly strong for human subjects studies.
 Because so many different factors influence human behavior, we generally do not see correlations above 0.5 on distinct tasks.*
+*This makes the results of the statistical analysis about as conclusive as the nature of the study allows - we have to keep in mind that this study setup only allows to show associations but no causality, and therefore the results allow for multiple interpretations in the same way as any such statistical analysis.*
 *We have provided both graphics and PCA results in order to demonstrate our results in more than one way; while these results are preliminary, the cost involved in designing and executing a study which would enable multivariate hypothesis testing (given the multiple correlations between dependent variables) is prohibitive.
 This study is the first to examine the visual and cognitive skills necessary to evaluate statistical lineups; follow-up studies may provide more statistically precise results.*
+
 4. **Does the revision contain additional discussion of why PCA is a proper approach to interpret the data for the VIS community members who do not have sufficient statistics background?**    
 **(refer to the concern of R2)**
+
 *The revision includes a reference to a PCA tutorial (for those unfamiliar with the procedure) and explains the rationale for utilizing PCA in some additional detail.
 Section 3.3 - Visual Abilities and Lineup Performance has been modified extensively to provide additional details about PCA.*
+
 5. **Does the revision has the clear description of "most different"?**    
 **(refer to the first concern of R3)**
+
 *The use of 'most different' is a feature of the lineup protocol, and does not bias the participants to examine specific distributional characteristics.
-While participants may choose different distributional features to identify the 'most different' plot, this mirrors how we perceive graphics (we cue in to different features).
+While participants may choose different distributional features to identify the 'most different' plot, this mirrors our perception of perceive graphics (viewers are bound to cue in to different features).
 In addition, many statistical tests react to different features of the data - a t-test may not produce significant results even with a truly significant difference in means if the two distributions have wildly different variances.
-Finally, the use of 'most different' allows the lineup protocol to be used even in populations which do not have mathematical training and might not understand the use of terms such as mean and variance.*
+Moreover, the use of 'most different' allows the lineup protocol to be used even in populations which do not have mathematical training and might not understand the use of terms such as mean and variance.*
 *We have added additional language in Section 2.1 to emphasize that the phrasing is a standard part of the lineup protocol.*
+
 6. **Does the revision has the proper discussion of visualization design**    
 **(refer to the second concern of R3)**
+
 *This study was not designed to directly impact visualization design recommendations.
 While the lineup protocol can easily be used to compare many different visualization designs, the goal of this experiment was to examine the cognitive requirements of lineup evaluation at the participant level.
-As such, it would not be appropriate to discuss visualization design in detail.*
+As such, it would not be appropriate to discuss visualization design in detail.
+Each of the sets of lineups shown, were part of studies to compare the designs.
+We have cited the corresponding papers accordingly.
+*
 *While there are differences in the correlation of scores on lineup sets 1, 2, and 3 with the cognitive tests, this does not provide any additional information about whether designs in lineup set 1 are superior to those in lineup set 2 and lineup set 3, and in fact, the sets of lineups are not comparable in this way.
 Each set of lineups was taken from an experiment which did aim to rank designs by effectiveness, but in this study we are instead examining individual differences in cognitive abilities and lineup performance.*
+
 7. **Does the revision resolve a potential concern of confounding factors**    
 **(refer to the conern of R3)**
+
 *The issue of confounding factors is related to the issue raised in #6, that is, that we have not discussed visualization design in this paper.
@@
 As our goal was not to examine the elements of visualization design experimentally, controlling those elements adheres to good experimental design principles.*
+
 8. **Does the revision have an explicit related works section?**    
 **(refer to the last concern of R3)**
+
 *An explicit related works section has been added at the conclusion of Sec. 1.*
+
 9. **Does the revision handle all the minor issues raised by reviewers?**
+
+We have detailed responses to all minor issues below.
 
@@
 - Equation (2) states the same as equation (1).    
-*The repetition has been removed due to space constraints.*
+*The repetition has been removed.*
 - In section 3, you mention one exception in relation to the participants’ age, please be more specific.    
@@
 - In section 3, I suppose you mean that the demographics are representative for the student population and not for the whole university incl staff.    
-*Indeed, though the wording has been modified for clarity in the revised draft.*
+*Indeed, we have modified the wording for clarity in the revision.*
 - With regard to your assessment of the correlation in Fig 6. I would discriminate more between the individual results rather than stating that they are all highly correlated.    

### 2015-06-26 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
added a few things that need to be done, but not tonight ...

@@
 Each of the sets of lineups shown, were part of studies to compare the designs.
-We have cited the corresponding papers accordingly.
-*
+We have cited the corresponding papers accordingly.*
 *While there are differences in the correlation of scores on lineup sets 1, 2, and 3 with the cognitive tests, this does not provide any additional information about whether designs in lineup set 1 are superior to those in lineup set 2 and lineup set 3, and in fact, the sets of lineups are not comparable in this way.
@@
 - Equation (2) states the same as equation (1).    
-*The repetition has been removed.*
+*Equation (2) is a re-formulation of equation (1) just for the convenience of the reader; it has now been stated as such.*
 - In section 3, you mention one exception in relation to the participants’ age, please be more specific.    

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
added a few things that need to be done, but not tonight ...

@@
 
+\hhnote{this next paragraph is a bit confusing ...
+it seems to suggest that we are using lineups to distinguish between different designs after all.
+Do we need this paragraph at all?
+I think it would flow directly from the sentence before to the paragraph after this one.}
 Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
 This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic; this examination utilizes data aggregated over all participants who evaluated a specific lineup image.
+
+\hhnote{XXX this procedural description needs a few more pieces: order of different tests (pilot study showed that order did not matter), 
+total experimental time, experimental setup: in a lab with paper copies;  }
 
@@
 Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set.
+\hhnote{citations?}
 \begin{figure*}[htbp]\centering

### 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
added a few things that need to be done, but not tonight ...

@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 14:47:21 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 00:07:56 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 14:47:21 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 00:07:56 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 14:47:21 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 00:07:56 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Fri Jun 26 14:47:22 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 00:07:57 2015
 \begin{table}[ht]

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Additional references, etc.

--- 2015-06-26 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
@@
 
-Statistical lineup experiments may include the same set of data and distractor plots, visualized using different plot types.
-This allows researchers to examine whether, for instance, a box plot is more effective than a dot plot at presenting a certain distributional characteristic; this examination utilizes data aggregated over all participants who evaluated a specific lineup image.
+% \hhnote{XXX this procedural description needs a few more pieces: order of different tests (pilot study showed that order did not matter), 
+% total experimental time, experimental setup: in a lab with paper copies;  
+% 
+% This information is actually present in the "Experimental procedure" section below. This section is just discussing lineups.}
 
@@
 
-Each of the three sets of 20 lineups was taken from previous experiments examining which type of plot most effectively conveyed important characteristics of the underlying data set. 
+Each of the three sets of 20 lineups was taken from previous experiments \cite{loy:2015, roy2011visual} examining which type of plot most effectively conveyed important characteristics of the underlying data set.
+%\hhnote{citations?}
 \begin{figure*}[htbp]\centering
@@
 
+\paragraph{Experimental Procedure}
 All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
@@
 The duration of testing for each participant was approximately 1 hour.
-All test documents (with redacted copyrighted material) are available as supplemental materials.
+All test documents (with copyrighted material redacted) are available as supplemental materials.
 

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Additional references, etc.

@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 00:07:56 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 15:16:35 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 00:07:56 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 15:16:36 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 00:07:56 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 15:16:36 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 00:07:57 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 15:16:36 2015
 \begin{table}[ht]

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
study procedure, start

@@
 
-% \hhnote{XXX this procedural description needs a few more pieces: order of different tests (pilot study showed that order did not matter), 
-% total experimental time, experimental setup: in a lab with paper copies;  
-% 
-% This information is actually present in the "Experimental procedure" section below. This section is just discussing lineups.}
-
-In this study, we utilize three blocks of statistical lineups; each block consists of 5-7 datasets, each displayed using three or four types of plots (examples of the visualization types presented in each block of lineups are shown in figure \ref{fig:lineups}).
-For example, in lineup set 1, the same five datasets are displayed as a dot plot, density plot, box plot, and histogram.
+In this study, we utilize three blocks of statistical lineups; each of these blocks consists of 5-7 datasets, that are displayed in three or four types of plots (examples of the visualization types presented in each block of lineups are shown in figure \ref{fig:lineups}).
+For example, in the lineups of set 1, the same five datasets are displayed as a dot plot, density plot, box plot, and histogram.
 Each block of lineups are a subset of stimuli used in experiments designed to compare the visual significance of different plot types, however, our purpose in this study is not to determine the efficacy of different types of plots.
@@
 \paragraph{Experimental Procedure}
-All tasks (cognitive and lineups) were provided to participants on paper, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; internal pilot tests did not indicate an effect of task order.
+All tasks (cognitive and lineups) were provided to participants on paper \newtext{in a supervised lab setting}, to conform to the procedure used in~\cite{ekstrom1976manual}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; results from an internal pilot tests with a counterbalanced design did not indicate an effect of task order, nor was there any indication in the results of an interaction/learning effect from one task to the next.
 
@@
 
-In addition to these tests, participants were asked to complete a questionnaire which includes questions about colorblindness, mathematical background, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
+In addition to these tests, participants completed a questionnaire on personal background and demographics, including questions about colorblindness, mathematical training, self-perceived verbal/mathematical/artistic skills, time spent playing video games, and undergraduate major.
 These questions are designed to assess different factors which may influence a participant's skill at reading graphs and performing spatial tasks.
 
-After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task.
+After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task \newtext{(which led us to interweave lineup evaluations)}.
 The duration of testing for each participant was approximately 1 hour.
-All test documents (with copyrighted material redacted) are available as supplemental materials.
+\newtext{None of the participants who showed up for their scheduled session asked to leave early.} A set of test documents (with copyrighted material redacted) is available as supplemental materials.
 

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Compiled files
study procedure, start

@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Sat Jun 27 15:16:35 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 22:46:07 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Sat Jun 27 15:16:36 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 22:46:08 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.2.1 by xtable 1.7-1 package
-% Sat Jun 27 15:16:36 2015
+% latex table generated in R 3.1.1 by xtable 1.7-3 package
+% Sat Jun 27 22:46:08 2015
 \begin{table}[ht]

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
study procedure a bit later

@@
 All tasks (cognitive and lineups) were provided to participants on paper \newtext{in a supervised lab setting}, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; results from an internal pilot tests with a counterbalanced design did not indicate an effect of task order, nor was there any indication in the results of an interaction/learning effect from one task to the next.
-
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; an internal pilot study did not indicate an effect of task order, \newtext{nor did the pilot results suggest an interaction/learning effect from one task to the next.}
+
+Between cognitive tasks, participants were asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
 Participants were given 5 minutes to complete each block of 20 lineups.

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
study procedure a bit later

@@
 All tasks (cognitive and lineups) were provided to participants on paper \newtext{in a supervised lab setting}, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; results from an internal pilot tests with a counterbalanced design did not indicate an effect of task order, nor was there any indication in the results of an interaction/learning effect from one task to the next.
+The order of tasks was the same for each participant due to the constraints of in-person testing; an internal pilot study did not indicate an effect of task order, \newtext{nor did the pilot results suggest an interaction/learning effect from one task to the next.}
 
-Between cognitive tasks, participants were also asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
+Between cognitive tasks, participants were asked to complete three blocks of 20 lineups each, assembled from previous studies~\cite{hofmann2012graphical,majumder2013validation}.
 Participants were given 5 minutes to complete each block of 20 lineups.
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:46:07 2015
+% Sat Jun 27 22:48:32 2015
 \begin{table}[htb]
@@
 % latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:46:08 2015
+% Sat Jun 27 22:48:33 2015
 \begin{table}[ht]

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Removing newtext, merge conflict

@@
 %\hhnote{citations?}
+
 \begin{figure*}[htbp]\centering
@@
 \paragraph{Experimental Procedure}
-All tasks (cognitive and lineups) were provided to participants on paper \newtext{in a supervised lab setting}, to conform to the procedure used in~\cite{ekstrom1976manual}.
-The order of tasks was the same for each participant due to the constraints of in-person testing; an internal pilot study did not indicate an effect of task order, \newtext{nor did the pilot results suggest an interaction/learning effect from one task to the next.}
+All tasks (cognitive and lineups) were provided to participants on paper in a supervised lab setting, to conform to the procedure used in~\cite{ekstrom1976manual}.
+The order of tasks was the same for each participant due to the constraints of in-person testing; an internal pilot study did not indicate an effect of task order, nor did the pilot results suggest an interaction/learning effect from one task to the next.
 
@@
 
-After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task \newtext{(which led us to interweave lineup evaluations)}.
+After each test or block of lineups, participants were given 1-2 minutes to rest and recover; the order of tests was designed to prevent fatigue on a single task (which led us to interweave lineup evaluations).
 The duration of testing for each participant was approximately 1 hour.
-\newtext{None of the participants who showed up for their scheduled session asked to leave early.} A set of test documents (with copyrighted material redacted) is available as supplemental materials.
+None of the participants who showed up for their scheduled session asked to leave early.
+A complete set of test documents (with copyrighted material redacted) is available as supplemental materials.
 

### 2015-06-27 Paper/InfoVis2015Revision/TestingVisualAptitude.tex
Removing newtext, merge conflict

@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:48:32 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 16:26:37 2015
 \begin{table}[htb]
@@
 Table~\ref{tab:PCAvariance5} shows the importance of each principal component; from the distribution of the variance components, we see that the lineup test spans an additional dimension within the space of the four established tests.
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:48:33 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 16:26:37 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:48:33 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 16:26:37 2015
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.1.1 by xtable 1.7-3 package
-% Sat Jun 27 22:48:33 2015
+% latex table generated in R 3.2.1 by xtable 1.7-1 package
+% Sat Jun 27 16:26:38 2015
 \begin{table}[ht]

### 2015-07-20 Paper/InfoVis2015Revision/TestingVisualAptitude.Rnw
Updated cls file

@@
 sanitize <- function(x){
-  str_replace(gsub("_", ".", x, fixed=T), "Pr\\((.{1,})\\)", "Pr$(\\1)$")
+  x %>% 
+    gsub(pattern="_", replacement=".", fixed=T) %>%
+    gsub(pattern="Pr\\((.*)\\)", replacement="Pr$(\\1)$")
 }
@@
 <<ResultsScaledScores, echo=F, include=F, fig.width=4.5, fig.height=4, warning=F, message=F>>=
-loadfonts()
+#loadfonts()
 score.summary <- melt(ans.summary, measure.vars=21:25, value.name = "Score", variable.name = "Test")
@@
   #coord_flip()
-if (file.exists("./Figure/fig-ResultsScaledScores-1.pdf")) 
-  embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
+# if (file.exists("./Figure/fig-ResultsScaledScores-1.pdf")) 
+  #embedFonts("./Figure/fig-ResultsScaledScores-1.pdf")
 @
@@
 <<VisReasoningCategorical, echo=F, include=F, eval=T, fig.width=6,fig.height=5, message=F, warning=F, fig.pos='ht'>>=
-loadfonts()
+#loadfonts()
 lineup.summary.categorical <- rbind(lineup.summary.categorical, data.frame(id=1, lineup=NA, variable="verbal_skills", value=1))
@@
   theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
-if (file.exists("./Figure/fig-VisReasoningCategorical-1.pdf"))
-  embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
+# if (file.exists("./Figure/fig-VisReasoningCategorical-1.pdf"))
+  #embedFonts("./Figure/fig-VisReasoningCategorical-1.pdf")
 @
@@
 }
-loadfonts()
+#loadfonts()
 require(grid)
@@
 rm(list=c("data", "all", "densities"))
-if (file.exists("./Figure/fig-VisReasoningSPM-1.pdf")) embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
+# if (file.exists("./Figure/fig-VisReasoningSPM-1.pdf")) 
+  #embedFonts("./Figure/fig-VisReasoningSPM-1.pdf")
 @
@@
 )
-loadfonts()
+#loadfonts()
 ggplot()+
@@
   theme(axis.title=element_blank())
-if (file.exists("./Figure/fig-LineupTypeSPM-1.pdf")) embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
+# if (file.exists("./Figure/fig-LineupTypeSPM-1.pdf")) 
+  # embedFonts("./Figure/fig-LineupTypeSPM-1.pdf")
 @
@@
 <<lineupblockpca, echo=F, include=F, warning=F, message=F, fig.width=6, fig.height=4>>=
-
 lineup.block.pca <- prcomp(lineup.section.summary[,c("lineup_1", "lineup_2", "lineup_3", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
@@
 rotation <- merge(rotation, melt(importance, variable.name="Var2", value.name="importance"))
+# rotation$Var3 <- str_replace(rotation$Var1, pattern="lineup_\\d", replacement="Lineup-Core")
+# rotation$Var3 <- factor(rotation$Var3, 
+#                         levels=c("vis_search", "folding", "card_rot", "fig_class", "Lineup-Core"),
+#                         labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class", "Lineup Core"))
+
 rotation$Var1 <- factor(rotation$Var1, 
-                        levels=c("vis_search", "folding", "card_rot", "fig_class", "Lineup-Core"),
-                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Core"))
-
-#rotation$Var1 <- factor(rotation$Var1, 
-#                        levels=c("vis_search", "folding", "card_rot", "fig_class", #"lineup_3", "lineup_2", "lineup_1"),
-#                        labels=c("Visual Search", "Paper Folding", "Card Rotation", #"Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
+                       levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_3", "lineup_2", "lineup_1"),
+                       labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 3", "Lineup Test 2", "Lineup Test 1"))
 library(grid) # needed for arrow function
-loadfonts()
 qplot(x=0,  y=Var1, xend=value*importance, yend=Var1, geom="segment", data=rotation, arrow=arrow(length=unit(0.1, "cm"))) + facet_wrap(~Var2, ncol=4) + xlab("Influence") + ggtitle("Lineup Types and Visual Aptitude") + ylab("") + theme_bw()
-if (file.exists("./Figure/fig-lineupblockpca-1.pdf")) embedFonts("./Figure/fig-lineupblockpca-1.pdf")
+# if (file.exists("./Figure/fig-lineupblockpca-1.pdf")) 
+  # embedFonts("./Figure/fig-lineupblockpca-1.pdf")
 @

### 2015-07-20 Paper/InfoVis2015Revision/appendix.Rnw
Updated cls file

@@
 lineup.type.sum2$testnum <- paste0("Set ", lineup.type.sum2$testnum)
-loadfonts()
+# loadfonts()
 qplot(data=subset(lineup.type.sum2, !is.na(plot.type)& testtype=="lineup"), 
@@
         axis.ticks.y=element_blank(), axis.text.y=element_blank())
-embedFonts("./Figure/fig-lineup-types-1.pdf")
+# embedFonts("./Figure/fig-lineup-types-1.pdf")
 @
@@
 cor.labels <- ddply(lineup.type.sum3, .(lineup.type, testtype), summarize, x=min(scaled.testscore), y=max(lineup.pct), label=paste0("cor = ", round(cor(testscore, lineup.pct), 2)))
-loadfonts()
+# loadfonts()
 qplot(data=subset(lineup.type.sum3, testtype!="vis_search"), 
